### 2014-08-13 introduction-omer.tex
introduction+moti updated by lpxz

\section{Introduction}\label{Se:introduction}

In a multithreaded application, a data race occurs when two (or more) concurrent threads access a mutual memory location with neither ordering constraints nor guarding synchronization. This form of unstructured access to shared memory is often unintended, indicating a bug in the code.

A main challenge in detection of data races is that they are elusive, arising only in certain concurrent schedules. Because thread interleavings are decided nondeterministically by the scheduler, it is hard to anticipate and recreate data races. Indeed, in practice data races often manifest rarely, leading e.g. to situations where the bug is missed during testing but surfaces up during production runs (due to its rarity combined with differences in hardware, usage scenario, etc).

\paragraph{Existing Approaches} Various approaches have been proposed to detect data races, which strike different tradeoffs between (i) \emph{soundness}, whereby no false races are reported to the user, and (ii) \emph{completeness}, whereby no true races are missed. Sound techniques are largely based on, or refine, the \emph{happens-before} (\emph{HB}) relation \cite{XXX}. Unsound techniques that opt for more complete detection are based on lockset analysis, as popularized by the Eraser tool \cite{XXX}.

Driven by the requirement for high usability, which is governed by low to zero user tolerance to false warnings \cite{XXX}, we focus on fully sound detection techniques. A recent result in this space is Smagardakis et al.'s \emph{causally-precedes} (\emph{CP}) relation, which soundly relaxes HB edges between critical sections that do not exhibit conflicting accesses. A refinement of the CP model, proposed by Huang et al. \cite{XXX}, is to factor in control-flow information: Inter-thread CP edges, such that statements succeeding the target statement are not control dependent on it, are soundly discarded.

Both of these techniques, as well as preceding HB approaches, ensure soundness via notions of \emph{dependence}. As long as certain data and/or control dependencies are respected, reported data races are guaranteed to be real. While these are useful steps to detect more data races, the reported warnings are restricted to (i) the given execution trace and (ii) reorderings that preserve essential (data and/or control) dependencies.

\paragraph{Motivation} As we illustrate in Figure \ref{Fi:illustrative}, taken from the benchmark suite of Huang et al. \cite{XXX}, these restrictions can potentially block real races from the analysis' view. In 
the {\tt main} method of class {\tt Example}, a new thread {\tt t} is spawned. Subsequently, and in parallel, variable {\tt y} is both assigned by the main thread and queried by the spawned thread, which checks whether {\tt y}'s value is at least $3$.

\begin{figure}
\begin{lstlisting}
public class Example {
  static int x = -1;
  static int y = -1;

public static void main(String[] args) {
  MyThread t = new MyThread();
  t.start();
  y = 3;
  x = 0;
  y = 5; }

static class MyThread extends Thread {
  public void run() {
    if (y >= 3) {
      // Race: is it possible to print -1? 
      System.out.println(x);
    } else {
      System.out.println(x + "else");	
    } } } }
\end{lstlisting}
\caption{\label{Fi:illustrative}Benchmark from Huang et al.'s suite \cite{XXX} (with slight revisions) that exhibits two potential data races}
\end{figure}

There are three possibilities: Either (i) the test at line 14 is reached before {\tt y} is first assigned within {\tt main}, and so the initial value of {\tt y} of $-1$ is used and the test evaluates to {\tt false}, or (ii) the test occurs after the first assignment and succeeds for {\tt y $\equiv$ 3}, or (iii) the test takes place after the second assignment, again succeeding bu this time for {\tt y $\equiv$ 5}. In specific, because the test on {\tt y} may either succeed or fail, both branches of the conditional block within the {\tt run} method of inner class {\tt MyThread} may execute. Similarly to the analysis for {\tt y}, we notice that the first {\tt println} statement can either print the initial value of {\tt x} ($-1$) or its value after the assignment inside {\tt main} ($0$). To conclude, there are two potential data races in the code.

The CP model is too restrictive to discover both of these races, even when control-flow information is factored into the analysis. To demonstrate this, let us consider trace $\left[ 6,7,8,14,16,9,10 \right]$. That is, the first context switch occurs after {\tt y} is first assigned, method {\tt run} then executes to completion, and subsequently {\tt main} finishes executing. There is genuine control dependence between the statements at lines 8 and 14, which disables any form of reordering according to CP, the result being that both races are missed. Indeed, if we restrict the analysis to the given trace and dependencies that arise within it, then the races in Figure \ref{Fi:illustrative} remain out of reach.

\paragraph{Our Approach} The main observation underlying our approach is that accounting for dependencies alone is insufficient. The departure point of this paper is in modeling the \emph{values} of shared variables explicitly, rather than stopping at the level of \emph{dependencies} within and between threads. Importantly, this enables us --- in certain cases ---  to step outside the boundaries of the given trace and explore additional program behaviors.

Applied to the example in Figure \ref{Fi:illustrative}, our approach permits us to exploit two important facts:
\begin{itemize}
	\item First, the test at line 14 is satisfied both after the first assignment and after the second assignment to {\tt y}. In both cases $y \geq 3$. Making this observation, which refers directly to the values assumed by {\tt y} and their use in control-flow decisions, lets us account for the assignment {\tt x=0}, thereby uncovering the race over {\tt x}.
	\item Second, and even more significantly, explicit modeling of values allows us to simulate a context switch before {\tt y} is first assigned inside {\tt main}. In this scenario, {\tt y} has the value $-1$, which leads to a negative evaluation of the test at line 14. Note that we are no longer reordering events within the same trace, but have switched to another trace that is still guaranteed to be feasible since we are able to precisely maintain the program's state and behavior along all transitions (including those within the {\tt else} branch, which was not executed in the input trace).
\end{itemize}
To summarize, relaxing the ability to reoder events --- to the point of simulating new execution traces beyond the input trace --- is the key to improving detection capabilities. The principal idea undetlying our approach is to model the shared state explicitly, rather than merely recording dependencies that arise within the given trace. 
As we demonstrate experimentally in Section \ref{Se:evaluation}, this form of refinement has a dramatic effect on detection capabilities compared to previous approaches. ** MORE EXPERIMENTAL DATA NEEDED **

At the technical level, we represent the shared state, and ensure the feasibility of reorderings and execution along new branches, via a novel encoding of relevant events as a constraint system. ** MORE DETAILS NEEDED **

\paragraph{Contributions} The principal contributions of this paper are the following:

### 2014-08-13 introduction.tex
introduction+moti updated by lpxz

\section{Introduction}\label{Se:introduction}
A race occurs when two concurrent threads access the same variable and when (i) at least one access is a write, (ii) the threads take no explicit mechanism to prevent the accesses from being simultaneous~\cite{eraser}. More intuitively, a race occurs when two accesses of the same variable (one is write) can happen in parallel.  Race detection is a fundamental analysis for concurrent programs, which underlies  many client applications including the atomicity/order violation detection~\cite{learning,avio,ordervio}, the dataflow analysis~\cite{dataflow} and the record and replay~\cite{leap,replay2012}.

%fasttrack pldi paper claims happens-before analysis never reports false positives. Check it,
Over the last decade, researchers have spent tremendous efforts on race detection and have gained relative success. 
The initial efforts center around the {\em synchronization primitives} including the locks and the scheduling control such as wait/notify and start/join. Lockset analysis, which considers only locks, produces many infeasible races that contradict the schedule control. Happens-before analysis reasons about all the primitives but imposes overly restrictive lock order constraints, i.e., two lock regions with the same lock should follow the execution order as observed. The following work further relaxes the lock constraints: Hybrid analysis allows the two lock regions to be ordered in either way; {\sf Universal Causal Graph (UCG)}~\cite{ucg} allows them to be ordered in either way only when this relaxation does not contradict the schedule control.

However, these seminal work misses the important {\em value constraints}, leading to false positives. Consider the example in Figure~\ref{fig:running}, 
where {\tt x} and {\tt y} are shared variables and no synchronizations are employed. The work that considers only synchronization constraints would report a race between line 2 and line 5 on x, which appear in the observed run. However, the race is a false alarm because line 5 can be executed only after line 4 reads the value of {\tt y} from line 3 (after line 2).

\begin{figure}
\centering
\includegraphics[width=0.38\textwidth]{figs/Visio-running.pdf}
\caption{Running Example}
\label{fig:running}
\end{figure}

%I　really think Yannis' original approach is incorrect as it only preserves the dependence imposed by lock regions, which is insufficient. Check it or Hide it?
The false positives greatly reduce the usability of the detection tool as programmers need to spend great efforts  in investigating each of them. 
Recently, the predictive analysis attracts increasing research attention for its soundness guarantee, i.e., no false alarms. Smagardakis et al. pioneered by proposing first of such work. Starting from an original trace, their sound predictive analysis ``predicts'' a new trace by relaxing the execution order 
 while preserving the inter-thread dependence constraints and the synchronization constraints.  The inter-thread dependence constraints require the predicted trace and the original trace to exhibit the same set of dependences between conflicting accesses. Preservation of them leads to the soundness 
 as every value remains the same. Huang et al.~\cite{} further optimizes by preserving only the inter-thread dependences that could affect the branch decisions in the following executions.

Although showing promises,  existing sound predictive analyses  are conservative as they satisfy the value constraints imposed by the branch conditions by preserving the inter-thread dependences and reading the same values at the branches. In this work, we propose the relaxed sound predictive analysis, which relaxes the value constraints to find more races without sacrificing the soundness. Our relaxation follows two key insights. First, we satisfy the value constraints computationally, rather than recording and preserving the dependences. The values read at the branches are allowed to change in the prediction as long as the change does not affect the branch decisions.  We realize the value constraints through  the satisfiability checker. Second, even if the change of the read values affects the branch decisions, we can still explore the un-executed branch  if it is ``simple'' enough. The underlying rationale is, when the executed branch contains the shared accesses, the sibling (un-executed) branch is likely to access the same shared variable. Reasoning about the un-executed branch is enabled by flexibility of  our computation-based approach, i.e., we just need to reverse the branch condition and resend it to the checker, while existing approaches cannot support such reasoning due to the lack of the computation power and the conservativeness of perservation-based approaches.

We highlight our approach with the example~\footnote{We omit the synchronization primitives purposely for simplification.} in Figure~\ref{fig:running}. Suppose the branch at line 4 changes to {\tt if(y>=3)}. In the original run, line 4 reads from the value defined at line 3. Therefore, Huang et al.'s approach preserves the same dependence in the predicted run, and therefore misses the race on x between line 2 and line 5. In contrast, our approach reports two more real races on x, (2,5) and (2,7).

We first demonstrate how we find the race (2,5). As shown in Figure~\ref{fig:demo}, we first encode trace in the SSA form, where each variable is defined exactly once (line number as the superscript). Besides, we introduce a new symbol, e.g., $R^4_{y}$,  to denote each shared read, e.g., the read of $y$ at   
 line 4. We also list the constraints for a valid predicted run. Race constraints specify line 2 and line 5 can run in parallel. Here $O_2$ and $O_5$ are  variables introduced to denote the orders. Program order constraints specify the intra-thread execution order. The rest  constraints model the value constraints: path condition constraints specify the branch decision for the branches preceding the potential racy event line 5; variable definition 
 constraints specify the intra-thread value flow due to each assignment; Inter-thread constraints specify the inter-thread value flow which is conditioned on the scheduling constraints. For example, $R^4_{y}=y^1 \wedge O_1< O_4 <O_3$ specifies that line 4 reads {\tt y} defined at line 1 if and only if line 4   
  happens after line 1 and before line 3.  Combined together, the constraints are sent to off-the-shelf solver such as z3 or yices, which then returns a solution: line 4 reads $y$ ($y$=3) defined at line 1, which satisfies the branch condition.

Finding the race (2,7) requires a slightly different set of constraints. Specifically, the race constraint becomes $O_2=O_7$, and the path condition constraint becomes $R^4_y<3$. The checker also returns a solution for these constraints: line 4 reads $y$ ($y$=0) from the initial value, which steers the execution towards the unexecuted path at the branch line 4.

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figs/Visio-demo.pdf}
\caption{Overview}
\label{fig:demo}
\end{figure}

\paragraph{Contributions} The principal contributions of this paper are the following:

### 2014-08-13 x.sh

@@ -1,0 +1,5 @@
+pdflatex -interaction nonstopmode paper
+bibtex paper
+pdflatex -interaction nonstopmode paper
+pdflatex -interaction nonstopmode paper
+evince paper.pdf

### 2014-08-16 overview.tex
revised intro and tech overview

\section{Technical Overview}\label{Se:techoverview}

In this section, we walk the reader through a detailed technical description of our approach based on the example in Figure \ref{fig:running}. 
%
As input, we assume (i) a program $P$ as well as (ii) a dynamic execution trace of $P$ in Static Single Assignment (SSA) form, such that every variable is defined exactly once.

We begin with an explanation of the encoding process:

\paragraph{Program Order} The first set of constraints reflects ordering constraints between program statements. We use the symbol $O_i$ to denote the $i$th program statement, which yields the following formula for Figure \ref{fig:running}:
$$
	O_1 < O_2 < O_3 \bigwedge O_4 < O_5 \bigwedge O_4 < O_7
$$
That is, the first 3 statements are totally ordered, and the {\tt if} statement executes before the body (either $O_5$ or $O_7$).

\paragraph{Variable Definitions} The next set of constraints, denoted ${\tt z}^i=k$, capture variable definitions: Variable {\tt z} is assigned value $k$ at statement $i$. For our running example, we obtain:
$$
	{\tt x}^0 = 0 \bigwedge {\tt y}^0 = 0 \bigwedge {\tt y}^1 = 3 \bigwedge 
			{\tt x}^2 = 1 \bigwedge {\tt y}^3 = 5
$$
As an example, ${\tt y}^3 = 5$ denotes that the value assigned to variable {\tt y} 
at line {\tt 3} is $5$.

\paragraph{Thread Interference} To express inter-thread flow constraints, we utilize expressions of the form $R_{\tt z}^i={\tt z}^j$, which denotes that line $i$ reads variable {\tt z}, and the definition it reads comes from line $j$. The resulting formula for our example is
$$
\begin{array}{lcl}
	& & (R_{\tt y}^4={\tt y}^0 \wedge O_4 < O_1)  \\
& \bigvee &
	(R_{\tt y}^4={\tt y}^1 \wedge O_1 < O_4 < O_3) \\
& \bigvee &
	(R_{\tt y}^4={\tt y}^3 \wedge O_3 < O_4)
\end{array}
$$    
Notice, importantly, that the formula combines flow constraints with order constraints, which are essential to determinize the value read at a given statement. As an example, 
$(R_{\tt y}^4={\tt y}^1 \wedge O_1 < O_4 < O_3)$ means that the value of {\tt y} read at 
line {\tt 4} is that set at line {\tt 1} assuming an execution order whereby the first statement executes followed by the fourth then third statements.

\paragraph{Path Conditions} To preserve path conditions while potentially permitting dependence-violating reordering (e.g., a context switch after line {\tt 1} in Figure \ref{fig:running}), we model explicitly the condition. For the running example (ignoring the statements in red), this yields:
$$
	(R_{\tt y}^4 > 2
$$
That is, the value of variable {\tt y} read at line {\tt 4} is greater than $2$. Indeed, this constraint is satisfied by the assignments to {\tt y} both at line {\tt 1} and at line {\tt 3}.

\paragraph{Race Condition} The final constraint, forcing the check whether a particular race is feasible, is to demand that two conflicting statements occur at the same time. For the potential race between lines {\tt 2} and {\tt 5}, we obtain:
$$
	O_2 = O_5
$$ 
This asserts that both statements occur simultaneously, which --- together with the other constraints --- guarantees the feasibility of the predicted race if a solution is found for the overall constraint system.

\paragraph{Unexplored Branches} Beyond the encoding steps so far, which focus on the given trace, we can often encode constraints along unexplored branches. In our running example, this is essential to discover the race between lines {\tt 2} and {\tt 7}.

The conflicting accesses in this case, determined based on static analysis of $P$, are expressed as $O_2=O_7$. In addition, we negate the path condition, thereby obtaining $R_{\tt y}^4 \leq 2$ in place of $R_{\tt y}^4 \leq 2$.

\paragraph{Constraint Solving} Having conjoined the formuals from the different encoding steps into (i) a global representation of all feasibility constraints (path, ordering, assignment and other constraints) and (ii) the requirement for a given race to occur (expressed as simultaneous execution of the conflicting accesses), we discharge the resulting formula to an off-the-shelf constraint solver. If successful, the solver returns a solution for the specified constraints, and in particular, we obtain a feasible trace that gives rise to the race at hand.

### 2014-08-17 impl.tex
tech+impl

\section{Implementation}\label{sec:impl}
For array access event $arr[i]=x$, we treat them as the field access event $arr.i=x$, where the index $i$ is treated as a field.

We distinguish the accesses of the shared fields and local fields. To collect the shared fields, we preprocess the trace and map each field to a set of accessing threads. If a field is written by two or more threads, it is shared. If a field is written by one thread and read by other threads, it is shared. Otherwise, it is a local field.

### 2014-08-17 tech.tex
tech+impl

\section{Basics}~\label{sec:basic}

\subsection{Trace Terminology}
Our analysis starts with a trace $\tau$,  a sequence of events, $e_0, e_1, \dots, e_n$.  
There are three types of events in general.
\begin{itemize}
\item the shared access, which includes the read and write of the shared fields, e.g., $o.f$=$x$ and $x$=$o.f$.
\item the local access, which includes the access of the local variable, e.g., $x=y+z$ or $v.f=x$ (where $v$ is a thread-local object).
\item the branch event, which evaluates the branch condition to true/false, e.g., $x>3$.
\item the synchronization event, which includes start/join, wait/notify, lock/unlock events, e.g., $lock(o)$
\end{itemize}

Each event $e_i \in \Sigma$ is a tuple, $<t, id, a, v, ins>$, where $t\in \mathcal{T}$ denotes the thread generating the event, $id\in \mathcal{ID}$ denotes the unique integer assigned to the event, $a\mathcal{A}$ denotes the address of the object or field (if any) accessed in the event,  $v\in \mathcal{V}$ denotes the value of the definition (if any) in the event, and $ins\in \mathcal{INS}$ denotes the three-address instruction generating the event.  Specifically, the address of the object $o$ is denoted as $id(o)\in \mathcal{ID}$, which is a string value representing $o$ uniquely, the address of the static field $f$ is  $id(f)$ and the address of the instance object field $o.f$ is  $id(o)\_id(f)$. Besides, as the event is derived by instrumenting the three-address code and monitoring the instrumented execution. Therefore, each event can involve at most three operands.

The trace supports its standard operations as follows.
\begin{itemize}
\item projection, e.g., $\tau|t$ returns~\footnote{This is the abbreviation for the complete form $\tau|Thread=t$ } the events from the thread $t$,  $\tau|a$ returns the event involving the address $a$.  
\item concatenation. $\tau'=\tau e$ represents the new trace by appending the event $e$ to $\tau$.
\item length $|\tau|$. 
\item selecting an element. $\tau[0]$ and $\tau[|\tau|-1]$ represents the first and last event in $\tau$.
\end{itemize}

%The modeling of synchronization event is standard and explained in existing work, so we focus on the rest two types of events in this paper.

In addition, we maintain auxiliary information as follows.
\begin{itemize}
\item $AT: \mathcal{A} \times \mathcal{T} \rightarrow \gamma$ is a function that returns a trace $\tau \in \gamma$ that contains only the accesses of the address $a \in \mathcal{A}$ by the thread $t\in \mathcal{T}$. Each trace $\tau$ in $\gamma$ is defined over the alphabet of events $\Sigma$, specifically, the trace is an empty trace $\epsilon$ or defined in this way:  $\forall 0\leq i\leq |\tau|,   \tau[i]\in \Sigma, and, \forall i\neq j, \tau[i]\neq \tau[j]$. 
\item $R: \mathcal{A} \rightarrow \gamma$ is a function that returns the read accesses of the address $a\in \mathcal{A}$.
\item $W: \mathcal{A} \rightarrow \gamma$ is a function that returns the write accesses of the address $a \in \mathcal{A}$.
\item $Sync: \mathcal{A} \rightarrow \gamma$ is a function that returns the synchronization events involving the address $a\in \mathcal{A}$. 
\end{itemize}

\subsection{Symbolic Trace}
To facilitate the symbolic analysis, we need to introduce symbols to represent the operands in each event. Symbols allow us to overcome the limitation of concrete dependences and allow us to explore more dependences symbolically.

%shared access only, local access only

Like other symbolic analysis~\cite{jeff,chao}, the symbolic trace should be in the SSA form, i.e., each variable is defined exactly once. This is because the constraint solver employed by the analysis does not allow the same variable to hold more than one value.  Besides, we introduce symbols to represent shared reads and shared writes which leave the inter-thread dependence between reads/writes undetermined.

\subsection{Method Calls}~\label{sec:basic}

\section{Time Window}~\label{sec:basic}

\section{Heap Invariant}~\label{sec:basic}

### 2014-08-17 introduction.tex

@@
 \section{Introduction}\label{Se:introduction}
-A race occurs when two concurrent threads access the same variable and when (i) at least one access is a write, (ii) the threads take no explicit mechanism to prevent the accesses from being simultaneous~\cite{eraser}.
-More intuitively, a race occurs when two accesses of the same variable (one is write) can happen in parallel.
-Race detection is a fundamental analysis for concurrent programs, which underlies  many client applications including the atomicity/order violation detection~\cite{learning,avio,ordervio}, the dataflow analysis~\cite{dataflow} and the record and replay~\cite{leap,replay2012}.
 
-%fasttrack pldi paper claims happens-before analysis never reports false positives. Check it,
-Over the last decade, researchers have spent tremendous efforts on race detection and have gained relative success. 
-The initial efforts center around the {\em synchronization primitives} including the locks and the scheduling control such as wait/notify and start/join.
-Lockset analysis, which considers only locks, produces many infeasible races that contradict the schedule control.
-Happens-before analysis reasons about all the primitives but imposes overly restrictive lock order constraints, i.e., two lock regions with the same lock should follow the execution order as observed.
-The following work further relaxes the lock constraints: Hybrid analysis allows the two lock regions to be ordered in either way; {\sf Universal Causal Graph (UCG)}~\cite{ucg} allows them to be ordered in either way only when this relaxation does not contradict the schedule control.
+In a multithreaded application, a data race occurs when two concurrent threads access the same memory location, such that (i) at least one of the accesses is write access, and (ii) no explicit mechanism is enforced to prevent the threads from simultaneous access to the location \cite{eraser}.
+Race detection is important not only because races often reveal bugs in the implementation, but also as the basis of other techniques and analyses like atomicity checking \cite{learning,avio,ordervio}, data-flow analysis \cite{dataflow} and record/replay \cite{leap,replay2012}.
 
-However, these seminal work misses the important {\em value constraints}, leading to false positives. Consider the example in Figure~\ref{fig:running}, 
-where {\tt x} and {\tt y} are shared variables and no synchronizations are employed.
-The work that considers only synchronization constraints would report a race between line 2 and line 5 on x, which appear in the observed run.
-However, the race is a false alarm because line 5 can be executed only after line 4 reads the value of {\tt y} from line 3 (after line 2).
+In this paper, we focus on \emph{sound} race detection, whereby reported races are guaranteed to be real.
+This is the key requirement for adoption of the tool by developers \cite{XXX}.
+Ensuring soundness is a difficult challenge, which mandates dynamic forms of analysis.
+Indeed, extensive research has been carried out on dynamic race detection.
+The general goal has been to derive constraints from a given execution trace on event reordering, and check for the remaining reorderings whether they disclose data races.
+The key question then becomes about constraint extraction.
+
+%Fasttrack PLDI paper claims happens-before analysis never reports false positives. Check.
+\paragraph{Existing Approaches} Initial attempts to address this challenge focused on built-in synchronization primitives.
+These include locks as well as the wait/notify and start/join scheduling controls.
+Notable among these efforts is \emph{lockset} analysis, which considers only locks.
+Because the derived constraints are partial, permitting certain infeasible event reorderings, lockset analysis cannot guarantee soundness.
+
+A different tradeoff is struck by the \emph{happens-before (HB)} approach.
+In this style of analysis, all synchronization primitives are accounted for, though reordering constraints are conservative.
+As an example, HB inhibits reordering of two synchronized blocks governed by the same lock.
+Recently there have been successful attempts to relax HB constraints.
+Among these are hybrid analysis, which permits both orderings of lock-synchronized blocks, and the \emph{Universal Causal Graph (UCG)}~\cite{ucg} representation, which also enables both orderings but only if these are consistent with wait/notify- and start/join-induced constraints.
+
+Unfortunately, even full consideration of synchronization constructs is insufficient.
+Branching decisions must also be respected to ensure the feasibility of a reported race.
+As an example, we refer to the trace in Figure \ref{fig:running}.
+(For now, ignore the statements in red.) This trace is free of any explicit synchronization statements, and so a race between the assignment to variable {\tt x} at line {\tt 2} and the use of {\tt x} at line {\tt 5} may appear possible.
+This race is, however, spurious, since the use of {\tt x} at line {\tt 5} is governed by positive evaluation of the test at line {\tt 4}, which is conditioned on the assignment at line {\tt 3}.
+Hence, line {\tt 2} necessarily preceds line {\tt 5}.
 
@@
 \centering
-\includegraphics[width=0.38\textwidth]{figs/Visio-running.pdf}
-\caption{Running Example}
+\begin{tabular}{ll}
+\multicolumn{2}{c}{{\tt x = 0; y = 0; z = 3;}} \\
+\multicolumn{2}{c}{{\color{red} {\tt z = 2;}}} \\
+\hline
+\multicolumn{1}{c}{$T_1$} & \multicolumn{1}{c}{$T_2$} \\
+\hline
+{\tt 1: y = 3;} & \\
+{\tt 2: x = 1;} & \\
+{\tt 3: y = 5;} & \\
+& {\tt 4: if (y > z)} \\
+& {\tt 5:~~print(1/x);} \\	
+& {\color{red} {\tt 6: else}} \\
+& {\color{red} {\tt 7:~~print(2/x);}}
+\end{tabular}
+\caption{Example illustrating ordering constraints beyond synchronization primitives}
 \label{fig:running}
@@
 
-%I　really think Yannis' original approach is incorrect as it only preserves the dependence imposed by lock regions, which is insufficient. Check it or Hide it?
-The false positives greatly reduce the usability of the detection tool as programmers need to spend great efforts  in investigating each of them. 
-Recently, the predictive analysis attracts increasing research attention for its soundness guarantee, i.e., no false alarms.
+As an alternative, driven by the requirement for soundness, 
 Smagardakis et al.
-pioneered by proposing first of such work.
-Starting from an original trace, their sound predictive analysis ``predicts'' a new trace by relaxing the execution order
- while preserving the inter-thread dependence constraints and the synchronization constraints.
-The inter-thread dependence constraints require the predicted trace and the original trace to exhibit the same set of dependences between conflicting accesses.
-Preservation of them leads to the soundness
- as every value remains the same. Huang et al.~\cite{} further optimizes by preserving only the inter-thread dependences that could affect the branch decisions in the following executions.
+have recently proposed \emph{predictive analysis}.
+In this approach, both synchronization constraints and inter-thread dependencies are preserved, where inter-thread dependencies are respected by
+only allowing reorderings that leave the dependence structure exhibited by the original trace in tact.
+This ensures that the values of shared memory locations remain the same, which secures the soundness argument.
+As a soundness-preserving relaxation, Huang et al.
+\cite{XXX} permit reorderings as long as control dependencies are respected.
 
-Although showing promises,  existing sound predictive analyses  are conservative as they satisfy the value constraints imposed by the branch conditions by preserving the inter-thread dependences and reading the same values at the branches.
-In this work, we propose the relaxed sound predictive analysis, which relaxes the value constraints to find more races without sacrificing the soundness.
-Our relaxation follows two key insights.
-First, we satisfy the value constraints computationally, rather than recording and preserving the dependences.
-The values read at the branches are allowed to change in the prediction as long as the change does not affect the branch decisions.
-We realize the value constraints through  the satisfiability checker.
-Second, even if the change of the read values affects the branch decisions, we can still explore the un-executed branch  if it is ``simple'' enough.
-The underlying rationale is, when the executed branch contains the shared accesses, the sibling (un-executed) branch is likely to access the same shared variable.
-Reasoning about the un-executed branch is enabled by flexibility of  our computation-based approach, i.e., we just need to reverse the branch condition and resend it to the checker, while existing approaches cannot support such reasoning due to the lack of the computation power and the conservativeness of perservation-based approaches.
+While being sound and able to find real races, existing predictive analyses are conservative in two respects:
+\begin{enumerate}
+	\item Constraints imposed by branch conditions are abided by preserving the exact values flowing into the condition as in the original trace.
+	\item A second limitation, implied by the first, is that traces that diverge from the original trace in their branching history are strictly out of scope. 
+\end{enumerate}
+ Due to these restrictions, enforced as a conservative means of ensuring soundness, existing predictive analysis suffer from limited coverage.
 
-We highlight our approach with the example~\footnote{We omit the synchronization primitives purposely for simplification.} in Figure~\ref{fig:running}.
-Suppose the branch at line 4 changes to {\tt if(y>=3)}.
-In the original run, line 4 reads from the value defined at line 3.
-Therefore, Huang et al.'s approach preserves the same dependence in the predicted run, and therefore misses the race on x between line 2 and line 5.
-In contrast, our approach reports two more real races on x, (2,5) and (2,7).
+ \paragraph{Our Approach} The departure point of our approach, yielding strictly better coverage, is to consider the explicit values of shared memory locations instead of dependencies between statements.
+The gain in coverage is twofold:
+ \begin{enumerate}
+ 	\item Reorderings that violate the original dependence structure but preserve the branching history become possible.
+The values read by a branching statement may change so long as the branch condition evaluates to its original truth value.
+ 	\item In addition, in certain cases we can guarantee soundness while directing execution toward branches that diverge from the original trace.
+This mandates that the effects of an unexplored branch can be modeled precisely, which holds frequently in practice because the original and new branchees often access the same set of shared memory locations.
+ \end{enumerate}
 
-We first demonstrate how we find the race (2,5).
-As shown in Figure~\ref{fig:demo}, we first encode trace in the SSA form, where each variable is defined exactly once (line number as the superscript).
-Besides, we introduce a new symbol, e.g., $R^4_{y}$,  to denote each shared read, e.g., the read of $y$ at
- line 4.
-We also list the constraints for a valid predicted run.
-Race constraints specify line 2 and line 5 can run in parallel.
-Here $O_2$ and $O_5$ are  variables introduced to denote the orders.
-Program order constraints specify the intra-thread execution order.
-The rest  constraints model the value constraints: path condition constraints specify the branch decision for the branches preceding the potential racy event line 5; variable definition
- constraints specify the intra-thread value flow due to each assignment; Inter-thread constraints specify the inter-thread value flow which is conditioned on the scheduling constraints.
-For example, $R^4_{y}=y^1 \wedge O_1< O_4 <O_3$ specifies that line 4 reads {\tt y} defined at line 1 if and only if line 4
-  happens after line 1 and before line 3.
-Combined together, the constraints are sent to off-the-shelf solver such as z3 or yices, which then returns a solution: line 4 reads $y$ ($y$=3) defined at line 1, which satisfies the branch condition.
+ Returning to the example in Figure \ref{fig:running}, this time also with reference to the statements in red, we demonstrate how value-based reasoning improves coverage.
+First, unlike predictive analysis, which is unable to detect the race between lines {\tt 2} and {\tt 5} because of the dependence between lines {\tt 3} and {\tt 4}, value-based reasoning reveals that trace
+ [{\tt 1}, {\tt 4}, {\tt 5}, {\tt 2}, {\tt 3}] is feasible, since the first assignment, {\tt y = 3}, satisfies the condition {\tt y > z}.
+Hence the race is discovered.
+A second race, between lines {\tt 2} and {\tt 7}, is detected by negating the condition.
+This is possible because the effects of the {\tt else} branch can be modeled precisely.
+ Negation leads to an execution starting at line {\tt 4} (where {\tt y $\equiv$ 0}). The race between lines {\tt 2} and {\tt 7} then becomes visible.
 
-Finding the race (2,7) requires a slightly different set of constraints.
-Specifically, the race constraint becomes $O_2=O_7$, and the path condition constraint becomes $R^4_y<3$.
-The checker also returns a solution for these constraints: line 4 reads $y$ ($y$=0) from the initial value, which steers the execution towards the unexecuted path at the branch line 4.
+ Concretely, we achieve value-level (rather than dependence-level) granularity computationally via a unique encoding of the input execution trace as a constraint system.
+The constraints are then processed by a satisfiability checker.
+This gives us the flexibility to explore nontrivial constraints, such as
+ forcing a context switch after the first assignment to {\tt y} or negating the condition in Figure \ref{fig:running}.
+Importantly, these manipulations are beyond what dependence-based reasoning can achieve, since this view of the trace is too conservative.
 
-\begin{figure}
-\centering
-\includegraphics[width=0.5\textwidth]{figs/Visio-demo.pdf}
-\caption{Overview}
-\label{fig:demo}
-\end{figure}
-
-\paragraph{Contributions} The principal contributions of this paper are the following:
+\paragraph{Contributions} This paper makes the following principal contributions:

### 2014-08-21 tech.tex
now

@@
 \item the shared access, which includes the read and write of the shared fields, e.g., $o.f$=$x$ and $x$=$o.f$.
-\item the local access, which includes the access of the local variable, e.g., $x=y+z$ or $v.f=x$ (where $v$ is a thread-local object).
+\item the local access, which includes only the access of the local variable, e.g., $x=y+z$ or $v.f=x$ (where $v$ is a thread-local object).
 \item the branch event, which evaluates the branch condition to true/false, e.g., $x>3$.
@@
 
-Each event $e_i \in \Sigma$ is a tuple, $<t, id, a, v, ins>$, where $t\in \mathcal{T}$ denotes the thread generating the event, $id\in \mathcal{ID}$ denotes the unique integer assigned to the event, $a\mathcal{A}$ denotes the address of the object or field (if any) accessed in the event,  $v\in \mathcal{V}$ denotes the value of the definition (if any) in the event, and $ins\in \mathcal{INS}$ denotes the three-address instruction generating the event.
+Each event $e_i \in \Sigma$ is a tuple, $<t, id, a, v, ins>$, where $t\in \mathcal{T}$ denotes the thread generating the event, $id\in \mathcal{ID}$ denotes the unique integer assigned to the event (event id), $a\mathcal{A}$ denotes the address of the object or field (if any) accessed in the event,  $v\in \mathcal{V}$ denotes the value of the definition (if any) in the event, and $ins\in \mathcal{INS}$ denotes the three-address instruction generating the event.
 Specifically, the address of the object $o$ is denoted as $id(o)\in \mathcal{ID}$, which is a string value representing $o$ uniquely, the address of the static field $f$ is  $id(f)$ and the address of the instance object field $o.f$ is  $id(o)\_id(f)$.
@@
 
+{\bf Local Variables\  }
 Like other symbolic analysis~\cite{jeff,chao}, the symbolic trace should be in the SSA form, i.e., each variable is defined exactly once.
-This is because the constraint solver employed by the analysis does not allow the same variable to hold more than one value.
-Besides, we introduce symbols to represent shared reads and shared writes which leave the inter-thread dependence between reads/writes undetermined.
+This is because the constraint solver employed by the analysis requires each variable to hold only one value.
+Besides, we need to make sure each use still reads from the same definition thread-locally.
 
-\subsection{Method Calls}~\label{sec:basic}
+The  simple procedure shows the construction of the symbolic trace for local variables defined or used.
+The symbols are constructed by combining the static instruction and the runtime event id.
+Lines 6-10 handles the local variable definition.
+We build a symbolic variable $s$ for it by combining the variable name and the event id.
+The uniqueness of the event id guarantees that each symbolic variable is defined exactly once.
+In addition, we replace the variable to the symbolic variable in the instruction and record the replacement in $table$.
+Lines 3-5 updates the local variable used in each event so that it is replaced with the symbolic variable for the corresponding definition.
+Here, the corresponding definition and the use share the same variable name, therefore, we can easily find out the symbolic variable through looking up the $table$.
+
+The SSA form of the trace is different from the SSA form of the instruction as the SSA instruction can only distinguish definitions at different program points but cannot distinguish the definitions at different execution points that share the same program point.
+
+\begin{algorithmic}[3]
+\For {$e: \tau$}
+ \State $ins\gets e.ins$ 
+ \For {$ins.use:ins.uses$}
+ \State $ins.use \gets table(ins.use)$
+ \EndFor
+  \If {$ins.def\neq null$}
+    \State $ s\gets ins.def^{e.id}$
+	\State $ table [ins.def \rightarrow s]$
+	\State $ins.def\gets s$
+ \EndIf
+\EndFor
+\end{algorithmic}
+
+%TODO update the intro+moti, make sure the same style
+\begin{figure}
+\centering
+\begin{tabular}{ll}
+\multicolumn{2}{c}{{\tt {\bf x} = 0; {\bf y} = 0;}} \\
+\hline
+\multicolumn{1}{c}{$T_1$} & \multicolumn{1}{c}{$T_2$} \\
+\hline
+{\tt 1: s=0; } & \\
+{\tt 2: for(i=1;i<3;i++)} & \\
+{\tt 3: \ \ \ s+=i;} & \\
+{\tt 4: {\bf y} = s;} & \\
+{\tt 5: {\bf x} = 1;} & \\
+{\tt 6: {\bf y} = 5;} & \\
+& {\tt 7: if ({\bf y} > 2)} \\
+& {\tt 8:~~print({\bf x}+1);} \\	
+& {\color{Gray} {\tt 9: else}} \\
+& {\color{Gray} {\tt 10:~~print({\bf x}+2);}}
+\end{tabular}
+\caption{Running Example (shared variables are in bold font). }
+\label{fig:running2}
+\end{figure}
+
+{\bf Shared Accesses}   Besides, we introduce symbols to represent shared reads and shared writes which leave the inter-thread dependence between reads/writes undetermined.
+For each read (or write) of shared variable $x$, we introduce $R^{id}_x$ (or $W^{id}_x$) to denote it, where $id$ is the event id.
+Consider the code in Figure~\ref{fig:running2}, which resembles the example in Figure~\ref{fig:running} except that it includes a for loop at lines 1-2.
+The symbolic trace is produced in Figure~\ref{fig:t4running2}. For simplicity, we only show the symbolic variables, while omitting other information such as thread information.
+
+%TODO distinguish T1 and T2 in the text. 
+\begin{figure}
+\centering
+\begin{tabular}{l|l}
+\hline
+\multicolumn{1}{c}{$Trace$} & \multicolumn{1}{c}{$Symbolic\  Trace$} \\
+\hline
+{\tt 0: {\bf x}=0} &  {\tt 0: $W^0_x$=0}    \\
+{\tt 1: {\bf y}=0} &   {\tt 1: $W^1_y$=0}   \\
+{\tt 2: s=0} &  {\tt 2: $s^2$=0}   \\
+{\tt 3: i=1} &     {\tt 3: $i^3$=1}   \\
+{\tt 4: i<3} &    {\tt 4: $i^3$<3} \\
+{\tt 5: s=s+i} & {\tt 5: $s^5$=$s^2$+$i^3$}   \\
+{\tt 6: i=2} &       {\tt 6: $i^6$=2}  \\
+{\tt 7: i<3} &      {\tt 7: $i^6$<3}  \\
+{\tt 8: s=s+i} &  {\tt 8: $s^8$=$s^5$+$i^6$}  \\
+{\tt 9: i=3} &     {\tt 9: $i^9$=3}  \\
+{\tt 10: i<3} &    {\tt 10: $i^9$<3}  \\
+{\tt 11: {\bf y} = s;} &  {\tt 11: $W^{11}_y$ = $s^8$;}  \\
+{\tt 12: {\bf x} = 1;} &    {\tt 12: $W^{12}_x$ = 1;}   \\
+{\tt 13: {\bf y} = 5;} &    {\tt 13: $W^{13}_y$ = 5;}  \\
+{\tt 14: {\bf y} > 2}  &    {\tt 14: $R^{14}_y$ > 2} \\
+{\tt 15: tmp={\bf x}+1;}  & {\tt 15: tmp=$R^{15}_x$+1;}   \\	
+{\tt 16: print(tmp);} &  {\tt 16: print(tmp);}  \\
+\end{tabular}
+\caption{Trace}
+\label{fig:t4running2}
+\end{figure}
+
+{\bf Method Calls\ } The key to supporting the method calls is to capture the value flow the actual argument to the formal argument, and the flow from the return statement to the LHS variable of the method call.
+To explicitly model the value flow, record two additional events for each method call.
+Consider the example in Figure~\ref{fig:methcall},   we record the local access event $y1=y;$ for the argument value flow and record the  local access event $x=i2;$.
+Recording the additional events is achieved through instrumenting the call site and the callee method statically.
+
+The above simple strategy however hides the complexity of the virtual method calls.
+At a call site of a virtual method, the static instrumentation cannot know precisely which method would be called.
+Therefore, we do not know what formal argument the actual argument flows to.
+Consider the example in Figure~\ref{fig:methcall}, suppose another implementation of the virtual method exists (in the comments).
+We do not know how to instrument the code statically, $y1=y$ or $y2=y$.
+
+To avoid the problem, we have to combine the runtime knowledge. 
+Our strategy is as follows: rather than directly record direct value flow from actual argument to formal argument, we introduce an artificial variable during the static instrumentation.
+Then we  insert the instrumentation $record(ARG0=y;)$ at call site, and insert  $record(y1=ARG0)$ at the entry of the method $func$ declared in the first class, and insert $record(y2=ARG0);$ at the entry of the method $func$ declared in the second class.
+At runtime, depending on which $func$ method is invoked, we record either the event sequence $ARG0=y; y1=ARG0$ or the sequence $ARG0=y; y2=ARG0$, which precisely captures the value flow.
+We model the return value flow similarly.
+
+Note that although different methods use the same names for the artificial variable,  they are translated to different variables after we get the SSA form of symbolic trace.
+
+\begin{figure}
+\centering
+\begin{tabular}{ll}
+{\tt  x=o.func(y) } &  \\ 
+ {\tt  func(y1)\{ // class O1} &  \\
+ {\tt  \ \      i=y1; } & \\
+ {\tt  \ \      i2=2*i; } &  \\
+ {\tt  \  \      return i2;} & \\ 
+ {\tt           \}} & \\
+
+  {\tt //  func(y2)\{ // class O2} &  \\
+ {\tt  // \ \      j=y2; } & \\
+ {\tt  // \ \      j2=3*j; } &  \\
+ {\tt  // \  \      return j;} & \\ 
+ {\tt  //         \}} & \\ 
+\end{tabular}
+\caption{Method Calls}
+\label{fig:methcall}
+\end{figure}
+
+\subsection{Constraints}
+The constraints
 

### 2014-08-21 overview.tex

@@
 
-We begin with an explanation of the encoding process:
+We begin with an explanation of the encoding process. The resulting formula is provided in Figure \ref{fig:encoding}. We describe the conjuncts comprising the formula one by one.
+
+\begin{figure*}
+	\begin{center}
+$$
+	\begin{array}{rcl}
+	& \left( O_1 < O_2 < O_3 \wedge O_4 < O_5 \wedge O_4 < O_7 \right) & \textbf{(program order)} \\
+\bigwedge & \left( {\tt x}^0 = 0 \wedge {\tt y}^0 = 0 \wedge {\tt y}^1 = 3 \wedge 
+	{\tt x}^2 = 1 \wedge {\tt y}^3 = 5 \right) & \textbf{(variable definitions)} \\
+\bigwedge & \left(		(R_{\tt y}^4={\tt y}^0 \wedge O_4 < O_1) \bigvee	
+(R_{\tt y}^4={\tt y}^1 \wedge O_1 < O_4 < O_3) \bigvee
+(R_{\tt y}^4={\tt y}^3 \wedge O_3 < O_4)
+		\right) & \textbf{(thread interference)} \\
+\bigwedge & R_{\tt y}^4 > 2 & \textbf{(path conditions)} \\
+\bigwedge & O_2 = O_5 & \textbf{(race condition)}
+	\end{array} 
+$$
+\end{center}
+\caption{\label{fig:encoding}\tool\ encoding of the trace in Figure \ref{fig:running} as a constraint system}
+\end{figure*}
 
@@
 $$
-	O_1 < O_2 < O_3 \bigwedge O_4 < O_5 \bigwedge O_4 < O_7
+	O_1 < O_2 < O_3 \wedge O_4 < O_5 \wedge O_4 < O_7
 $$
@@
 $$
-	{\tt x}^0 = 0 \bigwedge {\tt y}^0 = 0 \bigwedge {\tt y}^1 = 3 \bigwedge 
-			{\tt x}^2 = 1 \bigwedge {\tt y}^3 = 5
+	{\tt x}^0 = 0 \wedge {\tt y}^0 = 0 \wedge {\tt y}^1 = 3 \wedge 
+			{\tt x}^2 = 1 \wedge {\tt y}^3 = 5
 $$
@@
 $$
-	(R_{\tt y}^4 > 2
+	R_{\tt y}^4 > 2
 $$
@@
 
-\paragraph{Constraint Solving} Having conjoined the formuals from the different encoding steps into (i) a global representation of all feasibility constraints (path, ordering, assignment and other constraints) and (ii) the requirement for a given race to occur (expressed as simultaneous execution of the conflicting accesses), we discharge the resulting formula to an off-the-shelf constraint solver.
-If successful, the solver returns a solution for the specified constraints, and in particular, we obtain a feasible trace that gives rise to the race at hand.
+\paragraph{Constraint Solving} Having conjoined the formuals from the different encoding steps into (i) a global representation of all feasibility constraints (path, ordering, assignment and other constraints) and (ii) the requirement for a given race to occur (expressed as simultaneous execution of the conflicting accesses), we discharge the resulting formula to an off-the-shelf constraint solver, such as Z3 or Yices.
+If successful, the solver returns a solution for the specified constraints.
+
+In particular, the solution discloses a feasible trace that gives rise to the race at hand.
+The trace is identified uniquely via the order enforced in the solution over the variables $O_i$, which represent scheduling order.

### 2014-08-23 tech.tex
tech/impl

--- 2014-08-17 tech.tex
@@
-\section{Basics}~\label{sec:basic}
+%\section{Basics}~\label{sec:basic}
 
-\subsection{Trace Terminology}
+\section{Trace Terminology}
 Our analysis starts with a trace $\tau$,  a sequence of events, $e_0, e_1, \dots, e_n$.  
-There are three types of events in general.
+There are  types of events in general.
 \begin{itemize}
 \item the shared access, which includes the read and write of the shared fields, e.g., $o.f$=$x$ and $x$=$o.f$.
-\item the local access, which includes the access of the local variable, e.g., $x=y+z$ or $v.f=x$ (where $v$ is a thread-local object).
+\item the local access, which includes only the access of the local variable, e.g., $x=y+z$ or $v.f=x$ (where $v$ is a thread-local object).
 \item the branch event, which evaluates the branch condition to true/false, e.g., $x>3$.
-\item the synchronization event, which includes start/join, wait/notify, lock/unlock events, e.g., $lock(o)$
+\item the method call event, which includes the method call, e.g., $o.method()$.
+\item the synchronization event, which includes start/join, wait/notify, lock/unlock events, e.g., $lock(o)$.
 \end{itemize}
 
-Each event $e_i \in \Sigma$ is a tuple, $<t, id, a, v, ins>$, where $t\in \mathcal{T}$ denotes the thread generating the event, $id\in \mathcal{ID}$ denotes the unique integer assigned to the event, $a\mathcal{A}$ denotes the address of the object or field (if any) accessed in the event,  $v\in \mathcal{V}$ denotes the value of the definition (if any) in the event, and $ins\in \mathcal{INS}$ denotes the three-address instruction generating the event.
-Specifically, the address of the object $o$ is denoted as $id(o)\in \mathcal{ID}$, which is a string value representing $o$ uniquely, the address of the static field $f$ is  $id(f)$ and the address of the instance object field $o.f$ is  $id(o)\_id(f)$.
-Besides, as the event is derived by instrumenting the three-address code and monitoring the instrumented execution.
-Therefore, each event can involve at most three operands.
+%TODO synchronization constraints.
+
+Each event $e_i \in \Sigma$ is a tuple, $<t, id, a, v, type, ins, \_def, \_uses, \_ins>$, where $t\in \mathcal{T}$ denotes the thread generating the event, $id\in \mathcal{ID}$ denotes the unique integer assigned to the event (event id).
+We say $e_1 < e_2$ iff $e_1.id<e_2.id$.
+$a\in \mathcal{A}$ denotes the address of the object or field (if any) accessed in the event,  $v\in \mathcal{V}$ denotes the value of the definition (if any) in the event, $type$ denotes the type of the event.
+
+Specifically, the address of the object $o$ is denoted as $id(o)\in \mathcal{ID}$, which is a string value representing $o$ uniquely, the address of the static field $f$ is  $id(f)$ and the address of the instance object field $o.f$ is  $id(o).id(f)$.
+
+$ins\in \mathcal{INS}$ denotes the static three-address instruction corresponding to the event.
+$\_def \in \mathcal{S}$ is the symbolic representation of the variable definition  in the event (We will explain the symbolic representation in Section~\ref{sec:symtrace}).
+$\_uses$ are the symbolic representations of the variable uses.
+$\_ins$ is the symbolic form of the instruction after changing the variables into symbolic variables.
 
@@
 \begin{itemize}
-\item projection, e.g., $\tau|t$ returns~\footnote{This is the abbreviation for the complete form $\tau|Thread=t$ } the events from the thread $t$,  $\tau|a$ returns the event involving the address $a$.
+\item projection, e.g., $\tau|Thread=t$ returns the events from the thread $t$,  $\tau|Address=a$ returns the event involving the address $a$, $\tau|Type=branch$ returns all branch events.  
 \item concatenation. $\tau'=\tau e$ represents the new trace by appending the event $e$ to $\tau$.
 \item length $|\tau|$. 
-\item selecting an element. $\tau[0]$ and $\tau[|\tau|-1]$ represents the first and last event in $\tau$.
+\item selection.
+$\tau[i]$ and $\tau[j]$ denotes the $i$th and $j$th event in $\tau$.
+$\tau[i\dots j]$ denotes the event sequence from the $i$th event to the $j$th event.
+$\tau.before(e)$ and $\tau.after(e)$ are the helper function for selecting the event sequences before or after $e$.
+$\tau.between(e_1, e_2)$  selects the event sequence between $e_1$ and $e_2$ (inclusively).
 \end{itemize}
@@
 \begin{itemize}
-\item $AT: \mathcal{A} \times \mathcal{T} \rightarrow \gamma$ is a function that returns a trace $\tau \in \gamma$ that contains only the accesses of the address $a \in \mathcal{A}$ by the thread $t\in \mathcal{T}$.
-Each trace $\tau$ in $\gamma$ is defined over the alphabet of events $\Sigma$, specifically, the trace is an empty trace $\epsilon$ or defined in this way:  $\forall 0\leq i\leq |\tau|,   \tau[i]\in \Sigma, and, \forall i\neq j, \tau[i]\neq \tau[j]$.
-\item $R: \mathcal{A} \rightarrow \gamma$ is a function that returns the read accesses of the address $a\in \mathcal{A}$.
-\item $W: \mathcal{A} \rightarrow \gamma$ is a function that returns the write accesses of the address $a \in \mathcal{A}$.
-\item $Sync: \mathcal{A} \rightarrow \gamma$ is a function that returns the synchronization events involving the address $a\in \mathcal{A}$. 
+%\item $AT: \mathcal{A} \times \mathcal{T} \rightarrow \Gamma$ is a function that returns a trace $\tau \in \gamma$ that contains only the accesses of the address $a \in \mathcal{A}$ by the thread $t\in \mathcal{T}$.
+\item $R: \mathcal{A} \rightarrow \Gamma$ is a function that returns the read accesses of the address $a\in \mathcal{A}$.
+Each trace $\tau$ in $\Gamma$ is defined over the alphabet of events $\Sigma$, specifically, the trace is an empty trace $\epsilon$ or defined in this way:  $\forall 0\leq i\leq |\tau|,   \tau[i]\in \Sigma, and, \forall i\neq j, \tau[i]\neq \tau[j]$.
+\item $W: \mathcal{A} \rightarrow \Gamma$ is a function that returns the write accesses of the address $a \in \mathcal{A}$.
+\item $Sync: \mathcal{A} \rightarrow \Gamma$ is a function that returns the synchronization events involving the address $a\in \mathcal{A}$. 
 \end{itemize}
 
-\subsection{Symbolic Trace}
-To facilitate the symbolic analysis, we need to introduce symbols to represent the operands in each event.
+\section{Symbolic Trace}~\label{sec:symtrace}
+To facilitate the symbolic analysis, we need to introduce symbols to represent the the variables in each event.
 Symbols allow us to overcome the limitation of concrete dependences and allow us to explore more dependences symbolically.
+We refer to this process as the symbolicization.
 
@@
 
+{\bf Local Variables\  }
 Like other symbolic analysis~\cite{jeff,chao}, the symbolic trace should be in the SSA form, i.e., each variable is defined exactly once.
-This is because the constraint solver employed by the analysis does not allow the same variable to hold more than one value.
-Besides, we introduce symbols to represent shared reads and shared writes which leave the inter-thread dependence between reads/writes undetermined.
+This is because the constraint solver employed by the analysis requires each variable to hold only one value.
+Besides, we need to preserve the local data dependences, i.e., to guarantee each use still reads from the same definition thread-locally.
 
-\subsection{Method Calls}~\label{sec:basic}
+The  simple procedure shows the construction of the symbolic trace for local variables defined or used. The symbols are constructed by combining the 
+name of the variable in the static instruction $e.ins$ and the runtime event id $e.id$.
+Lines 6-10 handles the local variable definition.
+The symbolic representation $e.\_def$ for a variable definition is the combination of the variable name and the event id, where  the unique event id guarantees the uniqueness of the symbolic representation.
+In addition, we record the mapping in $table$.
+Lines 3-5 build the symbolic representations for the variables used in the instruction, which are computed through looking up $table$ with the variable name as the key.
+For a local stack variable, its uses and definition involve the same variable name.
+However, for a local heap variable, i.e.,  a field, its uses and definition may not involve the same variable name.
+For example, $o.f=5$ and $y=r.f$ ($o=f$) involve the same field but involve different names.
+To simplify the representation, we introduce a stack variable $l_{address(o.f)}$ to replace the heap variable. The replacement is conducted before the construction of the symbolic trace.
 
-\section{Time Window}~\label{sec:basic}
+An important assumption for the replacement is that the address of the heap variable will remain the same in the predicted run.
+We will guarantee the validity of the assumption, as explained in Section.
 
-\section{Heap Invariant}~\label{sec:basic}
+%The SSA form of the trace is different from the SSA form of the instruction as the SSA instruction can only distinguish definitions at different program points but cannot distinguish the definitions at different execution points that share the same program point.
+
+\begin{algorithmic}[3]
+\State $ins=e.ins$
+\For {$e: \tau$}
+  \For {$i: 0\dots ins.uses.length-1$}
+ \State $e.\_uses[i] \gets table.get(ins.uses[i])$
+ \EndFor
+  \If {$e.\_def\neq null$}
+    \State $ e.\_def \gets ins.def^{e.id}$
+	\State $ table.put(ins.def,  e.\_def)$
+ \EndIf
+\EndFor
+\end{algorithmic}
+
+%TODO update the intro+moti, make sure the same style
+\begin{figure}
+\centering
+\begin{tabular}{ll}
+\multicolumn{2}{c}{{\tt {\bf x} = 0; {\bf y} = 0;}} \\
+\hline
+\multicolumn{1}{c}{$T_1$} & \multicolumn{1}{c}{$T_2$} \\
+\hline
+{\tt 1: s=0; } & \\
+{\tt 2: for(i=1;i<3;i++)} & \\
+{\tt 3: \ \ \ s+=i;} & \\
+{\tt 4: {\bf y} = s;} & \\
+{\tt 5: {\bf x} = 1;} & \\
+{\tt 6: {\bf y} = 5;} & \\
+& {\tt 7: if ({\bf y} > 2)} \\
+& {\tt 8:~~print({\bf x}+1);} \\	
+& {\color{Gray} {\tt 9: else}} \\
+& {\color{Gray} {\tt 10:~~print({\bf x}+2);}}
+\end{tabular}
+\caption{Running Example (shared variables are in bold font). }
+\label{fig:running2}
+\end{figure}
+
+{\bf Shared Variables}   Besides, we introduce symbols to represent shared variables read or written.
+For each shared variable $x$ read (or written), we introduce $R^{id}_x$ (or $W^{id}_x$) to denote it, where $id$ is the event id.
+We denote the read and the write with different symbols, which allow us to reason about different dependence relations among them flexibly.
+
+Consider the code in Figure~\ref{fig:running2}, which resembles the example in Figure~\ref{fig:running} except that it includes a for loop at lines 1-2.
+Figure~\ref{fig:t4running2} shows the instruction $e.ins$ of each event $e$ and its symbolic instruction $e.\_ins$, which is essentially replaces the variables in $e.ins$ to the symbolic variables.
+
+%TODO distinguish T1 and T2 in the text. 
+\begin{figure}
+\centering
+\begin{tabular}{l|l}
+\hline
+\multicolumn{1}{c}{$Trace$} & \multicolumn{1}{c}{$Symbolic\  Trace$} \\
+\hline
+{\tt 0: {\bf x}=0} &  {\tt 0: $W^0_x$=0}    \\
+{\tt 1: {\bf y}=0} &   {\tt 1: $W^1_y$=0}   \\
+{\tt 2: s=0} &  {\tt 2: $s^2$=0}   \\
+{\tt 3: i=1} &     {\tt 3: $i^3$=1}   \\
+{\tt 4: i<3} &    {\tt 4: $i^3$<3} \\
+{\tt 5: s=s+i} & {\tt 5: $s^5$=$s^2$+$i^3$}   \\
+{\tt 6: i=2} &       {\tt 6: $i^6$=2}  \\
+{\tt 7: i<3} &      {\tt 7: $i^6$<3}  \\
+{\tt 8: s=s+i} &  {\tt 8: $s^8$=$s^5$+$i^6$}  \\
+{\tt 9: i=3} &     {\tt 9: $i^9$=3}  \\
+{\tt 10: i<3} &    {\tt 10: $i^9$<3}  \\
+{\tt 11: {\bf y} = s;} &  {\tt 11: $W^{11}_y$ = $s^8$;}  \\
+{\tt 12: {\bf x} = 1;} &    {\tt 12: $W^{12}_x$ = 1;}   \\
+{\tt 13: {\bf y} = 5;} &    {\tt 13: $W^{13}_y$ = 5;}  \\
+{\tt 14: {\bf y} > 2}  &    {\tt 14: $R^{14}_y$ > 2} \\
+{\tt 15: tmp={\bf x}+1;}  & {\tt 15: tmp=$R^{15}_x$+1;}   \\	
+{\tt 16: print(tmp);} &  {\tt 16: print(tmp);}  \\
+\end{tabular}
+\caption{Trace}
+\label{fig:t4running2}
+\end{figure}
+
+{\bf Method Calls\ } To support the method calls, we need to capture the value flow from the actual argument to the formal argument, and the flow from the return variable to the variable definition at the call site.
+We capture the value flow by recording additional events for each method call.
+Consider the example in Figure~\ref{fig:methcall},   we need to record events to capture the implicit assignment $y1=y$ and $x=i2$ due to the method call.
+
+Our strategy is as follows.
+To capture the argument flow $y1=y$, we record two events, one capturing $ARG0=y$ and the other capturing $y1=ARG0$, where $ARG0$ is an artificial variable denoting the first argument of any method.
+This simple strategy helps handle the virtual method calls.
+At the call site, we do not know which target method would be called, the first event records  information about the actual argument.
+Then, a target method is invoked, the second event records information about the formal argument of the resolved target method.
+Combined together, the two events record the argument value flow.
+The recorded events also undergo the above symbolicization process for gaining the SSA form.
+
+%The above simple strategy however hides the complexity of the virtual method calls.
+At a call site of a virtual method, the static instrumentation cannot know precisely which method would be called.
+Therefore, we do not know what formal argument the actual argument flows to.
+Consider the example in Figure~\ref{fig:methcall}, suppose another implementation of the virtual method exists (in the comments).
+We do not know how to instrument the code statically, $y1=y$ or $y2=y$.
+%
+%
+%To avoid the problem, we have to combine the runtime knowledge. 
+%Our strategy is as follows: rather than directly record direct value flow from actual argument to formal argument, we introduce an artificial variable during the static instrumentation.
+Then we  insert the instrumentation $record(ARG0=y;)$ at call site, and insert  $record(y1=ARG0)$ at the entry of the method $func$ declared in the first class, and insert $record(y2=ARG0);$ at the entry of the method $func$ declared in the second class.
+At runtime, depending on which $func$ method is invoked, we record either the event sequence $ARG0=y; y1=ARG0$ or the sequence $ARG0=y; y2=ARG0$, which precisely captures the value flow.
+We model the return value flow similarly.
+%
+%Note that although different methods use the same names for the artificial variable,  they are translated to different variables after we get the SSA form of symbolic trace.
+
+\begin{figure}
+\centering
+\begin{tabular}{ll}
+{\tt  x=o.func(y) } &  \\ 
+ {\tt  func(y1)\{ // class O1} &  \\
+ {\tt  \ \      i=y1; } & \\
+ {\tt  \ \      i2=2*i; } &  \\
+ {\tt  \  \      return i2;} & \\ 
+ {\tt           \}} & \\ 
+  {\tt //  func(y2)\{ // class O2} &  \\
+   {\tt //    ...     } & \\
+
+% {\tt  // \ \      j=y2; } & \\
+% {\tt  // \ \      j2=3*j; } &  \\
+% {\tt  // \  \      return j;} & \\ 
+% {\tt  //         \}} & \\ 
+\end{tabular}
+\caption{Method Calls}
+\label{fig:methcall}
+\end{figure}
+
+Until now, we get the trace with the symbolic variables. The following section uses these symbolic representations to compute races. 
+Hereafter, we always refer to the symbolic form, such as symbolic variable or symbolic instruction, unless otherwise specified. 
+%Unless otherwise specified, we refer to the symbolic variables
+
+\section{Constraints}~\label{sec:constraints}
+First we collect the candidate racy pairs and apply the constraint solver to verify each of them. 
+In general, the candidate racy pairs are collected from the trace as  pairs of shared accesses (one is a write) of the same variable from different threads.
+
+$$
+\{(e_1, e_2| e_1.a=e_2.a \vee e_1.t\neq e_2.t \vee (e_1\in W(e_1.a) \wedge e_2\in W(e_2.a) ))\}
+$$
+
+For each race pair ($e_1$, $e_2$), we construct the constraints $\Phi$ that encode the feasibility requirement and apply the solver to verify the feasibility.
+The constraints $\Phi$ are the conjunction of the following categories of constraints: local order constraints $\Phi_{Local}$, variable definition constraints $\Phi_{VDef}$, the thread interference constraints $\Phi_{Interference}$, the path condition constraints $\Phi_{Path}$ and the race condition constraints $\Phi_{Race}$, and the synchronization constraints $\Phi_{Sync}$.
+In the following, we discuss each of them in  order.
+We omit the synchronization constraints as they are identical to ~\cite{jeff}.
+
+{\bf Race condition constraints \ } The race condition specifies that the racy pair may happen at the same time, i.e., they own the same order. 
+Given two events in the racy pair, ($e_1$, $e_2$), suppose $O(e)$ denotes the order variable associated with the event, we have,
+
+$$
+\Phi_{Race}= (O(e_1)=O(e_2))
+$$
+
+{\bf Path condition constraints \ } The path condition constraints specify that the predicted run should follow the same paths to produce the same set of events (Section~\ref{sec:unexplored} will relax these constraints to explore more paths).
+It is directly derived from the branch conditions taking the symbolic variables, as illustrated at line 10 in Figure~\ref{fig:t4running2}.
+Given a racy pair, ($e_1$, $e_2$), suppose $e_1$ is after $e_2$ in the trace, i.e., $e_2< e_1$.
+
+$$
+\Phi_{Path}=\bigwedge_{e\in (\tau.before(e_1)|Type=branch)} e.\_ins 
+$$
+
+The branches do not need to include those after both racy events as we only need to guarantee the feasibility of the  execution leading to the racy events.
+The branches include not only those from the two threads executing $e_1$ and $e_2$ but also those from other threads, as other threads the feasibility may require the interference from other threads.
+
+{\bf Variable definition constraints\ } The variable definition constraint captures the value flow due to each statement from the right hand side to the left hand side.
+It is directly derived from the symbolic form of the instruction in the each event, as illustrated at line 5 in Figure~\ref{fig:t4running2}.
+Suppose $e_1$ is after $e_2$ in the trace,
+
+ $$
+\Phi_{VDef}=\bigwedge_{e\in (\tau.before(e_1)|(Type=shardAccess\vee Type=localAccess))} e.\_ins 
+$$
+
+Similar to the branches, we include also the accesses from other threads that precede $e_1$.
+
+{\bf Intra-thread order constraints\ } During the predictive analysis, each thread produces the same event sequence in the original run and the predicted run (Section~\ref{sec:unexplored} will relax these constraints to explore more event sequences).
+The constraints are encoded as follows.
+
+$$
+\forall t,  0<i<|\tau'|-1, where\  \tau'=\tau|t, O(\tau'[i])<O(\tau'[i+1])
+$$
+
+Intuitively, the formula specifies that there is an order between any two consecutive events from the same thread.
+
+ In our settings, the order variable is denoted with the event id, e.g., the order variable for the event with id $15$ is denoted as $O_{15}$
+For the example in Figure~\ref{fig:t4running2}, inside the execution of thread $T1$, we have $O_1<O_2 < \dots < O_{13}$. 
+Inside the execution of thread $T2$, we have $O_{14}<O_{15}<O_{16}$.
+
+\subsection{Relaxation for the schedules breaking dependences}
+In addition to the above constraints, we also need to reason about the dependences between the shared reads and shared writes. 
+ Previous predictive analysis requires the predicted run to share many scheduling decisions with the original run so that each shared read reads from the same write (or the same written value) in two runs.
+Our approach achieves the relaxation, without the need for preserving such scheduling decisions.
+In our settings, each shared read can read from any write of the same variable from a different thread, or from the preceding write from the same thread, as long as the resultant execution is feasible.
+
+{\bf Thread interference constraints \ }  Thread interference constraints model the value flows between the writes and reads of the same shared variable. 
+Consider Figure~\ref{fig:t4running2},  the read $R^{14}_y$ may read from the writes $W^1_y$, $W^{11}_y$ or $W^{13}_y$.
+Each read-write correlation further requires certain scheduling constraints.
+For example, to read from the write $W^1_y$, the read should happen after $W^1_y$, and no other writes of the same variable interleave between them.
+The constraint can be captured as $R^{14}_y=W^1_y \wedge O_1<O_{14} \wedge (O_{14}<O_{11} \vee O_{11} < O_1) \wedge (O_{14}<O_{13} \vee O_{13}<O_{1})$.
+
+In general,  given an event $e_R$ that contains the read $R$, and the set $S$ of candidate matching write events, we have the following formula.
+Here, $W_e$ denotes the shared write access included in the event $e$.
+
+% explain W_e.
+$$
+\Phi_{Interference}= \bigvee_{e\in S} (R=W_e) \wedge O(e)<O(e_R)
+\wedge \bigwedge_{e'\in S\setminus e} (O(e')<O(e) \vee O(e_R)<O(e'))
+$$
+
+% candidate matching operations
+
+The candidate matching operations $S$ include both all the writes of the same variable/address from other threads and the preceding write of the same variable from the current thread. 
+%Therefore, given the event $e_R$, 
+%$$S= \{e| e \in W(e_R.a) \wedge ((e.t\neq e_R.t) \vee (e.t=e_R.t \wedge ( \nexists e' \in W(e_R.a) s.t., e'.t=e_R.t \wedge e.id < e'.id < e_R.id)) )\}
+%$$.
+
+%We demonstrate in Section~\ref{sec:constraints} that the address plays an important role in correlating the shared reads and shared writes.
+Intuitively speaking, a remote write matches a read  if and only if they access the same address.
+In the predicted run, we use the address from the original run to judge the matching relations between writes and reads, which is however not always valid.
+
+{\bf Heap invariant\ } In the above, we find the matching operations for the predicted run based on the address collected in the original run.
+However, the matching may be invalid because the address of a field may change if the base reference variable reads a different value, which is allowed in our relaxed analysis.
+
+Consider the following example, in the original run following the line number order, thread $T_2$ reads $o_2$ at line 4 and uses the field $o_2.f$ at line 5. 
+Therefore, following the address value, the write at line 3 is a candidate match for the read at line 5.
+However, in the predicted run, the match may be invalid if the relaxation allows line 4 to read from a different write at line 1.
+
+\begin{figure}
+\centering
+\begin{tabular}{ll}
+\hline
+\multicolumn{1}{c}{$T_1$} & \multicolumn{1}{c}{$T_2$} \\
+\hline
+{\tt 1: {\bf $o_2$} = $new ClassA()$;} & \\
+{\tt 1: {\bf A} = $o_1$;} & \\
+{\tt 2: {\bf A} = $o_2$;} & \\
+{\tt 3: {\bf $o_2$.f} = 5;} & \\
+& {\tt 4: {\bf r} ={\bf A}; } \\
+& {\tt 5: x={\bf r}.f;} \\	
+\end{tabular}
+\caption{Heap Invariant}
+\label{fig:heapinv}
+\end{figure}
+
+We need to guarantee that the address collected for each event in the original run remains the same in the predicted run. 
+Therefore, we add the heap invariant constraint $\Phi_{Heap}$.
+Given an event $e$ that defines the object reference that is used as the base for a field reference or a method call (before the racy pair), we have,
+
+$$
+\Phi_{Heap}=(e.\_def=e.v)
+$$
+$base(e_{inv})=baseDefEvent(e_{inv}).v$, requires the symbol representing the base object to read the original value.
+
+%\begin{theorem}
+%Given a shared read event $e_r$ and a shared write event $e_w$, if they involve the same address $a$ in the original run, 
+%they should involve the same address $a$ in the predicted run; if they involve different addresses in the original run, they should involve different addresses in the predicted run. 
+%\end{theorem}
+
+%Besides, we need to guarantee the base of the method invocation to be identical between the predicted run and the original run.
+Otherwise, the method call may be dispatched to a different target method.
+The constraint,
+In this way, we guarantees three properties: (1) the matching events in the original run are also the matching events in the predicted run, i.e., our above strategy for finding candidate matching events is valid.
+(2) the shared accesses forming a racy pair in the original run also form a racy pair in the predicted run as the base object references do not change from original run to predicted run.
+(3) the dispatching of method calls remains the same in both runs as the base object reference remains the same.
+
+Suppose $e_1$ is after $e_2$ in the trace, the following algorithm judges whether an event $e$ defines such a base reference.
+The branch at line 4 returns false because we encounter a redefinition before the use as the base reference.
+
+\begin{algorithmic}[3]
+\State $o = e.ins.def$
+\State $\tau'=\tau.before(e_1)|e.t$
+\For {$e_i \in \tau'.after(e)$}
+\If {$e_i.ins.def=o$}
+\State $return false$
+\Else
+{
+    \If {$e_i.ins \ references\  o$}
+    \State $ return \ true$
+    \EndIf
+}
+\EndIf
+\EndFor
+\State $return \ false$
+\end{algorithmic}
+
+%We add heap invariant constraints to guarantee the theorem, $base(e_r)=base(e_w)$.
+Here $base(e_r)$ (or $base(e_w)$) is the symbol representing the base reference of the field accessed in $e_r$.
+The following algorithm finds such a symbol through backward analysis over the trace.
+%We can get the base from the address of a field at line 2. Remember that the address of a field is a concatenation of the base and the field. 
+%Through the backward analysis, we can find the definition of the base and return the symbol for the definition. 
+%TODO do not use value for finding base, unnecessary
+
+%Note that we do not require the address  $a'$ in the predicted run to be identical to the address $a$ in the original run.
+Differently, existing approaches require them to be the same.
+Therefore, our approach relaxes the heap constraints too.
+
+%\begin{algorithmic}[3]
+%\State $a=e_r.a$
+%\State $base=getBase(a)$
+%\For {$i: \tau.indexOf(e_r)\dots 0$}
+%  \If {$\tau[i].v=base$}
+%    \State $ return \tau[i].ins.left$
+% \EndIf
+%\EndFor
+%\end{algorithmic}
+
+%In addition to the matching of the read and the write, we need to also guarantee the racy pair still involves the same address in the predicted run. 
+%Consider the same example in Figure~\ref{fig:heapinv}, suppose line 3 and line 5 form a race pair.
+Our analysis does not preserve the addresses, therefore, may lead to different addresses at line 3 and line 5.
+To guarantee that they still form a race, we need to add the heap constraint similar to the above.
+
+\subsection{Relaxation for the un-explored branches}~\label{sec:unexplored}
+To reason about the un-explored branches, we conduct the symbolic execution to collect the symbolic trace for the unexplored branches. 
+At the high level, the symbolic execution starts from any branch event $e$ in the original trace $\tau$, and symbolically executes the un-executed branch. 
+The un-explored branch may contains branches itself, therefore, we conduct the depth-first search (DFS) to collect the symbolic traces $\Gamma_e$ for all paths inside the branch.
+We truncate the original trace at the branch event,  append it with the negation of the branch event and the symbolic trace derived during the DFS.
+The new trace is $\tau.before(e) neg(e) \tau_e$, where $\tau_e$ is any symbolic trace from $\Gamma_e$.
+We apply the constraint solver to the new trace to find potential races.
+
+Loop issues and object creation issues are known limitations of symbolic execution.
+For simplicity, we assume the un-explored branch is simple enough, i.e., it does not contain the loops and every object reference is fully resolved at the branch point.
+In case that the assumption does not hold, we avoid exploring such a branch, sacrificing the coverage in general.
+
+Besides, as mentioned above, we need to know the address which guides the matching of reads and writes. As we assume the unexplored branch has no object creation, all addresses are fully resolved.
+
+\section{Implementation}\label{sec:impl}
+
+\subsection{Time Window and Initial Value constraints}~\label{sec:timewindow}
+The trace is typically long and the constraint solver cannot scale to the whole trace. Instead, we adopt the notion of time window. 
+We divide the trace into $N$ traces of the same length, where $N$ is configurable parameter (1000 in our experiment).
+To support the time window, we need to store the values of the shared variables and local variables at the end of a window and encode such values as the inputs of the next window.
+
+We maintain two maps,  $sstore$ and $lstore$, for the shared variables and local variables respectively.
+Suppose the trace for the last window is $\tau$.
+The following code illustrates how we maintain the $sstore$, i.e., storing the value written by the last write with each address.
+Beside, the variable $e.ins.left$ represents the left hand variable inside the instruction $e.ins$.
+
+\begin{algorithmic}[3]
+\For {$a \in \mathcal{A}$}
+ \State $\tau_a=W(a)$
+ \State $sstore.put(a,\tau_a[|\tau_a|-1].v)$ 
+\EndFor
+\end{algorithmic}
+
+%TODO define left:
+\begin{algorithmic}[3]
+\For {$e \in \tau$}
+  \If {$e.isLocalAcc()$}
+    \State $ lstore.put(e.\_def, e.v)$
+  \Else
+     \If{$e.isRead()$}      
+        \State $ lstore.put(e.\_def, e.v)$
+     \EndIf
+  \EndIf 
+\EndFor
+\end{algorithmic}
+
+When constructing the constraints for the next window, we specify the following constraints to encode the input values.
+
+The following formula enhances the thread interfernce constraint to include the case that the read reads from the initial values.
+
+$$\Phi_{TI}=\Phi_{TI} \wedge  R=sstore.get(e_R.a) \wedge \bigvee_{e\in S} O(e_R)<O(e)$$.
+
+The following formula encodes the input values of local variables.
+
+$$\Phi_{IN}=\bigvee_{l\in lstore.keys()} l=lstore.get(l)$$.
+
+{\bf Object Creation\ } Given the event in the form of $x=new Object()$, we encode the created object as a unique integer and treats the instruction as a normal integer assignment.
+The unique integer is the address of the object, calculated by $System.identityHashcode(x)$.
+
+{\bf Array access\ } For array access event $arr[i]=x$, we treat them as the field access event $arr.i=x$, where the index $i$ is treated as a field.
+
+{\bf Shared Analysis\ } We distinguish the accesses of the shared fields and local fields.
+To collect the shared fields, we preprocess the trace and map each field to a set of accessing threads.
+If a field is written by two or more threads, it is shared.
+If a field is written by one thread and read by other threads, it is shared.
+Otherwise, it is a local field.
+
+{\bf Optimization\ }.
+Before the symbolic analysis, we conduct the backward slicing by treating the racy pair as the seeds.
+Only those selected in the slices are preserved while the rest events are removed.

### 2014-08-24 trace.tex
new stuff

\section{Trace Encoding}

In this section, we describe how a concrete execution trace is encoded as a symbolic trace, which enables \tool\ to 
derive reordering constraints.

\subsection{Preliminaries}

Throughout this paper, we assume a standard operational semantics, which defines (i) a mapping ${\sf thr}$ between
execution threads, each having a unique identifier $i \in \mathbb{N}$, and their respective code, as well as (ii) per-thread stack and shared heap memory. The code executed by a given thread follows the syntax in Table \ref{Ta:syntax}. In the starting state, each thread points to a parameter-free method.

As is standard, we assume an interleaved semantics of concurrency. A transition is of the form
$\sigma \stackrel{i / {\tt s}}{\longrightarrow} \sigma'$, denoting that thread $i$ took an evaluation step
in prestate $\sigma$, wherein atomic
statement {\tt s} was executed, which resulted in poststate $\sigma'$. We refer to a sequence of transitions from
the starting state to either an exceptional state (e.g., due to null dereference) or a state where all the threads have 
reduced their respective code to $\epsilon$ as a \emph{trace}.

\begin{table}
	\begin{center}
	\begin{tabular}{rl}
						\multicolumn{1}{l}{{\tt p} $::=$} & \\
					    $\overline{\texttt{r m($\overline{{\tt x}}$) \{ s \}}}$ & {\bf (method)} \\
						\multicolumn{1}{l}{{\tt s} $::=$} & \\
						{\tt x = y}\ $|$\ {\tt x = $c$} $|$\ & {\bf (assignment)} \\
						{\tt x = new()} $|$\ & {\bf (allocation)} \\
						{\tt x = f($\overline{{\tt y}}$)} $|$\ & {\bf (invocation)} \\
						{\tt z = x \{ $+$,$-$,$\times$,$/$ \} y}\ $|$\ & {\bf (aexp)} \\
						{\tt b = x \{ $<$,$\leq$,$\equiv$,$\geq$,$>$ \} y}\ $|$\ & {\bf (comparison)} \\
						{\tt b = bx $\wedge$ by}\ $|$\ {\tt b = bx $\vee$ by} $|$\ & {\bf (bexp)} \\
						{\tt y = x.f} $|$\ {\tt x.f = y}\ $|$\ & {\bf (heap)} \\
						{\tt if (b) \{ s \} $|$\ {\tt while (b) \{  s \}} $|$\ {\tt s ; s }} $|$\ & {\bf (control)} \\
						{\tt lock(x)}\ $|$\ {\tt unlock(x)} & {\bf (sync)}
	\end{tabular}
	\end{center}
	\caption{\label{Ta:syntax}Language syntax}
\end{table}

We make use of the following helper functions:
\begin{itemize}
	\item ${\sf proj}\ t\ i$ projects trace $t$ onto all transitions involving thread $i$.
	\item $t[k]$ obtains the $k$th transition within trace $t$.
	\item ${\sf pre}\ t\ \tau$ is the prefix of trace $t$ preceding transition $\tau$. For the suffix beyond $\tau$, we 
	use ${\sf post}\ t\ \tau$. Finally, ${\sf bet}\ t\ \tau_1\ \tau_2$ returns the transitions delimited by $\tau_1$ and $\tau_2$.
\end{itemize}

\subsection{Basic Encoding}

\subsection{Advanced Encoding}

### 2014-08-24 trace.tex
more

@@
 
-As is standard, we assume an interleaved semantics of concurrency. A transition is of the form
+As is standard, we assume an interleaved semantics of concurrency. A \emph{transition}, or \emph{event}, is of the form
 $\sigma \stackrel{i / {\tt s}}{\longrightarrow} \sigma'$, denoting that thread $i$ took an evaluation step
@@
 
+We describe how \tool\ encodes a trace into symbolic form in two steps. We first start with ``local'' intraprocedural events --- i.e., 
+transitions that manipulate memory accessed by at most a single thread during an invocation-free run --- and then extend our encoding scheme to the entire set of possible events.
+
+The fundamental encoding transformation is to induce Static Single Assignment (SSA) form on the raw trace, such that
+a variable is defined exactly once. In this way, def/use chains become explicit, and encoding of trace events as constraints is simplified. As an illustration, trace
+\begin{quote}
+	{\tt x=1; x<3; x=3; y=1; z=x+y}
+\end{quote}
+becomes
+\begin{quote}
+	{\tt x$^0$=1; x$^0$<3; x$^1$=3; y$^0$=1; z$^0$=x$^1$+y$^0$}
+\end{quote}
+
+Beyond the default SSA rewriting algorithm, we apply a specialized transformation to handle local heap accesses.\footnote{
+	Note that while in general determining whether a given heap access is local is an undecidable problem, in the context
+	of a concrete execution trace this determination is straightforward.
+} We replace accesses to field {\tt f} of local object {\tt x} with a fresh local variable {\tt l$_{\tt x.f}$}.
+This is done prior to the standard SSA transformation, and under the assumption that the heap object pointed-to by {\tt x} will remain the same in the predicted trace.
+
 \subsection{Advanced Encoding}

### 2014-08-24 trace.tex
more

@@
 \begin{quote}
-	{\tt x=1; x<3; x=3; y=1; z=x+y}
+	{\tt 1: x=1; 2: x<3; 3: x=3;} \\
+	{\tt 4: y=1;} \\
+	{\tt 5: z=x+y}
 \end{quote}
@@
 \begin{quote}
-	{\tt x$^0$=1; x$^0$<3; x$^1$=3; y$^0$=1; z$^0$=x$^1$+y$^0$}
+	{\tt 1: x$^1$=1; 2: x$^1$<3; 3: x$^3$=3;} \\
+	{\tt 4: y$^4$=1;} \\
+	{\tt 5: z$^5$=x$^3$+y$^4$}
 \end{quote}
+(For readability, we version variables according to the line number of their definition.)
 
@@
 	of a concrete execution trace this determination is straightforward.
-} We replace accesses to field {\tt f} of local object {\tt x} with a fresh local variable {\tt l$_{\tt x.f}$}.
-This is done prior to the standard SSA transformation, and under the assumption that the heap object pointed-to by {\tt x} will remain the same in the predicted trace.
+} In the baseline SSA form, def/use chains over heap accesses are ignored.
+To ensure feasibility, we must account for such accesses.
+Hence, we replace accesses to field {\tt f} of local object $o$ with a fresh local variable {\tt l$_{o{\tt .f}}$}, where $o{\tt .f}$ denotes the memory location itself and not its value.
+This is done prior to the standard SSA transformation, and under the assumption that in the predicted run, the base object $o$ remains the same.
 

### 2014-08-24 trace.tex
stuff

@@
 execution threads, each having a unique identifier $i \in \mathbb{N}$, and their respective code, as well as (ii) per-thread stack and shared heap memory.
+
 The code executed by a given thread follows the syntax in Table \ref{Ta:syntax}.
-In the starting state, each thread points to a parameter-free method.
+The text of a program is a sequence of zero or more method declarations, as given in the definition of symbol {\tt p}.
+Methods accept zero or more arguments $\overline{{\tt x}}$, have a body ${\tt s}$, and may have a return value (which we leave implicit).
+For simplicity, we avoid from static typing as well as virtual methods.
+The body of a method consists of the core grammar for symbol {\tt s}.
+We avoid from specifying syntax checking rules, as the grammar is fully standard.
+
+For simplicity, we assume that in the starting state each thread points to a parameter-free method.
+In this way, we can simply assume an empty starting state (i.e., an empty heap), and eliminate complexities such as user-provided inputs and initialization of arguments with default values.
+These extensions are of course possible, and in practice we handle these cases, but reflecting them in the formalism would result in needless complications.
 
@@
 						\multicolumn{1}{l}{{\tt p} $::=$} & \\
-					    $\overline{\texttt{r m($\overline{{\tt x}}$) \{ s \}}}$ & {\bf (method)} \\
+					    $\overline{\texttt{m($\overline{{\tt x}}$) \{ s \}}}$ & {\bf (method)} \\
 						\multicolumn{1}{l}{{\tt s} $::=$} & \\
@@
 						{\tt x = f($\overline{{\tt y}}$)} $|$\ & {\bf (invocation)} \\
+						{\tt return x} $|$\ & {\bf (return)} \\
 						{\tt z = x \{ $+$,$-$,$\times$,$/$ \} y}\ $|$\ & {\bf (aexp)} \\
@@
 
-\subsection{Basic Encoding}
+\subsection{Basic Encoding: Local Accesses}
 
@@
 
-The fundamental encoding transformation is to induce Static Single Assignment (SSA) form on the raw trace, such that
+\paragraph{SSA Form} The fundamental encoding transformation is to induce Static Single Assignment (SSA) form on the raw trace, such that
 a variable is defined exactly once. In this way, def/use chains become explicit, and encoding of trace events as constraints is simplified. As an illustration, trace
@@
 
-Beyond the default SSA rewriting algorithm, we apply a specialized transformation to handle local heap accesses.\footnote{
-	Note that while in general determining whether a given heap access is local is an undecidable problem, in the context
+\paragraph{Local Heap Accesses} Beyond the default SSA rewriting algorithm, we apply a specialized transformation to handle local heap accesses.
+While in general determining whether a given heap access is local is an undecidable problem, in the context
 of a concrete execution trace this determination is straightforward.
-} In the baseline SSA form, def/use chains over heap accesses are ignored.
-To ensure feasibility, we must account for such accesses.
+We can thus improve upon the baseline SSA form, where def/use chains over heap accesses are ignored.
+Indeed, to ensure feasibility, we must account for such accesses.
 Hence, we replace accesses to field {\tt f} of local object $o$ with a fresh local variable {\tt l$_{o{\tt .f}}$}, where $o{\tt .f}$ denotes the memory location itself and not its value.
@@
 
-\subsection{Advanced Encoding}
+\subsection{Full Encoding: Method Calls and Shared Accesses}
+
+To complete the encoding algorithm, we next explain how shared heap accesses and method calls are dealt with.
+Similarly to the case of local heap accesses, we take advantage of the fact that the full execution trace is available for analysis to detect shared accesses (the dual of local accesses) as well as call-site resolutions (which is more relevant in practice, where we support virtual method calls).
+
+\paragraph{Method Invocations} To account for method invocations, we induce additional context on variable accesses.
+Instead of recording only the version of a variable, we also represent as part of the variable's identifier its enclosing method and its invocation counter.
+Thus, symbol ${\tt m}^2 \colon {\tt v}^3$ is interpreted as the definition of variable ${\tt v}$ at offset 3 within method ${\tt m}$ during its second invocation.
+
+In this way, the trace is flattened.
+Method invocations are substituted with qualified variable names, and the rest of the encoding steps, described above, remain unchanged.
+Because the trace is finite and fully resolved, challenges such as looping, recursion and mutual recursion are all obviated.
+
+\paragraph{Shared Heap Accesses} The final aspect of our encoding process is representation of shared heap accesses.
+As illustrated in Section \ref{Se:techoverview}, this is done via designated symbols:
+\begin{itemize}
+	\item $W^{k}_{\tt v}$ denotes write access to variable ${\tt v}$ at offset $k$ (where $k$ is a qualified offset to account  for the interprocedural setting).
+	\item Analogously, $R^{k}_{\tt v}$ denotes that ${\tt v}$ is read at offset $k$.
+\end{itemize}
+
+The main motivation to encode shared heap accesses using special symbols is to simplify downstream processing, and in particular, the definition of candidate races.
+We could encode the question of shared versus local as additional constraints on variable accesses by different threads, but that would be more complicated and less efficient.

### 2014-08-24 trace.tex
trace

@@
 \section{Trace Encoding}
+
+% % 
+% TODO: Explain the format of the raw trace in impl.tex.
+% %
 
@@
 	\item $t[k]$ obtains the $k$th transition within trace $t$.
+	\item ${\sf index}\ t\ \tau$ retrieves the index, or offset, of transition $\tau$ within trace $t$. When simply writing
+	 ${\sf index}\ \tau$ (while omitting the trace parameter) we refer to the index of $\tau$ within the original trace. 
 	\item ${\sf pre}\ t\ \tau$ is the prefix of trace $t$ preceding transition $\tau$. For the suffix beyond $\tau$, we 
@@
 \begin{itemize}
-	\item $W^{k}_{\tt v}$ denotes write access to variable ${\tt v}$ at offset $k$ (where $k$ is a qualified offset to account  for the interprocedural setting).
-	\item Analogously, $R^{k}_{\tt v}$ denotes that ${\tt v}$ is read at offset $k$.
+	\item $W^{k}_{\tt v}$ denotes write access to variable ${\tt v}$ at trace index $k$ (where ${\tt v}$ is assumed to be a qualified identifier to account for the interprocedural setting).
+	\item Analogously, $R^{k}_{\tt v}$ denotes that ${\tt v}$ is read at index $k$.
 \end{itemize}
@@
 The main motivation to encode shared heap accesses using special symbols is to simplify downstream processing, and in particular, the definition of candidate races.
-We could encode the question of shared versus local as additional constraints on variable accesses by different threads, but that would be more complicated and less efficient.
+We could encode the question of shared versus local accesses as additional constraints (requiring that a variable be accessed by more than one thread), but that would be more complicated and less efficient.

### 2014-08-28 constraints.tex
some changes

\section{Constraint Derivation and Resolution}

We now describe in detail the constraint encoding and solving steps.

\subsection{Same Trace}

We begin with the basic setting, which enforces strict reordering of the events along the input trace without diverging into new execution paths. We relax this requirement later, such that different branches can be followed as long as feasibility is retained.

\paragraph{Intra-thread Order}

The first set of constraints reflects control flow within the individual threads, which must remain unchanged under the reordering transformation. Given input trace $t$, this is expressed as the following formula:
$$
\begin{array}{rl}
	\forall \tau,\tau' \in t. & {\sf thread}\ \tau \equiv {\sf thread}\ \tau'. \\
										& {\sf index}\ \tau < {\sf index}\ \tau' \Rightarrow O_{\tau} < O_{\tau'}
\end{array}
$$ 
The logical variables $O_x$ express ordering constraints. The requirement, as stated above, is that these variables
reflect the same order as the projection of ${\sf index}$ onto individual threads.

\paragraph{Race Condition}

Given pair $\tau$ and $\tau'$ of events that both access a common memory location $\ell$, we demand that
$$
\begin{array}{rl}
					& {\sf thread}\ \tau \neq {\sf thread}\ \tau' \\
\bigwedge 	& ({\sf writes}\ \tau\ \ell \vee \tau'\ {\tt s}'\ \ell) \\
\bigwedge   & O_{\tau} = O_{\tau'}
\end{array}
$$
That is, (i) events $\tau$ and $\tau'$ are executed by different threads, (ii) at least one of the events performs write access to $\ell$, and (iii) the events occur at the simultanesouly. This is a direct logical encoding of the definition of a race condition.

\paragraph{Path Constraints}

The requirement with respect to branching is that the prefix of the original trace $t$ up to the pair $\tau$ and $\tau'$ of candidate racing events remains identical in the predicted trace $t'$. More accurately, under the assumption that ${\sf index}\ \tau < {\sf index}\ \tau'$, 
we require that 
$$
	\bigwedge_{\tau'' \in t \cap {\bf bexp}.\
		\tau'' \in {\sf pre}\ \tau'} \lsyn {\sf stmt}\ \tau'' \rsyn\ t \equiv \lsyn {\sf stmt}\ \tau'' \rsyn\ t'
$$ 
where ${\sf stmt}$ is a helper function that obtains the code statement incident in a given transition. That is, all branching transitions up to $\tau'$ (which occurs after $\tau$) preserve their boolean interpretation under $t'$. This ensures that there are no divergences from the path containing the racing events, though beyond that path any feasible continuation is permitted. Importantly, contrary to \cite{JEFF-PLDI14}, we do not pose the requirement that the values flowing into branching statements remain the same, but suffice with the relaxed requirement that the evaluation of branching expressions is invariant under the input and predicted traces.

\paragraph{Variable Definitions}

The final requirement for the basic setting is that left-hand variables are defined according to the same right-hand variables as before. That is, version $i$ of variable ${\tt u}$ is defined as version $j$ of variable ${\tt v}$ in input trace $t$, then the same remains true in predicted  trace $t'$. This is enforced as the formula
$$
	\bigwedge_{\tau'' \in t \cap {\bf asgn}.\
		\tau'' \in {\sf pre}\ \tau'} {\sf stmt}\ \tau'' \in t' 
$$
where we again assume that ${\sf index}\ \tau < {\sf index}\ \tau'$. That is, the same statement occurring in $t$ is also present in $t'$ (though the transitions may differ). Since the statements of $t'$ are a permutation of the statements of $t$, we are assured that use/def flow is constrained appropriately.

\subsection{Unexplored Paths}

### 2014-08-29 eval.tex
eval update

\section{Evaluation}~\label{sec:eval}

{\bf Evaluation Method} We conduct our experiment with a set of large applications, e.g.,  those from the {\sf Dacapo} benchmark suite. 
The benchmarks are shown in Table. We compare our approach with the state-of-the-art technique proposed by Huang et al. As the capability of finding races depends on the monitored run, for fair comparison, we 
 apply both techniques to the same monitored run. Specifically, we use two threads in the monitored run as two threads are sufficient to manifest most of races~\cite{shanlu}.

The {\sf Dacapo} relies on the reflection functionality to execute, which imposes challenges to static analysis and transformation. We apply the tool {\sf Tamiflex}~\cite{} to overcome the problems, which resolves the reflection using the classes observed in the dynamic run. The window size is 10000.

Our experiments and measurements 
were all conducted on an x86 64 Thinkpad W530 worksta- 
tion with eight 2.30GHz Intel Core i7-3610QM processors, 
16GB of RAM and 6M caches. The workstation runs version 12.04 of the Ubuntu Linux distribution, and has the Sun 
64-Bit 1.6.0\_26 Java virtual machine (JVM) installed~\footnote{Note that Sun JDK 1.7 does not support the transformation of dacapo applications with {\sf tamiflex}.}.

\subsection{Scalability Issues}
Our approach suffers from scalability issues for two reasons. First, as we record local access events in the monitoring run,  the monitoring run takes long time, e.g., more than 2 hours, and our trace is typically large, e.g., at the GB level. To avoid the out of memory issue during the monitoring run, we store the event to the database on the fly, rather than storing them in a buffer. Besides, we terminate the monitoring run with a shutdown hook if it is beyond 30 minutes. Second,  the prediction phase needs to load the trace to memory, which easily causes out of memory problem. To solve the problem, we separate the analysis into several runs, each run restarting the JVM and resuming from the window of last run. Different runs communicate the window id using the external file. These runs are organized together automatically using the ant. The outputs of the files are merged into one summary finally. Besides,

\begin{table}[htbp]
\caption{haha}
\begin{tabular}{|l|l|}
\hline
ad & fda \\ \hline
\multicolumn{1}{|r|}{1} & \multicolumn{1}{r|}{2} \\ \hline
\end{tabular}
\label{tb:ref}
\end{table}

\section{Case studies}
We manually inspect the reported races in small applications to gain better understanding about our approach.

{\bf Relaxing the Inter-thread dependence\ } Figure~\ref{fig:relax1} demonstrates the relaxation of inter-thread dependence enabled by our approach. 
The code is from the benchmark {\sf bbuffer}, where the line number is marked. Huang et al.~\cite{} detects the race between line 291 and line 400, but fails to detect the race between line 294 and line 400. The reason is as follows. In the observation run, the execution follows the order,  lines 400, 291 and 294. For the event at line 294, its preceding branch at line 291 reads from line 400. Therefore, Huang et al.~\cite{}
 requires the predicted run to preserve the dependence between line 291 and line 400 so that line 291 reads exactly the same value and the branch takes the same branch decision. The dependence enforces the order constraint, $400 \rightarrow 291$, which further enforces the order $400\rightarrow 291 \rightarrow 294$.    Our approach does not require the existence of such dependence. Specifically, we allow line 291 to happen before line 400 in the predicted run as long as the value read by it leads to the same branch decision, which is true in this case.  As a result, there is no order constraint between line 294 and line 400, and the two forms a race. We re-replay such race  easily using the eclipse IDE breakpoints.

\begin{figure}[htp]
\centering
\includegraphics[scale=0.7]{cases/Visio-bbuffer.pdf}
\caption{Relaxation of Inter-thread dependence}\label{fig:relax1}
\end{figure}

\section{Validity Threat}
Java method can have 65535 bytecode instructions maximally. Therefore, we count the number of bytecode instructions inside each method, if the number exceeds 65525, we avoid instrumenting the method. The consequence is that, we will miss the races inside the method. In this case, we specify the variables read from the method to be equal to their concrete values in the constraints. THe constraint solver can proceed safely without being affected by such methods.

We do not support the boolean operations such as $\&$, bit operators $<<$, which contributes to most of our misses.

### 2014-08-29 tech.tex

@@
-%\section{Basics}~\label{sec:basic}
-
-\section{Trace Terminology}
-Our analysis starts with a trace $\tau$,  a sequence of events, $e_0, e_1, \dots, e_n$.  
-There are  types of events in general.
-\begin{itemize}
-\item the shared access, which includes the read and write of the shared fields, e.g., $o.f$=$x$ and $x$=$o.f$.
-\item the local access, which includes only the access of the local variable, e.g., $x=y+z$ or $v.f=x$ (where $v$ is a thread-local object).
-\item the branch event, which evaluates the branch condition to true/false, e.g., $x>3$.
-\item the method call event, which includes the method call, e.g., $o.method()$.
-\item the synchronization event, which includes start/join, wait/notify, lock/unlock events, e.g., $lock(o)$.
-\end{itemize}
-
-%TODO synchronization constraints.
-
-Each event $e_i \in \Sigma$ is a tuple, $<t, id, a, v, type, ins, \_def, \_uses, \_ins>$, where $t\in \mathcal{T}$ denotes the thread generating the event, $id\in \mathcal{ID}$ denotes the unique integer assigned to the event (event id).
+%%\section{Basics}~\label{sec:basic}
+%
+%
+%\section{Trace Terminology}
+%Our analysis starts with a trace $\tau$,  a sequence of events, $e_0, e_1, \dots, e_n$.  
+%There are  types of events in general.
+%\begin{itemize}
+%\item the shared access, which includes the read and write of the shared fields, e.g., $o.f$=$x$ and $x$=$o.f$.
+%\item the local access, which includes only the access of the local variable, e.g., $x=y+z$ or $v.f=x$ (where $v$ is a thread-local object).
+%\item the branch event, which evaluates the branch condition to true/false, e.g., $x>3$.
+%\item the method call event, which includes the method call, e.g., $o.method()$.
+%\item the synchronization event, which includes start/join, wait/notify, lock/unlock events, e.g., $lock(o)$.
+%\end{itemize}
+%
+%%TODO synchronization constraints.
+%
+%Each event $e_i \in \Sigma$ is a tuple, $<t, id, a, v, type, ins, \_def, \_uses, \_ins>$, where $t\in \mathcal{T}$ denotes the thread generating the event, $id\in \mathcal{ID}$ denotes the unique integer assigned to the event (event id).
 We say $e_1 < e_2$ iff $e_1.id<e_2.id$.
 $a\in \mathcal{A}$ denotes the address of the object or field (if any) accessed in the event,  $v\in \mathcal{V}$ denotes the value of the definition (if any) in the event, $type$ denotes the type of the event.
-
-Specifically, the address of the object $o$ is denoted as $id(o)\in \mathcal{ID}$, which is a string value representing $o$ uniquely, the address of the static field $f$ is  $id(f)$ and the address of the instance object field $o.f$ is  $id(o).id(f)$.
-
-$ins\in \mathcal{INS}$ denotes the static three-address instruction corresponding to the event.
+%
+%Specifically, the address of the object $o$ is denoted as $id(o)\in \mathcal{ID}$, which is a string value representing $o$ uniquely, the address of the static field $f$ is  $id(f)$ and the address of the instance object field $o.f$ is  $id(o).id(f)$.
+%
+%
+%$ins\in \mathcal{INS}$ denotes the static three-address instruction corresponding to the event.
 $\_def \in \mathcal{S}$ is the symbolic representation of the variable definition  in the event (We will explain the symbolic representation in Section~\ref{sec:symtrace}).
@@
 $\_ins$ is the symbolic form of the instruction after changing the variables into symbolic variables.
-
-The trace supports its standard operations as follows.
-\begin{itemize}
-\item projection, e.g., $\tau|Thread=t$ returns the events from the thread $t$,  $\tau|Address=a$ returns the event involving the address $a$, $\tau|Type=branch$ returns all branch events.  
-\item concatenation. $\tau'=\tau e$ represents the new trace by appending the event $e$ to $\tau$.
-\item length $|\tau|$. 
-\item selection.
+%
+%
+%The trace supports its standard operations as follows.
+%\begin{itemize}
+%\item projection, e.g., $\tau|Thread=t$ returns the events from the thread $t$,  $\tau|Address=a$ returns the event involving the address $a$, $\tau|Type=branch$ returns all branch events.  
+%\item concatenation. $\tau'=\tau e$ represents the new trace by appending the event $e$ to $\tau$.
+%\item length $|\tau|$. 
+%\item selection.
 $\tau[i]$ and $\tau[j]$ denotes the $i$th and $j$th event in $\tau$.
@@
 $\tau.between(e_1, e_2)$  selects the event sequence between $e_1$ and $e_2$ (inclusively).
-\end{itemize}
-
-%The modeling of synchronization event is standard and explained in existing work, so we focus on the rest two types of events in this paper.
-
-In addition, we maintain auxiliary information as follows.
-\begin{itemize}
-%\item $AT: \mathcal{A} \times \mathcal{T} \rightarrow \Gamma$ is a function that returns a trace $\tau \in \gamma$ that contains only the accesses of the address $a \in \mathcal{A}$ by the thread $t\in \mathcal{T}$.
-\item $R: \mathcal{A} \rightarrow \Gamma$ is a function that returns the read accesses of the address $a\in \mathcal{A}$.
+%\end{itemize}
+%
+%
+%%The modeling of synchronization event is standard and explained in existing work, so we focus on the rest two types of events in this paper.
+%
+%In addition, we maintain auxiliary information as follows.
+%\begin{itemize}
+%%\item $AT: \mathcal{A} \times \mathcal{T} \rightarrow \Gamma$ is a function that returns a trace $\tau \in \gamma$ that contains only the accesses of the address $a \in \mathcal{A}$ by the thread $t\in \mathcal{T}$.
+%\item $R: \mathcal{A} \rightarrow \Gamma$ is a function that returns the read accesses of the address $a\in \mathcal{A}$.
 Each trace $\tau$ in $\Gamma$ is defined over the alphabet of events $\Sigma$, specifically, the trace is an empty trace $\epsilon$ or defined in this way:  $\forall 0\leq i\leq |\tau|,   \tau[i]\in \Sigma, and, \forall i\neq j, \tau[i]\neq \tau[j]$.
-\item $W: \mathcal{A} \rightarrow \Gamma$ is a function that returns the write accesses of the address $a \in \mathcal{A}$.
-\item $Sync: \mathcal{A} \rightarrow \Gamma$ is a function that returns the synchronization events involving the address $a\in \mathcal{A}$. 
-\end{itemize}
-
-\section{Symbolic Trace}~\label{sec:symtrace}
-To facilitate the symbolic analysis, we need to introduce symbols to represent the the variables in each event.
+%\item $W: \mathcal{A} \rightarrow \Gamma$ is a function that returns the write accesses of the address $a \in \mathcal{A}$.
+%\item $Sync: \mathcal{A} \rightarrow \Gamma$ is a function that returns the synchronization events involving the address $a\in \mathcal{A}$. 
+%\end{itemize}
+%
+%
+%\section{Symbolic Trace}~\label{sec:symtrace}
+%To facilitate the symbolic analysis, we need to introduce symbols to represent the the variables in each event.
 Symbols allow us to overcome the limitation of concrete dependences and allow us to explore more dependences symbolically.
 We refer to this process as the symbolicization.
-
-%shared access only, local access only
-
-{\bf Local Variables\  }
-Like other symbolic analysis~\cite{jeff,chao}, the symbolic trace should be in the SSA form, i.e., each variable is defined exactly once.
+%
+%%shared access only, local access only
+%
+%
+%
+%
+%{\bf Local Variables\  }
+%Like other symbolic analysis~\cite{jeff,chao}, the symbolic trace should be in the SSA form, i.e., each variable is defined exactly once.
 This is because the constraint solver employed by the analysis requires each variable to hold only one value.
 Besides, we need to preserve the local data dependences, i.e., to guarantee each use still reads from the same definition thread-locally.
-
-The  simple procedure shows the construction of the symbolic trace for local variables defined or used. The symbols are constructed by combining the 
-name of the variable in the static instruction $e.ins$ and the runtime event id $e.id$.
+%
+%The  simple procedure shows the construction of the symbolic trace for local variables defined or used. The symbols are constructed by combining the 
+%name of the variable in the static instruction $e.ins$ and the runtime event id $e.id$.
 Lines 6-10 handles the local variable definition.
@@
 In addition, we record the mapping in $table$.
-Lines 3-5 build the symbolic representations for the variables used in the instruction, which are computed through looking up $table$ with the variable name as the key.
+%Lines 3-5 build the symbolic representations for the variables used in the instruction, which are computed through looking up $table$ with the variable name as the key.
 For a local stack variable, its uses and definition involve the same variable name.
@@
 For example, $o.f=5$ and $y=r.f$ ($o=f$) involve the same field but involve different names.
-To simplify the representation, we introduce a stack variable $l_{address(o.f)}$ to replace the heap variable. The replacement is conducted before the construction of the symbolic trace.
-
-An important assumption for the replacement is that the address of the heap variable will remain the same in the predicted run.
+%To simplify the representation, we introduce a stack variable $l_{address(o.f)}$ to replace the heap variable. The replacement is conducted before the construction of the symbolic trace.
+%
+%An important assumption for the replacement is that the address of the heap variable will remain the same in the predicted run.
 We will guarantee the validity of the assumption, as explained in Section.
-
-%The SSA form of the trace is different from the SSA form of the instruction as the SSA instruction can only distinguish definitions at different program points but cannot distinguish the definitions at different execution points that share the same program point.
-
-\begin{algorithmic}[3]
-\State $ins=e.ins$
-\For {$e: \tau$}
-  \For {$i: 0\dots ins.uses.length-1$}
- \State $e.\_uses[i] \gets table.get(ins.uses[i])$
- \EndFor
-  \If {$e.\_def\neq null$}
-    \State $ e.\_def \gets ins.def^{e.id}$
-	\State $ table.put(ins.def,  e.\_def)$
- \EndIf
-\EndFor
-\end{algorithmic}
-
-%TODO update the intro+moti, make sure the same style
-\begin{figure}
-\centering
-\begin{tabular}{ll}
-\multicolumn{2}{c}{{\tt {\bf x} = 0; {\bf y} = 0;}} \\
-\hline
-\multicolumn{1}{c}{$T_1$} & \multicolumn{1}{c}{$T_2$} \\
-\hline
-{\tt 1: s=0; } & \\
-{\tt 2: for(i=1;i<3;i++)} & \\
-{\tt 3: \ \ \ s+=i;} & \\
-{\tt 4: {\bf y} = s;} & \\
-{\tt 5: {\bf x} = 1;} & \\
-{\tt 6: {\bf y} = 5;} & \\
-& {\tt 7: if ({\bf y} > 2)} \\
-& {\tt 8:~~print({\bf x}+1);} \\	
-& {\color{Gray} {\tt 9: else}} \\
-& {\color{Gray} {\tt 10:~~print({\bf x}+2);}}
-\end{tabular}
-\caption{Running Example (shared variables are in bold font). }
-\label{fig:running2}
-\end{figure}
-
-{\bf Shared Variables}   Besides, we introduce symbols to represent shared variables read or written.
+%
+%%The SSA form of the trace is different from the SSA form of the instruction as the SSA instruction can only distinguish definitions at different program points but cannot distinguish the definitions at different execution points that share the same program point.
+%
+%
+%
+%
+%
+%\begin{algorithmic}[3]
+%\State $ins=e.ins$
+%\For {$e: \tau$}
+%  \For {$i: 0\dots ins.uses.length-1$}
+% \State $e.\_uses[i] \gets table.get(ins.uses[i])$
+% \EndFor
+%  \If {$e.\_def\neq null$}
+%    \State $ e.\_def \gets ins.def^{e.id}$
+%	\State $ table.put(ins.def,  e.\_def)$
+% \EndIf
+%\EndFor
+%\end{algorithmic}
+%
+%
+%%TODO update the intro+moti, make sure the same style
+%\begin{figure}
+%\centering
+%\begin{tabular}{ll}
+%\multicolumn{2}{c}{{\tt {\bf x} = 0; {\bf y} = 0;}} \\
+%\hline
+%\multicolumn{1}{c}{$T_1$} & \multicolumn{1}{c}{$T_2$} \\
+%\hline
+%{\tt 1: s=0; } & \\
+%{\tt 2: for(i=1;i<3;i++)} & \\
+%{\tt 3: \ \ \ s+=i;} & \\
+%{\tt 4: {\bf y} = s;} & \\
+%{\tt 5: {\bf x} = 1;} & \\
+%{\tt 6: {\bf y} = 5;} & \\
+%& {\tt 7: if ({\bf y} > 2)} \\
+%& {\tt 8:~~print({\bf x}+1);} \\	
+%& {\color{Gray} {\tt 9: else}} \\
+%& {\color{Gray} {\tt 10:~~print({\bf x}+2);}}
+%\end{tabular}
+%\caption{Running Example (shared variables are in bold font). }
+%\label{fig:running2}
+%\end{figure}
+%
+%
+%
+%{\bf Shared Variables}   Besides, we introduce symbols to represent shared variables read or written.
 For each shared variable $x$ read (or written), we introduce $R^{id}_x$ (or $W^{id}_x$) to denote it, where $id$ is the event id.
 We denote the read and the write with different symbols, which allow us to reason about different dependence relations among them flexibly.
-
-Consider the code in Figure~\ref{fig:running2}, which resembles the example in Figure~\ref{fig:running} except that it includes a for loop at lines 1-2.
-Figure~\ref{fig:t4running2} shows the instruction $e.ins$ of each event $e$ and its symbolic instruction $e.\_ins$, which is essentially replaces the variables in $e.ins$ to the symbolic variables.
-
-%TODO distinguish T1 and T2 in the text. 
-\begin{figure}
-\centering
-\begin{tabular}{l|l}
-\hline
-\multicolumn{1}{c}{$Trace$} & \multicolumn{1}{c}{$Symbolic\  Trace$} \\
-\hline
-{\tt 0: {\bf x}=0} &  {\tt 0: $W^0_x$=0}    \\
-{\tt 1: {\bf y}=0} &   {\tt 1: $W^1_y$=0}   \\
-{\tt 2: s=0} &  {\tt 2: $s^2$=0}   \\
-{\tt 3: i=1} &     {\tt 3: $i^3$=1}   \\
-{\tt 4: i<3} &    {\tt 4: $i^3$<3} \\
-{\tt 5: s=s+i} & {\tt 5: $s^5$=$s^2$+$i^3$}   \\
-{\tt 6: i=2} &       {\tt 6: $i^6$=2}  \\
-{\tt 7: i<3} &      {\tt 7: $i^6$<3}  \\
-{\tt 8: s=s+i} &  {\tt 8: $s^8$=$s^5$+$i^6$}  \\
-{\tt 9: i=3} &     {\tt 9: $i^9$=3}  \\
-{\tt 10: i<3} &    {\tt 10: $i^9$<3}  \\
-{\tt 11: {\bf y} = s;} &  {\tt 11: $W^{11}_y$ = $s^8$;}  \\
-{\tt 12: {\bf x} = 1;} &    {\tt 12: $W^{12}_x$ = 1;}   \\
-{\tt 13: {\bf y} = 5;} &    {\tt 13: $W^{13}_y$ = 5;}  \\
-{\tt 14: {\bf y} > 2}  &    {\tt 14: $R^{14}_y$ > 2} \\
-{\tt 15: tmp={\bf x}+1;}  & {\tt 15: tmp=$R^{15}_x$+1;}   \\	
-{\tt 16: print(tmp);} &  {\tt 16: print(tmp);}  \\
-\end{tabular}
-\caption{Trace}
-\label{fig:t4running2}
-\end{figure}
-
-{\bf Method Calls\ } To support the method calls, we need to capture the value flow from the actual argument to the formal argument, and the flow from the return variable to the variable definition at the call site.
+%
+%
+%Consider the code in Figure~\ref{fig:running2}, which resembles the example in Figure~\ref{fig:running} except that it includes a for loop at lines 1-2.
+%Figure~\ref{fig:t4running2} shows the instruction $e.ins$ of each event $e$ and its symbolic instruction $e.\_ins$, which is essentially replaces the variables in $e.ins$ to the symbolic variables. 
+% 
+%
+%
+%%TODO distinguish T1 and T2 in the text. 
+%\begin{figure}
+%\centering
+%\begin{tabular}{l|l}
+%\hline
+%\multicolumn{1}{c}{$Trace$} & \multicolumn{1}{c}{$Symbolic\  Trace$} \\
+%\hline
+%{\tt 0: {\bf x}=0} &  {\tt 0: $W^0_x$=0}    \\
+%{\tt 1: {\bf y}=0} &   {\tt 1: $W^1_y$=0}   \\
+%{\tt 2: s=0} &  {\tt 2: $s^2$=0}   \\
+%{\tt 3: i=1} &     {\tt 3: $i^3$=1}   \\
+%{\tt 4: i<3} &    {\tt 4: $i^3$<3} \\
+%{\tt 5: s=s+i} & {\tt 5: $s^5$=$s^2$+$i^3$}   \\
+%{\tt 6: i=2} &       {\tt 6: $i^6$=2}  \\
+%{\tt 7: i<3} &      {\tt 7: $i^6$<3}  \\
+%{\tt 8: s=s+i} &  {\tt 8: $s^8$=$s^5$+$i^6$}  \\
+%{\tt 9: i=3} &     {\tt 9: $i^9$=3}  \\
+%{\tt 10: i<3} &    {\tt 10: $i^9$<3}  \\
+%{\tt 11: {\bf y} = s;} &  {\tt 11: $W^{11}_y$ = $s^8$;}  \\
+%{\tt 12: {\bf x} = 1;} &    {\tt 12: $W^{12}_x$ = 1;}   \\
+%{\tt 13: {\bf y} = 5;} &    {\tt 13: $W^{13}_y$ = 5;}  \\
+%{\tt 14: {\bf y} > 2}  &    {\tt 14: $R^{14}_y$ > 2} \\
+%{\tt 15: tmp={\bf x}+1;}  & {\tt 15: tmp=$R^{15}_x$+1;}   \\	
+%{\tt 16: print(tmp);} &  {\tt 16: print(tmp);}  \\
+%\end{tabular}
+%\caption{Trace}
+%\label{fig:t4running2}
+%\end{figure}
+%
+%
+%
+%{\bf Method Calls\ } To support the method calls, we need to capture the value flow from the actual argument to the formal argument, and the flow from the return variable to the variable definition at the call site.
 We capture the value flow by recording additional events for each method call.
-Consider the example in Figure~\ref{fig:methcall},   we need to record events to capture the implicit assignment $y1=y$ and $x=i2$ due to the method call.
-
-Our strategy is as follows.
+%Consider the example in Figure~\ref{fig:methcall},   we need to record events to capture the implicit assignment $y1=y$ and $x=i2$ due to the method call. 
+%
+%Our strategy is as follows.
 To capture the argument flow $y1=y$, we record two events, one capturing $ARG0=y$ and the other capturing $y1=ARG0$, where $ARG0$ is an artificial variable denoting the first argument of any method.
 This simple strategy helps handle the virtual method calls.
-At the call site, we do not know which target method would be called, the first event records  information about the actual argument.
+%At the call site, we do not know which target method would be called, the first event records  information about the actual argument.
 Then, a target method is invoked, the second event records information about the formal argument of the resolved target method.
@@
 The recorded events also undergo the above symbolicization process for gaining the SSA form.
-
-%The above simple strategy however hides the complexity of the virtual method calls.
+%
+%
+%
+%
+%%The above simple strategy however hides the complexity of the virtual method calls.
 At a call site of a virtual method, the static instrumentation cannot know precisely which method would be called.
@@
 We do not know how to instrument the code statically, $y1=y$ or $y2=y$.
-%
-%
-%To avoid the problem, we have to combine the runtime knowledge. 
-%Our strategy is as follows: rather than directly record direct value flow from actual argument to formal argument, we introduce an artificial variable during the static instrumentation.
+%%
+%%
+%%To avoid the problem, we have to combine the runtime knowledge. 
+%%Our strategy is as follows: rather than directly record direct value flow from actual argument to formal argument, we introduce an artificial variable during the static instrumentation.
 Then we  insert the instrumentation $record(ARG0=y;)$ at call site, and insert  $record(y1=ARG0)$ at the entry of the method $func$ declared in the first class, and insert $record(y2=ARG0);$ at the entry of the method $func$ declared in the second class.
@@
 We model the return value flow similarly.
+%%
+%%Note that although different methods use the same names for the artificial variable,  they are translated to different variables after we get the SSA form of symbolic trace. 
 %
-%Note that although different methods use the same names for the artificial variable,  they are translated to different variables after we get the SSA form of symbolic trace.
-
-\begin{figure}
-\centering
-\begin{tabular}{ll}
-{\tt  x=o.func(y) } &  \\ 
- {\tt  func(y1)\{ // class O1} &  \\
- {\tt  \ \      i=y1; } & \\
- {\tt  \ \      i2=2*i; } &  \\
- {\tt  \  \      return i2;} & \\ 
- {\tt           \}} & \\ 
-  {\tt //  func(y2)\{ // class O2} &  \\
-   {\tt //    ...     } & \\
-
-% {\tt  // \ \      j=y2; } & \\
-% {\tt  // \ \      j2=3*j; } &  \\
-% {\tt  // \  \      return j;} & \\ 
-% {\tt  //         \}} & \\ 
-\end{tabular}
-\caption{Method Calls}
-\label{fig:methcall}
-\end{figure}
-
-Until now, we get the trace with the symbolic variables. The following section uses these symbolic representations to compute races. 
-Hereafter, we always refer to the symbolic form, such as symbolic variable or symbolic instruction, unless otherwise specified. 
-%Unless otherwise specified, we refer to the symbolic variables
+% 
+%
+%\begin{figure}
+%\centering
+%\begin{tabular}{ll}
+%{\tt  x=o.func(y) } &  \\ 
+% {\tt  func(y1)\{ // class O1} &  \\
+% {\tt  \ \      i=y1; } & \\
+% {\tt  \ \      i2=2*i; } &  \\
+% {\tt  \  \      return i2;} & \\ 
+% {\tt           \}} & \\ 
+%  {\tt //  func(y2)\{ // class O2} &  \\
+%   {\tt //    ...     } & \\ 
+%  
+%  
+%% {\tt  // \ \      j=y2; } & \\
+%% {\tt  // \ \      j2=3*j; } &  \\
+%% {\tt  // \  \      return j;} & \\ 
+%% {\tt  //         \}} & \\ 
+%\end{tabular}
+%\caption{Method Calls}
+%\label{fig:methcall}
+%\end{figure}
+%
+%
+%
+%
+%
+%
+%Until now, we get the trace with the symbolic variables. The following section uses these symbolic representations to compute races. 
+%Hereafter, we always refer to the symbolic form, such as symbolic variable or symbolic instruction, unless otherwise specified. 
+%%Unless otherwise specified, we refer to the symbolic variables
 

### 2014-08-29 trace.tex

@@
 						\multicolumn{1}{l}{{\tt p} $::=$} & \\
-					    $\overline{\texttt{m($\overline{{\tt x}}$) \{ s \}}}$ & {\bf (method)} \\
+					    $\overline{\texttt{m($\overline{{\tt x}}$) \{ s \}}}$ & {\bf (mdcl)} \\
 						\multicolumn{1}{l}{{\tt s} $::=$} & \\
-						{\tt x = y}\ $|$\ {\tt x = $c$} $|$\ & {\bf (assignment)} \\
-						{\tt x = new()} $|$\ & {\bf (allocation)} \\
-						{\tt x = f($\overline{{\tt y}}$)} $|$\ & {\bf (invocation)} \\
+						{\tt x = y}\ $|$\ {\tt x = $c$} $|$\ & {\bf (asgn)} \\
+						{\tt x = new()} $|$\ & {\bf (alloc)} \\
+						{\tt x = f($\overline{{\tt y}}$)} $|$\ & {\bf (call)} \\
 						{\tt return x} $|$\ & {\bf (return)} \\
 						{\tt z = x \{ $+$,$-$,$\times$,$/$ \} y}\ $|$\ & {\bf (aexp)} \\
-						{\tt b = x \{ $<$,$\leq$,$\equiv$,$\geq$,$>$ \} y}\ $|$\ & {\bf (comparison)} \\
+						{\tt b = x \{ $<$,$\leq$,$\equiv$,$\geq$,$>$ \} y}\ $|$\ & \\
 						{\tt b = bx $\wedge$ by}\ $|$\ {\tt b = bx $\vee$ by} $|$\ & {\bf (bexp)} \\
 						{\tt y = x.f} $|$\ {\tt x.f = y}\ $|$\ & {\bf (heap)} \\
-						{\tt if (b) \{ s \} $|$\ {\tt while (b) \{  s \}} $|$\ {\tt s ; s }} $|$\ & {\bf (control)} \\
+						{\tt if (b) \{ s \} $|$\ {\tt while (b) \{  s \}} $|$\ {\tt s ; s }} $|$\ & {\bf (ctrl)} \\
 						{\tt lock(x)}\ $|$\ {\tt unlock(x)} & {\bf (sync)}
@@
 Instead of recording only the version of a variable, we also represent as part of the variable's identifier its enclosing method and its invocation counter.
-Thus, symbol ${\tt m}^2 \colon {\tt v}^3$ is interpreted as the definition of variable ${\tt v}$ at offset 3 within method ${\tt m}$ during its second invocation.
+Thus, symbol ${\tt m}^2 \colon {\tt v}^7$ is interpreted as the definition of local variable ${\tt v}$ within the second invocation of method ${\tt m}$ that occurs at trace index 7.
 

### 2014-08-30 tech.tex
eval finished

@@
 
+One important requirement is that we need to preserve the same branch decisions for all preceding branches, rather than those of the guarding branches. 
+Consider the following code, where $sx$ and $sy$ are shared variables with the initial values 0 and $tmp$ is a local variable with initial value 0, there are two branches, one is the guarding branch for update of the shared variable at line 4, the other is not.
+Suppose in the observation run, the first branch takes the false branch (hence, $tmp=5$ is not included in the trace), and the second takes the true branch.
+If the solver only guarantees the same branch decision for the guarding branch, i.e., the second, it risks reporting the false feasibility of line 4.
+Specifically, line 4 is infeasible if the first branch is after some remote update that forces it to take the true branch.
+However, in this case, the solver still judges line 4 as feasible as it does not model the true branch of the first branch, which is not in the trace.
+Therefore, we need to respect all branch decisions so that the
+set of events under reasoning are all captured in the trace.
+
+ \begin{figure}
+\centering
+\begin{tabular}{ll}
+               & {\tt // sx=4;}  \\ 
+{\tt if(sx>3) } &  \\ 
+ {\tt \ \ \  {tmp=5;}} &  \\
+ {\tt if(tmp$\neq$ 5 ) } & \\
+  {\tt {sy=0;}} & \\
+\end{tabular}
+\caption{Method Calls}
+\label{fig:methcall}
+\end{figure}
+
 {\bf Variable definition constraints\ } The variable definition constraint captures the value flow due to each statement from the right hand side to the left hand side.
@@
 {\bf Array access\ } For array access event $arr[i]=x$, we treat them as the field access event $arr.i=x$, where the index $i$ is treated as a field. 
+We ensure the predicted run sees the same index at the same event by specifying the additional constraint $i==v$, where $v$ is the concrete index value observed.
 
@@
 Only those selected in the slices are preserved while the rest events are removed.
+
+\subsection{Scalability Issues}
+Our approach suffers from scalability issues for two reasons.
+First, as we record local access events in the monitoring run,  the monitoring run takes long time, e.g., more than 2 hours, and our trace is typically large, e.g., at the GB level.
+To avoid the out of memory issue during the monitoring run, we store the event to the database on the fly, rather than storing them in a buffer.
+Besides, we terminate the monitoring run with a shutdown hook if it is beyond 30 minutes.
+Second,  the prediction phase needs to load the trace to memory, which easily causes out of memory problem.
+To solve the problem, we separate the analysis into several runs, each run restarting the JVM and resuming from the window of last run.
+Different runs communicate the window id using the external file.
+These runs are organized together automatically using the ant.
+The outputs of the files are merged into one summary finally.
+Besides,

### 2014-09-01 constraints.tex
more changes to tech part

@@
 
-We begin with the basic setting, which enforces strict reordering of the events along the input trace without diverging into new execution paths.
-We relax this requirement later, such that different branches can be followed as long as feasibility is retained.
+We begin with the basic setting, which enforces conservative reordering of the events along the input trace without diverging into new dependence structures or execution paths.
+We relax this requirement later, such that different data-flow paths and branches can be followed as long as feasibility is retained.
 
@@
 
-The final requirement for the basic setting is that left-hand variables are defined according to the same right-hand variables as before.
+A final requirement for the basic setting is that left-hand variables are defined according to the same right-hand variables as before.
 That is, version $i$ of variable ${\tt u}$ is defined as version $j$ of variable ${\tt v}$ in input trace $t$, then the same remains true in predicted  trace $t'$.
@@
 
-\subsection{Unexplored Paths}
+\subsection{Relaxations}
+
+We now move to the novel feature of \tool, which is its ability to explore execution schedules that depart from the data flow exhibited in the original trace.
+More precisely, \tool\ is able to relax flow dependencies in the original trace, whereby a thread reads a shared memory location written by another thread, while enforcing feasibility.
+This is strictly beyond the coverage potential of existing predictive analyses, which restrict trace transformations to ones where any read access to a shared memory location must correspond to the same write access as in the original trace.
+
+\paragraph{Relaxation of Flow Dependencies}
+
+To ensure feasibility under relaxation of flow dependencies, we need to secure the link between the execution schedule and the write/read flow.
+As an illustration from the example in Figure \ref{fig:running}, $R_{\tt y}^4={\tt y}^1 \wedge O_1 < O_4 < O_3$ specifies that in a schedule where thread $T_1$ executes line {\tt 1}, then $T_2$ executes line {\tt 4}, and then the schedule switches again to $T_1$ to execute line {\tt 3}, the read access to ${\tt y}$ at line {\tt 4} obtains the value assigned to ${\tt y}$ at line {\tt 1}: $R_{\tt y}^4={\tt y}^1$.
+
+The full and general constraint formula, given read $R_{\ell}$ of location $\ell$ as part of event $e$ with set ${\cal W}$ of matching write events (i.e., events including write access to $\ell$), takes the following form:
+$$
+\begin{array}{rll}
+\bigvee_{e_w \in {\cal W}} &  & (R_{\ell} = {\ell}^{{\sf index}\ e_w}) \\
+&		\bigwedge 	&  O(e_w) < O(e) \\
+&		\bigwedge_{e' \in {\cal W} \setminus \{ e_w \}} & (O(e') < O(e_w) \vee O(e) < O(e'))
+\end{array}
+$$
+This disjunctive formula iterates over all matching write events, and demands for each that (i) it occurs prior to the read event ($O(e_w) < O(e)$) and (ii) all other write events either occur before ($O(e') < O(e_w)$) it or after the read event
+($O(e) < O(e')$).
+
+An important concern that arises due to relaxation of flow dependencies is that heap accesses may change their meaning.
+As an illustration, we refer to Figure \ref{fig:heapAccess}.
+While the read at line {\tt 7} appears to match the write at line {\tt 5}, this is conditioned on the read at line {\tt 6} being linked to the assignment at line {\tt 4}.
+If the predicted run violates this link, then feasibility is no longer guaranteed.
+In particular, if reordering results in {\tt z} being assigned the first rather than second allocated object, then the write at line {\tt 5} no longer matches the read at line {\tt 7}.
+
+\begin{figure}
+	\centering
+	\begin{tabular}{ll}
+		\hline
+		\multicolumn{1}{c}{$T_1$} & \multicolumn{1}{c}{$T_2$} \\
+		\hline
+		{\tt 1: x1 = new();} & \\
+		{\tt 2: x2 = new();} & \\
+		{\tt 3: y = x1;} & \\
+		{\tt 4: y = x2;} & \\
+		{\tt 5: x2.f = 5;} & \\	
+		& {\tt 6: z = y;} \\
+		& {\tt 7: w = z.f;} \\
+	\end{tabular}
+	\caption{\label{fig:heapAccess}Example illustrating the need to account for heap accesses during trace transformation}
+\end{figure}
+
+To address this challenge, we enhance the constraint system with the requirement that heap objects that are dereferenced before the candidate racing events retain their original address in the predicted trace.
+This achieves three guarantees: First, matching events in the original trace are guaranteed to also match in the predicted trace.
+Second, candidate races in the original trace remain viable in the predicted trace.
+Finally, in a practical setting involving virtual method calls (which goes beyond our core grammar in Table \ref{Ta:syntax}), call-site resolutions are the same across the original and predicted traces.
+Notice that in this setting, we consider as relevant not only the targets of field dereferences but also the targets of virtual method invocations.
+
+Formally, given pair $\tau$ and $\tau'$ of candidate racing events such that ${\sf index}\ \tau < {\sf index}\ \tau'$, 
+we require that
+$$
+\bigwedge_{\tau'' \in t \cap {\bf heapr}.\
+	\tau'' \in {\sf pre}\ \tau'} \lsyn {\sf stmt}\ \tau'' \rsyn\ t \equiv \lsyn {\sf stmt}\ \tau'' \rsyn\ t'
+$$ 
+That is, all heap dereferences up to the later of the candidate racing events must retain their original meaning in the predicted trace $t'$.
+This ensures in particular that the base reference remains identical to the one in the original trace $t$.
+
+\paragraph{Relaxation of Branching Decisions}
+
+Beyond relaxing flow dependencies, \tool\ is also capable of exploring code branches that were not executed in the original trace.
+This is achieved via a symbolic representation of the input trace.
+Given branching event $e_b$,
+\begin{enumerate}
+	\item model the branching condition symbolically to execute along the negation of the original branch;
+	\item apply depth-first search (DFS) to uncover all execution suffixes under the unexplored branch (which may itself contain branching statements); and
+	\item for each suffix $t^s$,
+	\begin{enumerate}
+		\item truncate the original trace at $e_b$ {\tt s} yielding $t_s$; and
+		\item concatenate $t_s$ with the negation of $e_b$ followed by $t^s$.
+	\end{enumerate}
+\end{enumerate}
+Constraint solving is applied to each of the resulting traces analogously to the original trace.
+
+We emphasize that exploration of new execution paths is subject to all the known limitations of symbolic execution, including in particular loop structures and object allocation.
+\tool\ currently fails if (i) the unexplored branch containts loops or (ii) there are object references that cannot be fully resolved at the branching point.
+In case of failure, \tool\ moves on to other branches.
+We demonstrate in Section \ref{sec:eval} that despite these limitations, the increase in coverage thanks to exploration of new paths is significant.

### 2014-09-01 trace.tex
more changes to tech part

@@
 						{\tt b = bx $\wedge$ by}\ $|$\ {\tt b = bx $\vee$ by} $|$\ & {\bf (bexp)} \\
-						{\tt y = x.f} $|$\ {\tt x.f = y}\ $|$\ & {\bf (heap)} \\
+						{\tt y = x.f} $|$\ & {\bf (heapr)} \\ 
+						{\tt x.f = y}\ $|$\ {\tt x.f = $c$}\ $|$\ & {\bf (heapw)} \\
 						{\tt if (b) \{ s \} $|$\ {\tt while (b) \{  s \}} $|$\ {\tt s ; s }} $|$\ & {\bf (ctrl)} \\

### 2014-09-02 constraints.tex
more changes to tech part
edits

@@
 					& {\sf thread}\ \tau \neq {\sf thread}\ \tau' \\
-\bigwedge 	& ({\sf writes}\ \tau\ \ell \vee \tau'\ {\tt s}'\ \ell) \\
+\bigwedge 	& (\ell \in {\sf writeset}\ \tau \vee \ell \in {\sf writeset}\ \tau') \\
 \bigwedge   & O_{\tau} = O_{\tau'}
@@
 $$
-That is, (i) events $\tau$ and $\tau'$ are executed by different threads, (ii) at least one of the events performs write access to $\ell$, and (iii) the events occur at the simultanesouly.
+That is, (i) events $\tau$ and $\tau'$ are executed by different threads, (ii) at least one of the events performs write access to $\ell$ (as judged by the ${\sf writeset}$ membership check), and (iii) the events occur simultanesouly.
 This is a direct logical encoding of the definition of a race condition.
@@
 	\begin{enumerate}
-		\item truncate the original trace at $e_b$ {\tt s} yielding $t_s$; and
+		\item truncate the original trace at $e_b$ yielding prefix $t_s$; and
 		\item concatenate $t_s$ with the negation of $e_b$ followed by $t^s$.
@@
 We demonstrate in Section \ref{sec:eval} that despite these limitations, the increase in coverage thanks to exploration of new paths is significant.
+
+\subsection{Proof of Correctness}
+
+\input{proof}

### 2014-09-02 constraints.tex
edits

@@
 $$
-\bigwedge_{\tau'' \in t \cap {\bf heapr}.\
-	\tau'' \in {\sf pre}\ \tau'} \lsyn {\sf stmt}\ \tau'' \rsyn\ t \equiv \lsyn {\sf stmt}\ \tau'' \rsyn\ t'
+\bigwedge_{\tau'' \equiv {\tt y=x.f} \in t \cap {\bf heapr}.\
+	\tau'' \in {\sf pre}\ \tau'} {\sf env}\ \sigma(t,\tau'')\ {\tt x} = {\sf env}\ \sigma(t',\tau'')\ {\tt x}
 $$
-That is, all heap dereferences up to the later of the candidate racing events must retain their original meaning in the predicted trace $t'$.
-This ensures in particular that the base reference remains identical to the one in the original trace $t$.
+where ${\sf env}$ is the state mapping from local variables to their value and $\sigma(t,\tau'')$ 
+(resp. $\sigma(t',\tau'')$) is the state arising at trace $t$ (resp. $t'$) immediately before event $\tau''$.
+This constraint fixes that all heap dereferences up to the later of the candidate racing events retain their original base object as in the predicted trace $t'$.
 

### 2014-09-02 introduction.tex
intro

@@
 %Fasttrack PLDI paper claims happens-before analysis never reports false positives. Check.
-\paragraph{Existing Approaches} Initial attempts to address this challenge focused on built-in synchronization primitives.
+\paragraph{Existing Approaches} 
+Numerous techniques have been proposed to date for race detection, which either sacrifice soundness~\cite{ucg,XXX,YYY,ZZZ}  or have significant coverage limitations~\cite{XXX,YYY,ZZZ}.
+Recently \emph{predictive analysis} has emerged as a promising alternative, whereby a single trace is considered, and races are discovered by permuting the execution schedule governing the trace \cite{JEFF,YANNIS}.
+
+Predictive analysis guarantees soundness, and has the potential for high coverage, though current predictive analyses suffer from two major limitations:
+\begin{enumerate}
+	\item All the permutations attempted by the analysis must preserve the flow dependencies exhibited by the original trace.
+As an example, if in the original trace thread $t_1$ reads a shared variable that was written by thread $t_2$, then the same must hold in the permuted trace.
+	\item The analysis cannot step outside the boundaries of the input trace, e.g. by exploring branches that were not followed in that trace.
+\end{enumerate}
+Both of these constraints are a conservative means of ensuring soundness.
+In an empirical study we conducted, which we describe in Section \ref{sec:eval}, we found that these restrictions often lead to serious loss in coverage.
+
+As an illustration, we refer to the trace in Figure \ref{fig:running}, where the statements in red denote an unexplored branch outside the trace.
+Existing predictive analyses are unable to detect the race between lines {\tt 2} and {\tt 5} because of the dependence between lines {\tt 3} and {\tt 4}, which is a barrier to the needed reorderings.
+A second race that is missed, as it involves the statements in red, is between lines {\tt 2} and {\tt 7}.
+
+%Initial attempts to address this challenge focused on built-in synchronization primitives.
 These include locks as well as the wait/notify and start/join scheduling controls.
@@
 Because the derived constraints are partial, permitting certain infeasible event reorderings, lockset analysis cannot guarantee soundness.
-
-A different tradeoff is struck by the \emph{happens-before (HB)} approach.
+%
+%A different tradeoff is struck by the \emph{happens-before (HB)} approach.
 In this style of analysis, all synchronization primitives are accounted for, though reordering constraints are conservative.
@@
 Among these are hybrid analysis, which permits both orderings of lock-synchronized blocks, and the \emph{Universal Causal Graph (UCG)}~\cite{ucg} representation, which also enables both orderings but only if these are consistent with wait/notify- and start/join-induced constraints.
-
-Unfortunately, even full consideration of synchronization constructs is insufficient.
+%
+%Unfortunately, even full consideration of synchronization constructs is insufficient.
 Branching decisions must also be respected to ensure the feasibility of a reported race.
@@
 
-As an alternative, driven by the requirement for soundness, 
-Smagardakis et al.
+%As an alternative, driven by the requirement for soundness, 
+%Smagardakis et al.
 have recently proposed \emph{predictive analysis}.
 In this approach, both synchronization constraints and inter-thread dependencies are preserved, where inter-thread dependencies are respected by
-only allowing reorderings that leave the dependence structure exhibited by the original trace in tact.
+%only allowing reorderings that leave the dependence structure exhibited by the original trace in tact.
 This ensures that the values of shared memory locations remain the same, which secures the soundness argument.
@@
 \cite{XXX} permit reorderings as long as control dependencies are respected.
+%
+%While being sound and able to find real races, existing predictive analyses are conservative in two respects:
+%\begin{enumerate}
+%	\item Constraints imposed by branch conditions are abided by preserving the exact values flowing into the condition as in the original trace.
+%	\item A second limitation, implied by the first, is that traces that diverge from the original trace in their branching history are strictly out of scope. 
+%\end{enumerate}
+% Due to these restrictions, enforced as a conservative means of ensuring soundness, existing predictive analysis suffer from limited coverage.
 
-While being sound and able to find real races, existing predictive analyses are conservative in two respects:
-\begin{enumerate}
-	\item Constraints imposed by branch conditions are abided by preserving the exact values flowing into the condition as in the original trace.
-	\item A second limitation, implied by the first, is that traces that diverge from the original trace in their branching history are strictly out of scope. 
-\end{enumerate}
- Due to these restrictions, enforced as a conservative means of ensuring soundness, existing predictive analysis suffer from limited coverage.
+ \paragraph{Our Approach} We describe a novel approach to predictive race detection, which we have implemented as the \tool\ analysis tool, that is able to relax both of these constraints.
+Compared to the state of the art \cite{JEFF}, \tool\ is able to detect x$1.5$ more races with only the first restriction relaxed, and x$2.5$ more races with both restrictions relaxed.
 
- \paragraph{Our Approach} The departure point of our approach, yielding strictly better coverage, is to consider the explicit values of shared memory locations instead of dependencies between statements.
-The gain in coverage is twofold:
+ The departure point of our approach is to consider the explicit values of shared memory locations instead of dependencies between trace events. The gain in coverage is twofold:
  \begin{enumerate}
@@
 
- Returning to the example in Figure \ref{fig:running}, this time also with reference to the statements in red, we demonstrate how value-based reasoning improves coverage.
-First, unlike predictive analysis, which is unable to detect the race between lines {\tt 2} and {\tt 5} because of the dependence between lines {\tt 3} and {\tt 4}, value-based reasoning reveals that trace
+ Returning to the example in Figure \ref{fig:running}, we demonstrate how value-based reasoning improves coverage. First, value-based reasoning reveals that trace 
  [{\tt 1}, {\tt 4}, {\tt 5}, {\tt 2}, {\tt 3}] is feasible, since the first assignment, {\tt y = 3}, satisfies the condition {\tt y > z}.

### 2014-09-02 constraints.tex
changed structure

--- 2014-09-02 constraints.tex
@@
-\section{Constraint Derivation and Resolution}
+\section{Exploration of Unexecuted Branches}
 
-We now describe in detail the constraint encoding and solving steps.
-
-\subsection{Same Trace}
-
-We begin with the basic setting, which enforces conservative reordering of the events along the input trace without diverging into new dependence structures or execution paths.
-We relax this requirement later, such that different data-flow paths and branches can be followed as long as feasibility is retained.
-
-\paragraph{Intra-thread Order}
-
-The first set of constraints reflects control flow within the individual threads, which must remain unchanged under the reordering transformation.
-Given input trace $t$, this is expressed as the following formula:
-$$
-\begin{array}{rl}
-	\forall \tau,\tau' \in t. & {\sf thread}\ \tau \equiv {\sf thread}\ \tau'. \\
-										& {\sf index}\ \tau < {\sf index}\ \tau' \Rightarrow O_{\tau} < O_{\tau'}
-\end{array}
-$$ 
-The logical variables $O_x$ express ordering constraints. The requirement, as stated above, is that these variables
-reflect the same order as the projection of ${\sf index}$ onto individual threads.
-
-\paragraph{Race Condition}
-
-Given pair $\tau$ and $\tau'$ of events that both access a common memory location $\ell$, we demand that
-$$
-\begin{array}{rl}
-					& {\sf thread}\ \tau \neq {\sf thread}\ \tau' \\
-\bigwedge 	& (\ell \in {\sf writeset}\ \tau \vee \ell \in {\sf writeset}\ \tau') \\
-\bigwedge   & O_{\tau} = O_{\tau'}
-\end{array}
-$$
-That is, (i) events $\tau$ and $\tau'$ are executed by different threads, (ii) at least one of the events performs write access to $\ell$ (as judged by the ${\sf writeset}$ membership check), and (iii) the events occur simultanesouly.
-This is a direct logical encoding of the definition of a race condition.
-
-\paragraph{Path Constraints}
-
-The requirement with respect to branching is that the prefix of the original trace $t$ up to the pair $\tau$ and $\tau'$ of candidate racing events remains identical in the predicted trace $t'$.
-More accurately, under the assumption that ${\sf index}\ \tau < {\sf index}\ \tau'$,
-we require that 
-$$
-	\bigwedge_{\tau'' \in t \cap {\bf bexp}.\
-		\tau'' \in {\sf pre}\ \tau'} \lsyn {\sf stmt}\ \tau'' \rsyn\ t \equiv \lsyn {\sf stmt}\ \tau'' \rsyn\ t'
-$$ 
-where ${\sf stmt}$ is a helper function that obtains the code statement incident in a given transition.
-That is, all branching transitions up to $\tau'$ (which occurs after $\tau$) preserve their boolean interpretation under $t'$.
-This ensures that there are no divergences from the path containing the racing events, though beyond that path any feasible continuation is permitted.
-Importantly, contrary to \cite{JEFF-PLDI14}, we do not pose the requirement that the values flowing into branching statements remain the same, but suffice with the relaxed requirement that the evaluation of branching expressions is invariant under the input and predicted traces.
-
-\paragraph{Variable Definitions}
-
-A final requirement for the basic setting is that left-hand variables are defined according to the same right-hand variables as before.
-That is, version $i$ of variable ${\tt u}$ is defined as version $j$ of variable ${\tt v}$ in input trace $t$, then the same remains true in predicted  trace $t'$.
-This is enforced as the formula
-$$
-	\bigwedge_{\tau'' \in t \cap {\bf asgn}.\
-		\tau'' \in {\sf pre}\ \tau'} {\sf stmt}\ \tau'' \in t' 
-$$
-where we again assume that ${\sf index}\ \tau < {\sf index}\ \tau'$.
-That is, the same statement occurring in $t$ is also present in $t'$ (though the transitions may differ).
-Since the statements of $t'$ are a permutation of the statements of $t$, we are assured that use/def flow is constrained appropriately.
-
-\subsection{Relaxations}
-
-We now move to the novel feature of \tool, which is its ability to explore execution schedules that depart from the data flow exhibited in the original trace.
-More precisely, \tool\ is able to relax flow dependencies in the original trace, whereby a thread reads a shared memory location written by another thread, while enforcing feasibility.
-This is strictly beyond the coverage potential of existing predictive analyses, which restrict trace transformations to ones where any read access to a shared memory location must correspond to the same write access as in the original trace.
-
-\paragraph{Relaxation of Flow Dependencies}
-
-To ensure feasibility under relaxation of flow dependencies, we need to secure the link between the execution schedule and the write/read flow.
-As an illustration from the example in Figure \ref{fig:running}, $R_{\tt y}^4={\tt y}^1 \wedge O_1 < O_4 < O_3$ specifies that in a schedule where thread $T_1$ executes line {\tt 1}, then $T_2$ executes line {\tt 4}, and then the schedule switches again to $T_1$ to execute line {\tt 3}, the read access to ${\tt y}$ at line {\tt 4} obtains the value assigned to ${\tt y}$ at line {\tt 1}: $R_{\tt y}^4={\tt y}^1$.
-
-The full and general constraint formula, given read $R_{\ell}$ of location $\ell$ as part of event $e$ with set ${\cal W}$ of matching write events (i.e., events including write access to $\ell$), takes the following form:
-$$
-\begin{array}{rll}
-\bigvee_{e_w \in {\cal W}} &  & (R_{\ell} = {\ell}^{{\sf index}\ e_w}) \\
-&		\bigwedge 	&  O(e_w) < O(e) \\
-&		\bigwedge_{e' \in {\cal W} \setminus \{ e_w \}} & (O(e') < O(e_w) \vee O(e) < O(e'))
-\end{array}
-$$
-This disjunctive formula iterates over all matching write events, and demands for each that (i) it occurs prior to the read event ($O(e_w) < O(e)$) and (ii) all other write events either occur before ($O(e') < O(e_w)$) it or after the read event
-($O(e) < O(e')$).
-
-An important concern that arises due to relaxation of flow dependencies is that heap accesses may change their meaning.
-As an illustration, we refer to Figure \ref{fig:heapAccess}.
-While the read at line {\tt 7} appears to match the write at line {\tt 5}, this is conditioned on the read at line {\tt 6} being linked to the assignment at line {\tt 4}.
-If the predicted run violates this link, then feasibility is no longer guaranteed.
-In particular, if reordering results in {\tt z} being assigned the first rather than second allocated object, then the write at line {\tt 5} no longer matches the read at line {\tt 7}.
-
-\begin{figure}
-	\centering
-	\begin{tabular}{ll}
-		\hline
-		\multicolumn{1}{c}{$T_1$} & \multicolumn{1}{c}{$T_2$} \\
-		\hline
-		{\tt 1: x1 = new();} & \\
-		{\tt 2: x2 = new();} & \\
-		{\tt 3: y = x1;} & \\
-		{\tt 4: y = x2;} & \\
-		{\tt 5: x2.f = 5;} & \\	
-		& {\tt 6: z = y;} \\
-		& {\tt 7: w = z.f;} \\
-	\end{tabular}
-	\caption{\label{fig:heapAccess}Example illustrating the need to account for heap accesses during trace transformation}
-\end{figure}
-
-To address this challenge, we enhance the constraint system with the requirement that heap objects that are dereferenced before the candidate racing events retain their original address in the predicted trace.
-This achieves three guarantees: First, matching events in the original trace are guaranteed to also match in the predicted trace.
-Second, candidate races in the original trace remain viable in the predicted trace.
-Finally, in a practical setting involving virtual method calls (which goes beyond our core grammar in Table \ref{Ta:syntax}), call-site resolutions are the same across the original and predicted traces.
-Notice that in this setting, we consider as relevant not only the targets of field dereferences but also the targets of virtual method invocations.
-
-Formally, given pair $\tau$ and $\tau'$ of candidate racing events such that ${\sf index}\ \tau < {\sf index}\ \tau'$, 
-we require that
-$$
-\bigwedge_{\tau'' \in t \cap {\bf heapr}.\
-	\tau'' \in {\sf pre}\ \tau'} \lsyn {\sf stmt}\ \tau'' \rsyn\ t \equiv \lsyn {\sf stmt}\ \tau'' \rsyn\ t'
-$$ 
-That is, all heap dereferences up to the later of the candidate racing events must retain their original meaning in the predicted trace $t'$.
-This ensures in particular that the base reference remains identical to the one in the original trace $t$.
+We now switch to the second feature of \tool, which is its ability to reason about unexplored branches.
 
@@
 
-\subsection{Proof of Correctness}
+\paragraph{Proof of Correctness}
 

### 2014-09-02 overview.tex
changed structure

@@
 As input, we assume (i) a program $P$ as well as (ii) a dynamic execution trace of $P$ in Static Single Assignment (SSA) form, such that every variable is defined exactly once.
+
+\subsection{Preliminaries}
+
+Throughout this paper, we assume a standard operational semantics, which defines (i) a mapping ${\sf thr}$ between
+execution threads, each having a unique identifier $i \in \mathbb{N}$, and their respective code, as well as (ii) per-thread stack and shared heap memory.
+
+The code executed by a given thread follows the syntax in Table \ref{Ta:syntax}.
+The text of a program is a sequence of zero or more method declarations, as given in the definition of symbol {\tt p}.
+Methods accept zero or more arguments $\overline{{\tt x}}$, have a body ${\tt s}$, and may have a return value (which we leave implicit).
+For simplicity, we avoid from static typing as well as virtual methods.
+The body of a method consists of the core grammar for symbol {\tt s}.
+We avoid from specifying syntax checking rules, as the grammar is fully standard.
+
+For simplicity, we assume that in the starting state each thread points to a parameter-free method.
+In this way, we can simply assume an empty starting state (i.e., an empty heap), and eliminate complexities such as user-provided inputs and initialization of arguments with default values.
+These extensions are of course possible, and in practice we handle these cases, but reflecting them in the formalism would result in needless complications.
+
+As is standard, we assume an interleaved semantics of concurrency. A \emph{transition}, or \emph{event}, is of the form
+$\sigma \stackrel{i / {\tt s}}{\longrightarrow} \sigma'$, denoting that thread $i$ took an evaluation step
+in prestate $\sigma$, wherein atomic
+statement {\tt s} was executed, which resulted in poststate $\sigma'$. We refer to a sequence of transitions from
+the starting state to either an exceptional state (e.g., due to null dereference) or a state where all the threads have 
+reduced their respective code to $\epsilon$ as a \emph{trace}.
+
+\begin{table}
+	\begin{center}
+		\begin{tabular}{rl}
+			\multicolumn{1}{l}{{\tt p} $::=$} & \\
+			$\overline{\texttt{m($\overline{{\tt x}}$) \{ s \}}}$ & {\bf (mdcl)} \\
+			\multicolumn{1}{l}{{\tt s} $::=$} & \\
+			{\tt x = y}\ $|$\ {\tt x = $c$} $|$\ & {\bf (asgn)} \\
+			{\tt x = new()} $|$\ & {\bf (alloc)} \\
+			{\tt x = f($\overline{{\tt y}}$)} $|$\ & {\bf (call)} \\
+			{\tt return x} $|$\ & {\bf (return)} \\
+			{\tt z = x \{ $+$,$-$,$\times$,$/$ \} y}\ $|$\ & {\bf (aexp)} \\
+			{\tt b = x \{ $<$,$\leq$,$\equiv$,$\geq$,$>$ \} y}\ $|$\ & \\
+			{\tt b = bx $\wedge$ by}\ $|$\ {\tt b = bx $\vee$ by} $|$\ & {\bf (bexp)} \\
+			{\tt y = x.f} $|$\ & {\bf (heapr)} \\ 
+			{\tt x.f = y}\ $|$\ {\tt x.f = $c$}\ $|$\ & {\bf (heapw)} \\
+			{\tt if (b) \{ s \} $|$\ {\tt while (b) \{  s \}} $|$\ {\tt s ; s }} $|$\ & {\bf (ctrl)} \\
+			{\tt lock(x)}\ $|$\ {\tt unlock(x)} & {\bf (sync)}
+		\end{tabular}
+	\end{center}
+	\caption{\label{Ta:syntax}Language syntax}
+\end{table}
+
+We make use of the following helper functions:
+\begin{itemize}
+	\item ${\sf proj}\ t\ i$ projects trace $t$ onto all transitions involving thread $i$.
+	\item $t[k]$ obtains the $k$th transition within trace $t$.
+	\item ${\sf index}\ t\ \tau$ retrieves the index, or offset, of transition $\tau$ within trace $t$. When simply writing
+	${\sf index}\ \tau$ (while omitting the trace parameter) we refer to the index of $\tau$ within the original trace. 
+	\item ${\sf pre}\ t\ \tau$ is the prefix of trace $t$ preceding transition $\tau$. For the suffix beyond $\tau$, we 
+	use ${\sf post}\ t\ \tau$. Finally, ${\sf bet}\ t\ \tau_1\ \tau_2$ returns the transitions delimited by $\tau_1$ and $\tau_2$.
+\end{itemize}
+
+\subsection{Constraint System}
 

### 2014-09-02 trace.tex
changed structure

--- 2014-08-29 trace.tex
@@
-\section{Trace Encoding}
+\section{Relaxation of Flow Dependencies}
 
@@
 
-In this section, we describe how a concrete execution trace is encoded as a symbolic trace, which enables \tool\ to 
-derive reordering constraints.
+We begin by describing the first relaxation, whereby the dependence structure defined by the original trace is potentially violated while retaining soundness.
 
-\subsection{Preliminaries}
+\subsection{Trace Encoding}
 
-Throughout this paper, we assume a standard operational semantics, which defines (i) a mapping ${\sf thr}$ between
-execution threads, each having a unique identifier $i \in \mathbb{N}$, and their respective code, as well as (ii) per-thread stack and shared heap memory.
-
-The code executed by a given thread follows the syntax in Table \ref{Ta:syntax}.
-The text of a program is a sequence of zero or more method declarations, as given in the definition of symbol {\tt p}.
-Methods accept zero or more arguments $\overline{{\tt x}}$, have a body ${\tt s}$, and may have a return value (which we leave implicit).
-For simplicity, we avoid from static typing as well as virtual methods.
-The body of a method consists of the core grammar for symbol {\tt s}.
-We avoid from specifying syntax checking rules, as the grammar is fully standard.
-
-For simplicity, we assume that in the starting state each thread points to a parameter-free method.
-In this way, we can simply assume an empty starting state (i.e., an empty heap), and eliminate complexities such as user-provided inputs and initialization of arguments with default values.
-These extensions are of course possible, and in practice we handle these cases, but reflecting them in the formalism would result in needless complications.
-
-As is standard, we assume an interleaved semantics of concurrency. A \emph{transition}, or \emph{event}, is of the form
-$\sigma \stackrel{i / {\tt s}}{\longrightarrow} \sigma'$, denoting that thread $i$ took an evaluation step
-in prestate $\sigma$, wherein atomic
-statement {\tt s} was executed, which resulted in poststate $\sigma'$. We refer to a sequence of transitions from
-the starting state to either an exceptional state (e.g., due to null dereference) or a state where all the threads have 
-reduced their respective code to $\epsilon$ as a \emph{trace}.
-
-\begin{table}
-	\begin{center}
-	\begin{tabular}{rl}
-						\multicolumn{1}{l}{{\tt p} $::=$} & \\
-					    $\overline{\texttt{m($\overline{{\tt x}}$) \{ s \}}}$ & {\bf (mdcl)} \\
-						\multicolumn{1}{l}{{\tt s} $::=$} & \\
-						{\tt x = y}\ $|$\ {\tt x = $c$} $|$\ & {\bf (asgn)} \\
-						{\tt x = new()} $|$\ & {\bf (alloc)} \\
-						{\tt x = f($\overline{{\tt y}}$)} $|$\ & {\bf (call)} \\
-						{\tt return x} $|$\ & {\bf (return)} \\
-						{\tt z = x \{ $+$,$-$,$\times$,$/$ \} y}\ $|$\ & {\bf (aexp)} \\
-						{\tt b = x \{ $<$,$\leq$,$\equiv$,$\geq$,$>$ \} y}\ $|$\ & \\
-						{\tt b = bx $\wedge$ by}\ $|$\ {\tt b = bx $\vee$ by} $|$\ & {\bf (bexp)} \\
-						{\tt y = x.f} $|$\ {\tt x.f = y}\ $|$\ & {\bf (heap)} \\
-						{\tt if (b) \{ s \} $|$\ {\tt while (b) \{  s \}} $|$\ {\tt s ; s }} $|$\ & {\bf (ctrl)} \\
-						{\tt lock(x)}\ $|$\ {\tt unlock(x)} & {\bf (sync)}
-	\end{tabular}
-	\end{center}
-	\caption{\label{Ta:syntax}Language syntax}
-\end{table}
-
-We make use of the following helper functions:
-\begin{itemize}
-	\item ${\sf proj}\ t\ i$ projects trace $t$ onto all transitions involving thread $i$.
-	\item $t[k]$ obtains the $k$th transition within trace $t$.
-	\item ${\sf index}\ t\ \tau$ retrieves the index, or offset, of transition $\tau$ within trace $t$. When simply writing
-	 ${\sf index}\ \tau$ (while omitting the trace parameter) we refer to the index of $\tau$ within the original trace. 
-	\item ${\sf pre}\ t\ \tau$ is the prefix of trace $t$ preceding transition $\tau$. For the suffix beyond $\tau$, we 
-	use ${\sf post}\ t\ \tau$. Finally, ${\sf bet}\ t\ \tau_1\ \tau_2$ returns the transitions delimited by $\tau_1$ and $\tau_2$.
-\end{itemize}
-
-\subsection{Basic Encoding: Local Accesses}
-
-We describe how \tool\ encodes a trace into symbolic form in two steps. We first start with ``local'' intraprocedural events --- i.e., 
+The first step of the process is to rewrite the concrete execution trace into a symbolic form that enables \tool\ to 
+derive reordering constraints. We describe how \tool\ encodes a trace into symbolic form in two steps. We first start with ``local'' intraprocedural events --- i.e., 
 transitions that manipulate memory accessed by at most a single thread during an invocation-free run --- and then extend our encoding scheme to the entire set of possible events.
 
-\paragraph{SSA Form} The fundamental encoding transformation is to induce Static Single Assignment (SSA) form on the raw trace, such that
+\paragraph{Basic Encoding: Local Accesses}
+
+The fundamental encoding transformation is to induce Static Single Assignment (SSA) form on the raw trace, such that
 a variable is defined exactly once. In this way, def/use chains become explicit, and encoding of trace events as constraints is simplified. As an illustration, trace
@@
 
-\paragraph{Local Heap Accesses} Beyond the default SSA rewriting algorithm, we apply a specialized transformation to handle local heap accesses.
+Beyond the default SSA rewriting algorithm, we apply a specialized transformation to handle local heap accesses.
 While in general determining whether a given heap access is local is an undecidable problem, in the context
@@
 
-\subsection{Full Encoding: Method Calls and Shared Accesses}
+\paragraph{Full Encoding: Method Calls and Shared Accesses}
 
@@
 
-\paragraph{Method Invocations} To account for method invocations, we induce additional context on variable accesses.
+To account for method invocations, we induce additional context on variable accesses.
 Instead of recording only the version of a variable, we also represent as part of the variable's identifier its enclosing method and its invocation counter.
@@
 
-\paragraph{Shared Heap Accesses} The final aspect of our encoding process is representation of shared heap accesses.
-As illustrated in Section \ref{Se:techoverview}, this is done via designated symbols:
+The final aspect of our encoding process is representation of shared heap accesses. As illustrated in Section \ref{Se:techoverview}, this is done via designated symbols:
 \begin{itemize}
@@
 We could encode the question of shared versus local accesses as additional constraints (requiring that a variable be accessed by more than one thread), but that would be more complicated and less efficient.
+
+\subsection{Constraint System}
+
+Having explained how the trace is encoded, we now describe in detail how constraints are derived from the trace, such that any permutation considered by the analysis is guaranteed to represent a feasible execution schedule.
+
+\paragraph{Intra-thread Order}
+
+The first set of constraints reflects control flow within the individual threads, which must remain unchanged under the reordering transformation.
+Given input trace $t$, this is expressed as the following formula:
+$$
+\begin{array}{rl}
+\forall \tau,\tau' \in t. & {\sf thread}\ \tau \equiv {\sf thread}\ \tau'. \\
+& {\sf index}\ \tau < {\sf index}\ \tau' \Rightarrow O_{\tau} < O_{\tau'}
+\end{array}
+$$ 
+The logical variables $O_x$ express ordering constraints. The requirement, as stated above, is that these variables
+reflect the same order as the projection of ${\sf index}$ onto individual threads.
+
+\paragraph{Race Condition}
+
+Given pair $\tau$ and $\tau'$ of events that both access a common memory location $\ell$, we demand that
+$$
+\begin{array}{rl}
+& {\sf thread}\ \tau \neq {\sf thread}\ \tau' \\
+\bigwedge 	& (\ell \in {\sf writeset}\ \tau \vee \ell \in {\sf writeset}\ \tau') \\
+\bigwedge   & O_{\tau} = O_{\tau'}
+\end{array}
+$$
+That is, (i) events $\tau$ and $\tau'$ are executed by different threads, (ii) at least one of the events performs write access to $\ell$ (as judged by the ${\sf writeset}$ membership check), and (iii) the events occur simultanesouly.
+This is a direct logical encoding of the definition of a race condition.
+
+\paragraph{Path Constraints}
+
+The requirement with respect to branching is that the prefix of the original trace $t$ up to the pair $\tau$ and $\tau'$ of candidate racing events remains identical in the predicted trace $t'$.
+More accurately, under the assumption that ${\sf index}\ \tau < {\sf index}\ \tau'$,
+we require that 
+$$
+\bigwedge_{\tau'' \in t \cap {\bf bexp}.\
+	\tau'' \in {\sf pre}\ \tau'} \lsyn {\sf stmt}\ \tau'' \rsyn\ t \equiv \lsyn {\sf stmt}\ \tau'' \rsyn\ t'
+$$ 
+where ${\sf stmt}$ is a helper function that obtains the code statement incident in a given transition.
+That is, all branching transitions up to $\tau'$ (which occurs after $\tau$) preserve their boolean interpretation under $t'$.
+This ensures that there are no divergences from the path containing the racing events, though beyond that path any feasible continuation is permitted.
+Importantly, contrary to \cite{JEFF-PLDI14}, we do not pose the requirement that the values flowing into branching statements remain the same, but suffice with the relaxed requirement that the evaluation of branching expressions is invariant under the input and predicted traces.
+
+\paragraph{Variable Definitions}
+
+A final requirement for the basic setting is that left-hand variables are defined according to the same right-hand variables as before.
+That is, version $i$ of variable ${\tt u}$ is defined as version $j$ of variable ${\tt v}$ in input trace $t$, then the same remains true in predicted  trace $t'$.
+This is enforced as the formula
+$$
+\bigwedge_{\tau'' \in t \cap {\bf asgn}.\
+	\tau'' \in {\sf pre}\ \tau'} {\sf stmt}\ \tau'' \in t' 
+$$
+where we again assume that ${\sf index}\ \tau < {\sf index}\ \tau'$.
+That is, the same statement occurring in $t$ is also present in $t'$ (though the transitions may differ).
+Since the statements of $t'$ are a permutation of the statements of $t$, we are assured that use/def flow is constrained appropriately.
+
+\paragraph{Relaxation of Flow Dependencies}
+
+We now move to the novel feature of \tool, which is its ability to explore execution schedules that depart from the data flow exhibited in the original trace.
+More precisely, \tool\ is able to relax flow dependencies in the original trace, whereby a thread reads a shared memory location written by another thread, while enforcing feasibility.
+This is strictly beyond the coverage potential of existing predictive analyses, which restrict trace transformations to ones where any read access to a shared memory location must correspond to the same write access as in the original trace.
+
+To ensure feasibility under relaxation of flow dependencies, we need to secure the link between the execution schedule and the write/read flow.
+As an illustration from the example in Figure \ref{fig:running}, $R_{\tt y}^4={\tt y}^1 \wedge O_1 < O_4 < O_3$ specifies that in a schedule where thread $T_1$ executes line {\tt 1}, then $T_2$ executes line {\tt 4}, and then the schedule switches again to $T_1$ to execute line {\tt 3}, the read access to ${\tt y}$ at line {\tt 4} obtains the value assigned to ${\tt y}$ at line {\tt 1}: $R_{\tt y}^4={\tt y}^1$.
+
+The full and general constraint formula, given read $R_{\ell}$ of location $\ell$ as part of event $e$ with set ${\cal W}$ of matching write events (i.e., events including write access to $\ell$), takes the following form:
+$$
+\begin{array}{rll}
+\bigvee_{e_w \in {\cal W}} &  & (R_{\ell} = {\ell}^{{\sf index}\ e_w}) \\
+&		\bigwedge 	&  O(e_w) < O(e) \\
+&		\bigwedge_{e' \in {\cal W} \setminus \{ e_w \}} & (O(e') < O(e_w) \vee O(e) < O(e'))
+\end{array}
+$$
+This disjunctive formula iterates over all matching write events, and demands for each that (i) it occurs prior to the read event ($O(e_w) < O(e)$) and (ii) all other write events either occur before ($O(e') < O(e_w)$) it or after the read event
+($O(e) < O(e')$).
+
+An important concern that arises due to relaxation of flow dependencies is that heap accesses may change their meaning.
+As an illustration, we refer to Figure \ref{fig:heapAccess}.
+While the read at line {\tt 7} appears to match the write at line {\tt 5}, this is conditioned on the read at line {\tt 6} being linked to the assignment at line {\tt 4}.
+If the predicted run violates this link, then feasibility is no longer guaranteed.
+In particular, if reordering results in {\tt z} being assigned the first rather than second allocated object, then the write at line {\tt 5} no longer matches the read at line {\tt 7}.
+
+\begin{figure}
+	\centering
+	\begin{tabular}{ll}
+		\hline
+		\multicolumn{1}{c}{$T_1$} & \multicolumn{1}{c}{$T_2$} \\
+		\hline
+		{\tt 1: x1 = new();} & \\
+		{\tt 2: x2 = new();} & \\
+		{\tt 3: y = x1;} & \\
+		{\tt 4: y = x2;} & \\
+		{\tt 5: x2.f = 5;} & \\	
+		& {\tt 6: z = y;} \\
+		& {\tt 7: w = z.f;} \\
+	\end{tabular}
+	\caption{\label{fig:heapAccess}Example illustrating the need to account for heap accesses during trace transformation}
+\end{figure}
+
+To address this challenge, we enhance the constraint system with the requirement that heap objects that are dereferenced before the candidate racing events retain their original address in the predicted trace.
+This achieves three guarantees: First, matching events in the original trace are guaranteed to also match in the predicted trace.
+Second, candidate races in the original trace remain viable in the predicted trace.
+Finally, in a practical setting involving virtual method calls (which goes beyond our core grammar in Table \ref{Ta:syntax}), call-site resolutions are the same across the original and predicted traces.
+Notice that in this setting, we consider as relevant not only the targets of field dereferences but also the targets of virtual method invocations.
+
+Formally, given pair $\tau$ and $\tau'$ of candidate racing events such that ${\sf index}\ \tau < {\sf index}\ \tau'$, 
+we require that
+$$
+\bigwedge_{\tau'' \equiv {\tt y=x.f} \in t \cap {\bf heapr}.\
+	\tau'' \in {\sf pre}\ \tau'} {\sf env}\ \sigma(t,\tau'')\ {\tt x} = {\sf env}\ \sigma(t',\tau'')\ {\tt x}
+$$
+where ${\sf env}$ is the state mapping from local variables to their value and $\sigma(t,\tau'')$ 
+(resp. $\sigma(t',\tau'')$) is the state arising at trace $t$ (resp. $t'$) immediately before event $\tau''$.
+This constraint fixes that all heap dereferences up to the later of the candidate racing events retain their original base object as in the predicted trace $t'$.
+
+\subsection{Correctness Proof}
+
+{\bf MISSING!}

### 2014-09-02 proof.tex
proof

Researchers~\cite{pldi14,maximal} propose two axioms, {\em prefix closedness} and {\em local determinism}, which we adapt to establish our feasibility guarantee. 
Suppose $\mathcal{F}$ denotes the domain of feasible traces.

Intuitively, Prefix closedness means that if a trace is feasible, then the trace is still feasible after we discard the events after a point. The underlying reason is that the events after a point cannot affect the events before it. More formally,

\begin{myaxiom}[Prefix closedness]~\label{axiom:prefix}
If $t_1$ $t_2$ $\in \mathcal{F}$, then $t_1 \in \mathcal{F}$. 
\end{myaxiom}

Intuitively, local determinism says that only the previous events of the same thread determine the existence of an event. For example, the existence of an event is only determined by the branch  thread-locally. However, the value read by an event may be affected by another thread because a shared read reads from the most recent write, which may come from a different thread. The value in other events, such as branch or write,  are locally determined too.
More formally,

\begin{myaxiom}[Local determinism]~\label{axiom:local}
Assume $t_1e_1, t_2 \in \mathcal{F}$, and $t_1|thread(e_1)\approx t_2|thread(e_1)$, where $\approx$ denotes the two traces are equal if the data values in the read and write events are ignored, 
Each event is determined by the previous events in the same thread. There are four cases:
\begin{itemize}
%TODO refine it.
\item {\bf Branch\ }. If op($e_1$)=branch, and $e_1$ is in the form of $x<y$ (without loss of generality), then there exists values $v_x$, $v_y$ such that $t_2 e_1 [v_x/data_x, v_y/data_y] \in \mathcal{F}$. Here $e_1 [v_x/data_x]$ represents that we replace the value of $x$ from $data_x$ to $v_x$ in $e_1$.
\item {\bf Read\ }. If op($e_1$)=read, and $e_2$ is a read event that is identical to $e_1$ except that it may read a different value, and $t_2 e_2$ is consistent, then $t_2 e_2 \in \mathcal{F}$.
\item {\bf Write\ }. If op($e_1$)=write, and there is a value $v$ such that $t_2 e_1[v/data]\in \mathcal{F}$. 
\item {\bf Others \ }. If op($e_1$) is of type different from above and $t_2 e_1$ is consistent, then $t_2 e_1 \in \mathcal{F}$.
\end{itemize}

Here, the consistency is defined as follows. 
A trace is consistency iff it satisfies (1) read-write consistency, i.e., read event should contain the value written by the most recent write event, and (2) the synchronization consistency, e.g., the lock acquire and release of a lock should not be interleaved by the lock operations of the same lock, the begin event of a thread should follow the fork event of the parent thread.
\end{myaxiom}

The local determinism resembles the version in ~\cite{pldi14,maximal}. A crucial difference is that, for the branch event, we do not require all reads thread locally before it to read the identical values as in the original trace $t_1 e_1$. Instead, we allow the previous reads to read different values as long as the branch event is evaluated to the same boolean value, which guarantees the existence of the following events. Such relaxation allows us to find more races, while still preserving the feasibility.

An important assumption of ours is, we assume all local accesses are recorded in the trace so that we can determine the boolean status of a branch event computationally. ~\cite{pldi14}, which assume the local accesses are not recorded and cannot determine the status computationally, have to enforce the reads before a branch to read the same values, leading to a weaker feasibility criterion. We believe our axiom captures the real-world scenario more faithfully.
Another important assumption is for the read-write consistency, i.e., we assume the heap invariant, so that we know exactly which reads correspond to which writes in the predicted run.

%\begin{myaxiom}
%A set of traces $\mathcal{F}$ is feasible if it satisfies prefix closedness and local determinism axioms. 
%\end{myaxiom}

They are called axioms because they are the rules that the sequentially consistent system~\cite{maximal} should follow.
We refer the readers to the detailed discussion~\cite{maximal, pldi14}.

%TODO what is \Phi
\begin{mytheorem}[Soundness]
All traces in the closure $closure(t, \Phi)$, which includes $t$ and is closed under the derivations in Axioms~\ref{axiom:prefix}~\ref{axiom:local}, are feasible.
\end{mytheorem}

%$\theta_1(expr(e_1))=\theta_2(expr_{e_1})$, then $t_2 \e_1 \in \mathcal{F}$. Here $\theta_1, \theta_2$ represent the mappings between variables and values established prior to $e_1$ by the same thread $thread(e_1)$. $expr(e_1)$ is a helper function that returns the expression in the branch condition. 
\begin{proof}
We order the traces in $closure(t)$ as, $t_0=t, t_1, t_2, t_3, \dots t_n$. Clearly, $t_0$ is feasible. In the following, we prove by induction, i.e., assume $t_{n}$ is feasible, then $t_{n+1}$ which is derived from $S=\{t_0, \dots t_n\}$ should be feasible. There are several possible derivations.
\begin{itemize}
\item if $t_{n+1}$ is a prefix of $t'\in S$, then $t_{n+1}\in\mathcal{F}$ because of the prefix closedness. $t_{n+1}$ shares the same mappings as $t'$.
\item if $t_{n+1}$ is derived from $t^1 e^1, t^2\in \mathcal{F}$, we have $t^1|thread(e^1)\approx t^2|thread(e^1)$.
\begin{itemize}
\item $op(e^1)=branch$ and $e^1= x<y$. Our solver computes a mapping $\theta^d$ of all relevant variables such that $\theta(x<y)=\theta^d(x<y)$, then 
$t^2 e_1[\theta^d(x)/\theta(x),\theta^d(y)/\theta(y) ]  \in \mathcal{F}$. Note that the mapping computed by solver may be different from the mapping in the trace $t^1$.
\item $op(e^1)=read$. Our solver ensures the read-write consistency, i.e., $t^2 e^2$ is consistent, then $t^2 e^2 \in \mathcal{F}$. Note that $e^2$ is the same as $e^1$ except that it may read a different value.
\item $op(e^1)=write$. There exists a value $v$ such that $t^2 e^2[v/data]\in \mathcal{F}$, where $v$ is included in the mapping computed by the solver. 
\item $op(e^1)$ is of other types. Our solver ensures the consistency, i.e., $t^2 e^1$ is consistent, then $t^2 e^2 \in \mathcal{F}$.
\end{itemize}
\end{itemize}
\end{proof}

\begin{mytheorem}[Maximality]
$\forall t', s.t., t'|th\approx t|th for any thread th, if t'\notin closure(t,\Phi), t' is infeasible.$
\end{mytheorem}

\begin{proof}
We sketch the proof. All traces that satisfy  $t'|th\approx t|th for any thread th$ 
\end{proof}

%discussion.

### 2014-09-02 overview.tex
fix typo

--- 2014-08-21 overview.tex
@@
 %
-As input, we assume (i) a program $P$ as well as (ii) a dynamic execution trace of $P$ in Static Single Assignment (SSA) form, such that every variable is defined exactly once.
+As input, we assume (i) a program $P$ as well as (ii) a trace of $P$ recorded during a dynamic execution.
+
+% in Static Single Assignment (SSA) form, such that every variable is defined exactly once.
+
+\subsection{Preliminaries}
+To facilitate our presentation, we first introduce some terminologies used throughout this paper.
+
+{\bf Trace \ } A trace $\tau$ is a sequence of events recorded during the observation run.
+
+{\bf Event\ } An event, $e=<t, id, inst, map>$, is a concrete representation that captures the details about the runtime execution of a static instruction $inst$,
+where $t$ denotes the thread that issues the event, $id$ is the id associated with each event, $inst$ is the static instruction and $map$ maps the variable in the instruction to the runtime values.
+The key property of  $id$ is {\em uniqueness}, i.e., any two events in the trace own different ids.
+A natural id for each event is its index in the trace.
+We denote the thread and id as $t^e$ and $i^e$ respectively.
+
+The instructions are three-address instructions involving at most three operands, which modern compilers commonly support.  Specifically, we are interested in the following types of instructions.
+
+\begin{table}
+	\begin{center}
+		\begin{tabular}{rl}
+			\multicolumn{1}{l}{{\tt s} $::=$} & \\
+			{\tt y = x.f} $|$\ & {\bf (heapr)} \\ 
+			{\tt x.f = y} $|$\ & {\bf (heapw)} \\ %\ $|$\ {\tt x.f = $c$}\
+			{\tt z = x $bop$ y}\  $|$ & {\bf (assign)} \\ %$|$\ {\tt z = $c$} $|$ {\tt z = new()}
+			{\tt if (x $bop$  y) goto ...  $|$} &  {\bf (branch)} \\
+			{\tt lock(l)}\ $|$\ {\tt unlock(l)}  $|$& {\bf (sync)} \\
+			{\tt fork(t)}\ $|$\ {\tt join(t)}$|$ {\tt begin(t)}\ $|$\ {\tt end(t)}$|$  & {\bf (thread)}
+		\end{tabular}
+	\end{center}
+	\caption{\label{Ta:syntax}Language syntax}
+\end{table}
+
+When the variable does not appear on the left hand of an equation, such as $y$ in $x.f=y$, it may refer to a variable, a constant or event object creation expression $new (...)$.
+The $bop$ stands for the binary operator, which may refer to $+, -, *, /, \%, \wedge, \vee$ in the assignment, or refer to $<, >, =$ in the branch.
+The target of the branch event is not important in our scope, therefore, we may omit the $goto$ part.
+The listed instructions are so basic that it can capture most events in a common imperative programming language.
+For example, the method call can be modeled by the above list, where the argument passing and value return are modeled as assignment events, and the virtual function resolution is modeled as the branch event conditioned on the type of the caller object.
+
+The map stores the value for each variable observed in the dynamic run. For simplicity, we use the notion $\lsyn x \rsyn^e$ to retrieve the value associated with the variable $x$ in $e$.
+
+In addition, we propose the projection operation  $|$  to help the reasoning of the trace: $\tau|t$ contains only the (ordered) events from the thread $t$; $\tau|i$ contains only the event of which the id equals $i$; $\tau|\leq e$ denotes the prefix of $e$, i.e., the events preceding the event $e$; $\tau|\geq e$ denotes the events after $e$; $\tau|>=e_1\wedge <=e_2$ denotes the events between $e_1$ and $e_2$.
+
+%We make use of the following helper functions:
+%\begin{itemize}
+%	\item ${\sf proj}\ t\ i$ projects trace $t$ onto all transitions involving thread $i$.
+%	\item $t[k]$ obtains the $k$th transition within trace $t$.
+%	\item ${\sf index}\ t\ \tau$ retrieves the index, or offset, of transition $\tau$ within trace $t$. When simply writing
+%	${\sf index}\ \tau$ (while omitting the trace parameter) we refer to the index of $\tau$ within the original trace. 
+%	\item ${\sf pre}\ t\ \tau$ is the prefix of trace $t$ preceding transition $\tau$. For the suffix beyond $\tau$, we 
+%	use ${\sf post}\ t\ \tau$. Finally, ${\sf bet}\ t\ \tau_1\ \tau_2$ returns the transitions delimited by $\tau_1$ and $\tau_2$.
+%\end{itemize}
+
+%Throughout this paper, we assume a standard operational semantics, which defines (i) a mapping ${\sf thr}$ between
+%execution threads, each having a unique identifier $i \in \mathbb{N}$, and their respective code, as well as (ii) per-thread stack and shared heap memory.
+
+%The code executed by a given thread follows the syntax in Table \ref{Ta:syntax}.
+The text of a program is a sequence of zero or more method declarations, as given in the definition of symbol {\tt p}.
+Methods accept zero or more arguments $\overline{{\tt x}}$, have a body ${\tt s}$, and may have a return value (which we leave implicit).
+For simplicity, we avoid from static typing as well as virtual methods.
+The body of a method consists of the core grammar for symbol {\tt s}.
+We avoid from specifying syntax checking rules, as the grammar is fully standard.
+
+%For simplicity, we assume that in the starting state each thread points to a parameter-free method.
+In this way, we can simply assume an empty starting state (i.e., an empty heap), and eliminate complexities such as user-provided inputs and initialization of arguments with default values.
+These extensions are of course possible, and in practice we handle these cases, but reflecting them in the formalism would result in needless complications.
+
+%As is standard, we assume an interleaved semantics of concurrency. A \emph{transition}, or \emph{event}, is of the form
+%$\sigma \stackrel{i / {\tt s}}{\longrightarrow} \sigma'$, denoting that thread $i$ took an evaluation step
+%in prestate $\sigma$, wherein atomic
+%statement {\tt s} was executed, which resulted in poststate $\sigma'$. We refer to a sequence of transitions from
+%the starting state to either an exceptional state (e.g., due to null dereference) or a state where all the threads have 
+%reduced their respective code to $\epsilon$ as a \emph{trace}.
+
+%TODO this section is inconsistent with others. change it!
+\subsection{Constraint System}
 

### 2014-09-03 overview.tex
SSA form of trace

@@
 \subsection{Preliminaries}
-To facilitate our presentation, we first introduce some terminologies used throughout this paper.
+To facilitate our presentation, we first introduce some terminologies used throughout this paper. At the high level, a trace is a sequence of events recorded during the observation run.
 
-{\bf Trace \ } A trace $\tau$ is a sequence of events recorded during the observation run.
+{\bf Event\ } An event, $e=<t, id, inst, map>$, is a concrete representation that captures the details about the runtime execution of a static instruction $inst$.
 
-{\bf Event\ } An event, $e=<t, id, inst, map>$, is a concrete representation that captures the details about the runtime execution of a static instruction $inst$,
-where $t$ denotes the thread that issues the event, $id$ is the id associated with each event, $inst$ is the static instruction and $map$ maps the variable in the instruction to the runtime values.
+\begin{itemize}
+\item $t$ refers to the thread that issues the event, denoted as $t^e$.
+\item  $id$ refers to the id associated with each event, denoted as $i^e$.
 The key property of  $id$ is {\em uniqueness}, i.e., any two events in the trace own different ids.
 A natural id for each event is its index in the trace.
-We denote the thread and id as $t^e$ and $i^e$ respectively.
-
-The instructions are three-address instructions involving at most three operands, which modern compilers commonly support.  Specifically, we are interested in the following types of instructions.
+\item $inst$ is the static instruction.
+The instructions are three-address instructions involving at most three operands, which modern compilers commonly support.
+Specifically, we are interested in the types of instructions listed in Table~\ref{Ta:syntax}.
+When the variable does not appear on the left hand of an equation, such as $y$ in $x.f=y$, it may refer to a variable, a constant or event object creation expression $new (...)$.
+The $bop$ stands for the binary operator, which may refer to $+, -, *, /, \%, \wedge, \vee$ in the assignment, or refer to $<, >, =$ in the branch.
+The target of the branch event is not important in our scope, therefore, we may omit the $goto$ part.
+The listed instructions are so basic that it can capture most events in a common imperative programming language.
+For example, the method call can be modeled by the above list, where the argument passing and value return are modeled as assignment events, and the virtual function resolution is modeled as the branch event conditioned on the type of the caller object.
+\item $map$ maps the variable in the instruction to the runtime values.
+The map stores the value for each variable observed in the dynamic run.
+For simplicity, we use the notion $\lsyn x \rsyn^e$ to retrieve the value associated with the variable $x$ in $e$.
+Specially, $\lsyn x.f \rsyn^e$ returns the location of $x.f$ (rather than  the value stored in the location), which is denoted as a pair $(\lsyn x \rsyn^e, f)$.
+\end{itemize}
 
@@
 			{\tt x.f = y} $|$\ & {\bf (heapw)} \\ %\ $|$\ {\tt x.f = $c$}\
-			{\tt z = x $bop$ y}\  $|$ & {\bf (assign)} \\ %$|$\ {\tt z = $c$} $|$ {\tt z = new()}
-			{\tt if (x $bop$  y) goto ...  $|$} &  {\bf (branch)} \\
+			{\tt z = x $bop$ y}\  $|$& {\bf (assign)} \\ %$|$\ {\tt z = $c$} $|$ {\tt z = new()}
+			{\tt if (x $bop$  y) goto ...  $|$&  {\bf (branch)} \\
 			{\tt lock(l)}\ $|$\ {\tt unlock(l)}  $|$& {\bf (sync)} \\
@@
 
-When the variable does not appear on the left hand of an equation, such as $y$ in $x.f=y$, it may refer to a variable, a constant or event object creation expression $new (...)$.
-The $bop$ stands for the binary operator, which may refer to $+, -, *, /, \%, \wedge, \vee$ in the assignment, or refer to $<, >, =$ in the branch.
-The target of the branch event is not important in our scope, therefore, we may omit the $goto$ part.
-The listed instructions are so basic that it can capture most events in a common imperative programming language.
-For example, the method call can be modeled by the above list, where the argument passing and value return are modeled as assignment events, and the virtual function resolution is modeled as the branch event conditioned on the type of the caller object.
-
-The map stores the value for each variable observed in the dynamic run. For simplicity, we use the notion $\lsyn x \rsyn^e$ to retrieve the value associated with the variable $x$ in $e$.
-
-In addition, we propose the projection operation  $|$  to help the reasoning of the trace: $\tau|t$ contains only the (ordered) events from the thread $t$; $\tau|i$ contains only the event of which the id equals $i$; $\tau|\leq e$ denotes the prefix of $e$, i.e., the events preceding the event $e$; $\tau|\geq e$ denotes the events after $e$; $\tau|>=e_1\wedge <=e_2$ denotes the events between $e_1$ and $e_2$.
+{\bf Trace \ } 
+We propose the projection operation and domains to facilitate the reasoning of the trace.
+\begin{itemize}
+\item projection operation  ``$|$'': $\tau|t$ contains only the (ordered) events from the thread $t$; $\tau|l$ contains the events involving the location $l$; $\tau|i$ contains only the event of which the id equals $i$; $\tau|\leq e$ denotes the prefix of $e$, i.e., the events preceding the event $e$; $\tau|\geq e$ denotes the events after $e$; $\tau|>=e_1\wedge <=e_2$ denotes the events between $e_1$ and $e_2$.
+\item sets: $\mathcal{T}$ denotes the set of threads involved;  $\mathcal{L}$ denotes the set of memory locations involved; 
+$\mathcal{SV}$ denotes the set of shared variables, which reference the shared locations, the shared location $l$ can be judged by counting the number of threads in $\tau|l$;  $\mathcal{E}$ denotes the set of events involved.
+\end{itemize}
 

### 2014-09-03 trace.tex
changed structure
all
SSA form of trace

@@
 \section{Relaxation of Flow Dependencies}
+The main goal of our analysis is to, given an original trace $\tau$ over the events $\mathcal{E}$, derive a trace $\tau'$ over the same set of events, which has a new scheduling $\pi$ of the events (the order of events inside each thread should remain unchanged) and a new mapping $\theta$ from variables to values in the events.
+Our first relaxation comes from the insight that, we allow the variables to read different values than in the original trace even if such variables are used to determine the control flow, while existing approaches enforce them to read the same values as in the original trace.
+As a result, our relaxation allows more schedules, which are likely to expose more races.
+
+Not every scheduling or every mapping $\theta$ guarantees a new feasible trace $\tau'$, therefore, we need to compute the scheduling and mapping that lead to feasible trace.
+We explain the technique details in the following.
 
@@
 
-We begin by describing the first relaxation, whereby the dependence structure defined by the original trace is potentially violated while retaining soundness.
+%We begin by describing the first relaxation, whereby the dependence structure defined by the original trace is potentially violated while retaining soundness.
 
-\subsection{Trace Encoding}
+\subsection{SSA Form of the Trace}
+First of all, we require the SSA form of the trace, i.e., each variable in defined exactly once in the trace.
+Such requirement is a prerequisite for the computation of the mapping $\theta$, as it can map each variable to only one value.
+Another side effect is that the def/use chains become explicit in the SSA form, which simplifies the following analysis steps.
 
-The first step of the process is to rewrite the concrete execution trace into a symbolic form that enables \tool\ to 
-derive reordering constraints. We describe how \tool\ encodes a trace into symbolic form in two steps. We first start with ``local'' intraprocedural events --- i.e., 
-transitions that manipulate memory accessed by at most a single thread during an invocation-free run --- and then extend our encoding scheme to the entire set of possible events.
-
-\paragraph{Basic Encoding: Local Accesses}
-
-The fundamental encoding transformation is to induce Static Single Assignment (SSA) form on the raw trace, such that
-a variable is defined exactly once. In this way, def/use chains become explicit, and encoding of trace events as constraints is simplified. As an illustration, trace
-\begin{quote}
-	{\tt 1: x=1; 2: x<3; 3: x=3;} \\
-	{\tt 4: y=1;} \\
-	{\tt 5: z=x+y}
-\end{quote}
-becomes
-\begin{quote}
-	{\tt 1: x$^1$=1; 2: x$^1$<3; 3: x$^3$=3;} \\
-	{\tt 4: y$^4$=1;} \\
-	{\tt 5: z$^5$=x$^3$+y$^4$}
-\end{quote}
-(For readability, we version variables according to the line number of their definition.)
-
-Beyond the default SSA rewriting algorithm, we apply a specialized transformation to handle local heap accesses.
-While in general determining whether a given heap access is local is an undecidable problem, in the context
-of a concrete execution trace this determination is straightforward.
-We can thus improve upon the baseline SSA form, where def/use chains over heap accesses are ignored.
-Indeed, to ensure feasibility, we must account for such accesses.
-Hence, we replace accesses to field {\tt f} of local object $o$ with a fresh local variable {\tt l$_{o{\tt .f}}$}, where $o{\tt .f}$ denotes the memory location itself and not its value.
-This is done prior to the standard SSA transformation, and under the assumption that in the predicted run, the base object $o$ remains the same.
-
-\paragraph{Full Encoding: Method Calls and Shared Accesses}
-
-To complete the encoding algorithm, we next explain how shared heap accesses and method calls are dealt with.
-Similarly to the case of local heap accesses, we take advantage of the fact that the full execution trace is available for analysis to detect shared accesses (the dual of local accesses) as well as call-site resolutions (which is more relevant in practice, where we support virtual method calls).
-
-To account for method invocations, we induce additional context on variable accesses.
-Instead of recording only the version of a variable, we also represent as part of the variable's identifier its enclosing method and its invocation counter.
-Thus, symbol ${\tt m}^2 \colon {\tt v}^7$ is interpreted as the definition of local variable ${\tt v}$ within the second invocation of method ${\tt m}$ that occurs at trace index 7.
-
-In this way, the trace is flattened.
-Method invocations are substituted with qualified variable names, and the rest of the encoding steps, described above, remain unchanged.
-Because the trace is finite and fully resolved, challenges such as looping, recursion and mutual recursion are all obviated.
-
-The final aspect of our encoding process is representation of shared heap accesses. As illustrated in Section \ref{Se:techoverview}, this is done via designated symbols:
+\tool\ handles the local assignment and heap accesses differently.
 \begin{itemize}
-	\item $W^{k}_{\tt v}$ denotes write access to variable ${\tt v}$ at trace index $k$ (where ${\tt v}$ is assumed to be a qualified identifier to account for the interprocedural setting).
-	\item Analogously, $R^{k}_{\tt v}$ denotes that ${\tt v}$ is read at index $k$.
+\item {\bf Local Assignment\ } The SSA form for local assignment resembles the SSA form of static instructions in compiler optimization, except that the loops and recursions are fully resolved in a concrete trace and there is no need for the Phi node.
+More concretely, we replace the variable $v$ defined in an event, as well as the following uses of the definition, to a new variable $v^{id}$, where $id$ is the unique id of the event.
+The uniqueness of the id guarantees that no two events define the same variable, i.e., each variable is defined exactly once.
+Note that the uses of a definition can be computed easily by scanning the trace afterwards for the variables of the same name before the next redefinition.
+\item {\bf Heap accesses\ } The accesses of the local heap locations and the accesses of shared locations behave differently, therefore, we introduce different SSA form for them. 
+\begin{itemize}
+\item {\bf Local Heap Accesses\ } The accesses of local heap locations behave similarly to the local assignment as both the definition and uses belong to the single thread.
+Therefore we model them as local assignments: We introduce a fresh local variable {\tt l$_{\lsyn o.f \rsyn}^{id}$} to replace each definition and  the corresponding uses of a  field {\tt f} of the local object $o$.
+An assumption here is the uses and the definition still refer to the same base object $o$ in the predicted run, which is enforced through the constraint specification (Section~\ref{sec:...}).
+\item  {\bf Shared Heap Accesses\ } The accesses of shared locations are more complex.
+Each shared read may read from one of multiple writes that define the shared location, under different schedules.
+To correlate the reads and the writes, we introduce two symbols, $W^{id}_{\tt \lsyn o.f\rsyn }$, which denotes write access to
+the field $f$ of the shared object $o$ by the event $id$, and $R^{id}_{\tt \lsyn o.f\rsyn }$,  which denotes the read access of the location $o.f$ by the event $id$. 
+Similar to the local heap accesses, an assumption for the correlation is that the writes and reads still refer to the same base object $o$ in the predicted run, which we guarantee by specifying additional constraints (Section~\ref{sec:...}).
+\end{itemize}
 \end{itemize}
 
-The main motivation to encode shared heap accesses using special symbols is to simplify downstream processing, and in particular, the definition of candidate races.
-We could encode the question of shared versus local accesses as additional constraints (requiring that a variable be accessed by more than one thread), but that would be more complicated and less efficient.
+% We could encode the question of shared versus local accesses as additional constraints (requiring that a variable be accessed by more than one thread), but that would be more complicated and less efficient.
+
+%\paragraph{Basic Encoding: Local Accesses}
+%
+%The fundamental encoding transformation is to induce Static Single Assignment (SSA) form on the raw trace, such that
+%a variable is defined exactly once. In this way, , and encoding of trace events as constraints is simplified. As an illustration, trace
+%\begin{quote}
+%	{\tt 1: x=1; 2: x<3; 3: x=3;} \\
+%	{\tt 4: y=1;} \\
+%	{\tt 5: z=x+y}
+%\end{quote}
+%becomes
+%\begin{quote}
+%	{\tt 1: x$^1$=1; 2: x$^1$<3; 3: x$^3$=3;} \\
+%	{\tt 4: y$^4$=1;} \\
+%	{\tt 5: z$^5$=x$^3$+y$^4$}
+%\end{quote}
+%(For readability, we version variables according to the line number of their definition.)
+
+%%TODO update the intro+moti, make sure the same style
+\begin{figure}
+\centering
+\begin{tabular}{ll}
+\multicolumn{2}{c}{{\tt {\bf x} = 0; {\bf y} = 0;}} \\
+\hline
+\multicolumn{1}{c}{$T_1$} & \multicolumn{1}{c}{$T_2$} \\
+\hline
+{\tt 1: s=0; } & \\
+{\tt 2: for(i=1;i<3;i++)} & \\
+{\tt 3: \ \ \ s+=i;} & \\
+{\tt 4: {\bf y} = s;} & \\
+{\tt 5: {\bf x} = 1;} & \\
+{\tt 6: {\bf y} = 5;} & \\
+& {\tt 7: if ({\bf y} > 2)} \\
+& {\tt 8:~~print({\bf x}+1);} \\	
+& {\color{Gray} {\tt 9: else}} \\
+& {\color{Gray} {\tt 10:~~print({\bf x}+2);}}
+\end{tabular}
+\caption{Running Example (shared variables are in bold font). }
+\label{fig:running2}
+\end{figure}
+
+\begin{figure}
+\centering
+\begin{tabular}{l|l}
+\hline
+\multicolumn{1}{c}{$Trace$} & \multicolumn{1}{c}{$SSA form of  Trace$} \\
+\hline
+{\tt 0: {\bf x}=0} &  {\tt 0: $W^0_x$=0}    \\
+{\tt 1: {\bf y}=0} &   {\tt 1: $W^1_y$=0}   \\
+{\tt 2: s=0} &  {\tt 2: $s^2$=0}   \\
+{\tt 3: i=1} &     {\tt 3: $i^3$=1}   \\
+{\tt 4: i<3} &    {\tt 4: $i^3$<3} \\
+{\tt 5: s=s+i} & {\tt 5: $s^5$=$s^2$+$i^3$}   \\
+{\tt 6: i=2} &       {\tt 6: $i^6$=2}  \\
+{\tt 7: i<3} &      {\tt 7: $i^6$<3}  \\
+{\tt 8: s=s+i} &  {\tt 8: $s^8$=$s^5$+$i^6$}  \\
+{\tt 9: i=3} &     {\tt 9: $i^9$=3}  \\
+{\tt 10: i<3} &    {\tt 10: $i^9$<3}  \\
+{\tt 11: {\bf y} = s;} &  {\tt 11: $W^{11}_y$ = $s^8$;}  \\
+{\tt 12: {\bf x} = 1;} &    {\tt 12: $W^{12}_x$ = 1;}   \\
+{\tt 13: {\bf y} = 5;} &    {\tt 13: $W^{13}_y$ = 5;}  \\
+{\tt 14: {\bf y} > 2}  &    {\tt 14: $R^{14}_y$ > 2} \\
+{\tt 15: tmp={\bf x}+1;}  & {\tt 15: tmp=$R^{15}_x$+1;}   \\	
+{\tt 16: print(tmp);} &  {\tt 16: print(tmp);}  \\
+\end{tabular}
+\caption{Trace}
+\label{fig:t4running2}
+\end{figure}
+
+Figure~\ref{fig:t4running2} exemplifies a trace and its SSA form, where the trace is generated from the program in Figure~\ref{fig:running2}.
+Each label in  Figure~\ref{fig:t4running2} on the left hand denotes the id of the event.
+As seen, event 5 uses the variable $s^2$ defined at event 2 and defines the variable $s^5$, which is used afterwards.
+The read/write of the shared variables are denoted in the special form, such as $R^{14}_y$ and $W^{13}_y$.
+We use $x, y$ to represent the location for simplicity.
 
@@
 This constraint fixes that all heap dereferences up to the later of the candidate racing events retain their original base object as in the predicted trace $t'$.
-
-\subsection{Correctness Proof}
-
-{\bf MISSING!}

### 2014-09-03 constraints.tex
changed structure
tech

@@
 \section{Exploration of Unexecuted Branches}
+We now switch to the second feature of \tool, which is its ability to reason about unexplored branches.
+At the high level, we simply need to derive a trace that contains the unexplored branch, after that, we can reduce the analysis to the analysis in Section~\ref{sec:relax} (starting from the SSA processing).
+\tool\ derives such traces by symbolically executing the unexplored branches to record the events and including these events in the new traces.
 
-We now switch to the second feature of \tool, which is its ability to reason about unexplored branches.
+%TODO | and projection, abuse the terms?
+As shown in the following, suppose the symbolic execution  $symEngine(\tau, e)$ returns the single-thread traces starting from the branch $e$, which now takes a different decision.
+The resultant trace is computed by first removing the events thread-locally after $e$ (inclusively) and then appending one of the single-thread traces that represents a path in the unexplored branch of $e$.
+The negation $neg(e)$ of $e$, which is the same as $e$ except the evaluation result of the boolean expression differs, is also appended.
 
-\paragraph{Relaxation of Branching Decisions}
+\begin{algorithmic}[3]
+\For {$e: \tau$}
+  \If {$type(e)=branch$}
+    \State $S=symEngine(\tau, e)$
+	\For {$\tau': S$}
+	\State $\tau_r=\tau-((\tau|t^e)|>=e)$
+	\State $\tau_r=\tau_r.replace(e, neg(e) + \tau')$
+	\EndFor
+ \EndIf
+\EndFor
+\end{algorithmic}
 
-Beyond relaxing flow dependencies, \tool\ is also capable of exploring code branches that were not executed in the original trace.
+{\bf Symbolic Execution\ } Our symbolic execution is realized on top of the dataflow analysis, which generates events for each instruction within the unexplored branch.
+The symbolic execution has known limitations in reasoning about the schedules and loops (and recursion).
+Therefore, our analysis assumes all the events in the unexplored branch happen atomically at the branch event $e$ without interleavings from other threads, as illustrated also in the above algorithm (line 6).
+In this way, our analysis adopts the sequential reasoning.
+\tool\ will reschedule the events based on the relaxation.
+As for loop (or recursion), the analysis terminates the current data flow immediately after one iteration, i.e., we expand the loop for only once.
+Otherwise, the analysis terminates the data flow normally if the flow goes out of the scope of the unexplored branch.
+
+The analysis starts with the event $e$ (exclusively), with the current state $\sigma$ at $e$ as the initial state. 
+Remember that we use the runtime values for two types of events: the heap value of the base objects and the evaluation result of the boolean expression in the branch. 
+Other values are maintained but not used. Therefore, we also need to maintain the heap value and the boolean expression value in the symbolic execution.
+
+First, we propose standard kill/gen rules for three basic heap instructions to maintain the heap update along a given path.
+Specially, we separate the object creation from the local assignment.
+In the object creation, we assign a unique integer to represent the newly created object and store it into the state $\sigma$.
+For field reference, $\sigma[x.f]$ is actually short for two-level interpretation $\sigma[\sigma[x].f]$, where the first finds out the base object $\sigma[x]$ and the second searches for the object referenced by its field $f$.
+The kill/gen rules take effect only when $x.f$ is not of the primitive type.
+
+The analysis also generates the events, where the heap values are stored in the map of the event for latter use. For the branch, we store its boolean evaluation result.
+As the boolean result is available only after the branch takes the decision, we delay the generation of the branch event until the first event after the branch.
+
+\begin{table*}
+\centering
+%\begin{small}
+\begin{tabular}{l|c|c}
+\multicolumn{1}{c|}{Operation} & {Kill} & {Gen} \\
+\hline
+{\tt $y$=$new (...)$} 			&  $\sigma[y] \mapsto \star$		&  $\sigma[y] \mapsto newI$, $<t, cnt, y=newI, \bot>$\\
+{\tt $y$=$x.f$} 			&   $\sigma[y]\mapsto \star$			&  $\sigma[y] \mapsto \sigma[x.f]$, $<t, cnt, y=x.f, \lsyn x \rsyn=\sigma[x]>$  \\
+{\tt $x.f=y$} & $\sigma[x.f] \mapsto \star$		&  $\sigma[x.f] \mapsto \sigma[y]$, $<t, cnt, x.f=y, \lsyn x \rsyn=\sigma[x]>$ \\
+{\tt $z=x\  bop\  y$} & 		&  $<t, cnt, z=x\  bop\  y, \bot>$ \\
+{\tt $if(x<y)$} & 	  &  $<t, cnt, if(x<y), \lsyn x<y \rsyn=bool>$\\
+\end{tabular}
+\caption{\label{table:killgen}Dataflow analysis}
+%\end{small}
+\end{table*}
+
+%Beyond relaxing flow dependencies, \tool\ is also capable of exploring code branches that were not executed in the original trace.
 This is achieved via a symbolic representation of the input trace.
 Given branching event $e_b$,
-\begin{enumerate}
-	\item model the branching condition symbolically to execute along the negation of the original branch;
-	\item apply depth-first search (DFS) to uncover all execution suffixes under the unexplored branch (which may itself contain branching statements); and
-	\item for each suffix $t^s$,
-	\begin{enumerate}
-		\item truncate the original trace at $e_b$ yielding prefix $t_s$; and
-		\item concatenate $t_s$ with the negation of $e_b$ followed by $t^s$.
-	\end{enumerate}
-\end{enumerate}
-Constraint solving is applied to each of the resulting traces analogously to the original trace.
+%\begin{enumerate}
+%	\item model the branching condition symbolically to execute along the negation of the original branch;
+%	\item apply depth-first search (DFS) to uncover all execution suffixes under the unexplored branch (which may itself contain branching statements); and
+%	\item for each suffix $t^s$,
+%	\begin{enumerate}
+%		\item truncate the original trace at $e_b$ yielding prefix $t_s$; and
+%		\item concatenate $t_s$ with the negation of $e_b$ followed by $t^s$.
+%	\end{enumerate}
+%\end{enumerate}
+%Constraint solving is applied to each of the resulting traces analogously to the original trace.
 
-We emphasize that exploration of new execution paths is subject to all the known limitations of symbolic execution, including in particular loop structures and object allocation.
+%We emphasize that exploration of new execution paths is subject to all the known limitations of symbolic execution, including in particular loop structures and object allocation.
 \tool\ currently fails if (i) the unexplored branch containts loops or (ii) there are object references that cannot be fully resolved at the branching point.
@@
 We demonstrate in Section \ref{sec:eval} that despite these limitations, the increase in coverage thanks to exploration of new paths is significant.
-
-\paragraph{Proof of Correctness}
-
-\input{proof}

### 2014-09-03 overview.tex
tech

@@
 The key property of  $id$ is {\em uniqueness}, i.e., any two events in the trace own different ids.
-A natural id for each event is its index in the trace.
+The important property of $id$ is {\em monotonicity}, i.e.,  the events from the same thread should own the strictly increasing ids.
+Throughout this paper, we use  the index of an event in the trace as its id, which satisfies the above properties.
 \item $inst$ is the static instruction.
@@
 When the variable does not appear on the left hand of an equation, such as $y$ in $x.f=y$, it may refer to a variable, a constant or event object creation expression $new (...)$.
-The $bop$ stands for the binary operator, which may refer to $+, -, *, /, \%, \wedge, \vee$ in the assignment, or refer to $<, >, =$ in the branch.
-The target of the branch event is not important in our scope, therefore, we may omit the $goto$ part.
+The $bop$ stands for the binary operator, which may refer to $+, -, *, /, \%, \wedge, \vee$ in the assignment, or refer to $<, >, =, \wedge, \vee$ in the branch.
+The target of the branch event is not important in our scope, therefore, we may abbreviate the branch instruction as the boolean expression afterwards.
 The listed instructions are so basic that it can capture most events in a common imperative programming language.
@@
 For simplicity, we use the notion $\lsyn x \rsyn^e$ to retrieve the value associated with the variable $x$ in $e$.
-Specially, $\lsyn x.f \rsyn^e$ returns the location of $x.f$ (rather than  the value stored in the location), which is denoted as a pair $(\lsyn x \rsyn^e, f)$.
 \end{itemize}
+
+%Specially, $\lsyn x.f \rsyn^e$ returns the location of $x.f$ (rather than  the value stored in the location), which is denoted as a pair $(\lsyn x \rsyn^e, f)$.
 

### 2014-09-03 trace.tex
tech

@@
 Each shared read may read from one of multiple writes that define the shared location, under different schedules.
-To correlate the reads and the writes, we introduce two symbols, $W^{id}_{\tt \lsyn o.f\rsyn }$, which denotes write access to
-the field $f$ of the shared object $o$ by the event $id$, and $R^{id}_{\tt \lsyn o.f\rsyn }$,  which denotes the read access of the location $o.f$ by the event $id$. 
+Given the write $x.f=y$ or the read $y=x.f$, we introduce two symbols, $W^{id}_{\tt x.f }$, which denotes the write access by the event $id$ to the field $f$ of the shared object referenced by $x$, and $R^{id}_{\tt  x.f }$,  which denotes the read access  by the event $id$ to the location $x.f$.
+Note that, $x.f$ in the subscript is interpreted as the location, denoted as a pair $(\lsyn x \rsyn^e, f)$, rather than the value stored in the location.
 Similar to the local heap accesses, an assumption for the correlation is that the writes and reads still refer to the same base object $o$ in the predicted run, which we guarantee by specifying additional constraints (Section~\ref{sec:...}).
@@
 \hline
-\multicolumn{1}{c}{$Trace$} & \multicolumn{1}{c}{$SSA form of  Trace$} \\
+\multicolumn{1}{c}{$Trace$} & \multicolumn{1}{c}{$SSA \ form \ of\   Trace$} \\
 \hline
@@
 
+%TODO value or symbol for o.f, rethink about it.
 \subsection{Constraint System}
+Based on the SSA form of the trace, we build the constraints to compute a new trace with the new schedule and new mapping of variables to values.
+To model the schedule, we introduce the order variable $O_{e}$ for each event $e$.
+Given two events $e_i, e_j$ from two different threads, $O_{e_i}<O_{e_j}$ in a trace means $e_i$ is scheduled before $e_j$.
+We omit the synchronization constraints purposely, which are well explained in all existing predictive analysis techniques~\cite{yannis, pldi14}.
 
-Having explained how the trace is encoded, we now describe in detail how constraints are derived from the trace, such that any permutation considered by the analysis is guaranteed to represent a feasible execution schedule.
+%TODO synchronization constraint
 
-\paragraph{Intra-thread Order}
+%Having explained how the trace is encoded, we now describe in detail how constraints are derived from the trace, such that any permutation considered by the analysis is guaranteed to represent a feasible execution schedule.
 
-The first set of constraints reflects control flow within the individual threads, which must remain unchanged under the reordering transformation.
-Given input trace $t$, this is expressed as the following formula:
+{\bf Race Condition\ } Given any pair  of events, $e$ and $e'$,  that both access a common memory location $l$, we demand that
 $$
 \begin{array}{rl}
-\forall \tau,\tau' \in t. & {\sf thread}\ \tau \equiv {\sf thread}\ \tau'. \\
-& {\sf index}\ \tau < {\sf index}\ \tau' \Rightarrow O_{\tau} < O_{\tau'}
+& t^e \neq t^{e'} \\
+\bigwedge 	& (e\  writes\  l\   \vee e'\  writes\  l) \\
+\bigwedge   & O_{e} = O_{e'}
 \end{array}
 $$
-The logical variables $O_x$ express ordering constraints. The requirement, as stated above, is that these variables
-reflect the same order as the projection of ${\sf index}$ onto individual threads.
+That is, (i) events $e$ and $e'$ are executed by different threads, (ii) at least one of the events performs write access to $l$, and (iii) the events run concurrently.
+This is a direct logical encoding of the definition of a race condition.
+We refer to both $e$ and $e'$ as the racy event.
 
-\paragraph{Race Condition}
+{\bf Intra-thread Constraints\ } Given a racy pair under analysis, $e$ and $e'$, suppose $e$ is after $e'$ in the original trace $\tau$, the predictive analysis reschedules the events in the prefix of $e$, i.e., prior to $e$, so that $e$ and $e'$ can run concurrently.
+A necessary condition applied to predictive analysis techniques is that each thread should follow the same event sequence prior to $e$ as in the original run.
+We explain the rationale underlying this requirement in Section~\ref{sec:proof}, but now proceed to explain how to guarantee it using constraints.
+The preservation of the event sequence for each thread requires the intra-thread constraints:
 
-Given pair $\tau$ and $\tau'$ of events that both access a common memory location $\ell$, we demand that
+\begin{itemize}
+\item {\bf Branch Decision Constraints\ } The branches in the prefix should take the same decisions to reproduce the same set of events. More formally,  
+we require that 
+$$
+\bigwedge_{e_i \in \tau|\leq e \wedge  inst^{e_i}=x<y.\ }  x<y 	\equiv \lsyn x<y \rsyn^{e_i}
+$$ 
+where $e_i$ is a branch event of which the boolean expression is $x<y$ (we use the binary operator $<$ for illustration) and  $\lsyn x<y \rsyn^{e_i}$ stores the  evaluation result of the boolean expression in the original trace $\tau$.
+The constraint specifies that the boolean expression in the derived trace should be evaluated to the same boolean value.
+Importantly, different from existing analyses \cite{yannis,pldi14}, we do not pose the requirement that the values flowing into branching statements remain the same, but suffice with the relaxed requirement that the evaluation of branching expressions remain the same.
+For example, suppose  the branch event $x<y$ takes the value $1<2$ in the original trace, existing analyses require the same values for the variables $x$ and $y$ in the predicted run, while we allow  the branch event to take other values of the variables, such as $3<4$.
+\item {\bf Intra-thread Order Constraints\ } The events in the sequence should follow the same order as in the original trace. More formally, we require that
 $$
 \begin{array}{rl}
-& {\sf thread}\ \tau \neq {\sf thread}\ \tau' \\
-\bigwedge 	& (\ell \in {\sf writeset}\ \tau \vee \ell \in {\sf writeset}\ \tau') \\
-\bigwedge   & O_{\tau} = O_{\tau'}
+\forall e_i, e_j \in \tau. & t^{e_i}\equiv t^{e_j}. \\
+& i^{e_i} < i^{e_j}\  \Rightarrow O_{e_i} < O_{e_j}
 \end{array}
 $$ 
-That is, (i) events $\tau$ and $\tau'$ are executed by different threads, (ii) at least one of the events performs write access to $\ell$ (as judged by the ${\sf writeset}$ membership check), and (iii) the events occur simultanesouly.
-This is a direct logical encoding of the definition of a race condition.
+The constraint specifies that the two events from the same thread should follow the same order as reflected by the ids of the events.
+\item {\bf Assignment Constraints\ } The mapping of local variables should not contradict each instruction.   More formally, we require that
+$$
+\bigwedge_{e_i \in \tau|<=e \wedge type(e_i)=assign|heapr|heapw}\
+	inst^{e_i}
+$$
+where $inst^{e_i}$, such as $x=y+z$ or $R^{id}_{x}=y$, captures the constraint over the values of the variables imposed by the instruction.
+For example, $x=y+z$ is not satisfied if the solver computes a mapping from the three variables to $3, 4, 5$ respectively.
+Again we consider only the events in the prefix $\tau|e$, where $e$ is the racy event that happens later.
+Specially, for the instruction involving object creation, $x=new (...)$, we encode the object as its address, i.e., a constant integer returned by $System.identityHashcode(x)$ at runtime.
+\end{itemize}
 
-\paragraph{Path Constraints}
+{\bf Rescheduling through Relaxation of Flow Dependencies\ } We now move to the novel feature of \tool, which is its ability to explore execution schedules that depart from the value flow exhibited in the original trace.
+More precisely, \tool\ is able to relax value flow dependencies in the original trace, whereby a thread reads a
+different value possibly written by other events (or even events from other threads) to the shared memory location, while enforcing feasibility.
+ This is strictly beyond the coverage potential of existing predictive analyses, which restrict trace transformations to ones where any read access to a shared memory location must correspond to the same written values as in the original trace, which are usually from the same write events.
 
-The requirement with respect to branching is that the prefix of the original trace $t$ up to the pair $\tau$ and $\tau'$ of candidate racing events remains identical in the predicted trace $t'$.
-More accurately, under the assumption that ${\sf index}\ \tau < {\sf index}\ \tau'$,
-we require that 
-$$
-\bigwedge_{\tau'' \in t \cap {\bf bexp}.\
-	\tau'' \in {\sf pre}\ \tau'} \lsyn {\sf stmt}\ \tau'' \rsyn\ t \equiv \lsyn {\sf stmt}\ \tau'' \rsyn\ t'
-$$ 
-where ${\sf stmt}$ is a helper function that obtains the code statement incident in a given transition.
-That is, all branching transitions up to $\tau'$ (which occurs after $\tau$) preserve their boolean interpretation under $t'$.
-This ensures that there are no divergences from the path containing the racing events, though beyond that path any feasible continuation is permitted.
-Importantly, contrary to \cite{JEFF-PLDI14}, we do not pose the requirement that the values flowing into branching statements remain the same, but suffice with the relaxed requirement that the evaluation of branching expressions is invariant under the input and predicted traces.
+To ensure feasibility under relaxation of flow dependencies, we need to secure the flow between the read/write with the execution schedule. 
+For example, in Figure \ref{fig:running}, for the read access to ${\tt y}$ at line {\tt 4} obtains the value assigned to ${\tt y}$ at line {\tt 1}, we need the schedule where  thread $T_1$ executes line {\tt 1}, then $T_2$ executes line {\tt 4}, and then the schedule switches again to $T_1$ to execute line {\tt 3}.
 
-\paragraph{Variable Definitions}
+% $R_{\tt y}^4={\tt y}^1 \wedge O_1 < O_4 < O_3$ specifies that in a schedule , the $R_{\tt y}^4={\tt y}^1$.
 
-A final requirement for the basic setting is that left-hand variables are defined according to the same right-hand variables as before.
-That is, version $i$ of variable ${\tt u}$ is defined as version $j$ of variable ${\tt v}$ in input trace $t$, then the same remains true in predicted  trace $t'$.
-This is enforced as the formula
-$$
-\bigwedge_{\tau'' \in t \cap {\bf asgn}.\
-	\tau'' \in {\sf pre}\ \tau'} {\sf stmt}\ \tau'' \in t' 
-$$
-where we again assume that ${\sf index}\ \tau < {\sf index}\ \tau'$.
-That is, the same statement occurring in $t$ is also present in $t'$ (though the transitions may differ).
-Since the statements of $t'$ are a permutation of the statements of $t$, we are assured that use/def flow is constrained appropriately.
-
-\paragraph{Relaxation of Flow Dependencies}
-
-We now move to the novel feature of \tool, which is its ability to explore execution schedules that depart from the data flow exhibited in the original trace.
-More precisely, \tool\ is able to relax flow dependencies in the original trace, whereby a thread reads a shared memory location written by another thread, while enforcing feasibility.
-This is strictly beyond the coverage potential of existing predictive analyses, which restrict trace transformations to ones where any read access to a shared memory location must correspond to the same write access as in the original trace.
-
-To ensure feasibility under relaxation of flow dependencies, we need to secure the link between the execution schedule and the write/read flow.
-As an illustration from the example in Figure \ref{fig:running}, $R_{\tt y}^4={\tt y}^1 \wedge O_1 < O_4 < O_3$ specifies that in a schedule where thread $T_1$ executes line {\tt 1}, then $T_2$ executes line {\tt 4}, and then the schedule switches again to $T_1$ to execute line {\tt 3}, the read access to ${\tt y}$ at line {\tt 4} obtains the value assigned to ${\tt y}$ at line {\tt 1}: $R_{\tt y}^4={\tt y}^1$.
-
-The full and general constraint formula, given read $R_{\ell}$ of location $\ell$ as part of event $e$ with set ${\cal W}$ of matching write events (i.e., events including write access to $\ell$), takes the following form:
+%TODO l and ell
+In general, the constraint formula, given read $R_{\ell}$ of location $\ell$ as part of event $e$ with set ${\cal W}$ of matching write events (i.e., events including write access to $\ell$), takes the following form:
 $$
 \begin{array}{rll}
-\bigvee_{e_w \in {\cal W}} &  & (R_{\ell} = {\ell}^{{\sf index}\ e_w}) \\
+\bigvee_{e_w \in {\cal W}} &  & (R^{e}_{\ell} = W^{e_w}_{\ell}) \\
 &		\bigwedge 	&  O(e_w) < O(e) \\
@@
 
-An important concern that arises due to relaxation of flow dependencies is that heap accesses may change their meaning.
+An important concern that arises due to relaxation of flow dependencies is that heap accesses may change their meaning, i.e., they involve different base objects and do no longer match with each other.
 As an illustration, we refer to Figure \ref{fig:heapAccess}.
-While the read at line {\tt 7} appears to match the write at line {\tt 5}, this is conditioned on the read at line {\tt 6} being linked to the assignment at line {\tt 4}.
+While the read at event {\tt 7} appears to match the write at event {\tt 5}, this is conditioned on the read at event {\tt 6} being linked to the assignment at event {\tt 4}.
 If the predicted run violates this link, then feasibility is no longer guaranteed.
-In particular, if reordering results in {\tt z} being assigned the first rather than second allocated object, then the write at line {\tt 5} no longer matches the read at line {\tt 7}.
+In particular, if reordering results in {\tt z} being assigned the first rather than second allocated object, then the read at event {\tt 7} no longer matches the write  at event {\tt 5}.
 
@@
 	\end{tabular}
-	\caption{\label{fig:heapAccess}Example illustrating the need to account for heap accesses during trace transformation}
+	\caption{\label{fig:heapAccess}Example illustrating the need to account for heap accesses during trace transformation, where the labelled number denotes the event id.}
 \end{figure}
 
-To address this challenge, we enhance the constraint system with the requirement that heap objects that are dereferenced before the candidate racing events retain their original address in the predicted trace.
-This achieves three guarantees: First, matching events in the original trace are guaranteed to also match in the predicted trace.
-Second, candidate races in the original trace remain viable in the predicted trace.
-Finally, in a practical setting involving virtual method calls (which goes beyond our core grammar in Table \ref{Ta:syntax}), call-site resolutions are the same across the original and predicted traces.
-Notice that in this setting, we consider as relevant not only the targets of field dereferences but also the targets of virtual method invocations.
+To address this challenge, we enhance the constraint system with the requirement that heap objects that are dereferenced in the following field reference before the candidate racing events retain their original address in the predicted trace.
+This achieves two guarantees: First, matching heap access events in the original trace are guaranteed to also match in the predicted trace.
+Second, candidate races in the original trace remain viable in the predicted trace as they still refer to same location.
+Note that, the resolution of the virtual method call is modeled as the branch events that check the type of the target object, for which the constraints naturally guarantee the branch to take the same decision, i.e., the call to be resolved to the same method implementation.
 
-Formally, given pair $\tau$ and $\tau'$ of candidate racing events such that ${\sf index}\ \tau < {\sf index}\ \tau'$, 
+%Finally, in a practical setting involving virtual method calls, call-site resolutions are the same across the original and predicted traces. 
+%Notice that in this setting, we consider as relevant not only the targets of field dereferences but also the targets of virtual method invocations.
+
+Formally, given pair $e$ and $e'$ of candidate racing events such that $e$ happens after $e'$ in the trace $\tau$, 
 we require that
 $$
-\bigwedge_{\tau'' \equiv {\tt y=x.f} \in t \cap {\bf heapr}.\
-	\tau'' \in {\sf pre}\ \tau'} {\sf env}\ \sigma(t,\tau'')\ {\tt x} = {\sf env}\ \sigma(t',\tau'')\ {\tt x}
+\bigwedge_{ e_i \equiv {\tt y=x.f}|x.f=y \wedge e_i \in \tau|e }\
+	x \equiv  \lsyn x \rsyn^{e_i}
 $$
-where ${\sf env}$ is the state mapping from local variables to their value and $\sigma(t,\tau'')$ 
-(resp. $\sigma(t',\tau'')$) is the state arising at trace $t$ (resp. $t'$) immediately before event $\tau''$.
-This constraint fixes that all heap dereferences up to the later of the candidate racing events retain their original base object as in the predicted trace $t'$.
+%where ${\sf env}$ is the state mapping from local variables to their value and $\sigma(t,\tau'')$ 
+%(resp. $\sigma(t',\tau'')$) is the state arising at trace $t$ (resp. $t'$) immediately before event $\tau''$.
+This constraint fixes that all heap dereferences  prior to the racy event retain their original base object as in the original trace $\tau$.
+
+By sending the above constraints to a solver, we compute the necessary schedule orders among the events as well as the mapping of the variables.
+The necessary schedule orders define a partial order among the events, which permit a set of schedules that define the complete order that complies with it.

### 2014-09-03 related.tex
related work

\section{Related Work}

In this section, we survey related research on race detection with special emphasis on predictive trace analysis (PTA).

\subsection{Race Detection Techniques: Broad Survey}

Initial attempts to address the challenge of race detection focused on built-in synchronization primitives \cite{XXX,YYY}. These include locks as well as the wait/notify and start/join scheduling controls. Notable among these efforts is \emph{lockset} analysis, which considers only locks \cite{XXX}. Because the derived constraints are partial, permitting certain infeasible event reorderings, lockset analysis cannot guarantee soundness \cite{Huang}.

A different tradeoff is struck by the \emph{happens-before (HB)} approach \cite{Lamport}. In this style of analysis, all synchronization primitives are accounted for, though reordering constraints are conservative. As an example, HB inhibits reordering of two synchronized blocks governed by the same lock. PTA is an instance  of the HB approach, whose starting point is a concrete execution trace.

Recently there have been successful attempts to relax HB constraints. Among these are hybrid analysis \cite{XXX}, which permits both orderings of lock-synchronized blocks, as well as the \emph{Universal Causal Graph (UCG)} representation~\cite{ucg}, which also enables both orderings but only if these are consistent with wait/notify- and start/join-induced constraints.

\subsection{Predictive Trace Analysis}

Given concurrent execution trace $t$, a \emph{maximal and sound causal model} based on $t$ defines the set of all traces that a program that is able to generate $t$ can generate (maximality), and only those traces (soundness). PTA is founded on the notion of sound causality, as it considers feasible reorderings of the input trace that prove a candidate data race as such.

The first to propose PTA are Smagardakis et al. \cite{XXX}. In their original work, both synchronization constraints and inter-thread dependencies are preserved, where inter-thread dependencies are respected by 
only allowing reorderings that leave the dependence structure exhibited by the original trace in tact. This ensures that the values of shared memory locations remain the same, which secures the soundness argument, though maximality is not guaranteed.

Said et al. \cite{XXX} describe a PTA that is also sound albeit not maximal. Similarly to our analysis, Said et al. perform symbolic analysis of the input trace, and then utilize an SMT solver to search for interleaved schedules that establish the presence of data races. Soundsness is guaranteed by their ability to precisely encode the semantics of sequential consistency. ExceptionNULL \cite{XXX} is another example of a sound PTA without maximality guarantees, where the goal is to detect null dereferences rather than data races.

Serbanuta et al. \cite{XXX} have recently shown that it is possible to build a model that is not only sound but also maximal: Any extension of the model with a new trace renders the model unsound. This provides a foundation for different forms of PTA, including data races, but also atomicity, serializability and other properties.

Inspired by this result and closer to \tool\ is the PTA technique of Huang et al. \cite{XXX}, which is both sound and maximal. This technique, too, makes use of an SMT solver based on symbolic encoding of the input trace. As part of the encoding, control-flow information is taken into account to enable reorderings that do not violate control dependencies. Still, the two relaxations that \tool\ features, which enable (i) value- rather than dependence-based reasoning about data flow and (ii) consideration of unexplored branches, are strictly outside the scope of Huang et al.'s technique. In Section \ref{sec:eval}, we demonstrate via direct comparison with Huang et al. the dramatic improvement in coverage thanks to these relaxation methods, which we prove to handle in a sound and maximal manner.

### 2014-09-03 overview.tex
overview

@@
 Throughout this paper, we use  the index of an event in the trace as its id, which satisfies the above properties.
+Throughout this paper, we use the event $e$ and its id $id$ interchangeably in our notation given the one-to-one correspondence between $e$ and $id$.
+Thus, for example, if the index of event $e$ is 3, then $t^e$ and $t^3$ refer to the same thread.
 \item $inst$ is the static instruction.
@@
 The target of the branch event is not important in our scope, therefore, we may abbreviate the branch instruction as the boolean expression afterwards.
-The listed instructions are so basic that it can capture most events in a common imperative programming language.
-For example, the method call can be modeled by the above list, where the argument passing and value return are modeled as assignment events, and the virtual function resolution is modeled as the branch event conditioned on the type of the caller object.
-\item $map$ maps the variable in the instruction to the runtime values.
+The listed instructions suffice to represent all trace events of interest.
+This is because a concrete finite execution trace can be reduced to a straight-line loop-free intraprocedural path program (argument passing modeled via assignments to fresh variables).
+This standard form of simplification preserves all the data-race-related information contained in the original trace.
+\item $map$ is a mapping from live expressions at the current event to their value (and thus a partial mapping from expressions to values).
+Live expressions include program variables, object fields, boolean expressions, etc.
+We use the notation $\lsyn exp \rsyn^e$ to retrieve the value associated with expression $exp$ at $e$.
+
+maps the variable in the instruction to the runtime values.
 The map stores the value for each variable observed in the dynamic run.
@@
 	\begin{center}
-		\begin{tabular}{rl}
-			\multicolumn{1}{l}{{\tt s} $::=$} & \\
-			{\tt y = x.f} $|$\ & {\bf (heapr)} \\ 
-			{\tt x.f = y} $|$\ & {\bf (heapw)} \\ %\ $|$\ {\tt x.f = $c$}\
-			{\tt z = x $bop$ y}\  $|$& {\bf (assign)} \\ %$|$\ {\tt z = $c$} $|$ {\tt z = new()}
-			{\tt if (x $bop$  y) goto ...  $|$&  {\bf (branch)} \\
-			{\tt lock(l)}\ $|$\ {\tt unlock(l)}  $|$& {\bf (sync)} \\
-			{\tt fork(t)}\ $|$\ {\tt join(t)}$|$ {\tt begin(t)}\ $|$\ {\tt end(t)}$|$  & {\bf (thread)}
+		\begin{tabular}{rcl}
+			\multicolumn{1}{l}{{\tt s} $::=$} & & \\
+			{\tt y = x.f} & $|$ & {\bf (heapr)} \\ 
+			{\tt x.f = y}  & $|$ & {\bf (heapw)} \\ %\ $|$\ {\tt x.f = $c$}\
+			{\tt z = x $bop$ y}\  & $|$ & {\bf (assign)} \\ %$|$\ {\tt z = $c$} $|$ {\tt z = new()}
+			{\tt if (x $bop$  y) goto ...} & $|$ &  {\bf (branch)} \\
+			{\tt lock(l)}\ $|$\ {\tt unlock(l)}  & $|$& {\bf (sync)} \\
+			{\tt fork(t)}\ $|$\ {\tt join(t)}$|$ {\tt begin(t)}\ $|$\ {\tt end(t)} &  & {\bf (thread)}
 		\end{tabular}
@@
 	& \left( O_1 < O_2 < O_3 \wedge O_4 < O_5 \wedge O_4 < O_7 \right) & \textbf{(program order)} \\
-\bigwedge & \left( {\tt x}^0 = 0 \wedge {\tt y}^0 = 0 \wedge {\tt y}^1 = 3 \wedge 
-	{\tt x}^2 = 1 \wedge {\tt y}^3 = 5 \right) & \textbf{(variable definitions)} \\
-\bigwedge & \left(		(R_{\tt y}^4={\tt y}^0 \wedge O_4 < O_1) \bigvee	
-(R_{\tt y}^4={\tt y}^1 \wedge O_1 < O_4 < O_3) \bigvee
-(R_{\tt y}^4={\tt y}^3 \wedge O_3 < O_4)
+\bigwedge & \left( W_{\tt x}^0 = 0 \wedge W_{\tt y}^0 = 0 \wedge W_{\tt z}^0 = 3 \wedge W_{\tt y}^1 = 3 \wedge 
+	W_{\tt x}^2 = 1 \wedge W_{\tt y}^3 = 5 \right) & \textbf{(variable definitions)} \\
+\bigwedge & \left(		(R_{\tt y}^4=W_{\tt y}^0 \wedge O_4 < O_1) \vee	
+(R_{\tt y}^4=W_{\tt y}^1 \wedge O_1 < O_4 < O_3) \vee
+(R_{\tt y}^4=W_{\tt y}^3 \wedge O_3 < O_4)
 		\right) & \textbf{(thread interference)} \\
-\bigwedge & R_{\tt y}^4 > 2 & \textbf{(path conditions)} \\
-\bigwedge & O_2 = O_5 & \textbf{(race condition)}
+\bigwedge & R_{\tt y}^4 > 2 \equiv {\bf true} & \textbf{(path conditions)} \\
+\bigwedge & t^2 \neq t^5 \wedge (2\ writes\ {\tt x} \vee 5\ writes\ {\tt x}) \wedge O_2 = O_5 & \textbf{(race condition)}
 	\end{array} 
@@
 
-\paragraph{Variable Definitions} The next set of constraints, denoted ${\tt z}^i=k$, capture variable definitions: Variable {\tt z} is assigned value $k$ at statement $i$.
+\paragraph{Variable Definitions} The next set of constraints, denoted ${\tt w}^i=k$ for local variables and $W_{\tt w}^i=k$ for shared variables, capture variable definitions: Variable {\tt w} is assigned value $k$ at statement $i$.
 For our running example, we obtain:
 $$
-	{\tt x}^0 = 0 \wedge {\tt y}^0 = 0 \wedge {\tt y}^1 = 3 \wedge 
-			{\tt x}^2 = 1 \wedge {\tt y}^3 = 5
+	W_{\tt x}^0 = 0 \wedge W_{\tt y}^0 = 0 \wedge W_{\tt z}^0 = 3 \wedge W_{\tt y}^1 = 3 \wedge
+			W_{\tt x}^2 = 1 \wedge W_{\tt y}^3 = 5
 $$
@@
 \begin{array}{lcl}
-	& & (R_{\tt y}^4={\tt y}^0 \wedge O_4 < O_1)  \\
+	& & (R_{\tt y}^4=W_{\tt y}^0 \wedge O_4 < O_1)  \\
 & \bigvee &
-	(R_{\tt y}^4={\tt y}^1 \wedge O_1 < O_4 < O_3) \\
+	(R_{\tt y}^4=W_{\tt y}^1 \wedge O_1 < O_4 < O_3) \\
 & \bigvee &
-	(R_{\tt y}^4={\tt y}^3 \wedge O_3 < O_4)
+	(R_{\tt y}^4=W_{\tt y}^3 \wedge O_3 < O_4)
 \end{array}
@@
 Notice, importantly, that the formula combines flow constraints with order constraints, which are essential to determinize the value read at a given statement. As an example, 
-$(R_{\tt y}^4={\tt y}^1 \wedge O_1 < O_4 < O_3)$ means that the value of {\tt y} read at 
+$(R_{\tt y}^4=W_{\tt y}^1 \wedge O_1 < O_4 < O_3)$ means that the value of ${\tt y}$ read at 
 line {\tt 4} is that set at line {\tt 1} assuming an execution order whereby the first statement executes followed by the fourth then third statements.
@@
 $$
-	R_{\tt y}^4 > 2
+	R_{\tt y}^4 > 2 \equiv {\bf true}
 $$
@@
 The conflicting accesses in this case, determined based on static analysis of $P$, are expressed as $O_2=O_7$.
-In addition, we negate the path condition, thereby obtaining $R_{\tt y}^4 \leq 2$ in place of $R_{\tt y}^4 \leq 2$.
+In addition, we negate the path condition, thereby obtaining $R_{\tt y}^4 > 2 \equiv {\bf false}$ in place of $R_{\tt y}^4 > 2 \equiv {\bf true}$.
 

### 2014-09-04 overview.tex
overview constraint renamed. error fixed

@@
 
+%TODO denote the event_{id} format?
 \begin{itemize}
@@
 The key property of  $id$ is {\em uniqueness}, i.e., any two events in the trace own different ids.
-The important property of $id$ is {\em monotonicity}, i.e.,  the events from the same thread should own the strictly increasing ids.
-Throughout this paper, we use  the index of an event in the trace as its id, which satisfies the above properties.
-Throughout this paper, we use the event $e$ and its id $id$ interchangeably in our notation given the one-to-one correspondence between $e$ and $id$.
-Thus, for example, if the index of event $e$ is 3, then $t^e$ and $t^3$ refer to the same thread.
+The other important property of $id$ is {\em monotonicity}, i.e.,  the events from the same thread should own the strictly increasing ids.
+Throughout this paper, unless otherwise specified, we use  the index of an event in the trace as its id, which satisfies the above properties.
+Throughout this paper,  we denote an event as $e_{id}$ with the id as the subscript.
+We may use the terms $e_{id}$ and $id$ interchangeably given their one-to-one correspondence, e.g., $t^{e_3}$ and $t^3$.
 \item $inst$ is the static instruction.
@@
 The listed instructions suffice to represent all trace events of interest.
-This is because a concrete finite execution trace can be reduced to a straight-line loop-free intraprocedural path program (argument passing modeled via assignments to fresh variables).
+This is because a concrete finite execution trace can be reduced to a straight-line loop-free call-free path program  (argument passing and value return of the method call are modeled as assignments).
 This standard form of simplification preserves all the data-race-related information contained in the original trace.
@@
 We use the notation $\lsyn exp \rsyn^e$ to retrieve the value associated with expression $exp$ at $e$.
-
-maps the variable in the instruction to the runtime values.
-The map stores the value for each variable observed in the dynamic run.
-For simplicity, we use the notion $\lsyn x \rsyn^e$ to retrieve the value associated with the variable $x$ in $e$.
+Specially,  we also store the heap location value  $\lsyn \overline{x.f} \rsyn^e$ for a field $x.f$.
+Note the difference between $\lsyn \overline{x.f} \rsyn^e$ and $\lsyn x.f \rsyn^e$, which represent the location and the value stored in it respectively.
+The location value $\lsyn \overline{x.f} \rsyn^e$ is computed as a pair $(\lsyn x \rsyn^e, f)$.
 \end{itemize}
 
-%Specially, $\lsyn x.f \rsyn^e$ returns the location of $x.f$ (rather than  the value stored in the location), which is denoted as a pair $(\lsyn x \rsyn^e, f)$.
+%TODO virtual method call
 
@@
 \subsection{Constraint System}
+The predictive analysis takes the program and a trace as the input.
+Suppose it starts with the following trace of the program in  Figure \ref{fig:running}:  $e_1 e_2 e_3 e_4 e_5$, where the line number is used as the event id (subscript).
+A potential race is between the pair, $(e_2, e_5)$, as they access the shared variable $x$ from different threads and one is a write.
 
-We begin with an explanation of the encoding process. The resulting formula is provided in Figure \ref{fig:encoding}. We describe the conjuncts comprising the formula one by one.
+Given the trace, the predictive analysis preserves the same event sequence for each thread, i.e., $T_1$ still executes $e_1 e_2 e_3$ and $T_2$ still executes $e_4 e_5$, 
+while rescheduling the events from different threads so that the potential race pair can run concurrently, i.e., $e_2$ and $e_5$ run concurrently.
+Therefore, our goal is to compute a new schedule that witnesses a race, under the (both inter-thread and intra-thread) data flow constraints, control flow constraints and synchronization constraints.
+In the following, we explain only the necessary constraints (the full description of constraints is explained in Section~\ref{sec:tech}).
 
-\begin{figure*}
-	\begin{center}
+% We purposely omit the synchronization constraints, as they are already well explained in existing techniques~\cite{yannis, pecan}.
+
+To model the schedule, we assign an order (integer) variable to each event, e.g., $O_{e_1}$. The event with smaller order variable is scheduled earlier.
+Also, we introduce symbols to denote shared accesses, e.g., $W^{id}_{x}$ ($R^{id}_{x}$) denotes the write (read) of the shared variable $x$ by the event $e_{id}$.
+The local accesses are not used in this example and therefore omitted.
+
+%We begin with an explanation of the encoding process. The resulting formula is provided in Figure \ref{fig:encoding}. We describe the conjuncts comprising the formula one by one.
+
+%\begin{figure*}
+%	\begin{center}
+%$$
+%	\begin{array}{rcl}
+%	& \left( O_1 < O_2 < O_3 \wedge O_4 < O_5 \wedge O_4 < O_7 \right) & \textbf{(program order)} \\
+%\bigwedge & \left( W_{\tt x}^0 = 0 \wedge W_{\tt y}^0 = 0 \wedge W_{\tt z}^0 = 3 \wedge W_{\tt y}^1 = 3 \wedge 
+%	W_{\tt x}^2 = 1 \wedge W_{\tt y}^3 = 5 \right) & \textbf{(variable definitions)} \\
+%\bigwedge & \left(		(R_{\tt y}^4=W_{\tt y}^0 \wedge O_4 < O_1) \vee	
+%(R_{\tt y}^4=W_{\tt y}^1 \wedge O_1 < O_4 < O_3) \vee
+%(R_{\tt y}^4=W_{\tt y}^3 \wedge O_3 < O_4)
+%		\right) & \textbf{(thread interference)} \\
+%\bigwedge & R_{\tt y}^4 > 2 \equiv {\bf true} & \textbf{(path conditions)} \\
+%\bigwedge & t^2 \neq t^5 \wedge (2\ writes\ {\tt x} \vee 5\ writes\ {\tt x}) \wedge O_2 = O_5 & \textbf{(race condition)}
+%	\end{array} 
+%$$
+%\end{center}
+%\caption{\label{fig:encoding}\tool\ encoding of the trace in Figure \ref{fig:running} as a constraint system}
+%\end{figure*}
+
+{\bf Race Condition\ } The potential race pair involves two accesses of the same shared variable from different threads, where one access is a write.
+Given any potential race pair, e.g., ($e_2$, $e_5$), it is a real race if and only if the two accesses can occur at the same time (race condition), which is captured by the race condition constraint
 $$
-	\begin{array}{rcl}
-	& \left( O_1 < O_2 < O_3 \wedge O_4 < O_5 \wedge O_4 < O_7 \right) & \textbf{(program order)} \\
-\bigwedge & \left( W_{\tt x}^0 = 0 \wedge W_{\tt y}^0 = 0 \wedge W_{\tt z}^0 = 3 \wedge W_{\tt y}^1 = 3 \wedge 
-	W_{\tt x}^2 = 1 \wedge W_{\tt y}^3 = 5 \right) & \textbf{(variable definitions)} \\
-\bigwedge & \left(		(R_{\tt y}^4=W_{\tt y}^0 \wedge O_4 < O_1) \vee	
-(R_{\tt y}^4=W_{\tt y}^1 \wedge O_1 < O_4 < O_3) \vee
-(R_{\tt y}^4=W_{\tt y}^3 \wedge O_3 < O_4)
-		\right) & \textbf{(thread interference)} \\
-\bigwedge & R_{\tt y}^4 > 2 \equiv {\bf true} & \textbf{(path conditions)} \\
-\bigwedge & t^2 \neq t^5 \wedge (2\ writes\ {\tt x} \vee 5\ writes\ {\tt x}) \wedge O_2 = O_5 & \textbf{(race condition)}
-	\end{array} 
+	O_2 = O_5
+$$.
+The race condition constraint--- together with the other constraints --- guarantees the feasibility of the predicted race if a solution is found for the overall constraint system.
+
+{\bf Intra-thread Constraints\ } The predictive analysis needs to preserve the same event sequence for each thread, which further requires
+
+\begin{itemize}
+\item {\bf Control Flow Constraints\ } Each thread takes the same control flows (or branch decisions) to reproduce the events.
+ For the running example (ignoring the statements in red), this yields:
 $$
-\end{center}
-\caption{\label{fig:encoding}\tool\ encoding of the trace in Figure \ref{fig:running} as a constraint system}
-\end{figure*}
+	R_{\tt y}^4 > 2 \equiv {\bf true}
+$$
+That is, the value of variable {\tt y} read at $e_4$ is greater than $2$.
 
-\paragraph{Program Order} The first set of constraints reflects ordering constraints between program statements.
-We use the symbol $O_i$ to denote the $i$th program statement, which yields the following formula for Figure \ref{fig:running}:
+\item {\bf Intra-thread Order Constraints\ } The events should follow the same thread-local order as in the original trace.
+This constraint is imposed by the fact that the two runs share the same instruction sequence and take the same control flows.
+We yield the following formula:
 $$
-	O_1 < O_2 < O_3 \wedge O_4 < O_5 \wedge O_4 < O_7
+	O_1 < O_2 < O_3 \wedge O_4 < O_5
 $$
-That is, the first 3 statements are totally ordered, and the {\tt if} statement executes before the body (either $O_5$ or $O_7$).
+That is, the three events from $T_1$ are totally ordered, and the branch event $e_4$ in $T_2$ executes before the event $e_5$ inside the branch body (We do not include $e_7$ as it is not in the trace).
 
-\paragraph{Variable Definitions} The next set of constraints, denoted ${\tt w}^i=k$ for local variables and $W_{\tt w}^i=k$ for shared variables, capture variable definitions: Variable {\tt w} is assigned value $k$ at statement $i$.
+\item {\bf Intra-thread Data Flow Constraints\ } The events should respect the data flow constraints imposed by each instruction.
  For our running example, we obtain:
@@
 $$
-As an example, ${\tt y}^3 = 5$ denotes that the value assigned to variable {\tt y} 
-at line {\tt 3} is $5$.
+As an example, $W_{\tt y}^3= 5$ denotes that the value assigned to variable {\tt y}  by event $e_3$ is $5$.
+\end{itemize}
 
-\paragraph{Thread Interference} To express inter-thread flow constraints, we utilize expressions of the form $R_{\tt z}^i={\tt z}^j$, which denotes that line $i$ reads variable {\tt z}, and the definition it reads comes from line $j$.
+%The next set of constraints, denoted ${\tt w}^i=k$ for local variables and $W_{\tt w}^i=k$ for shared variables, capture variable definitions: Variable {\tt w} is assigned value $k$ at statement $i$.
+
+Indeed, this constraint is satisfied by the assignments to {\tt y} both at line {\tt 1} and at line {\tt 3}.
+
+{\bf Inter-thread Data Flow Constraints\ } The inter-thread data flow constraints capture what writes the read events read from and under what scheduling condition the data flow occurs.
 The resulting formula for our example is
@@
 $$    
-Notice, importantly, that the formula combines flow constraints with order constraints, which are essential to determinize the value read at a given statement. As an example, 
-$(R_{\tt y}^4=W_{\tt y}^1 \wedge O_1 < O_4 < O_3)$ means that the value of ${\tt y}$ read at 
-line {\tt 4} is that set at line {\tt 1} assuming an execution order whereby the first statement executes followed by the fourth then third statements.
+Notice, importantly, that the formula associates the data flow constraints with the scheduling order constraints. 
+ As an example, 
+$(R_{\tt y}^4=W_{\tt y}^1 \wedge O_1 < O_4 < O_3)$ means that the read $e_4$ reads from the write $e_1$, under the condition that $e_4$ happens after $e_1$ and no other     writes (such as $e_3$) interleave them.
 
-\paragraph{Path Conditions} To preserve path conditions while potentially permitting dependence-violating reordering (e.g., a context switch after line {\tt 1} in Figure \ref{fig:running}), we model explicitly the condition.
-For the running example (ignoring the statements in red), this yields:
-$$
-	R_{\tt y}^4 > 2 \equiv {\bf true}
-$$
-That is, the value of variable {\tt y} read at line {\tt 4} is greater than $2$. Indeed, this constraint is satisfied by the assignments to {\tt y} both at line {\tt 1} and at line {\tt 3}.
-
-\paragraph{Race Condition} The final constraint, forcing the check whether a particular race is feasible, is to demand that two conflicting statements occur at the same time.
-For the potential race between lines {\tt 2} and {\tt 5}, we obtain:
-$$
-	O_2 = O_5
-$$ 
-This asserts that both statements occur simultaneously, which --- together with the other constraints --- guarantees the feasibility of the predicted race if a solution is found for the overall constraint system.
-
-\paragraph{Unexplored Branches} Beyond the encoding steps so far, which focus on the given trace, we can often encode constraints along unexplored branches.
+{\bf Unexplored Branches\ } Beyond the encoding steps so far, which focus on the given trace, we can often encode constraints along unexplored branches.
 In our running example, this is essential to discover the race between lines {\tt 2} and {\tt 7}.
@@
 In addition, we negate the path condition, thereby obtaining $R_{\tt y}^4 > 2 \equiv {\bf false}$ in place of $R_{\tt y}^4 > 2 \equiv {\bf true}$.
+We also model the other necessary constraints such as the intra-thread order constraint $O_4 < O_7$.
 
-\paragraph{Constraint Solving} Having conjoined the formuals from the different encoding steps into (i) a global representation of all feasibility constraints (path, ordering, assignment and other constraints) and (ii) the requirement for a given race to occur (expressed as simultaneous execution of the conflicting accesses), we discharge the resulting formula to an off-the-shelf constraint solver, such as Z3 or Yices.
+{\bf Constraint Solving\ } Having conjoined the formuals from the different encoding steps into (i) a global representation of all feasibility constraints (path, ordering, assignment and other constraints) and (ii) the requirement for a given race to occur (expressed as simultaneous execution of the conflicting accesses), we discharge the resulting formula to an off-the-shelf constraint solver, such as Z3 or Yices.
 If successful, the solver returns a solution for the specified constraints.

### 2014-09-04 guarantee.tex
proof, tech, overview

\section{Theoretical Guarantees}~\label{sec:guarantee}
In this section, we first prove the soundness of \tool and then discuss the detection capability as compared to existing approaches.
Both the proof and discussion are in the context of sequential consistency model~\cite{scmodel}.

\begin{mytheorem}[Soundness]
The trace $\tau$ returned by our solver is a feasible trace for the concurrent program.
\end{mytheorem}

\begin{proof}
The proof consists of two parts: (1) the trace is feasible, and (2) the trace can be generated by the program. We sketch the proof as follows.
\begin{itemize}
\item  $\tau$ is feasible. Researchers~\cite{chao,maximal} point out a trace is feasible iff it satisfies the sequential consistency, i.e., $\forall object\ o, \tau\downarrow_{o}$ satisfies the serial specification of the object $o$. If  all the events considered are of the types in Table~\ref{Ta:syntax}, then the sequential consistency precisely means the following: (1) read-write consistency, i.e., each read event of a variable should contain the value written by the most recent write event, and (2) the synchronization consistency, e.g., the lock acquire and release of a lock should not be interleaved by the lock operations of the same lock, the begin event of a thread should follow the fork event of the parent thread. Our inter-thread data flow constraints guarantee the read-write consistency among shared locations. The SSA form encodes the read-write consistency among local variables. The synchronization consistency is also captured as constraints in our solver. 
\item $\tau$ can be generated by the program, i.e., for each thread $t$, $\tau\downarrow_{t}$ can be generated by the corresponding thread code.
This claim requires that,  for any two adjacent events $e$ and $e'$ in  $\tau\downarrow_{t}$, (1) if $e$ is not a branch, $inst^e$ and $inst^{e'}$ should be adjacent in the code; (2) if $e$ is a branch, $inst^e$ and $inst^{e'}$ should be adjacent and the $inst^e$ should be evaluated as the boolean value indicated by $inst^{e'}$. 
We derive the trace  through either concrete execution or symbolic execution, which satisfies the requirements and guarantees the correctness of the claim.
\end{itemize}
\end{proof}

\begin{mydiscussion}[Detection Capability]
Our technique has stronger detection capability than existing techniques~\cite{yannis,jeff,said}.
\end{mydiscussion}

First, to the best of our knowledge, none of the existing techniques automatically explore the un-executed branches by relaxing the schedules, while ours enables this.
Second, assume we disable the exploration of un-executed branches and all techniques start with the same trace. Our technique still explores more traces. 
Figure~\ref{fig:discuss} illustrates the difference pictorially, where two vertical lines represent the progress of two threads and the circles with numbers denote the events with ids. 
Existing techniques require the reads to read the same values, therefore, enforce the schedule order $e_5 \rightarrow e_6$ and $e_2\rightarrow e_3$. We allow the reads to read different values from different writes as long as the feasibility is preserved. Comparatively, we  enforce a weaker schedule order  $e_4 \rightarrow e_6$ and $e_2\rightarrow e_3$, which allows more scheduling, e.g., the concurrent execution of $e_7$ (or some event after $e_7$) and $e_5$.

However, our technique does not identify the maximal set of traces from a single trace. The underlying reason is that, we still enforce the schedule order for preserving the base object, e.g., $e_2\rightarrow e_3$. It is possible to further relax the schedule so that $e_3$ reads from $e_1$, but this requires challenging reasoning of the field accesses as their base objects become unknown. We plan to explore this idea in future work.

\begin{figure}[htp]
\centering
\includegraphics[scale=0.7]{figs/Visio-discuss.pdf}
\caption{Illustration of Detection Capability.}\label{fig:discuss}
\end{figure}

%
%In this section, we prove formally the soundness of \tool.
%
%Researchers~\cite{pldi14,maximal} propose two axioms, {\em prefix closedness} and {\em local determinism}, which we adapt to establish our feasibility guarantee. 
%Suppose $\mathcal{F}$ denotes the domain of feasible traces.
%
%
%Intuitively, Prefix closedness means that if a trace is feasible, then the trace is still feasible after we discard the events after a point. The underlying reason is that the events after a point cannot affect the events before it. More formally,
%
%\begin{myaxiom}[Prefix closedness]~\label{axiom:prefix}
%If $t_1$ $t_2$ $\in \mathcal{F}$, then $t_1 \in \mathcal{F}$. 
%\end{myaxiom}
%
%
%
%
%Intuitively, local determinism says that only the previous events of the same thread determine the existence of an event. For example, the existence of an event is only determined by the branch  thread-locally. However, the value read by an event may be affected by another thread because a shared read reads from the most recent write, which may come from a different thread. The value in other events, such as branch or write,  are locally determined too.
%More formally,
%
%
%
%
%
%\begin{myaxiom}[Local determinism]~\label{axiom:local}
%Assume $t_1e_1, t_2 \in \mathcal{F}$, and $t_1|thread(e_1)\approx t_2|thread(e_1)$, where $\approx$ denotes the two traces are equal if the data values in the read and write events are ignored, 
%Each event is determined by the previous events in the same thread. There are four cases:
%\begin{itemize}
%%TODO refine it.
%\item {\bf Branch\ }. If op($e_1$)=branch, and $e_1$ is in the form of $x<y$ (without loss of generality), then there exists values $v_x$, $v_y$ such that $t_2 e_1 [v_x/data_x, v_y/data_y] \in \mathcal{F}$. Here $e_1 [v_x/data_x]$ represents that we replace the value of $x$ from $data_x$ to $v_x$ in $e_1$.
%\item {\bf Read\ }. If op($e_1$)=read, and $e_2$ is a read event that is identical to $e_1$ except that it may read a different value, and $t_2 e_2$ is consistent, then $t_2 e_2 \in \mathcal{F}$.
%\item {\bf Write\ }. If op($e_1$)=write, and there is a value $v$ such that $t_2 e_1[v/data]\in \mathcal{F}$. 
%\item {\bf Others \ }. If op($e_1$) is of type different from above and $t_2 e_1$ is consistent, then $t_2 e_1 \in \mathcal{F}$.
%\end{itemize}
%
%Here, the consistency is defined as follows. 
%A trace is consistency iff it satisfies 
%\end{myaxiom}
%
%The local determinism resembles the version in ~\cite{pldi14,maximal}. A crucial difference is that, for the branch event, we do not require all reads thread locally before it to read the identical values as in the original trace $t_1 e_1$. Instead, we allow the previous reads to read different values as long as the branch event is evaluated to the same boolean value, which guarantees the existence of the following events. Such relaxation allows us to find more races, while still preserving the feasibility. 
%
%
%An important assumption of ours is, we assume all local accesses are recorded in the trace so that we can determine the boolean status of a branch event computationally. ~\cite{pldi14}, which assume the local accesses are not recorded and cannot determine the status computationally, have to enforce the reads before a branch to read the same values, leading to a weaker feasibility criterion. We believe our axiom captures the real-world scenario more faithfully.
%Another important assumption is for the read-write consistency, i.e., we assume the heap invariant, so that we know exactly which reads correspond to which writes in the predicted run.
%
%%\begin{myaxiom}
%%A set of traces $\mathcal{F}$ is feasible if it satisfies prefix closedness and local determinism axioms. 
%%\end{myaxiom}
%
%
%
%They are called axioms because they are the rules that the sequentially consistent system~\cite{maximal} should follow.
%We refer the readers to the detailed discussion~\cite{maximal, pldi14}.
%
%
%
%%TODO what is \Phi
%\begin{mytheorem}[Soundness]
%All traces in the closure $closure(t, \Phi)$, which includes $t$ and is closed under the derivations in Axioms~\ref{axiom:prefix}~\ref{axiom:local}, are feasible.
%\end{mytheorem}
%
%%$\theta_1(expr(e_1))=\theta_2(expr_{e_1})$, then $t_2 \e_1 \in \mathcal{F}$. Here $\theta_1, \theta_2$ represent the mappings between variables and values established prior to $e_1$ by the same thread $thread(e_1)$. $expr(e_1)$ is a helper function that returns the expression in the branch condition. 
%\begin{proof}
%We order the traces in $closure(t)$ as, $t_0=t, t_1, t_2, t_3, \dots t_n$. Clearly, $t_0$ is feasible. In the following, we prove by induction, i.e., assume $t_{n}$ is feasible, then $t_{n+1}$ which is derived from $S=\{t_0, \dots t_n\}$ should be feasible. There are several possible derivations.
%\begin{itemize}
%\item if $t_{n+1}$ is a prefix of $t'\in S$, then $t_{n+1}\in\mathcal{F}$ because of the prefix closedness. $t_{n+1}$ shares the same mappings as $t'$.
%\item if $t_{n+1}$ is derived from $t^1 e^1, t^2\in \mathcal{F}$, we have $t^1|thread(e^1)\approx t^2|thread(e^1)$.
%\begin{itemize}
%\item $op(e^1)=branch$ and $e^1= x<y$. Our solver computes a mapping $\theta^d$ of all relevant variables such that $\theta(x<y)=\theta^d(x<y)$, then 
%$t^2 e_1[\theta^d(x)/\theta(x),\theta^d(y)/\theta(y) ]  \in \mathcal{F}$. Note that the mapping computed by solver may be different from the mapping in the trace $t^1$.
%\item $op(e^1)=read$. Our solver ensures the read-write consistency, i.e., $t^2 e^2$ is consistent, then $t^2 e^2 \in \mathcal{F}$. Note that $e^2$ is the same as $e^1$ except that it may read a different value.
%\item $op(e^1)=write$. There exists a value $v$ such that $t^2 e^2[v/data]\in \mathcal{F}$, where $v$ is included in the mapping computed by the solver. 
%\item $op(e^1)$ is of other types. Our solver ensures the consistency, i.e., $t^2 e^1$ is consistent, then $t^2 e^2 \in \mathcal{F}$.
%\end{itemize}
%\end{itemize}
%\end{proof}
%
%
%
%\begin{mytheorem}[Maximality]
%$\forall t', s.t., t'|th\approx t|th for any thread th, if t'\notin closure(t,\Phi), t' is infeasible.$
%\end{mytheorem}
%
%\begin{proof}
%We sketch the proof. All traces that satisfy  $t'|th\approx t|th for any thread th$ 
%\end{proof}

%discussion.

### 2014-09-04 introduction.tex
proof, tech, overview

@@
 \begin{tabular}{ll}
-\multicolumn{2}{c}{{\tt x = 0; y = 0; z = 3;}} \\
-\multicolumn{2}{c}{{\color{red} {\tt z = 2;}}} \\
+\multicolumn{2}{c}{{\tt x = 0; y = 0;}} \\  % z = 3;
+%\multicolumn{2}{c}{{\color{red} {\tt z = 2;}}} \\
 \hline
@@
 {\tt 3: y = 5;} & \\
-& {\tt 4: if (y > z)} \\
+& {\tt 4: if (y > 2)} \\
 & {\tt 5:~~print(1/x);} \\	
@@
  \paragraph{Our Approach} We describe a novel approach to predictive race detection, which we have implemented as the \tool\ analysis tool, that is able to relax both of these constraints.
-Compared to the state of the art \cite{JEFF}, \tool\ is able to detect x$1.5$ more races with only the first restriction relaxed, and x$2.5$ more races with both restrictions relaxed.
+Compared to the state of the art \cite{pldi14}, \tool\ is able to detect x$1.5$ more races with only the first restriction relaxed, and x$2.5$ more races with both restrictions relaxed.
 
@@
  Returning to the example in Figure \ref{fig:running}, we demonstrate how value-based reasoning improves coverage. First, value-based reasoning reveals that trace 
- [{\tt 1}, {\tt 4}, {\tt 5}, {\tt 2}, {\tt 3}] is feasible, since the first assignment, {\tt y = 3}, satisfies the condition {\tt y > z}.
+ [{\tt 1}, {\tt 4}, {\tt 5}, {\tt 2}, {\tt 3}] is feasible, since the first assignment, {\tt y = 3}, satisfies the condition {\tt y > 2}.
 Hence the race is discovered.

### 2014-09-04 overview.tex
proof, tech, overview

@@
-\section{Technical Overview}\label{Se:techoverview}
+\section{Technical Overview}\label{sec:overview}
 
@@
 \item $t$ refers to the thread that issues the event, denoted as $t^e$.
-\item  $id$ refers to the id associated with each event, denoted as $i^e$.
+\item  $id$ refers to the id associated with each event.
 The key property of  $id$ is {\em uniqueness}, i.e., any two events in the trace own different ids.
@@
 The listed instructions suffice to represent all trace events of interest.
-This is because a concrete finite execution trace can be reduced to a straight-line loop-free call-free path program  (argument passing and value return of the method call are modeled as assignments).
+This is because a concrete finite execution trace can be reduced to a straight-line loop-free call-free path program  (argument passing of the method call is modeled as assignments and the virtual call resolution is modeled as branches).
 This standard form of simplification preserves all the data-race-related information contained in the original trace.
@@
 	\end{center}
-	\caption{\label{Ta:syntax}Language syntax}
+	\caption{\label{Ta:syntax}Instruction Types}
 \end{table}
@@
 {\bf Trace \ } 
-We propose the projection operation and domains to facilitate the reasoning of the trace.
-\begin{itemize}
-\item projection operation  ``$|$'': $\tau|t$ contains only the (ordered) events from the thread $t$; $\tau|l$ contains the events involving the location $l$; $\tau|i$ contains only the event of which the id equals $i$; $\tau|\leq e$ denotes the prefix of $e$, i.e., the events preceding the event $e$; $\tau|\geq e$ denotes the events after $e$; $\tau|>=e_1\wedge <=e_2$ denotes the events between $e_1$ and $e_2$.
-\item sets: $\mathcal{T}$ denotes the set of threads involved;  $\mathcal{L}$ denotes the set of memory locations involved; 
-$\mathcal{SV}$ denotes the set of shared variables, which reference the shared locations, the shared location $l$ can be judged by counting the number of threads in $\tau|l$;  $\mathcal{E}$ denotes the set of events involved.
-\end{itemize}
+We introduce the projection operation ``$\downarrow$'' to facilitate the reasoning of the trace:
+$\tau\downarrow_{t}$ contains only the (ordered) events from the thread $t$; $\tau\downarrow_{\ell}$ contains the events involving the location $\ell$; $\tau\downarrow_{\leq id}$ denotes the prefix of $e_{id}$, i.e., the events preceding the event $e_{id}$; $\tau\downarrow_{\geq id}$ denotes the events after $e_{id}$; $\tau\downarrow_{>=id1\wedge <=id2}$ denotes the events between $e_{id1}$ and $e_{id2}$.
+
+%\begin{itemize}
+%\item projection operation  : 
+%%\item sets: $\mathcal{T}$ denotes the set of threads involved;  $\mathcal{L}$ denotes the set of memory locations involved; 
+%%$\mathcal{SV}$ denotes the set of shared variables, which reference the shared locations, the shared location $l$ can be judged by counting the number of threads in $\tau\downarrow_{\ell}$;  $\mathcal{E}$ denotes the set of events involved.
+%\end{itemize}
 
@@
 Therefore, our goal is to compute a new schedule that witnesses a race, under the (both inter-thread and intra-thread) data flow constraints, control flow constraints and synchronization constraints.
-In the following, we explain only the necessary constraints (the full description of constraints is explained in Section~\ref{sec:tech}).
+In the following, we explain only the necessary constraints (the full description of constraints is explained in Section~\ref{sec:relax1}).
 
@@
 
-To model the schedule, we assign an order (integer) variable to each event, e.g., $O_{e_1}$. The event with smaller order variable is scheduled earlier.
-Also, we introduce symbols to denote shared accesses, e.g., $W^{id}_{x}$ ($R^{id}_{x}$) denotes the write (read) of the shared variable $x$ by the event $e_{id}$.
+To model the schedule, we assign an (integer) order  variable to each event, e.g., $O_{e_1}$. The event with smaller order variable is scheduled earlier.
+Also, we introduce symbols to denote shared accesses, e.g., $W^{id}_{x}$/$R^{id}_{x}$ denotes the write/read of the shared variable $x$ by the event $e_{id}$.
 The local accesses are not used in this example and therefore omitted.
@@
 
-{\bf Race Condition\ } The potential race pair involves two accesses of the same shared variable from different threads, where one access is a write.
+{\bf Race Condition\ } The potential race pair involves two accesses of the same shared location from different threads, where at least one access is a write.
 Given any potential race pair, e.g., ($e_2$, $e_5$), it is a real race if and only if the two accesses can occur at the same time (race condition), which is captured by the race condition constraint
@@
 $$
-	W_{\tt x}^0 = 0 \wedge W_{\tt y}^0 = 0 \wedge W_{\tt z}^0 = 3 \wedge W_{\tt y}^1 = 3 \wedge
+	W_{\tt x}^0 = 0 \wedge W_{\tt y}^0 = 0 \wedge W_{\tt y}^1 = 3 \wedge
 			W_{\tt x}^2 = 1 \wedge W_{\tt y}^3 = 5
@@
 
-Indeed, this constraint is satisfied by the assignments to {\tt y} both at line {\tt 1} and at line {\tt 3}.
+%Indeed, this constraint is satisfied by the assignments to {\tt y} both at line {\tt 1} and at line {\tt 3}.
 
@@
 In our running example, this is essential to discover the race between lines {\tt 2} and {\tt 7}.
+We conduct the symbolic execution to exercise the unexplored branches, which returns a set of traces representing possible executions of the unexplored branch.
+Each trace is combined with the original trace to find races involving the accesses in the unexplored path.
+For the racy events, $e_2$ and $e_7$, we first specify the race condition constraint  $O_2=O_7$.
+In addition, we negate the path condition, thereby obtaining $R_{\tt y}^4 > 2 \equiv {\bf false}$ in place of $R_{\tt y}^4 > 2 \equiv {\bf true}$.
+We also model the other necessary constraints such as the intra-thread order constraint $O_4 < O_7$, similar to the above analysis.
 
-The conflicting accesses in this case, determined based on static analysis of $P$, are expressed as $O_2=O_7$.
-In addition, we negate the path condition, thereby obtaining $R_{\tt y}^4 > 2 \equiv {\bf false}$ in place of $R_{\tt y}^4 > 2 \equiv {\bf true}$.
-We also model the other necessary constraints such as the intra-thread order constraint $O_4 < O_7$.
+{\bf Constraint Solving\ } The formuals from the different encoding steps are conjoined and sent to the off-the-shelf solver such as {\sf z3}.
+The race under analysis is real if the solver returns a solution, which includes the scheduling order and the values for the variables.
 
-{\bf Constraint Solving\ } Having conjoined the formuals from the different encoding steps into (i) a global representation of all feasibility constraints (path, ordering, assignment and other constraints) and (ii) the requirement for a given race to occur (expressed as simultaneous execution of the conflicting accesses), we discharge the resulting formula to an off-the-shelf constraint solver, such as Z3 or Yices.
-If successful, the solver returns a solution for the specified constraints.
+{\bf Highlights \ } In this example, our analysis finds two more pairs of races than existing analyses~\cite{yannis, jeff}, which include $(e_2, e_5)$ and $(e_2, e_7)$.
+The key insight for finding the first race is, we do not require $e_4$ to read from $e_3$ to retain the original value, as what existing approaches do, rather, we allow $e_4$ to read from $e_1$ while still preserving the feasibility of the branch.
+This relaxation allows the events after $e_4$ to run concurrently with the events before $e_3$, leading to the detection of the first race.
+The key insight for finding the second race is, we explore the unexplored branches and ask the solver to compute a feasible schedule that exercises them.
+The neighboring branches are likely to contain racy events if the executed branches contain racy events.
 
-In particular, the solution discloses a feasible trace that gives rise to the race at hand.
+%In particular, the solution discloses a feasible trace that gives rise to the race at hand.
 The trace is identified uniquely via the order enforced in the solution over the variables $O_i$, which represent scheduling order.

### 2014-09-04 relax1.tex
proof, tech, overview

\section{Relaxation of Flow Dependencies}~\label{sec:relax1}
The main goal of our analysis is to, given an original trace $\tau$, derive a trace $\tau'$ over the same set of events in $\tau$, which has a new scheduling  of the events (the order of events inside each thread should remain unchanged) and a new mapping from variables to values in the events. Our first relaxation comes from the insight that, we allow the variables to read different values than in the original trace even if such variables are used to determine the control flow, while existing approaches enforce them to read the same values as in the original trace. As a result, our relaxation allows more schedules, which are likely to expose more races.

Not every scheduling or every mapping  leads to a new feasible trace $\tau'$, therefore, we need to compute the scheduling and mapping that lead to feasible trace. We explain the technique details in the following.

% % 
% TODO: Explain the format of the raw trace in impl.tex.
% %

%We begin by describing the first relaxation, whereby the dependence structure defined by the original trace is potentially violated while retaining soundness.

\subsection{SSA Form of the Trace}~\label{sec:ssa}
First of all, we require the SSA form of the trace, i.e., each variable in defined exactly once in the trace. Such requirement is a prerequisite for the computation of the mapping, in which each variable is mapped to exactly one value. Another side effect is that the def/use chains become explicit in the SSA form, which simplifies the following analysis steps.

\tool\ handles the local assignment and heap accesses differently.
\begin{itemize}
\item {\bf Local Assignment\ } The SSA form for local assignment resembles the SSA form of static instructions in compiler optimization, except that the loops and recursions are fully resolved in a concrete trace, obviating the need for the {\sf Phi} node. More concretely, we replace the variable $v$ defined in an event, as well as the following uses of the definition, to a new variable $v^{id}$, where $id$ is the unique id of the event. The uniqueness of the id guarantees that no two events define the same variable, i.e., each variable is defined exactly once. Note that the uses of a definition can be computed easily by scanning the trace afterwards for the variables of the same name before the next redefinition. 
\item {\bf Heap accesses\ } The accesses of the local heap locations and the accesses of shared locations behave differently, therefore, we introduce different SSA form for them. 
\begin{itemize}
\item {\bf Local Heap Accesses\ } The accesses of local heap locations behave similarly to the local assignment as both the definition and uses belong to the single thread. Therefore we model them as local assignments: We introduce a fresh local variable {\tt $l_{o.f }$} to replace each definition and  the corresponding uses of the location of $o.f$ ($o$ denotes an object referenced by the variable $x$ in $x.f$).  
%An assumption here is the uses and the definition still refer to the same base object $o$ in the predicted run, which is enforced through the constraint specification %(Section~\ref{sec:relax1}).
\item  {\bf Shared Heap Accesses\ } The accesses of shared locations are more complex. Each shared read may read from one of multiple writes that update the shared location, under different schedules.     Given the write $x.f=y$ or the read $y=x.f$, we introduce two symbols, $W^{id}_{\tt o.f }$, which denotes the write access by the event $id$ to the  shared heap location  $o.f$  ($o$ denotes an object referenced by the variable $x$ in $x.f$), and $R^{id}_{\tt  o.f }$,  which denotes the read access  by the event $id$ to the location $o.f$. 
\end{itemize}
We use the heap location such as $o.f$ in the the symbolic form of heap accesses. An underlying assumption is that, the heap location remains unchanged in our predictive analysis, which we ensure through additional constraints (Section~\ref{sec:relax1}).
\end{itemize}

% Similar to the local heap accesses, an assumption for the correlation is that the writes and reads still refer to the same base object $o$ in the predicted run, which we guarantee by specifying additional constraints (Section~\ref{sec:relax1}).

% We could encode the question of shared versus local accesses as additional constraints (requiring that a variable be accessed by more than one thread), but that would be more complicated and less efficient.

%\paragraph{Basic Encoding: Local Accesses}
%
%The fundamental encoding transformation is to induce Static Single Assignment (SSA) form on the raw trace, such that
%a variable is defined exactly once. In this way, , and encoding of trace events as constraints is simplified. As an illustration, trace
%\begin{quote}
%	{\tt 1: x=1; 2: x<3; 3: x=3;} \\
%	{\tt 4: y=1;} \\
%	{\tt 5: z=x+y}
%\end{quote}
%becomes
%\begin{quote}
%	{\tt 1: x$^1$=1; 2: x$^1$<3; 3: x$^3$=3;} \\
%	{\tt 4: y$^4$=1;} \\
%	{\tt 5: z$^5$=x$^3$+y$^4$}
%\end{quote}
%(For readability, we version variables according to the line number of their definition.)

%%TODO update the intro+moti, make sure the same style
\begin{figure}
\centering
\begin{tabular}{ll}
\multicolumn{2}{c}{{\tt {\bf x} = 0; {\bf y} = 0;}} \\
\hline
\multicolumn{1}{c}{$T_1$} & \multicolumn{1}{c}{$T_2$} \\
\hline
{\tt 1: s=0; } & \\
{\tt 2: for(i=1;i<2;i++)} & \\
{\tt 3: \ \ \ s+=i;} & \\
{\tt 4: {\bf y} = s;} & \\
& {\tt 5: if ({\bf y} > 2)} \\
& {\tt 6:~~print({\bf x});} \\	
\end{tabular}
\caption{Running Example (shared variables are in bold font). }
\label{fig:running2}
\end{figure}

\begin{figure}
\centering
\begin{tabular}{l|l}
\hline
\multicolumn{1}{c}{$Trace$} & \multicolumn{1}{c}{$SSA \ form \ of\   Trace$} \\
\hline
{\tt 0: {\bf x}=0} &  {\tt 0: $W^0_x$=0}    \\
{\tt 1: {\bf y}=0} &   {\tt 1: $W^1_y$=0}   \\
{\tt 2: s=0} &  {\tt 2: $s^2$=0}   \\
{\tt 3: i=1} &     {\tt 3: $i^3$=1}   \\
{\tt 4: i<2} &    {\tt 4: $i^3$<3} \\
{\tt 5: s=s+i} & {\tt 5: $s^5$=$s^2$+$i^3$}   \\
{\tt 6: i=2} &       {\tt 6: $i^6$=2}  \\
{\tt 7: i<2} &      {\tt 7: $i^6$<3}  \\
{\tt 8: {\bf y} = s;} &  {\tt 8: $W^{8}_y$ = $s^5$;}  \\
{\tt 9: {\bf y} > 2}  &    {\tt 9: $R^{9}_y$ > 2} \\
{\tt 10: print({\bf x});} &  {\tt 10: print($R^{10}_x$);}  \\
\end{tabular}
\caption{Trace}
\label{fig:t4running2}
\end{figure}

Figure~\ref{fig:t4running2} exemplifies a trace and its SSA form, where the trace is generated from the program in Figure~\ref{fig:running2}.
Each label in  Figure~\ref{fig:t4running2} on the left hand denotes the id of the event. As seen, event $e_5$ uses the variable $s^2$ defined at event 2 and defines the variable $s^5$, which is used at $e_8$. The read/write of the shared variables are denoted in the special form, such as $W^{8}_y$ and  $R^{9}_y$.

\subsection{Constraint System}~\label{sec:constraints}
Based on the SSA form of the trace, we build the constraints to compute a new trace with the new schedule and new mapping of variables to values.
To model the schedule, we introduce the order variable $O_{id}$ for each event $e_{id}$. Given two events $e_i, e_j$ from two different threads, $O_{e_i}<O_{e_j}$ in a trace means $e_i$ is scheduled before $e_j$. We omit the synchronization constraints purposely, as they are well explained in all existing predictive analysis techniques~\cite{yannis, pldi14}.

%Having explained how the trace is encoded, we now describe in detail how constraints are derived from the trace, such that any permutation considered by the analysis is guaranteed to represent a feasible execution schedule.

{\bf Race Condition\ } Following the definition, a pair, ($e_i$, $e_j$), forms a race iff (1) $e_i$ and $e_j$ are accesses of the same location $\ell$ by different threads, (2) at least one of them writes to $\ell$, and (3) $e_i$ and $e_j$ run concurrently. We refer to the candidate pair throughout this section and call the two events as the racy events.

We first identify all candidate pairs of events that access the same location from different threads (and at least one is a write), then we check each pair  separately.   The checking is achieved by encoding all necessary constraints and invoking a constraint solver.  The first constraint asserts the feasibility of the concurrent execution of the racy events. 
$$
O_{i} = O_{j}
$$

{\bf Intra-thread Constraints\ } Given the candidate race pair, $e_i$ and $e_j$, suppose $e_i$ is after $e_j$ in the original trace $\tau$, the predictive analysis reschedules the events in the prefix of $e_i$, i.e., prior to $e_i$, so that $e_i$ and $e_j$ can run concurrently. The predictive analysis by design requires that each thread should follow the same event sequence prior to $e$ as in the original run.  The preservation of the event sequence for each thread requires the following intra-thread constraints:

\begin{itemize}
\item {\bf Control Flow Constraints\ } The branches in the predicted run should take the same decisions as in the original trace so that each thread reproduces the same set of events.  Specifically, we only need to reason about the branches prior to $e_i$.

More formally,  we require that 
$$
\bigwedge_{e_k \in \tau\downarrow_{\leq i} \wedge  type^{e_k}=branch.\ }  inst^{e_k} 	\equiv \lsyn inst \rsyn^{e_k}
$$ 
where  the branch event $e_k$ is used as a boolean expression. For example, given the branch {\tt if(x<y)\dots}, the constraint is in the form of $(x<y)\equiv true$, suppose the branch is evaluated as true in the original trace. The constraint specifies that the boolean expression in the predicted trace should be evaluated to the same boolean value as in the original trace. Importantly, different from existing analyses \cite{yannis,pldi14}, we do not pose the requirement that the values flowing into branching statements remain the same, but adopt the relaxed requirement that the evaluation of branching expressions remains the same. For example, suppose  the branch event $x<y$ takes the value $1<2$ in the original trace, existing analyses require the same values for the variables $x$ and $y$ in the predicted run, while we allow  the branch event to take other values of the variables, such as $3<4$. 
\item {\bf Intra-thread Order Constraints\ } The events in the sequence should follow the same order as in the original trace. More formally, we require that
$$
\begin{array}{rl}
\forall e_m, e_n \in \tau\downarrow_{\leq i}, s.t., & t^{e_m} = t^{e_n}. \\
& m < n}\  \Rightarrow O_{m} < O_{n}
\end{array}
$$ 
The constraint specifies that the two events from the same thread should follow the same order as reflected by the ids of the events. Note that $m$ in $e_m$ denotes its id. Again, we only consider the events in the prefix of $e_i$.
\item {\bf Intra-thread Data Flow Constraints\ } The mapping of local variables should not contradict the data flow semantics of each instruction.   More formally, we require that
$$
\bigwedge_{e_k \in \tau\downarrow_{\leq i} \wedge type(e_k)=assign|heapr|heapw}\
	inst^{e_k}
$$
where $inst^{e_k}$, such as $x=y+z$ or $R^{id}_{x}=y$, captures the constraint over the values of the variables imposed by the instruction. For example, $x=y+z$ is not satisfied if the solver maps the three variables to $3, 4, 5$ respectively. Again we consider only the events in the prefix of  $e_i$. Specially, for the instruction involving object creation, $x=new (...)$, we encode the object as a unique integer that denotes its heap address, calculated by $System.identityHashcode(x)$ at runtime.
\end{itemize}

%Rescheduling through Relaxation of Flow Dependencies
{\bf Inter-thread Data Flow Constraints for Relaxation\ } We now move to the novel feature of \tool, which is its ability to explore execution schedules that depart from the value flow exhibited in the original trace. More precisely, \tool\ is able to relax value flow dependencies in the original trace:  a read access may read a
different value from other write events, as long as the read value enforces feasibility.
 This is strictly beyond the coverage potential of existing predictive analyses, which restrict trace transformations to ones where any read access must read the same  value (often from the same write event) as in the original trace.

To ensure feasibility under relaxation of flow dependencies, we need to secure the flow between the read/write with the execution schedule. 
For example, in Figure \ref{fig:running}, for the read access to ${\tt y}$ at $e_4$ obtains the value assigned to ${\tt y}$ at $e_1$, we need the schedule where 
$e_4$ happens after $e_1$ and other writes such as $e_3$ do not interleave them ($e_3$ can only happen after $e_4$ in this case).

% $R_{\tt y}^4={\tt y}^1 \wedge O_1 < O_4 < O_3$ specifies that in a schedule , the $R_{\tt y}^4={\tt y}^1$.

%TODO l and ell
In general, the constraint formula, given read $R^{m}_{\ell}$ of location $\ell$ at the event $e_m$ with set ${\cal W}$ of matching write events (i.e., events including write access to $\ell$), takes the following form:
$$
\begin{array}{rll}
\bigvee_{e_n \in {\cal W}} &  & (R^{m}_{\ell} = W^{n}_{\ell}) \\
&		\bigwedge 	&  O_n < O_m \\
&		\bigwedge_{e_p \in {\cal W} \setminus \{ e_n \}} & (O_p < O_n \vee O_m < O_p)
\end{array}
$$
This disjunctive formula iterates over all matching write events, and demands for each that (i) it occurs prior to the read event ($O_n < O_m$) and (ii) all other write events either occur before ($O_p < O_n$) it or after the read event
($O_m < O_p$).

An important concern that arises due to relaxation of flow dependencies is that heap accesses may change their meaning, i.e., they involve different base objects and do no longer match with each other. As an illustration, we refer to Figure \ref{fig:heapAccess}. While the read at the event $e_7$ appears to match the write at $e_5$, this is conditioned on the read at $e_6$ being linked to the assignment at $e_4$. Suppose  the event $e_6$ reads from $e_3$ due to the relaxation, then $e_7$ and $e_5$ no longer share the same base object (or the same location). Even worse, we do not know what events $e_7$ match because the base object becomes unknown/symbolic.

\begin{figure}
	\centering
	\begin{tabular}{ll}
		\hline
		\multicolumn{1}{c}{$T_1$} & \multicolumn{1}{c}{$T_2$} \\
		\hline
		{\tt 1: $x^1$ = new();// creates o1} & \\
		{\tt 2: $x^2$= new();// creates o2} & \\
		{\tt 3: $W^3_{y}$ = $x^1$;} & \\
		{\tt 4: $W^4_{y}$ = $x^2$;} & \\
		{\tt 5: $W^5_{o2.f}$ = 5;} & \\	
		& {\tt 6: $z^6$ = $R^6_{y}$;} \\
		& {\tt 7: $w^7$ = $R^7_{o2.f}$;} \\
	\end{tabular}
	\caption{\label{fig:heapAccess} A trace.}
\end{figure}

To address this challenge, we enhance the constraint system with the requirement that heap objects that are dereferenced in the field reference before the racy events retain their original address in the predicted trace.
This achieves two guarantees: First, matching heap access events in the original trace are guaranteed to also match in the predicted trace. Second, candidate races in the original trace remain viable in the predicted trace as they still refer to same location. Third, the sharedness nature of each heap access remains the same because
each location is accessed by the same number of threads given that the base object is unchanged.

%Note that, the resolution of the virtual method call is modeled as the branch events that check the type of the target object, for which the constraints naturally guarantee the branch to take the same decision, i.e., the call to be resolved to the same method implementation.

%Finally, in a practical setting involving virtual method calls, call-site resolutions are the same across the original and predicted traces. 
%Notice that in this setting, we consider as relevant not only the targets of field dereferences but also the targets of virtual method invocations.

Formally, we require that
$$
\bigwedge_{ e_k \ has\  x.f \wedge e_i \in \tau\downarrow_{\leq i} }\
	x \equiv  \lsyn x \rsyn^{e_k}
$$
%where ${\sf env}$ is the state mapping from local variables to their value and $\sigma(t,\tau'')$ 
%(resp. $\sigma(t',\tau'')$) is the state arising at trace $t$ (resp. $t'$) immediately before event $\tau''$.
This constraint fixes that all heap dereferences  prior to the racy event retain their original base object as in the original trace $\tau$. 
In general, we need to specify such heap constraints when we hardcode the local heap accesses as local variables (Section~\ref{sec:ssa}) to ensure that the hardcoded data flow remains valid  after the relaxation. However,  most local heap accesses in the form of $x.f$ never read from any shared locations according to the static analysis, and therefore, cannot be affected by the relaxation. As an optimization, we do not specify the heap constraints for such local heap accesses.

By sending the above constraints to a solver, we compute the necessary schedule orders among the events as well as the mapping of the variables. The necessary schedule orders define a partial order among the events, which permit a set of schedules that define the complete order that complies with it.

### 2014-09-04 relax2.tex
proof, tech, overview

--- 2014-09-03 constraints.tex
@@
-\section{Exploration of Unexecuted Branches}
+\section{Exploration of Unexecuted Branches}~\label{sec:relax2}
 We now switch to the second feature of \tool, which is its ability to reason about unexplored branches.
@@
 	\For {$\tau': S$}
-	\State $\tau_r=\tau-((\tau|t^e)|>=e)$
+	\State $\tau_r=\tau-\tau\downarrow_{t^e \wedge \geq e}$
 	\State $\tau_r=\tau_r.replace(e, neg(e) + \tau')$
@@
 
-The analysis starts with the event $e$ (exclusively), with the current state $\sigma$ at $e$ as the initial state. 
-Remember that we use the runtime values for two types of events: the heap value of the base objects and the evaluation result of the boolean expression in the branch. 
-Other values are maintained but not used. Therefore, we also need to maintain the heap value and the boolean expression value in the symbolic execution.
-
-First, we propose standard kill/gen rules for three basic heap instructions to maintain the heap update along a given path.
-Specially, we separate the object creation from the local assignment.
-In the object creation, we assign a unique integer to represent the newly created object and store it into the state $\sigma$.
-For field reference, $\sigma[x.f]$ is actually short for two-level interpretation $\sigma[\sigma[x].f]$, where the first finds out the base object $\sigma[x]$ and the second searches for the object referenced by its field $f$.
-The kill/gen rules take effect only when $x.f$ is not of the primitive type.
+The analysis starts at the event $e$ (exclusively), with the state $\sigma$.
+Initially, the state includes the runtime heap value for the base objects that are already resolved.
+During the symbolic execution, the state is updated to maintain the heap value and also the boolean expression values for branches.
+The values for other variables are not used indeed and therefore treated as symbolic.
+The update of the state is realized through the   standard kill/gen rules for three basic heap instructions.
+In the object creation, which we separate from the local assignment, we assign a unique integer to represent the newly created object and store it into the state $\sigma$.
+For field access, $\sigma[x.f]$ actually denotes $\sigma[o.f]$, where  $o$ is the base object resolved in the state.
+These kill/gen rules take effect only when $x.f$  is not of the primitive type.
 

### 2014-09-04 eval.tex
edits

@@
 \section{Evaluation}~\label{sec:eval}
+Our evaluation focuses on the effectiveness  and scalability of our approach.
+To measure the effectiveness, we compare with the predictive analysis, {\sf RV}~\cite{}.
+We chose RV for two main reasons: (1) {\sf RV} is the only open source predictive analysis, (2) {\sf RV} represents the state-of-the-art approach, which is theoretically proven to have higher detection capability than other approaches.
 
-{\bf Evaluation Method} We conduct our experiment with a set of large applications, e.g.,  those from the {\sf Dacapo} benchmark suite. 
-The benchmarks are shown in Table.
-We compare our approach with the state-of-the-art technique proposed by Huang et al.
-As the capability of finding races depends on the monitored run, for fair comparison, we
- apply both techniques to the same monitored run. Specifically, we use two threads in the monitored run as two threads are sufficient to manifest most of races~\cite{shanlu}.
+{\bf Evaluation Method} We conduct our experiment on a large set of applications, which are also used to evaluate {\sf RV}.
+Specifically, the set includes large applications such as {\sf Jigsaw}, {\sf Xalan}, {\sf Lusearch}.
+To handle large applications that make use of reflection, we applied {\sf Tamiflex}~\cite{}, which replaces reflective calls with the concrete method invocations recorded in the observed run for the purpose of effective static analysis.
+We omit the benchmark {\sf eclipse} because the current version of {\sf Tamiflex}  leads to abnormal execution after the instrumentation, which throws an exception when the main service is started.
+By applying our tool to such abnormal execution, we identify only 3 races, similarly to the report of RV~\cite{}.
+In addition, the {\sf montecarlo} benchmark requires an external input that we downloaded from the internet and simplified.
+For the large applications, we utilized the most lightweight available configuration.
 
-The {\sf Dacapo} relies on the reflection functionality to execute, which imposes challenges to static analysis and transformation.
-We apply the tool {\sf Tamiflex}~\cite{} to overcome the problems, which resolves the reflection using the classes observed in the dynamic run.
-The window size is 10000.
+ As the detection capability  depends on the observed run, for fair comparison we monitor the execution once by recording all necessary information required by both approaches, and then apply both techniques to the monitored run.
+Besides, our reported data for RV may be different from the original report for two reasons: (1) the different observed runs lead to different sets of races, (2) the original implementation of RV contains a bug in identification of the branches, which leads to  misses of many branches and  incorrect reduction of dependences during the analysis.
+We confirmed the bug with the author and fixed the bug in our experiments.\footnote{We have created an in-depth report on the bug, along with a witness test case.
+(See:  \url{https://sites.google.com/site/recipe3141/}.)}
 
@@
 
-\subsection{Scalability Issues}
-Our approach suffers from scalability issues for two reasons.
-First, as we record local access events in the monitoring run,  the monitoring run takes long time, e.g., more than 2 hours, and our trace is typically large, e.g., at the GB level.
-To avoid the out of memory issue during the monitoring run, we store the event to the database on the fly, rather than storing them in a buffer.
-Besides, we terminate the monitoring run with a shutdown hook if it is beyond 30 minutes.
-Second,  the prediction phase needs to load the trace to memory, which easily causes out of memory problem.
-To solve the problem, we separate the analysis into several runs, each run restarting the JVM and resuming from the window of last run.
-Different runs communicate the window id using the external file.
-These runs are organized together automatically using the ant.
-The outputs of the files are merged into one summary finally.
-Besides,
+%Specifically, we use two threads in the monitored run as two threads are sufficient to manifest most of races~\cite{shanlu}.
 
-\begin{table}[htbp]
-\caption{haha}
-\begin{tabular}{|l|l|}
+\subsection{Effectiveness}
+Table~\ref{tab:main} shows the main results of our analysis, which includes four sections: Trace (details about the trace), Races (detected races), Difference (comparison with RV) and Running time (the time taken by the analysis).
+
+The Trace section includes the number of threads ($Th$), the number of shared reads ($Reads$) and shared writes ($Writes$), the subset of shared reads that read the base object references  ($Base$), where the  base object references are references used as the base/target in the following field reference or the method invocation,  the number of branches ($Br$), the synchronization events ($Sync$) and the total number of events ($Total$), which includes local accesses in addition to the aforementioned events.
+Specifically, for the branch, we report it in the form of $A/B$, where $A$ refers to the number of branches used by our analysis and $B$ refers to the number of branches used by RV.
+To ensure that the predicted run sees the same base object at each shared read/write, {\sf RV} inserts the artificial branch immediately in front of each shared access (and array accesses).
+We do not use such artificial branches.
+
+We make interesting observations about the Trace section.
+The non-local events, i.e., all the events listed in Table~\ref{tab:main}, occupy around 30\% of the total trace in the first 7 benchmarks, but occupy less than 1\% of the trace in the rest benchmarks, which have relatively more complex logic.
+The reads of the shared base objects occupy a small portion ($1/3$-$1/10$) of the total shared reads.
+The rest shared reads read only the primitive values or
+the references that are not the base references, e.g., the references involved only in the nullness check.
+Besides, our analysis involves significantly less branch events as compared to the {\sf RV} approach.
+The difference plays an important role in the detection, which we will explain shortly.
+
+The $Races$ section shows the number of races detected by {\sf Recipe-s}, i.e., {\sf Recipe} without exploring un-executed paths, {\sf Recipe}, i.e., the fully-fledged version, and {\sf RV}.
+By comparing the {\sf Recipe-s} and {\sf Recipe}, we find {\sf Recipe} finds 100+ more races, which demonstrates the strength of exploring  un-executed paths.
+Intuitively,  {\sf Recipe-s} predicts based on a single trace, while {\sf Recipe} predicts based on multiple traces containing different execution paths.
+We also compare the {\sf Recipe} version with {\sf RV} version, as illustrated in the $Difference$ section, where $Diff$ shows the races found by {\sf Recipe} but missed by {\sf RV}, $Diff'$ shows the races found by {\sf RV} but missed by {\sf Recipe}.
+
+First of all, {\sf Recipe} finds many 150+ more races than {\sf RV}.
+The reasons are multifold: (1) {\sf Recipe} can reason about the accesses in the un-executed paths, while {\sf RV} and {\sf Recipe-s} can only reason about the accesses in the executed paths.
+(2) {\sf Recipe} or {\sf Recipe-s} allows the relaxation of the scheduling even if it breaks the inter-thread read/write dependence in the observed run.
+{\sf Recipe} allows the majority of the shared reads, i.e., the reads of primitive values, to freely read from a different value from a different write as long as the value leads to the same branch decisions.
+{\sf RV}, however, requires them to read the same values as in the observed run.
+(3) An critical optimization proposed by {\sf RV} is to preserve dependences only for the reads before the preceding branches  of the racy events, rather than all the reads.
+However, this optimization is underplayed by the fact that {\sf RV} introduces huge amount of artificial branches, i.e., one before each shared field access, which ensures the use of the same base objects.
+We get rid of such artificial branches and instead, rely on the small amount of base read events to ensure the use of the same base objects (Section~\ref{}).
+Our strategy reduces the number of branches greatly and amplifies the effectiveness of the optimization.
+We also conduct case studies (Section~\ref{}) to better illustrate the scenarios.
+
+Another interesting observation is, although {\sf Recipe} should produce all races found by {\sf RV} in theory, {\sf Recipe} may miss races found by {\sf RV} in practice (Column $Diff'$), i.e., {\sf Recipe} is not strictly more effective than {\sf RV} in practice.
+The underlying reason is due to the limits of the constraint solver: (1) the solver cannot compute constraints with very complex arithmatic operations (2) the solver does not support some program constants such as the scientific notation, $3E-10$.
+
+The last section, Running time, compares the analysis time for both approaches.
+We find our approach is significantly slower than {\sf RV}, e.g., {\sf RV} often finishes within 200 seconds, while our approach may take more than 1 hour.
+This is because our approach needs to reason about the computation among variables inside the local access events, while {\sf RV} needs to only reason about the order relations among the events.
+
+{\bf Summary of advantage and weakness\ } In general, our approach is flexible, which allows the relaxation of schedules and paths based on the fine-grained reasoning of value flows.
+As a result, it detects many more races compared to existing approaches.
+On the other hand, it is heavy weighted.
+The suggestion is to combine it with the lightweight approach with soundness guarantee such as {\sf RV}, by treating {\sf RV} as the preprocessing and instructing {\sf Recipe} to skip those confirmed by {\sf RV}.
+In this way, we can complement {\sf Recipe} by finding those missed by it due to the limit of the constraint solver.
+
+%TODO Reorder the columns: RV ReceipeS Recipe
+
+\begin{table*}[htbp]
+\caption{Relaxed Analysis}
+\begin{flushleft}
+\begin{tabular}{|l|r|r|r|r|l|r|r|r|r|r|r|r|r|r|}
 \hline
-ad & fda \\ \hline
-\multicolumn{1}{|r|}{1} & \multicolumn{1}{r|}{2} \\ \hline
+ & \multicolumn{ 7}{c|}{Trace} & \multicolumn{ 3}{c|}{Races} & \multicolumn{ 2}{c|}{Difference} & \multicolumn{ 2}{c|}{Running time (sec)} \\ \hline
+Benchmarks & \multicolumn{1}{l|}{Th} & \multicolumn{1}{l|}{Reads} & \multicolumn{1}{l|}{Writes} & \multicolumn{1}{l|}{Base} & Br & \multicolumn{1}{l|}{sync} & \multicolumn{1}{l|}{Total} & \multicolumn{1}{l|}{Recipe-s} & \multicolumn{1}{l|}{Recipe} & \multicolumn{1}{l|}{RV} & \multicolumn{1}{l|}{Diff} & \multicolumn{1}{l|}{Diff'} & \multicolumn{1}{l|}{Recipe} & \multicolumn{1}{l|}{RV} \\ \hline
+critical & 3 & 13 & 7 & 5 & 2/13 & 6 & 78 & 8 & 8 & 8 & \multicolumn{1}{c|}{0} & 0 & 8 & \multicolumn{1}{c|}{2} \\ \hline
+airline & 11 & 45 & 15 & 4 & 32/82 & 30 & 317 & 9 & 9 & 9 & 0 & 0 & 490 & 4 \\ \hline
+account & 3 & 46 & 21 & 10 & 3/47 & 6 & 227 & 2 & 5 & 5 & 3 & 3 & 41 & 4 \\ \hline
+pingpong & 4 & 7 & 7 & 3 & 0/15 & 6 & 111 & 1 & 1 & 1 & 0 & 0 & 19 & 1 \\ \hline
+bbuffer & 4 & 640 & 118 & 10 & 634/1.1K & 217 & 3.3K & 13 & 25 & 9 & 21 & 5 & 62 & 5 \\ \hline
+bubblesort & 26 & 1.3k & 966 & 121 & 155/2.8K & 322 & 8.4K & 7 & 7 & 7 & 0 & 0 & 3295 & 3 \\ \hline
+bufwriter & 5 & 165 & 52 & 75 & 16/130 & 44 & 525 & 4 & 10 & 2 & 8 & 0 & 63 & 9 \\ \hline
+mergesort & 5 & 38 & 33 & 5 & 15/472 & 28 & 1.7K & 3 & 10 & 3 & 10 & 0 & 37 & 5 \\ \hline
+raytracer & 2 & 31 & 5 & 12 & 314/8,2K & 676 & 94.5K & 4 & 6 & 4 & 2 & 0 & 47 & 2 \\ \hline
+montecarlo & 2 & 5 & 86 & 2 & 1.9K/38.2K & 21.1K & 1.9M & 1 & 4 & 1 & 3 & 0 & 1 & 17 \\ \hline
+moldyn & 2 & 605 & 61 & 104 & 19.6K/52.6K & 62 & 203.4K & 6 & 14 & 2 & 12 & 0 & 2842 & 1 \\ \hline
+ftpserver & 28 & 684 & 299 & 71 & 4.4K/233.3K & 78.2K & 3.9M & 99 & 152 & 57 & 108 & 13 & 811 & 153 \\ \hline
+jigsaw & 12 & 525 & 702 & 211 & 63.2K/467.9K & 86.7K & 5.5M & 17 & 23 & 8 & 15 & 0 & 33 & 7 \\ \hline
+sunflow & 9 & \multicolumn{1}{l|}{2.1K} & \multicolumn{1}{l|}{1.3K} & 473 & 201.3K/827.0K & 50K & \multicolumn{1}{l|}{7.1M} & 38 & 78 & 20 & 69 & 11 & 4520 & 22 \\ \hline
+xalan & 9 & \multicolumn{1}{l|}{1.4K} & \multicolumn{1}{l|}{0.9K} & 209 & 15.7K/103.2K & \multicolumn{1}{l|}{190.1K} & \multicolumn{1}{l|}{6.6M} & 2 & 6 & 2 & 4 & 0 & 5317 & 10 \\ \hline
+lusearch & 10 & \multicolumn{1}{l|}{2.3K} & \multicolumn{1}{l|}{0.5K} & 715 & 22.2K/164K & \multicolumn{1}{l|}{93.2K} & \multicolumn{1}{l|}{9.1M} & 27 & 49 & 14 & 38 & 5 & 5430 & 8 \\ \hline
 \end{tabular}
-\label{tb:ref}
-\end{table}
+\end{flushleft}
+\label{tab:main}
+\end{table*}
 
-\section{Case studies}
+\subsection{Case studies}
 We manually inspect the reported races in small applications to gain better understanding about our approach.
@@
 
+{\bf Relaxing the Paths\ } Figure~\ref{fig:relax2} demonstrates how our approach relaxes the schedulings to account for the unexecuted paths. 
+Each thread invokes the method {\sf Sorting}, which recursively starts two children threads if there are two more available entries in the thread pool (lines 8-11), or  starts one child thread if there is only one available entry (lines 4-5).
+The availability is computed through the static method {\sf available} with the constant {\tt total}, which is equal to 5.
+The shared variable {\tt alive} denotes the used entries.
+
+Initially there is one sorting thread. After it starts Thread 1 and Thread 2, there are three threads alive and only two more entries are available.
+Suppose in the observation run, Thread 2 consumes both thread entries and starts the children Thread 3 and Thread 4 (not shown), updating {\tt alive} to 5, then Thread 1 cannot execute the branch at lines 4-5.
+Huang et al.~\cite{} require the predicted run to preserve the dependence denoted by the dotted line since the dependence affects the branch condition at line 2.
+As a result, the predicted run follows the same branch decision and cannot reason about the code at lines 4-5.
+Our analysis does not have such limitation.
+Instead, it allows the predicted run to reason about the un-explored code.
+Specifically, it does not need to preserve the dependence from line 11 to line 1 and it allows line 1 (Thread 1) to read from line 9 (Thread 2).
+As a result, the branch condition guarding the unexplored branch is evaluated to be true, enabling the unexplored path in the constraint solver.
+Finally, the solver identifies the race between line 5 (Thread 1) and line 1 (Thread 3).
+Note that the two lines are synchronized on different locks~\footnote{We abbreviate the {\tt synchronized} keyword as {\tt sync}.
+}.
+
+\begin{figure}[htp]
+\centering
+\includegraphics[scale=0.7]{cases/Visio-msort.pdf}
+\caption{Relaxation of Paths}\label{fig:relax2}
+\end{figure}
+
 \section{Validity Threat}

### 2014-09-04 guarantee.tex
edits

@@
 \section{Theoretical Guarantees}~\label{sec:guarantee}
-In this section, we first prove the soundness of \tool and then discuss the detection capability as compared to existing approaches.
+In this section, we first prove the soundness of \tool\ and then discuss the detection capability as compared to existing approaches.
 Both the proof and discussion are in the context of sequential consistency model~\cite{scmodel}.

### 2014-09-04 relax1.tex
edits

@@
 \section{Relaxation of Flow Dependencies}~\label{sec:relax1}
-The main goal of our analysis is to, given an original trace $\tau$, derive a trace $\tau'$ over the same set of events in $\tau$, which has a new scheduling  of the events (the order of events inside each thread should remain unchanged) and a new mapping from variables to values in the events.
-Our first relaxation comes from the insight that, we allow the variables to read different values than in the original trace even if such variables are used to determine the control flow, while existing approaches enforce them to read the same values as in the original trace.
+The main goal of our analysis is to derive, given an original trace $\tau$, a trace $\tau'$ over the same set of events in $\tau$, which has a new scheduling  of the events (the order of events inside each thread should remain unchanged) and a new mapping from variables to values in the events.
+Our first relaxation comes from the insight that we allow the variables to read different values than in the original trace even if such variables are used to determine the control flow, while existing approaches enforce them to read the same values as in the original trace.
 As a result, our relaxation allows more schedules, which are likely to expose more races.
 
-Not every scheduling or every mapping  leads to a new feasible trace $\tau'$, therefore, we need to compute the scheduling and mapping that lead to feasible trace.
-We explain the technique details in the following.
+Not every scheduling or every mapping  leads to a new feasible trace $\tau'$.
+Therefore, we need to compute scheduling and mapping instances that lead to feasible traces.
+We achieve this via constraint solving.
+We explain the details of our technique in the following.
 
@@
 \subsection{SSA Form of the Trace}~\label{sec:ssa}
-First of all, we require the SSA form of the trace, i.e., each variable in defined exactly once in the trace.
-Such requirement is a prerequisite for the computation of the mapping, in which each variable is mapped to exactly one value.
+First of all, we rewrite the trace into SSA form, such that every variable in defined exactly once in the trace.
+This requirement is a prerequisite for the computation of the mapping, in which each variable is mapped to exactly one value.
 Another side effect is that the def/use chains become explicit in the SSA form, which simplifies the following analysis steps.
@@
 The uniqueness of the id guarantees that no two events define the same variable, i.e., each variable is defined exactly once.
-Note that the uses of a definition can be computed easily by scanning the trace afterwards for the variables of the same name before the next redefinition.
-\item {\bf Heap accesses\ } The accesses of the local heap locations and the accesses of shared locations behave differently, therefore, we introduce different SSA form for them. 
+Note that uses of a definition can be computed easily by scanning the trace suffix for accesses to the same variable name before its next definition.
+\item {\bf Heap accesses\ } The accesses of the local heap locations and the accesses of shared locations play different roles in our analysis. We therefore introduce different SSA form for them. 
 \begin{itemize}
-\item {\bf Local Heap Accesses\ } The accesses of local heap locations behave similarly to the local assignment as both the definition and uses belong to the single thread.
+\item {\bf Local Heap Accesses\ } The accesses of local heap locations behave similarly to the local assignment as both the definition and uses belong to a single thread.
 Therefore we model them as local assignments: We introduce a fresh local variable {\tt $l_{o.f }$} to replace each definition and  the corresponding uses of the location of $o.f$ ($o$ denotes an object referenced by the variable $x$ in $x.f$).
@@
 \item  {\bf Shared Heap Accesses\ } The accesses of shared locations are more complex.
-Each shared read may read from one of multiple writes that update the shared location, under different schedules.
+Each shared read may read from one of multiple writes that update the shared location under different schedules (e.g., the read access to {\tt y} at line 4 in Figure \ref{fig:running}, which can either obtain the definition before line 1 or at line 1 or at line 3).
 Given the write $x.f=y$ or the read $y=x.f$, we introduce two symbols, $W^{id}_{\tt o.f }$, which denotes the write access by the event $id$ to the  shared heap location  $o.f$  ($o$ denotes an object referenced by the variable $x$ in $x.f$), and $R^{id}_{\tt  o.f }$,  which denotes the read access  by the event $id$ to the location $o.f$.
@@
 We use the heap location such as $o.f$ in the the symbolic form of heap accesses.
-An underlying assumption is that, the heap location remains unchanged in our predictive analysis, which we ensure through additional constraints (Section~\ref{sec:relax1}).
+An underlying assumption is that the heap location remains unchanged in our predictive analysis, which we ensure through additional constraints (Section~\ref{sec:relax1}).
 \end{itemize}
@@
 Each label in  Figure~\ref{fig:t4running2} on the left hand denotes the id of the event.
-As seen, event $e_5$ uses the variable $s^2$ defined at event 2 and defines the variable $s^5$, which is used at $e_8$.
-The read/write of the shared variables are denoted in the special form, such as $W^{8}_y$ and  $R^{9}_y$.
+As seen, event $e_5$ uses the variable $s^2$ defined at event $e_2$ and defines the variable $s^5$, which is used at $e_8$.
+Reads and writes of shared variables are denoted in the special form explained above (e.g.: $W^{8}_y$ and  $R^{9}_y$).
 
@@
 To model the schedule, we introduce the order variable $O_{id}$ for each event $e_{id}$.
-Given two events $e_i, e_j$ from two different threads, $O_{e_i}<O_{e_j}$ in a trace means $e_i$ is scheduled before $e_j$.
-We omit the synchronization constraints purposely, as they are well explained in all existing predictive analysis techniques~\cite{yannis, pldi14}.
+Given two events $e_i, e_j$ from two different threads, $O_{e_i}<O_{e_j}$ means that $e_i$ is scheduled before $e_j$.
+We omit the synchronization constraints intentionally, as they are well explained in all existing predictive analysis techniques~\cite{yannis, pldi14}.
 
@@
 
-{\bf Race Condition\ } Following the definition, a pair, ($e_i$, $e_j$), forms a race iff (1) $e_i$ and $e_j$ are accesses of the same location $\ell$ by different threads, (2) at least one of them writes to $\ell$, and (3) $e_i$ and $e_j$ run concurrently.
-We refer to the candidate pair throughout this section and call the two events as the racy events.
+{\bf Race Condition\ } Following the standard definition, pair ($e_i$, $e_j$) of events forms a race iff (1) $e_i$ and $e_j$ are accesses of the same location $\ell$ by different threads, (2) at least one of them writes to $\ell$, and (3) $e_i$ and $e_j$ run concurrently.
+We refer to the candidate pair throughout this section as (candidate) \emph{racy events}.
 
-We first identify all candidate pairs of events that access the same location from different threads (and at least one is a write), then we check each pair  separately.
+We first identify all candidate pairs of events that access the same location from different threads (and at least one is a write).
+We then check each pair  separately.
 The checking is achieved by encoding all necessary constraints and invoking a constraint solver.
-The first constraint asserts the feasibility of the concurrent execution of the racy events.
+The first constraint asserts the feasibility of the concurrent execution of the racy events:
 $$
@@
 
-{\bf Intra-thread Constraints\ } Given the candidate race pair, $e_i$ and $e_j$, suppose $e_i$ is after $e_j$ in the original trace $\tau$, the predictive analysis reschedules the events in the prefix of $e_i$, i.e., prior to $e_i$, so that $e_i$ and $e_j$ can run concurrently.
+{\bf Intra-thread Constraints\ } Given the candidate race pair, $e_i$ and $e_j$, suppose $e_i$ is after $e_j$ in the original trace $\tau$.
+The predictive analysis reschedules the events in the prefix of $e_i$, i.e., prior to $e_i$, so that $e_i$ and $e_j$ can run concurrently.
 The predictive analysis by design requires that each thread should follow the same event sequence prior to $e$ as in the original run.
@@
 where  the branch event $e_k$ is used as a boolean expression.
-For example, given the branch {\tt if(x<y)\dots}, the constraint is in the form of $(x<y)\equiv true$, suppose the branch is evaluated as true in the original trace.
+For example, given the branch {\tt if(x<y)\dots}, the constraint is in the form of $(x<y)\equiv {\bf true}$ assuming the branch evaluates to {\bf true} in the original trace.
 The constraint specifies that the boolean expression in the predicted trace should be evaluated to the same boolean value as in the original trace.
-Importantly, different from existing analyses \cite{yannis,pldi14}, we do not pose the requirement that the values flowing into branching statements remain the same, but adopt the relaxed requirement that the evaluation of branching expressions remains the same.
-For example, suppose  the branch event $x<y$ takes the value $1<2$ in the original trace, existing analyses require the same values for the variables $x$ and $y$ in the predicted run, while we allow  the branch event to take other values of the variables, such as $3<4$.
+Importantly, Unlike existing analyses \cite{yannis,pldi14}, we do not pose the requirement that the values flowing into branching statements remain the same, but adopt the relaxed requirement that the evaluation of branching expressions remains the same.
+For example, suppose  the branch event $x<y$ takes the value $1<2$ in the original trace.
+Existing analyses require the same values for the variables $x$ and $y$ in the predicted run, while we allow  the variables flowing into the branch event to assume other values, such as $3<4$, as long as the branch expression retains the same truth value.
 \item {\bf Intra-thread Order Constraints\ } The events in the sequence should follow the same order as in the original trace. More formally, we require that
@@
 \forall e_m, e_n \in \tau\downarrow_{\leq i}, s.t., & t^{e_m} = t^{e_n}. \\
-& m < n}\  \Rightarrow O_{m} < O_{n}
+& m < n\  \Rightarrow O_{m} < O_{n}
 \end{array}
 $$ 
-The constraint specifies that the two events from the same thread should follow the same order as reflected by the ids of the events.
+This constraint specifies that two events from the same thread should follow the same order as reflected by the ids of the events.
 Note that $m$ in $e_m$ denotes its id.
@@
 %Rescheduling through Relaxation of Flow Dependencies
-{\bf Inter-thread Data Flow Constraints for Relaxation\ } We now move to the novel feature of \tool, which is its ability to explore execution schedules that depart from the value flow exhibited in the original trace.
+{\bf Inter-thread Data Flow Constraints for Relaxation\ } We now move to the first novel feature of \tool, which is its ability to explore execution schedules that depart from the value flow exhibited in the original trace.
 More precisely, \tool\ is able to relax value flow dependencies in the original trace:  a read access may read a
@@
 To ensure feasibility under relaxation of flow dependencies, we need to secure the flow between the read/write with the execution schedule. 
-For example, in Figure \ref{fig:running}, for the read access to ${\tt y}$ at $e_4$ obtains the value assigned to ${\tt y}$ at $e_1$, we need the schedule where 
-$e_4$ happens after $e_1$ and other writes such as $e_3$ do not interleave them ($e_3$ can only happen after $e_4$ in this case).
+For example, in Figure \ref{fig:running} for the read access to ${\tt y}$ at $e_4$ to obtain the value assigned to ${\tt y}$ at $e_1$, we need a schedule where 
+$e_4$ happens after $e_1$ and other writes such as $e_3$ do not interleave them. ($e_3$ can only happen after $e_4$ in this case.)
 
@@
 
-An important concern that arises due to relaxation of flow dependencies is that heap accesses may change their meaning, i.e., they involve different base objects and do no longer match with each other.
+An important concern that arises due to relaxation of flow dependencies is that heap accesses may change their meaning, i.e., they involve different base objects and no longer match with each other.
 As an illustration, we refer to Figure \ref{fig:heapAccess}.
 While the read at the event $e_7$ appears to match the write at $e_5$, this is conditioned on the read at $e_6$ being linked to the assignment at $e_4$.
-Suppose  the event $e_6$ reads from $e_3$ due to the relaxation, then $e_7$ and $e_5$ no longer share the same base object (or the same location).
-Even worse, we do not know what events $e_7$ match because the base object becomes unknown/symbolic.
+Suppose  the event $e_6$ reads from $e_3$ due to the relaxation.
+Then $e_7$ and $e_5$ no longer share the same base object (or the same location).
+Even worse, we do not know which events $e_7$ matches, because the base object is no longer known.
 
@@
 
-To address this challenge, we enhance the constraint system with the requirement that heap objects that are dereferenced in the field reference before the racy events retain their original address in the predicted trace.
+To address this challenge, we enhance the constraint system with the requirement that heap objects that are dereferenced in field access statements before the racy events retain their original address in the predicted trace.
 This achieves two guarantees: First, matching heap access events in the original trace are guaranteed to also match in the predicted trace.
 Second, candidate races in the original trace remain viable in the predicted trace as they still refer to same location.
-Third, the sharedness nature of each heap access remains the same because
-each location is accessed by the same number of threads given that the base object is unchanged.
+Third, sharing between threads of heap locations remains unchanged  because
+each location is accessed by the same number of threads given that the base object is the same.
 

### 2014-09-05 overview.tex
relax1 ok

@@
 \subsection{Preliminaries}
-To facilitate our presentation, we first introduce some terminologies used throughout this paper. At the high level, a trace is a sequence of events recorded during the observation run.
+To facilitate our presentation, we first introduce some terminologies used throughout this paper. Intuitively, a trace is a sequence of events recorded during the observation run.
 
-{\bf Event\ } An event, $e=<t, id, inst, map>$, is a concrete representation that captures the details about the runtime execution of a static instruction $inst$.
+{\bf Event\ } An event, $e=<t, id, inst>$, is a concrete representation of the runtime execution of a static instruction $inst$.
 
-%TODO denote the event_{id} format?
 \begin{itemize}
-\item $t$ refers to the thread that issues the event, denoted as $t^e$.
+\item $t$ refers to the thread that issues the event $e$, denoted as $t^e$.
 \item  $id$ refers to the id associated with each event.
@@
 The other important property of $id$ is {\em monotonicity}, i.e.,  the events from the same thread should own the strictly increasing ids.
-Throughout this paper, unless otherwise specified, we use  the index of an event in the trace as its id, which satisfies the above properties.
+Unless otherwise specified, we use  the index of an event in the trace as its id, which satisfies the above properties.
 Throughout this paper,  we denote an event as $e_{id}$ with the id as the subscript.
-We may use the terms $e_{id}$ and $id$ interchangeably given their one-to-one correspondence, e.g., $t^{e_3}$ and $t^3$.
+%We may use the terms $e_{id}$ and $id$ interchangeably given their one-to-one correspondence, e.g., $t^{e_3}$ and $t^3$.
 \item $inst$ is the static instruction.
@@
 This standard form of simplification preserves all the data-race-related information contained in the original trace.
-\item $map$ is a mapping from live expressions at the current event to their value (and thus a partial mapping from expressions to values).
-Live expressions include program variables, object fields, boolean expressions, etc.
-We use the notation $\lsyn exp \rsyn^e$ to retrieve the value associated with expression $exp$ at $e$.
-Specially,  we also store the heap location value  $\lsyn \overline{x.f} \rsyn^e$ for a field $x.f$.
-Note the difference between $\lsyn \overline{x.f} \rsyn^e$ and $\lsyn x.f \rsyn^e$, which represent the location and the value stored in it respectively.
-The location value $\lsyn \overline{x.f} \rsyn^e$ is computed as a pair $(\lsyn x \rsyn^e, f)$.
 \end{itemize}
-
-%TODO virtual method call
 
@@
 
-{\bf Trace \ } 
-We introduce the projection operation ``$\downarrow$'' to facilitate the reasoning of the trace:
-$\tau\downarrow_{t}$ contains only the (ordered) events from the thread $t$; $\tau\downarrow_{\ell}$ contains the events involving the location $\ell$; $\tau\downarrow_{\leq id}$ denotes the prefix of $e_{id}$, i.e., the events preceding the event $e_{id}$; $\tau\downarrow_{\geq id}$ denotes the events after $e_{id}$; $\tau\downarrow_{>=id1\wedge <=id2}$ denotes the events between $e_{id1}$ and $e_{id2}$.
+{\bf Trace \ } A trace $\tau$ is a sequence of events. It can also be represented more comprehensively as,  $\tau=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \}, O, \theta>$.
+\begin{itemize}
+\item  $\Gamma$ refers to the set of threads involved in the trace, which include $t_1$, $t_2$, \dots, $t_n$.
+\item   $\{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \}$ denotes the event sequences produced by each thread.  
+\item  $O$ is a function that assigns  to every event an integer value that denotes its scheduling order. For example, $e_i$ is scheduled earlier than $e_j$ if $O(e_i)<O(e_j)$.
+\item  $\theta$ is a function that maps every variable to a value.
+For example, given an event $e$ executing the instruction $x=y+z$, the mapping recorded in the trace may be $\theta(x)=3$,  $\theta(y)=2$,  $\theta(z)=1$.
+If the variable $x$ is defined in two events, the mapping for $x$ becomes ambiguous.
+To avoid the ambiguity, we conduct a pre-processing  (Section~\ref{sec:relax1}) that produces the SSA form of the trace, in which each variable is defined exactly once.
+\end{itemize}
+
+%\item $map$ is a mapping from live expressions at the current event to their value (and thus a partial mapping from expressions to values).
+Live expressions include program variables, object fields, boolean expressions, etc.
+We use the notation $\lsyn exp \rsyn^e$ to retrieve the value associated with expression $exp$ at $e$.
+%Specially,  we also store the heap location value  $\lsyn \overline{x.f} \rsyn^e$ for a field $x.f$.
+Note the difference between $\lsyn \overline{x.f} \rsyn^e$ and $\lsyn x.f \rsyn^e$, which represent the location and the value stored in it respectively.
+The location value $\lsyn \overline{x.f} \rsyn^e$ is computed as a pair $(\lsyn x \rsyn^e, f)$.
+
+%We introduce the projection operation ``$\downarrow$'' to facilitate the reasoning of the trace:
+%$\tau\downarrow_{t}$ contains only the (ordered) events from the thread $t$; $\tau\downarrow_{\ell}$ contains the events involving the location $\ell$; $\tau\downarrow_{\leq id}$ denotes the prefix of $e_{id}$, i.e., the events preceding the event $e_{id}$; $\tau\downarrow_{\geq id}$ denotes the events after $e_{id}$; $\tau\downarrow_{>=id1\wedge <=id2}$ denotes the events between $e_{id1}$ and $e_{id2}$.
 
@@
 
-Given the trace, the predictive analysis preserves the same event sequence for each thread, i.e., $T_1$ still executes $e_1 e_2 e_3$ and $T_2$ still executes $e_4 e_5$, 
-while rescheduling the events from different threads so that the potential race pair can run concurrently, i.e., $e_2$ and $e_5$ run concurrently.
-Therefore, our goal is to compute a new schedule that witnesses a race, under the (both inter-thread and intra-thread) data flow constraints, control flow constraints and synchronization constraints.
-In the following, we explain only the necessary constraints (the full description of constraints is explained in Section~\ref{sec:relax1}).
+Given the trace, $\tau=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \}, O, \theta>$, the predictive analysis computes a new trace,  $\tau'=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \}, O', \theta'>$.
+$\tau'$ preserves the same event sequence for each thread, i.e., $T_1$ still executes $e_1 e_2 e_3$ and $T_2$ still executes $e_4 e_5$, but it reschedules the events from different threads so that the potential race pair runs concurrently, i.e., $e_2$ and $e_5$ run concurrently.
+When the schedule changes, the variables may get different values.
+Therefore, we need to compute both the new schedule $O'$ and the new value mapping $\theta'$, in order to witness a race.
+To guarantee that $\tau'$ is a feasible trace of the program $P$, the computation needs to respect a set of constraints, which include the value constraints, control flow constraints and synchronization constraints~\footnote{We omit the synchronization constraints in this paper, which are already well explained in existing techniques~\cite{yannis, pecan}.}.
+We explain them briefly in the following.
+The full details are explained in Section~\ref{sec:relax1} and the feasibility guarantee is explained in Section~\ref{sec:guarantee}.
 
-% We purposely omit the synchronization constraints, as they are already well explained in existing techniques~\cite{yannis, pecan}.
-
-To model the schedule, we assign an (integer) order  variable to each event, e.g., $O_{e_1}$. The event with smaller order variable is scheduled earlier.
-Also, we introduce symbols to denote shared accesses, e.g., $W^{id}_{x}$/$R^{id}_{x}$ denotes the write/read of the shared variable $x$ by the event $e_{id}$.
-The local accesses are not used in this example and therefore omitted.
+%The local accesses are not used in this example and therefore omitted.
 
@@
 
+We first preprocess the trace as follows.
+We introduce symbols to replace the shared accesses in the instructions, e.g., We use $W^{id}_{x}$/$R^{id}_{x}$ to replace the write/read of the shared variable $x$ by the event $e_{id}$.
+%The rationale is explained in Section~\ref{sec:relax1}.
+
 {\bf Race Condition\ } The potential race pair involves two accesses of the same shared location from different threads, where at least one access is a write.
@@
 $$
-	O_2 = O_5
+	O'(e_2) = O'(e_5)
 $$.
@@
 $$
-	R_{\tt y}^4 > 2 \equiv {\bf true}
+	\theta'(R_{\tt y}^4) > 2 \equiv {\bf true}
 $$
-That is, the value of variable {\tt y} read at $e_4$ is greater than $2$.
+That is, in the predicted trace $\tau'$, the value of variable {\tt y} read at $e_4$ should be greater than $2$.
 
@@
 $$
-	O_1 < O_2 < O_3 \wedge O_4 < O_5
+	O'(e_1) < O'(e_2) < O'(e_3) \wedge O'(e_4) < O'(e_5)
 $$
-That is, the three events from $T_1$ are totally ordered, and the branch event $e_4$ in $T_2$ executes before the event $e_5$ inside the branch body (We do not include $e_7$ as it is not in the trace).
+That is, in the predicted trace $\tau'$, the three events from $T_1$ are ordered in the same way as in $\tau$, and similarly, the branch event $e_4$ in $T_2$ still executes before the event $e_5$ inside the branch body.
 
-\item {\bf Intra-thread Data Flow Constraints\ } The events should respect the data flow constraints imposed by each instruction.
+\item {\bf Intra-thread Value Constraints\ } The events should respect the value constraints imposed by each instruction.
  For our running example, we obtain:
 $$
-	W_{\tt x}^0 = 0 \wedge W_{\tt y}^0 = 0 \wedge W_{\tt y}^1 = 3 \wedge
-			W_{\tt x}^2 = 1 \wedge W_{\tt y}^3 = 5
+\begin{array}{l}
+	\theta'(W_{\tt x}^0) = 0 \wedge \theta'(W_{\tt y}^0) = 0 \wedge \theta'(W_{\tt y}^1) = 3 \\ 
+	\wedge	\theta'(W_{\tt x}^2) = 1 \wedge \theta'(W_{\tt y}^3) = 5
+	\end{array}
 $$
-As an example, $W_{\tt y}^3= 5$ denotes that the value assigned to variable {\tt y}  by event $e_3$ is $5$.
+As an example, $\theta'(W_{\tt y}^3)= 5$ denotes that the predicted run still assigns the value 5 to the variable {\tt y}  at event $e_3$.
 \end{itemize}
@@
 
-{\bf Inter-thread Data Flow Constraints\ } The inter-thread data flow constraints capture what writes the read events read from and under what scheduling condition the data flow occurs.
+{\bf Inter-thread Value Constraints\ } The inter-thread value constraints capture what writes the read events read from and under what scheduling condition the value flow occurs.
 The resulting formula for our example is
@@
 \begin{array}{lcl}
-	& & (R_{\tt y}^4=W_{\tt y}^0 \wedge O_4 < O_1)  \\
+	& & (\theta'(R_{\tt y}^4)=\theta'(W_{\tt y}^0) \wedge O'(e_4) < O'(e_1) ) \\
 & \bigvee &
-	(R_{\tt y}^4=W_{\tt y}^1 \wedge O_1 < O_4 < O_3) \\
+	(\theta'(R_{\tt y}^4)=\theta'(W_{\tt y}^1) \wedge O'(e_1) < O'(e_4) < O'(e_3)) \\
 & \bigvee &
-	(R_{\tt y}^4=W_{\tt y}^3 \wedge O_3 < O_4)
+	(\theta'(R_{\tt y}^4)=\theta'(W_{\tt y}^3) \wedge O'(e_3) < O'(e_4))
 \end{array}
 $$    
-Notice, importantly, that the formula associates the data flow constraints with the scheduling order constraints. 
+Notice, importantly, that the formula associates the value constraints with the scheduling order constraints. 
  As an example, 
-$(R_{\tt y}^4=W_{\tt y}^1 \wedge O_1 < O_4 < O_3)$ means that the read $e_4$ reads from the write $e_1$, under the condition that $e_4$ happens after $e_1$ and no other     writes (such as $e_3$) interleave them.
+$\theta'(R_{\tt y}^4)=\theta'(W_{\tt y}^1) \wedge O'(e_1) < O'(e_4) < O'(e_3)$ means that, in the predicted run, the read $e_4$ reads from the write $e_1$, under the condition that $e_4$ happens after $e_1$ and no other  writes (such as $e_3$) interleave them.
 
@@
 Each trace is combined with the original trace to find races involving the accesses in the unexplored path.
-For the racy events, $e_2$ and $e_7$, we first specify the race condition constraint  $O_2=O_7$.
-In addition, we negate the path condition, thereby obtaining $R_{\tt y}^4 > 2 \equiv {\bf false}$ in place of $R_{\tt y}^4 > 2 \equiv {\bf true}$.
-We also model the other necessary constraints such as the intra-thread order constraint $O_4 < O_7$, similar to the above analysis.
+For the racy events, $e_2$ and $e_7$, we first specify the race condition constraint  $O'(e_2)=O'(e_7)$.
+In addition, we negate the path condition, thereby obtaining $\theta'(R_{\tt y}^4) > 2 \equiv {\bf false}$ in place of $\theta'(R_{\tt y}^4) > 2 \equiv {\bf true}$.
+We also model the other necessary constraints such as the intra-thread order constraint $O'(e_4) < O'(e_7)$, similar to the above analysis.
 

### 2014-09-05 relax1.tex
relax1 ok

@@
 \section{Relaxation of Flow Dependencies}~\label{sec:relax1}
-The main goal of our analysis is to derive, given an original trace $\tau$, a trace $\tau'$ over the same set of events in $\tau$, which has a new scheduling  of the events (the order of events inside each thread should remain unchanged) and a new mapping from variables to values in the events.
-Our first relaxation comes from the insight that we allow the variables to read different values than in the original trace even if such variables are used to determine the control flow, while existing approaches enforce them to read the same values as in the original trace.
-As a result, our relaxation allows more schedules, which are likely to expose more races.
-
-Not every scheduling or every mapping  leads to a new feasible trace $\tau'$.
-Therefore, we need to compute scheduling and mapping instances that lead to feasible traces.
-We achieve this via constraint solving.
-We explain the details of our technique in the following.
+Given the trace, $\tau=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \}, O, \theta>$, the main goal of our analysis is to derive a new trace,  $\tau'=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \}, O', \theta'>$, which reschedules the events from different threads (the order of events inside each thread should remain unchanged) to witness the race.
+The key insight underlying our first relaxation is that, the read events in $\tau'$ are allowed by our analysis to read different values from different writes (than in trace $\tau$), but they are required by existing approaches to read the same values as in $\tau$.
+Therefore, our relaxation allows more schedules, which are likely to expose more races.
+Note that we also need to compute the new value mapping $\theta'$ because the rescheduling may alter the value every variable is assigned to.
+The computation of the new schedule and new value mapping is explained in the following.
 
@@
 The uniqueness of the id guarantees that no two events define the same variable, i.e., each variable is defined exactly once.
-Note that uses of a definition can be computed easily by scanning the trace suffix for accesses to the same variable name before its next definition.
-\item {\bf Heap accesses\ } The accesses of the local heap locations and the accesses of shared locations play different roles in our analysis. We therefore introduce different SSA form for them. 
+Consider the example in Figure~\ref{fig:running2}, where the line number is also used as event id,  the local variable $x$ defined at line 4 is renamed as $x^4$ in the trace, while $x$ defined at line 9 is renamed as $x^9$, which is used in the following two events.
+At line 5, where the object $o3$ is created,  we define a reference variable $w^5$.
+Internally,  the object $o3$ is encoded as a unique integer that indicates its heap address.
+\item {\bf Heap accesses\ } The def/use relation between local heap accesses can be locally determined.
+Therefore, we treat them in the same way as local assignment.
+The def/use relations between shared heap accesses are more complex.
+We may link a read with different writes under different schedules, e.g., the read access to {\tt y} at line 4 in Figure \ref{fig:running} can either obtain the definition before line 1 or at line 1 or at line 3.
+Therefore, we introduce two symbols to represent the shared read and shared write respectively and use them to establish different links flexibly.
+The details are as follows.
 \begin{itemize}
-\item {\bf Local Heap Accesses\ } The accesses of local heap locations behave similarly to the local assignment as both the definition and uses belong to a single thread.
-Therefore we model them as local assignments: We introduce a fresh local variable {\tt $l_{o.f }$} to replace each definition and  the corresponding uses of the location of $o.f$ ($o$ denotes an object referenced by the variable $x$ in $x.f$).
+\item {\bf Local Heap Accesses\ } Given the write $x.f=y$ to a local heap location at event $e_{id}$,  we introduce a fresh local variable {\tt $l^{id}_{o.f}$} ($o$ is the runtime object referenced by $x$) to replace $x.f$.
+We also replace the uses of the location $o.f$ accordingly.
+Here  the subscript $o.f$ is interpreted as a heap location, i.e., the field $f$ of the object $o$.
+Consider the example in Figure~\ref{fig:running2}, at lines 6-7, we replace the local heap accesses as the local variable $l^6_{o3.f}$.
 %An assumption here is the uses and the definition still refer to the same base object $o$ in the predicted run, which is enforced through the constraint specification %(Section~\ref{sec:relax1}).
-\item  {\bf Shared Heap Accesses\ } The accesses of shared locations are more complex.
-Each shared read may read from one of multiple writes that update the shared location under different schedules (e.g., the read access to {\tt y} at line 4 in Figure \ref{fig:running}, which can either obtain the definition before line 1 or at line 1 or at line 3).
-Given the write $x.f=y$ or the read $y=x.f$, we introduce two symbols, $W^{id}_{\tt o.f }$, which denotes the write access by the event $id$ to the  shared heap location  $o.f$  ($o$ denotes an object referenced by the variable $x$ in $x.f$), and $R^{id}_{\tt  o.f }$,  which denotes the read access  by the event $id$ to the location $o.f$.
+\item  {\bf Shared Heap Accesses\ }    Given the write $x.f=y$ or the read $y=x.f$ of a shared location at event $e_{id}$, we introduce two symbols, $W^{id}_{\tt o.f }$ and $R^{id}_{\tt  o.f }$ respectively, where $o$ is the runtime object referenced by the $x$.
+Similarly, the subscript $o.f$ is interpreted as a heap location, i.e., the field $f$ of the object $o$.
 \end{itemize}
@@
 \centering
-\begin{tabular}{ll}
-\multicolumn{2}{c}{{\tt {\bf x} = 0; {\bf y} = 0;}} \\
-\hline
-\multicolumn{1}{c}{$T_1$} & \multicolumn{1}{c}{$T_2$} \\
-\hline
-{\tt 1: s=0; } & \\
-{\tt 2: for(i=1;i<2;i++)} & \\
-{\tt 3: \ \ \ s+=i;} & \\
-{\tt 4: {\bf y} = s;} & \\
-& {\tt 5: if ({\bf y} > 2)} \\
-& {\tt 6:~~print({\bf x});} \\	
+\begin{tabular}{ll|l}
+\multicolumn{3}{c}{{\tt {\bf s1} = new();//$o1$}} \\
+\multicolumn{3}{c}{{\tt {\bf s2} = new();//$o2$}} \\
+\multicolumn{1}{c}{$T_1$} & \multicolumn{1}{c}{$T_2$}  &  \multicolumn{1}{c}{$Trace$}\\
+{\tt 1: {\bf s1}.f=3; } &  & {\tt $W^{1}_{o1.f}$=3;}\\
+{\tt 2: {\bf s2}.f=1; } &  & {\tt $W^{2}_{o2.f}$=1;}\\
+{\tt 3: {\bf s1}.f=2; } & & {\tt $W^{3}_{o1.f}$=2;} \\
+& {\tt 4: x=0;} & {\tt $x^4$=0;}\\
+& {\tt 5: w=new();//$o3$} & {\tt $w^5$=$o3$;}\\
+& {\tt 6: w.f=10;}  & {\tt $l^6_{o3.f}$=10;}\\
+& {\tt 7: z=w.f;} & {\tt $z^7$=$l^6_{o3.f}$;}\\
+& {\tt 8: y={\bf s1}.f;} & {\tt $y^8$=$R^8_{o1.f}$;}\\
+& {\tt 9: x=y+z;} & {\tt $x^9$=$y^8$+$z^7$;}\\
+& {\tt 10: if(x>11)}& {\tt if($x^9$>11)}\\ 
+& {\tt 11: \ \ {\bf s2}.f=x;}  & {\tt $W^{11}_{o2.f}$=$x^9$;}\\
 \end{tabular}
-\caption{Running Example (shared variables are in bold font). }
+\caption{Running Example (shared variables are in bold font, the line number is used as the event id). }
 \label{fig:running2}
@@
 
-\begin{figure}
-\centering
-\begin{tabular}{l|l}
-\hline
-\multicolumn{1}{c}{$Trace$} & \multicolumn{1}{c}{$SSA \ form \ of\   Trace$} \\
-\hline
-{\tt 0: {\bf x}=0} &  {\tt 0: $W^0_x$=0}    \\
-{\tt 1: {\bf y}=0} &   {\tt 1: $W^1_y$=0}   \\
-{\tt 2: s=0} &  {\tt 2: $s^2$=0}   \\
-{\tt 3: i=1} &     {\tt 3: $i^3$=1}   \\
-{\tt 4: i<2} &    {\tt 4: $i^3$<3} \\
-{\tt 5: s=s+i} & {\tt 5: $s^5$=$s^2$+$i^3$}   \\
-{\tt 6: i=2} &       {\tt 6: $i^6$=2}  \\
-{\tt 7: i<2} &      {\tt 7: $i^6$<3}  \\
-{\tt 8: {\bf y} = s;} &  {\tt 8: $W^{8}_y$ = $s^5$;}  \\
-{\tt 9: {\bf y} > 2}  &    {\tt 9: $R^{9}_y$ > 2} \\
-{\tt 10: print({\bf x});} &  {\tt 10: print($R^{10}_x$);}  \\
-\end{tabular}
-\caption{Trace}
-\label{fig:t4running2}
-\end{figure}
+%\begin{figure}
+%\centering
+%\begin{tabular}{l|l}
+%\hline
+%\multicolumn{1}{c}{$Trace$} & \multicolumn{1}{c}{$SSA \ form \ of\   Trace$} \\
+%\hline
+%{\tt 0: {\bf x}=0} &  {\tt 0: $W^0_x$=0}    \\
+%{\tt 1: {\bf y}=0} &   {\tt 1: $W^1_y$=0}   \\
+%{\tt 2: s=0} &  {\tt 2: $s^2$=0}   \\
+%{\tt 3: i=1} &     {\tt 3: $i^3$=1}   \\
+%{\tt 4: i<2} &    {\tt 4: $i^3$<3} \\
+%{\tt 5: s=s+i} & {\tt 5: $s^5$=$s^2$+$i^3$}   \\
+%{\tt 6: i=2} &       {\tt 6: $i^6$=2}  \\
+%{\tt 7: i<2} &      {\tt 7: $i^6$<3}  \\
+%{\tt 8: {\bf y} = s;} &  {\tt 8: $W^{8}_y$ = $s^5$;}  \\
+%{\tt 9: {\bf y} > 2}  &    {\tt 9: $R^{9}_y$ > 2} \\
+%{\tt 10: print({\bf x});} &  {\tt 10: print($R^{10}_x$);}  \\
+%\end{tabular}
+%\caption{Trace}
+%\label{fig:t4running2}
+%\end{figure}
 
-Figure~\ref{fig:t4running2} exemplifies a trace and its SSA form, where the trace is generated from the program in Figure~\ref{fig:running2}.
-Each label in  Figure~\ref{fig:t4running2} on the left hand denotes the id of the event.
+%Figure~\ref{fig:t4running2} exemplifies a trace and its SSA form, where the trace is generated from the program in Figure~\ref{fig:running2}.
+%Each label in  Figure~\ref{fig:t4running2} on the left hand denotes the id of the event.
 As seen, event $e_5$ uses the variable $s^2$ defined at event $e_2$ and defines the variable $s^5$, which is used at $e_8$.
@@
 \subsection{Constraint System}~\label{sec:constraints}
-Based on the SSA form of the trace, we build the constraints to compute a new trace with the new schedule and new mapping of variables to values.
-To model the schedule, we introduce the order variable $O_{id}$ for each event $e_{id}$.
-Given two events $e_i, e_j$ from two different threads, $O_{e_i}<O_{e_j}$ means that $e_i$ is scheduled before $e_j$.
+Based on the SSA form of the original trace $\tau$, we build the constraints to compute a new trace $\tau'$ with the new schedule $O'$ and new value mapping $\theta'$.
+Consider the example in Figure~\ref{fig:running2}, in the original trace $\tau$, the event $e_8$ reads the value of 2 from $e_3$, therefore, $e_{11}$ must happen after $e_2$.
+Existing approaches enforce the same value dependence between $e_3$ and $e_8$, leading to the miss of the race between $e_2$ and $e_{11}$.
+Our analysis can derive a new trace $\tau'$ to find the race by relaxing the dependence.
+In $\tau'$, $e_8$ reads from a different event $e_1$, producing the new value mapping, $\theta'(y^8)=3$.
+Accordingly, at the event $e_9$, the value mapping for $x^9$ is also updated as 13 ($\theta'(y^8)$=3, $\theta'(z^7)=10$), enabling the true branch at line 10.
+Finally, the events $e_2$ and $e_{11}$ are scheduled to run concurrently.
+The relaxation needs to respect a set of constraints, which we explain in the following.
 We omit the synchronization constraints intentionally, as they are well explained in all existing predictive analysis techniques~\cite{yannis, pldi14}.
@@
 $$
-O_{i} = O_{j}
+O'(e_i) = O'(e_{j})
 $$
+In Figure~\ref{fig:running2}, the race condition for the race pair, ($e_2$, $e_{11}$), is $O'(e_2)=O'(e_{11})$
 
@@
 The predictive analysis reschedules the events in the prefix of $e_i$, i.e., prior to $e_i$, so that $e_i$ and $e_j$ can run concurrently.
+We refer to the prefix as $prefix(\tau, e_i)$.
+Throughout this section, we only consider the events in the prefix.
+
 The predictive analysis by design requires that each thread should follow the same event sequence prior to $e$ as in the original run.
@@
 
-More formally,  we require that 
+Without loss of generality, we assume the branch event $e_k$ is in the form of $if(x<y)$. Then  we require that 
 $$
-\bigwedge_{e_k \in \tau\downarrow_{\leq i} \wedge  type^{e_k}=branch.\ }  inst^{e_k} 	\equiv \lsyn inst \rsyn^{e_k}
+\bigwedge_{e_k \in prefix(\tau,e_i) \wedge  inst^{e_k}=if(x<y).\ }  \theta'(x)<\theta'(y) 	\equiv \theta(x)<\theta(y)
 $$ 
-where  the branch event $e_k$ is used as a boolean expression.
-For example, given the branch {\tt if(x<y)\dots}, the constraint is in the form of $(x<y)\equiv {\bf true}$ assuming the branch evaluates to {\bf true} in the original trace.
-The constraint specifies that the boolean expression in the predicted trace should be evaluated to the same boolean value as in the original trace.
+The constraint specifies that the branch condition should be evaluated as the same boolean value in the predicted run $\tau'$ as in the original trace $\tau$.
  Importantly, Unlike existing analyses \cite{yannis,pldi14}, we do not pose the requirement that the values flowing into branching statements remain the same, but adopt the relaxed requirement that the evaluation of branching expressions remains the same.
-For example, suppose  the branch event $x<y$ takes the value $1<2$ in the original trace.
-Existing analyses require the same values for the variables $x$ and $y$ in the predicted run, while we allow  the variables flowing into the branch event to assume other values, such as $3<4$, as long as the branch expression retains the same truth value.
+Consider the example in Figure~\ref{fig:running2}, given the branch if($x^9$>11) at $e_10$, the original mapping is $\theta(x^9)=12$.
+Existing analyses require the same value mapping for $x^9$  in the predicted run, while we allow  a different value mapping, e.g., $\theta'(x^9)=13$, which retains the same truth value for the branch expression.
 \item {\bf Intra-thread Order Constraints\ } The events in the sequence should follow the same order as in the original trace. More formally, we require that
@@
 \begin{array}{rl}
-\forall e_m, e_n \in \tau\downarrow_{\leq i}, s.t., & t^{e_m} = t^{e_n}. \\
-& m < n\  \Rightarrow O_{m} < O_{n}
+\forall e_m, e_n \in prefix(\tau,e_i), s.t., & t^{e_m} = t^{e_n}. \\
+ O(e_m) < O(e_n)\  & \Rightarrow O'(e_{m}) < O'(e_{n}) 
 \end{array}
@@
 This constraint specifies that two events from the same thread should follow the same order as reflected by the ids of the events.
-Note that $m$ in $e_m$ denotes its id.
-Again, we only consider the events in the prefix of $e_i$.
-\item {\bf Intra-thread Data Flow Constraints\ } The mapping of local variables should not contradict the data flow semantics of each instruction.   More formally, we require that
+\item {\bf Intra-thread Value Constraints\ } The value mapping of the variables should not contradict the value constraints imposed by each instruction.
+Without loss of generality, suppose the instruction is a local assignment in the form of $x=y+z$, then we require that
 $$
-\bigwedge_{e_k \in \tau\downarrow_{\leq i} \wedge type(e_k)=assign|heapr|heapw}\
-	inst^{e_k}
+\bigwedge_{e_k \in prefix(\tau,e_i) \wedge inst^{e_k}=x=y+z}\
+	\theta'(x)=\theta'(y)+\theta'(z)
 $$
-where $inst^{e_k}$, such as $x=y+z$ or $R^{id}_{x}=y$, captures the constraint over the values of the variables imposed by the instruction.
-For example, $x=y+z$ is not satisfied if the solver maps the three variables to $3, 4, 5$ respectively.
-Again we consider only the events in the prefix of  $e_i$.
-Specially, for the instruction involving object creation, $x=new (...)$, we encode the object as a unique integer that denotes its heap address, calculated by $System.identityHashcode(x)$ at runtime.
+Remember that every variable may obtain a different value in the new trace $\tau'$ because it may read from some shared variable affected by the rescheduling.
+The constraint specifies that, the variables should be re-assigned to some values that are consistent with the instruction.
+Consider the example in Figure~\ref{fig:running2}, given the event $x^9=y^8+z^7$ at line 9, if the mappings $\theta'(y^8)$ and $\theta'(z^7)$ change to 3 and 10 respectively, the mapping $\theta'(x^9)$ should change to 13 consistently.
+In the above, we use the local assignment event for demonstration, in general, the intra-thread value constraints should also be specified for the {\sf  heapr/heapw} events.
 \end{itemize}
@@
 %Rescheduling through Relaxation of Flow Dependencies
-{\bf Inter-thread Data Flow Constraints for Relaxation\ } We now move to the first novel feature of \tool, which is its ability to explore execution schedules that depart from the value flow exhibited in the original trace.
+{\bf Inter-thread Value Constraints for Relaxation\ } We now move to the first novel feature of \tool, which is its ability to explore execution schedules that depart from the value flow exhibited in the original trace.
 More precisely, \tool\ is able to relax value flow dependencies in the original trace:  a read access may read a
-different value from other write events, as long as the read value enforces feasibility.
+different value from other write events, as long as the read value enforces feasibility of the following execution.
  This is strictly beyond the coverage potential of existing predictive analyses, which restrict trace transformations to ones where any read access must read the same  value (often from the same write event) as in the original trace.
@@
 To ensure feasibility under relaxation of flow dependencies, we need to secure the flow between the read/write with the execution schedule. 
-For example, in Figure \ref{fig:running} for the read access to ${\tt y}$ at $e_4$ to obtain the value assigned to ${\tt y}$ at $e_1$, we need a schedule where 
-$e_4$ happens after $e_1$ and other writes such as $e_3$ do not interleave them. ($e_3$ can only happen after $e_4$ in this case.)
+For example, in Figure \ref{fig:running2}, for the read at $e_8$ to read from the write t $e_1$, we need a schedule where 
+$e_8$ happens after $e_1$ and other writes such as $e_3$ do not interleave them.
 
@@
 
-%TODO l and ell
 In general, the constraint formula, given read $R^{m}_{\ell}$ of location $\ell$ at the event $e_m$ with set ${\cal W}$ of matching write events (i.e., events including write access to $\ell$), takes the following form:
@@
 \begin{array}{rll}
-\bigvee_{e_n \in {\cal W}} &  & (R^{m}_{\ell} = W^{n}_{\ell}) \\
-&		\bigwedge 	&  O_n < O_m \\
-&		\bigwedge_{e_p \in {\cal W} \setminus \{ e_n \}} & (O_p < O_n \vee O_m < O_p)
+\bigvee_{e_n \in {\cal W}} &  & (\theta'(R^{m}_{\ell}) = \theta'(W^{n}_{\ell})) \\
+&		\bigwedge 	&  O'(e_n) < O'(e_m) \\
+&		\bigwedge_{e_p \in {\cal W} \setminus \{ e_n \}} & (O'(e_p) < O'(e_n) \vee O'(e_m) < O'(e_p))
 \end{array}
 $$
-This disjunctive formula iterates over all matching write events, and demands for each that (i) it occurs prior to the read event ($O_n < O_m$) and (ii) all other write events either occur before ($O_p < O_n$) it or after the read event
-($O_m < O_p$).
+This disjunctive formula iterates over all matching write events, and demands for each that (i) it occurs prior to the read event $O'(e_n) < O'(e_m)$ and (ii) all other write events either occur before $O'(e_p) < O'(e_n)$ it or after the read event
+$O'(e_m) < O'(e_p)$.
 
@@
 
-To address this challenge, we enhance the constraint system with the requirement that heap objects that are dereferenced in field access statements before the racy events retain their original address in the predicted trace.
+{\bf Invariant Base Object Constraints\ } To address this challenge, we enhance the constraint system with the requirement that heap objects that are dereferenced in field access statements before the racy events retain their original address in the predicted trace.
 This achieves two guarantees: First, matching heap access events in the original trace are guaranteed to also match in the predicted trace.
@@
 $$
-\bigwedge_{ e_k \ has\  x.f \wedge e_i \in \tau\downarrow_{\leq i} }\
-	x \equiv  \lsyn x \rsyn^{e_k}
+\bigwedge_{ e_k \ has\  x.f \wedge e_i \in prefix(\tau,e_i) }\
+	\theta'(x) \equiv  \theta(x)
 $$
@@
 %(resp. $\sigma(t',\tau'')$) is the state arising at trace $t$ (resp. $t'$) immediately before event $\tau''$.
+
 This constraint fixes that all heap dereferences  prior to the racy event retain their original base object as in the original trace $\tau$. 
-In general, we need to specify such heap constraints when we hardcode the local heap accesses as local variables (Section~\ref{sec:ssa}) to ensure that the hardcoded data flow remains valid  after the relaxation.
-However,  most local heap accesses in the form of $x.f$ never read from any shared locations according to the static analysis, and therefore, cannot be affected by the relaxation.
-As an optimization, we do not specify the heap constraints for such local heap accesses.
+The array accesses are handled similarly to the field accesses, but we need to additionally require the index variable to be the same as in the original trace.
+In addition, for local heap accesses, in case that the base variable reads a local object from a shared reference variable, the base variable may change due to the rescheduling.
+We also fix the base variable as a constant, in order to ensure the validity of the SSA encoding of the local heap accesses and preserve the def/use chains among them.
+The constraint is specified during the SSA encoding.
 

### 2014-09-05 introduction.tex
intro by xiangyu

--- 2014-08-17 introduction.tex
@@
+%Limitations of static analysis: scalability when it is path sensitive, difficulty in modeling heap, cannot reason about concrete values, have to use
+%abstraction, so that may not be sound, dynamic analysis is precise, easily scale to long and complex runs, but it only reasons about problems that
+%did occur.
+
+%Predictive analysis is so good
+
+%But it restricts too much, only relax schedule space and a little bit of the dependence space.
+
+%We observe many dynamic aspects can be relaxed without affecting soundness.  as a result, instead of reasoning one
+%execution, we can reason the neighborhood of the execution, in order words, we leverage dynamic analysis to provide
+%the information that is difficult to analyze statically such as aliasing and a long execution path that covers the
+%intended workflow, then we relax the predictive analysis to analyze neighborhood.
+
 \section{Introduction}\label{Se:introduction}
+Program analyses can be largely classified to static and dynamic analyses. Static analyses abstract programs statically.
+They are usually complete and do not require concrete program inputs to drive the analyses. However, they have difficulty in
+scaling to large and complex programs when the analyses are path-sensitive, due to the sheer number of paths and the 
+complexity of the paths. Path-insensitive analyses often do not have sufficient precision on the other hand. 
+Static analyses also have difficulty in modeling complex aliasing behavior. In contrast, dynamic analysese focus 
+on analyzing a few executions, usually just one. They are sound and easily scale to long and complex executions. But they 
+can only reason about properties that manifest themselves in the execution(s).
 
-In a multithreaded application, a data race occurs when two concurrent threads access the same memory location, such that (i) at least one of the accesses is write access, and (ii) no explicit mechanism is enforced to prevent the threads from simultaneous access to the location \cite{eraser}.
-Race detection is important not only because races often reveal bugs in the implementation, but also as the basis of other techniques and analyses like atomicity checking \cite{learning,avio,ordervio}, data-flow analysis \cite{dataflow} and record/replay \cite{leap,replay2012}.
+Recently, a novel kind of analysis called predictive analysis~\cite{...} was proposed. It has the capabilities of 
+achieving soundness as dynamic analyses and reasoning about properties that did not occur during execution 
+just like in static analysis. The basic idea is to leverage dynamic analysis to provide the information that 
+is difficult to acquire through static analysis, such as a complex execution path that covers important functionalities 
+and aliasing information. Note that during execution, we know precisely which write access reaches any given read 
+access. While such information is encoded in a trace, predictive analysis reaons about mutation of the trace to 
+expose problems or study properties that may not manifest during execution. It was used in sound data race detection. 
+In particular, it collects a trace of threaded execution that contains rich runtime information such as concrete values, 
+execution path, and memory accsses. It then leverages constraint solving to reason about if races can occur with
+a different thread schedule, {\em while enforcing the critical runtime information (e.g. values, thread
+local paths, and memory accesses) remains the same during schedule pertubation}. The rationale is that 
+it becomes as difficult as static analysis if we allow these critical runtime information change. 
+Preditive analysis has been shown to be very effective. It can detect real data races from complex programs.
+And more importantly, it guarantees the results are sound (i.e. no false positives).
 
-In this paper, we focus on \emph{sound} race detection, whereby reported races are guaranteed to be real.
-This is the key requirement for adoption of the tool by developers \cite{XXX}.
-Ensuring soundness is a difficult challenge, which mandates dynamic forms of analysis.
-Indeed, extensive research has been carried out on dynamic race detection.
-The general goal has been to derive constraints from a given execution trace on event reordering, and check for the remaining reorderings whether they disclose data races.
-The key question then becomes about constraint extraction.
-
-%Fasttrack PLDI paper claims happens-before analysis never reports false positives. Check.
-\paragraph{Existing Approaches} Initial attempts to address this challenge focused on built-in synchronization primitives.
-These include locks as well as the wait/notify and start/join scheduling controls.
-Notable among these efforts is \emph{lockset} analysis, which considers only locks.
-Because the derived constraints are partial, permitting certain infeasible event reorderings, lockset analysis cannot guarantee soundness.
-
-A different tradeoff is struck by the \emph{happens-before (HB)} approach.
-In this style of analysis, all synchronization primitives are accounted for, though reordering constraints are conservative.
-As an example, HB inhibits reordering of two synchronized blocks governed by the same lock.
-Recently there have been successful attempts to relax HB constraints.
-Among these are hybrid analysis, which permits both orderings of lock-synchronized blocks, and the \emph{Universal Causal Graph (UCG)}~\cite{ucg} representation, which also enables both orderings but only if these are consistent with wait/notify- and start/join-induced constraints.
-
-Unfortunately, even full consideration of synchronization constructs is insufficient.
-Branching decisions must also be respected to ensure the feasibility of a reported race.
-As an example, we refer to the trace in Figure \ref{fig:running}.
-(For now, ignore the statements in red.) This trace is free of any explicit synchronization statements, and so a race between the assignment to variable {\tt x} at line {\tt 2} and the use of {\tt x} at line {\tt 5} may appear possible.
-This race is, however, spurious, since the use of {\tt x} at line {\tt 5} is governed by positive evaluation of the test at line {\tt 4}, which is conditioned on the assignment at line {\tt 3}.
-Hence, line {\tt 2} necessarily preceds line {\tt 5}.
+Despite its effectiveness, we observe that existing predictive analyes are too restrictive. They 
+require too much runtime information to remain unchanged during analysis such that coverage is unnecessarily
+limited.
 
@@
 
-As an alternative, driven by the requirement for soundness, 
-Smagardakis et al.
-have recently proposed \emph{predictive analysis}.
-In this approach, both synchronization constraints and inter-thread dependencies are preserved, where inter-thread dependencies are respected by
-only allowing reorderings that leave the dependence structure exhibited by the original trace in tact.
-This ensures that the values of shared memory locations remain the same, which secures the soundness argument.
-As a soundness-preserving relaxation, Huang et al.
-\cite{XXX} permit reorderings as long as control dependencies are respected.
+Consider the example in Figure \ref{fig:running}, which shows an execution trace of two threads. The statements in red 
+denote an unexplored branch outside the trace. Existing predictive analyses are unable to detect the race between 
+lines {\tt 2} and {\tt 5} because they require that during trace permutations (by different schedules), a variable must not have
+a different value in a permuted trace\footnote{They allow the variable involved in a race to have different values.}.
+As such, they cannot execute line {\tt 4} before line {\tt 3}, which is needed to expose the race between lines {\tt 2}
+and {\tt 5}. 
+%the dependence between lines {\tt 3} and {\tt 4}, which is a barrier to the needed reorderings. 
+A second race between lines {\tt 2} and {\tt 7} is also missed as the mutated trace must have the same thread local
+path as the original trace.
 
-While being sound and able to find real races, existing predictive analyses are conservative in two respects:
-\begin{enumerate}
-	\item Constraints imposed by branch conditions are abided by preserving the exact values flowing into the condition as in the original trace.
-	\item A second limitation, implied by the first, is that traces that diverge from the original trace in their branching history are strictly out of scope. 
-\end{enumerate}
- Due to these restrictions, enforced as a conservative means of ensuring soundness, existing predictive analysis suffer from limited coverage.
+However, we observe that such restrictions are unnecessary. For example, we can allow line {\tt 4} to receive its
+value from line {\tt 1} as it has the same effect on thread local execution (i.e. the same branch is taken). 
+However, the relaxation would allow us to execute lines {\tt 4} and {\tt 5} before line {\tt 2}, exposing 
+the first missing race. We can also allow the trace to take the else branch such that the second race is also exposed. 
+Note that even though the else branch was not executed originally, its effect can be precisely modeled as the value 
+of $x$ is known from the original trace. As such, we can still perform sound analysis.
 
- \paragraph{Our Approach} The departure point of our approach, yielding strictly better coverage, is to consider the explicit values of shared memory locations instead of dependencies between statements.
-The gain in coverage is twofold:
- \begin{enumerate}
- 	\item Reorderings that violate the original dependence structure but preserve the branching history become possible.
-The values read by a branching statement may change so long as the branch condition evaluates to its original truth value.
- 	\item In addition, in certain cases we can guarantee soundness while directing execution toward branches that diverge from the original trace.
-This mandates that the effects of an unexplored branch can be modeled precisely, which holds frequently in practice because the original and new branchees often access the same set of shared memory locations.
- \end{enumerate}
+%that is missed, as it involves the statements in red, is between lines {\tt 2} and {\tt 7}.
+Therefore in this paper, we propose \sysname, a relaxed predictive analysis. We identify the critical runtime
+informantion that needs to be preserved during trace perturbation and relax the remaining information. 
+In particular, we preserve all the addresses de-referenced (e.g. heap and array accesses) in the original execution,
+and some of the branch outcomes such that part of the thread local paths stay intact. The criterion is that
+we forbid \sysname to explore an unexecuted branch if the effect of the branch cannot be precisely modeled
+(e.g. an array element is read while it was not defined in the original trace). The essence of our technique 
+is to {\em explore the neighborhood of the original execution as much as possible}. Due to the substantially
+enlarged search space, our results show that a race detector based on our relaxed predictive analysis 
+can detect XXX times more races than existing predictive analysis, while guaranteeing soundness.
 
- Returning to the example in Figure \ref{fig:running}, this time also with reference to the statements in red, we demonstrate how value-based reasoning improves coverage.
-First, unlike predictive analysis, which is unable to detect the race between lines {\tt 2} and {\tt 5} because of the dependence between lines {\tt 3} and {\tt 4}, value-based reasoning reveals that trace
- [{\tt 1}, {\tt 4}, {\tt 5}, {\tt 2}, {\tt 3}] is feasible, since the first assignment, {\tt y = 3}, satisfies the condition {\tt y > z}.
-Hence the race is discovered.
-A second race, between lines {\tt 2} and {\tt 7}, is detected by negating the condition.
-This is possible because the effects of the {\tt else} branch can be modeled precisely.
- Negation leads to an execution starting at line {\tt 4} (where {\tt y $\equiv$ 0}). The race between lines {\tt 2} and {\tt 7} then becomes visible.
-
- Concretely, we achieve value-level (rather than dependence-level) granularity computationally via a unique encoding of the input execution trace as a constraint system.
-The constraints are then processed by a satisfiability checker.
-This gives us the flexibility to explore nontrivial constraints, such as
- forcing a context switch after the first assignment to {\tt y} or negating the condition in Figure \ref{fig:running}.
-Importantly, these manipulations are beyond what dependence-based reasoning can achieve, since this view of the trace is too conservative.
-
-\paragraph{Contributions} This paper makes the following principal contributions:
+Our contrituions are summarized as follows.
+\begin{itemize}
+\item We propose to relax existing preditive analysis and identify a bound of relaxation (i.e. what can
+be relaxed).
+\item We develop a constraint encoding scheme that uniformally encodes the strict information (i.e. the
+part must be preserved) and the relaxed information. 
+\item We develop a propotype \sysname. Our results show XXX.
+\end{itemize}

### 2014-09-05 overview.tex
intro+relaxI

@@
 The key property of  $id$ is {\em uniqueness}, i.e., any two events in the trace own different ids.
-The other important property of $id$ is {\em monotonicity}, i.e.,  the events from the same thread should own the strictly increasing ids.
-Unless otherwise specified, we use  the index of an event in the trace as its id, which satisfies the above properties.
+Unless otherwise specified, we use  the index of an event in the trace as its id, which satisfies the above property.
 Throughout this paper,  we denote an event as $e_{id}$ with the id as the subscript.

### 2014-09-05 relax2.tex
edits
intro+relaxI

--- 2014-09-02 constraints.tex
@@
-\section{Exploration of Unexecuted Branches}
+\section{Exploration of Unexecuted Branches}~\label{sec:relax2}
+We now switch to the second feature of \tool, which is its ability to reason about neighboring branches that are not executed in the original trace $\tau$.
+At the high level, we replace the constraints that model the executed branch to the constraints that conservatively model the possible executions in both branches.
+We leverage the static analysis to  realize this relaxation. We explain the details below.
 
-We now switch to the second feature of \tool, which is its ability to reason about unexplored branches.
+\begin{figure}
+\centering
+\begin{tabular}{ll|l}
+\multicolumn{3}{c}{{\tt {\bf P} = new();//$o1$}} \\
+\multicolumn{1}{c}{$T_1$} & \multicolumn{1}{c}{$T_2$}  &  \multicolumn{1}{c}{$Trace$}\\
+{\tt 1: {\bf P}.f=100; } &  & {\tt $W^{1}_{o1.f}$=100;}\\
+{\tt 2: {\bf Q}=101; } &  & {\tt $W^{2}_{Q}$=101;}\\
+{\tt 3: {\bf Q}=11; } & & {\tt $W^{3}_{Q}$=11;} \\
+{\tt 4: {\bf Q}=1; } & & {\tt $W^{4}_{Q}$=1;} \\
+{\tt 5: {\bf S}=1; } & & {\tt $W^{5}_{S}$=1;} \\
+& {\tt 6: x={\bf Q};} & {\tt $x^6$=$R^6_{Q}$;}\\
+& {\tt 7: if(x<10)} & {\tt if($x^6$<10)}\\
+& {\tt 8: \ {\bf S}=0;}  & {\tt $W^8_{S}$=0;}\\
+& {\color{Gray}{\tt 9:  else}} &  {\color{Gray}{\tt if($x^6\geq$10)}}\\
+& {\color{Gray}{\tt 10: \ z={\bf P}.f;}} & {\color{Gray}{\tt $z^{10}$=$R^{10}_{o1.f}$;}}\\
+& {\color{Gray}{\tt 11: \ if(x<z)}} & {\color{Gray}{\tt if($x^6<z^{10}$)}}\\
+& {\color{Gray}{\tt 12: \ \ {\bf S}=0;}}& {\color{Gray}{\tt $W^{12}_{S}$=0;}}\\ 
+& {\color{Gray}{\tt 13: \ else }}  & {\color{Gray}{\tt if($x^6\geq z^{10}$)}}\\
+& {\color{Gray}{\tt 14: \ \ w=new(); }}  & {\color{Gray}{\tt $w^{14}$=$o2$;}}\\
+& {\color{Gray}{\tt   \ \ \ \ \ \  //o2 }}  & {}\\
+& {\color{Gray}{\tt 15: \ \ w.f=0; }}  & {\color{Gray}{\tt $l^{15}_{o2.f}$=0;}}\\
+& {\color{Gray}{\tt 16: \ \ {\bf S}=w.f; }}  & {\color{Gray}{\tt $W^{16}_{S}$=$l^{15}_{o2.f}$;}}\\
+\end{tabular}
+\caption{Running Example (shared variables are in bold font). }
+\label{fig:path}
+\end{figure}
 
-\paragraph{Relaxation of Branching Decisions}
+Consider the example in Figure~\ref{fig:path}, only the if-branch of $T_2$ is executed in the original trace $\tau$.
+Following the rules in Section~\ref{sec:relax1},   we derive a set of constraints to find the new trace $\tau'$ to witness the race pair ($e_5$, $e_8$). We list those specific to if-branch below.
 
-Beyond relaxing flow dependencies, \tool\ is also capable of exploring code branches that were not executed in the original trace.
+\begin{figure}
+	\begin{center}
+$$
+	\begin{array}{rcl}
+		&  O'_{e_5}=O'_{e_8} & \textbf{(race condition)} \\
+			& \theta'(x^6)<10 & \textbf{(Intra-thread control flow constraint)} \\
+	&  O'_{e_7}<O'_{e_8} & \textbf{(Intra-thread order constraint)} \\
+		& \theta'(W^8_S)=0 & \textbf{(Intra-thread value constraint)} \\
+	\end{array} 
+$$
+\end{center}
+caption{\label{fig:pathencode}
+\end{figure}
+
+We  replace these constraints to another set of constraints that account for both if-branch and else-branch from the event $e_7$.
+We refer to the branch event $e_7$ as the root branch event, denoted as $e_{root}$.
+
+First, we conduct the DFS traversal of the branch structure and produce an event, $<t, id, inst>$, to represent the potential execution of each instruction $inst$.
+Here, $t$ is the same as that of the root branch event as the exploration is conducted thread-locally, $id$ is a unique value different from existing ids.
+Specifically,
+ we maintain an increasing counter to assign the id following the traversal order. For this example, we simply use the line number as the event id.
+
+similar to Section~\ref{sec:relax1}, we need the SSA form of
+
+%%TODO | and projection, abuse the terms?
+%As shown in the following, suppose the symbolic execution  $symEngine(\tau, e)$ returns the single-thread traces starting from the branch $e$, which now takes a different decision.
+The resultant trace is computed by first removing the events thread-locally after $e$ (inclusively) and then appending one of the single-thread traces that represents a path in the unexplored branch of $e$.
+The negation $neg(e)$ of $e$, which is the same as $e$ except the evaluation result of the boolean expression differs, is also appended.
+%
+%
+%
+%\begin{algorithmic}[3]
+%\For {$e: \tau$}
+%  \If {$type(e)=branch$}
+%    \State $S=symEngine(\tau, e)$
+%	\For {$\tau': S$}
+%	\State $\tau_r=\tau-\tau\downarrow_{t^e \wedge \geq e}$
+%	\State $\tau_r=\tau_r.replace(e, neg(e) + \tau')$
+%	\EndFor
+% \EndIf
+%\EndFor
+%\end{algorithmic}
+%
+%
+%{\bf Symbolic Execution\ } Our symbolic execution is realized on top of the dataflow analysis, which generates events for each instruction within the unexplored branch.
+%The symbolic execution has known limitations in reasoning about the schedules and loops (and recursion).
+Therefore, our analysis assumes all the events in the unexplored branch happen atomically at the branch event $e$ without interleavings from other threads, as illustrated also in the above algorithm (line 6).
+In this way, our analysis adopts the sequential reasoning.
+\tool\ will reschedule the events based on the relaxation.
+As for loop (or recursion), the analysis terminates the current data flow immediately after one iteration, i.e., we expand the loop for only once.
+Otherwise, the analysis terminates the data flow normally if the flow goes out of the scope of the unexplored branch.
+%
+%The analysis starts at the event $e$ (exclusively), with the state $\sigma$.
+Initially, the state includes the runtime heap value for the base objects that are already resolved.
+During the symbolic execution, the state is updated to maintain the heap value and also the boolean expression values for branches.
+The values for other variables are not used indeed and therefore treated as symbolic.
+The update of the state is realized through the   standard kill/gen rules for three basic heap instructions.
+In the object creation, which we separate from the local assignment, we assign a unique integer to represent the newly created object and store it into the state $\sigma$.
+For field access, $\sigma[x.f]$ actually denotes $\sigma[o.f]$, where  $o$ is the base object resolved in the state.
+%These kill/gen rules take effect only when $x.f$  is not of the primitive type.
+%
+%The analysis also generates the events, where the heap values are stored in the map of the event for later use. For the branch, we store its boolean evaluation result.
+%As the boolean result is available only after the branch takes the decision, we delay the generation of the branch event until the first event after the branch. 
+%
+%
+%
+%
+%\begin{table*}
+%\centering
+%%\begin{small}
+%\begin{tabular}{l|c|c}
+%\multicolumn{1}{c|}{Operation} & {Kill} & {Gen} \\
+%\hline
+%{\tt $y$=$new (...)$} 			&  $\sigma[y] \mapsto \star$		&  $\sigma[y] \mapsto newI$, $<t, cnt, y=newI, \bot>$\\
+%{\tt $y$=$x.f$} 			&   $\sigma[y]\mapsto \star$			&  $\sigma[y] \mapsto \sigma[x.f]$, $<t, cnt, y=x.f, \lsyn x \rsyn=\sigma[x]>$  \\
+%{\tt $x.f=y$} & $\sigma[x.f] \mapsto \star$		&  $\sigma[x.f] \mapsto \sigma[y]$, $<t, cnt, x.f=y, \lsyn x \rsyn=\sigma[x]>$ \\
+%{\tt $z=x\  bop\  y$} & 		&  $<t, cnt, z=x\  bop\  y, \bot>$ \\
+%{\tt $if(x<y)$} & 	  &  $<t, cnt, if(x<y), \lsyn x<y \rsyn=bool>$\\
+%\end{tabular}
+%\caption{\label{table:killgen}Dataflow analysis}
+%%\end{small}
+%\end{table*}
+
+%Beyond relaxing flow dependencies, \tool\ is also capable of exploring code branches that were not executed in the original trace.
 This is achieved via a symbolic representation of the input trace.
 Given branching event $e_b$,
-\begin{enumerate}
-	\item model the branching condition symbolically to execute along the negation of the original branch;
-	\item apply depth-first search (DFS) to uncover all execution suffixes under the unexplored branch (which may itself contain branching statements); and
-	\item for each suffix $t^s$,
-	\begin{enumerate}
-		\item truncate the original trace at $e_b$ yielding prefix $t_s$; and
-		\item concatenate $t_s$ with the negation of $e_b$ followed by $t^s$.
-	\end{enumerate}
-\end{enumerate}
-Constraint solving is applied to each of the resulting traces analogously to the original trace.
+%\begin{enumerate}
+%	\item model the branching condition symbolically to execute along the negation of the original branch;
+%	\item apply depth-first search (DFS) to uncover all execution suffixes under the unexplored branch (which may itself contain branching statements); and
+%	\item for each suffix $t^s$,
+%	\begin{enumerate}
+%		\item truncate the original trace at $e_b$ yielding prefix $t_s$; and
+%		\item concatenate $t_s$ with the negation of $e_b$ followed by $t^s$.
+%	\end{enumerate}
+%\end{enumerate}
+%Constraint solving is applied to each of the resulting traces analogously to the original trace.
 
-We emphasize that exploration of new execution paths is subject to all the known limitations of symbolic execution, including in particular loop structures and object allocation.
+%We emphasize that exploration of new execution paths is subject to all the known limitations of symbolic execution, including in particular loop structures and object allocation.
 \tool\ currently fails if (i) the unexplored branch containts loops or (ii) there are object references that cannot be fully resolved at the branching point.

### 2014-09-05 eval.tex

@@
 
-{\bf Evaluation Method} We conduct our experiment on a large set of applications, which are also used to evaluate {\sf RV}.
-Specifically, the set includes large applications such as {\sf Jigsaw}, {\sf Xalan}, {\sf Lusearch}.
-To handle large applications that make use of reflection, we applied {\sf Tamiflex}~\cite{}, which replaces reflective calls with the concrete method invocations recorded in the observed run for the purpose of effective static analysis.
+{\bf Evaluation Method} We conducted our experiments on a set of 16 applications, which were also used to evaluate {\sf RV}.
+Specifically, the set includes large applications such as {\sf Jigsaw}, {\sf Xalan} and {\sf Lusearch}.
+To handle large applications that make use of reflection, we applied the {\sf Tamiflex} tool~\cite{}, which replaces reflective calls with the concrete method invocations recorded in the observed run for the purpose of effective static analysis.
 We omit the benchmark {\sf eclipse} because the current version of {\sf Tamiflex}  leads to abnormal execution after the instrumentation, which throws an exception when the main service is started.
@@
 16GB of RAM and 6M caches. The workstation runs version 12.04 of the Ubuntu Linux distribution, and has the Sun 
-64-Bit 1.6.0\_26 Java virtual machine (JVM) installed~\footnote{Note that Sun JDK 1.7 does not support the transformation of dacapo applications with {\sf tamiflex}.}.
+64-Bit 1.6.0\_26 Java virtual machine (JVM) installed.\footnote{We note that Sun JDK 1.7 does not support the transformation of dacapo applications with {\sf tamiflex}.}
 
@@
 The Trace section includes the number of threads ($Th$), the number of shared reads ($Reads$) and shared writes ($Writes$), the subset of shared reads that read the base object references  ($Base$), where the  base object references are references used as the base/target in the following field reference or the method invocation,  the number of branches ($Br$), the synchronization events ($Sync$) and the total number of events ($Total$), which includes local accesses in addition to the aforementioned events.
-Specifically, for the branch, we report it in the form of $A/B$, where $A$ refers to the number of branches used by our analysis and $B$ refers to the number of branches used by RV.
+Specifically, branch events are reported in the form $A/B$, where $A$ refers to the number of branches used by our analysis and $B$ refers to the number of branches used by RV.
 To ensure that the predicted run sees the same base object at each shared read/write, {\sf RV} inserts the artificial branch immediately in front of each shared access (and array accesses).
@@
 We make interesting observations about the Trace section.
-The non-local events, i.e., all the events listed in Table~\ref{tab:main}, occupy around 30\% of the total trace in the first 7 benchmarks, but occupy less than 1\% of the trace in the rest benchmarks, which have relatively more complex logic.
-The reads of the shared base objects occupy a small portion ($1/3$-$1/10$) of the total shared reads.
-The rest shared reads read only the primitive values or
-the references that are not the base references, e.g., the references involved only in the nullness check.
+The non-local events, i.e., all the events listed in Table~\ref{tab:main}, occupy around 30\% of the total trace in the first 7 benchmarks, but occupy less than 1\% of the trace in the rest of the benchmarks, which have relatively more complex logic.
+Reads of the shared base objects occupy a small portion ($\frac{1}{10}$-$\frac{1}{3}$) of the total shared reads.
+The rest of the shared reads read only primitive values or
+references not involved in field accesses (e.g., references involved in nullness checking).
 Besides, our analysis involves significantly less branch events as compared to the {\sf RV} approach.
-The difference plays an important role in the detection, which we will explain shortly.
+This difference plays an important role in the detection, which we will explain shortly.
 
-The $Races$ section shows the number of races detected by {\sf Recipe-s}, i.e., {\sf Recipe} without exploring un-executed paths, {\sf Recipe}, i.e., the fully-fledged version, and {\sf RV}.
-By comparing the {\sf Recipe-s} and {\sf Recipe}, we find {\sf Recipe} finds 100+ more races, which demonstrates the strength of exploring  un-executed paths.
+The $Races$ section shows the number of races detected by {\sf Recipe-s}, i.e., {\sf Recipe} without exploring un-executed paths; {\sf Recipe}, i.e., the full-fledged version; and {\sf RV}.
+Comparison between {\sf Recipe-s} and {\sf Recipe} reveals that {\sf Recipe} finds $>$100 more races, which demonstrates the strength of exploring  un-executed paths.
 Intuitively,  {\sf Recipe-s} predicts based on a single trace, while {\sf Recipe} predicts based on multiple traces containing different execution paths.
-We also compare the {\sf Recipe} version with {\sf RV} version, as illustrated in the $Difference$ section, where $Diff$ shows the races found by {\sf Recipe} but missed by {\sf RV}, $Diff'$ shows the races found by {\sf RV} but missed by {\sf Recipe}.
+We also compare between {\sf Recipe} and {\sf RV} version, as shown in the $Difference$ section, where $Diff$ shows the races found by {\sf Recipe} but missed by {\sf RV} while $Diff'$ shows the races found by {\sf RV} but missed by {\sf Recipe}.
 
-First of all, {\sf Recipe} finds many 150+ more races than {\sf RV}.
-The reasons are multifold: (1) {\sf Recipe} can reason about the accesses in the un-executed paths, while {\sf RV} and {\sf Recipe-s} can only reason about the accesses in the executed paths.
-(2) {\sf Recipe} or {\sf Recipe-s} allows the relaxation of the scheduling even if it breaks the inter-thread read/write dependence in the observed run.
-{\sf Recipe} allows the majority of the shared reads, i.e., the reads of primitive values, to freely read from a different value from a different write as long as the value leads to the same branch decisions.
+First of all, {\sf Recipe} finds $>$150 more races than {\sf RV}.
+There are several reasons for that: (1) {\sf Recipe} can reason about the accesses in the un-executed paths, while {\sf RV} and {\sf Recipe-s} can only reason about the accesses in the executed paths.
+(2) {\sf Recipe} or {\sf Recipe-s} relax the scheduling even if it breaks the inter-thread read/write dependence structure in the observed run.
+{\sf Recipe} allows the majority of the shared reads, i.e., the reads of primitive values, to freely read a different value from a different write as long as the value leads to the same branch decisions.
 {\sf RV}, however, requires them to read the same values as in the observed run.
@@
 Our strategy reduces the number of branches greatly and amplifies the effectiveness of the optimization.
-We also conduct case studies (Section~\ref{}) to better illustrate the scenarios.
+We also carried out case studies (Section~\ref{}) to better illustrate the scenarios.
 
-Another interesting observation is, although {\sf Recipe} should produce all races found by {\sf RV} in theory, {\sf Recipe} may miss races found by {\sf RV} in practice (Column $Diff'$), i.e., {\sf Recipe} is not strictly more effective than {\sf RV} in practice.
-The underlying reason is due to the limits of the constraint solver: (1) the solver cannot compute constraints with very complex arithmatic operations (2) the solver does not support some program constants such as the scientific notation, $3E-10$.
+Another interesting observation is that although {\sf Recipe} should produce all races found by {\sf RV} in theory, {\sf Recipe} may miss races found by {\sf RV} in practice (Column $Diff'$), i.e., {\sf Recipe} is not strictly more effective than {\sf RV} in practice.
+The reason for this is limitations of the underlying constraint solver: (1) the solver cannot compute constraints with very complex arithmatic operations, and (2) the solver does not support some program constants such as the scientific notation $3E-10$.
+In our empirical evaluation, we encounted only one benchmark, account, where these issues manifested.
 
@@
 
-{\bf Summary of advantage and weakness\ } In general, our approach is flexible, which allows the relaxation of schedules and paths based on the fine-grained reasoning of value flows.
+{\bf Summary of advantage and weakness\ } In general, our approach is flexible, which allows the relaxation of schedules and paths based on  fine-grained reasoning at the level of data values.
 As a result, it detects many more races compared to existing approaches.
-On the other hand, it is heavy weighted.
-The suggestion is to combine it with the lightweight approach with soundness guarantee such as {\sf RV}, by treating {\sf RV} as the preprocessing and instructing {\sf Recipe} to skip those confirmed by {\sf RV}.
-In this way, we can complement {\sf Recipe} by finding those missed by it due to the limit of the constraint solver.
+On the other hand, our approach is more heavyweight.
+An interesting topic for future research is how  to combine our approach with a lightweight approach with soundness guarantees like {\sf RV}.
+One option is to create a staged analysis, wherein {\sf RV} is first run.
+\tool\ is then left to run only on racy pairs not confirmed by {\sf RV}.
+In this way, we can also tackle misses due to practical limitations of the constraint solver.
 
@@
 \subsection{Case studies}
-We manually inspect the reported races in small applications to gain better understanding about our approach.
+We manually inspected the reported races over small applications to gain better understanding about our approach.
 
-{\bf Relaxing the Inter-thread dependence\ } Figure~\ref{fig:relax1} demonstrates the relaxation of inter-thread dependence enabled by our approach. 
+{\bf Relaxing Inter-thread dependencies\ } Figure~\ref{fig:relax1} demonstrates the relaxation of inter-thread dependence enabled by our approach. 
 The code is from the benchmark {\sf bbuffer}, where the line number is marked.
-Huang et al.~\cite{} detects the race between line 291 and line 400, but fails to detect the race between line 294 and line 400.
-The reason is as follows.
-In the observation run, the execution follows the order,  lines 400, 291 and 294.
+Huang et al.~\cite{} detect the race between lines 291 and 400, but fail to detect the race between lines 294 and 400.
+The reason is that in the observed run, the execution follows the order 400, 291 and 294.
 For the event at line 294, its preceding branch at line 291 reads from line 400.
 Therefore, Huang et al.~\cite{}
- requires the predicted run to preserve the dependence between line 291 and line 400 so that line 291 reads exactly the same value and the branch takes the same branch decision.
+ require the predicted run to preserve the dependence between line 291 and line 400, so that line 291 reads exactly the same value and the branching expression evaluates to the same truth value.
 The dependence enforces the order constraint, $400 \rightarrow 291$, which further enforces the order $400\rightarrow 291 \rightarrow 294$.
@@
 Specifically, we allow line 291 to happen before line 400 in the predicted run as long as the value read by it leads to the same branch decision, which is true in this case.
-As a result, there is no order constraint between line 294 and line 400, and the two forms a race.
-We re-replay such race  easily using the eclipse IDE breakpoints.
+As a result, there is no order constraint between line 294 and line 400, and the two form a race.
+We were able to replay this race  easily using the eclipse IDE breakpoints.
 
@@
 
-{\bf Relaxing the Paths\ } Figure~\ref{fig:relax2} demonstrates how our approach relaxes the schedulings to account for the unexecuted paths. 
-Each thread invokes the method {\sf Sorting}, which recursively starts two children threads if there are two more available entries in the thread pool (lines 8-11), or  starts one child thread if there is only one available entry (lines 4-5).
+{\bf Relaxing the Paths\ } Figure~\ref{fig:relax2} demonstrates how our approach relaxes the scheduling to account for  unexecuted paths. 
+All threads invoke the method {\sf Sorting}, which recursively starts two child threads if there are two more available entries in the thread pool (lines 8-11), or  starts one child thread if there is only one available entry (lines 4-5).
 The availability is computed through the static method {\sf available} with the constant {\tt total}, which is equal to 5.
@@
 Initially there is one sorting thread. After it starts Thread 1 and Thread 2, there are three threads alive and only two more entries are available.
-Suppose in the observation run, Thread 2 consumes both thread entries and starts the children Thread 3 and Thread 4 (not shown), updating {\tt alive} to 5, then Thread 1 cannot execute the branch at lines 4-5.
-Huang et al.~\cite{} require the predicted run to preserve the dependence denoted by the dotted line since the dependence affects the branch condition at line 2.
+Suppose that in the observed run Thread 2 consumes both thread entries and starts the child Thread 3 and Thread 4 (not shown), updating {\tt alive} to 5.
+Then Thread 1 cannot execute the branch at lines 4-5.
+Huang et al.~\cite{} require the predicted run to preserve the dependence denoted by the dotted line since this dependence affects the branch condition at line 2.
 As a result, the predicted run follows the same branch decision and cannot reason about the code at lines 4-5.
-Our analysis does not have such limitation.
+Our analysis does not suffer from this limitation.
 Instead, it allows the predicted run to reason about the un-explored code.
-Specifically, it does not need to preserve the dependence from line 11 to line 1 and it allows line 1 (Thread 1) to read from line 9 (Thread 2).
+Specifically, it does not need to preserve the dependence from line 11 to line 1, and it allows line 1 (Thread 1) to read from line 9 (Thread 2).
 As a result, the branch condition guarding the unexplored branch is evaluated to be true, enabling the unexplored path in the constraint solver.
@@
 
-\section{Validity Threat}
-Java method can have 65535 bytecode instructions maximally.
-Therefore, we count the number of bytecode instructions inside each method, if the number exceeds 65525, we avoid instrumenting the method.
-The consequence is that, we will miss the races inside the method.
+\section{Threats to Validity}
+A Java method can consist of up to 65535 bytecode instructions.
+Therefore, we count the number of bytecode instructions inside each method.
+If the resulting number exceeds 65525, we avoid instrumenting the method.
+The consequence is that we will potentially races inside a method for which we skipped instrumentation..
 In this case, we specify the variables read from the method to be equal to their concrete values in the constraints.
-THe constraint solver can proceed safely without being affected by such methods.
+The constraint solver can proceed safely without being affected by such methods.
 
-We do not support the boolean operations such as $\&$, bit operators $<<$, which contributes to most of our misses.
+A second point regards support for certain operators.
+In our current implementation, we do not support the boolean operators, including e.g.
+$\&$, as well as the bit operator $<<$, which accounts for most of our misses.

### 2014-09-05 related.tex
error fixes and related

@@
 
-Initial attempts to address the challenge of race detection focused on built-in synchronization primitives \cite{XXX,YYY}.
+Initial attempts to address the challenge of race detection focused on built-in synchronization primitives \cite{eraserPaper,Agarwal,vonPraun:2001,Choi:2002}.
 These include locks as well as the wait/notify and start/join scheduling controls.
-Notable among these efforts is \emph{lockset} analysis, which considers only locks \cite{XXX}.
-Because the derived constraints are partial, permitting certain infeasible event reorderings, lockset analysis cannot guarantee soundness \cite{Huang}.
+Notable among these efforts is \emph{lockset} analysis, which considers only locks \cite{eraserPaper}.
+Because the derived constraints are partial, permitting certain infeasible event reorderings, lockset analysis cannot guarantee soundness \cite{Naik:2006}.
 
-A different tradeoff is struck by the \emph{happens-before (HB)} approach \cite{Lamport}.
-In this style of analysis, all synchronization primitives are accounted for, though reordering constraints are conservative.
+A different tradeoff is struck by the \emph{happens-before (HB)} approach \cite{Christiaens:2001,Dinning:1990,Mellor-Crummey:1991}, which is inspired by and based on Lamport's HB relation \cite{Lamport}.
+In this style of analysis, all synchronization primitives are accounted for, though reordering constraints are conservative .
 As an example, HB inhibits reordering of two synchronized blocks governed by the same lock.
-PTA is an instance  of the HB approach, whose starting point is a concrete execution trace.
+
+%PTA is an instance  of the HB approach, whose starting point is a concrete execution trace.
 
@@
 The first to propose PTA are Smagardakis et al.
-\cite{XXX}.
+\cite{Smaragdakis:2012}.
 In their original work, both synchronization constraints and inter-thread dependencies are preserved, where inter-thread dependencies are respected by
@@
 Said et al.
-\cite{XXX} describe a PTA that is also sound albeit not maximal.
+\cite{Said:2011} describe a PTA that is also sound albeit not maximal.
 Similarly to our analysis, Said et al.
@@
 Soundsness is guaranteed by their ability to precisely encode the semantics of sequential consistency.
-ExceptionNULL \cite{XXX} is another example of a sound PTA without maximality guarantees, where the goal is to detect null dereferences rather than data races.
+ExceptionNULL \cite{Farzan:2012} is another example of a sound PTA without maximality guarantees, where the goal is to detect null dereferences rather than data races.
 
 Serbanuta et al.
-\cite{XXX} have recently shown that it is possible to build a model that is not only sound but also maximal: Any extension of the model with a new trace renders the model unsound.
-This provides a foundation for different forms of PTA, including data races, but also atomicity, serializability and other properties.
+\cite{SerbanutaCR12} have recently shown that it is possible to build a model that is not only sound but also maximal: Any extension of the model with a new trace renders the model unsound.
+The theoretical framework created by Serbanuta et al.
+provides a foundation for different forms of PTA, including data races, but also atomicity, serializability and other properties.
 
 Inspired by this result and closer to \tool\ is the PTA technique of Huang et al.
-\cite{XXX}, which is both sound and maximal.
-This technique, too, makes use of an SMT solver based on symbolic encoding of the input trace.
+\cite{HuangMR14}, which --- like our technique --- is sound and also maximal, but only under the assumption that local assignments are not modeled.
+Huang et al.s technique, too, makes use of an SMT solver based on symbolic encoding of the input trace.
 As part of the encoding, control-flow information is taken into account to enable reorderings that do not violate control dependencies.
@@
 In Section \ref{sec:eval}, we demonstrate via direct comparison with Huang et al.
-the dramatic improvement in coverage thanks to these relaxation methods, which we prove to handle in a sound and maximal manner.
+the dramatic improvement in coverage thanks to these relaxation methods, which we prove to handle in a sound manner.

### 2014-09-05 relax2.tex
error fixes and related
relaxII

@@
 We leverage the static analysis to  realize this relaxation. We explain the details below.
+
+As the static analysis has some known challenges, we make several simplifying assumptions: (1) the branch structure does not contain loops or recursions; (2) every base reference variables can be fully resolved to a concrete object, as required by our analysis (Section~\ref{sec:relax1}).
+% If the base reference variable is defined out of the branch and is not updated inside the unexplored branch, it is resolved to the object referenced by the variable at the beginning of the branch, which is computed from the trace.
+If the base reference variable is updated inside the unexplored branch to some shared access, the base reference variables cannot be fully resolved statically.
+However, if it is updated to an object created inside the unexplored branch, we can still model the object and treats the variable as fully resolved.
+ (3) we assume there are no array accesses in the branch structure, because the array accesses often use the index expressions that cannot be fully resolved statically.
+Similarly, we assume there are no use of collections.
+If any of the above assumptions is violated, we  avoid exploring the unexplored branches and adopt the purely dynamic solution in Section~\ref[sec:relax1}.
+ %TODO inline
+
+ For presentation, we refer to the branch structure as, $B=<header, B_{if}, B_{else}>$, which includes the header instruction and two branches.
+Without loss of generality, we assume the original trace takes the branch $B_{if}$.
+Our running example is shown in Figure~\ref{fig:path}.
+
+{\bf Events \ }.
+We traverse the control flow graph of the branch $B$ and produce an event, $<t, id, inst>$, to represent the possible execution of each instruction $inst$.
+Here, $t$ is the same as that of the root branch event as the exploration is conducted thread-locally, $id$ is a unique value different from existing ids.
+Specifically,  we maintain an increasing counter to assign the id following the traversal order.
+For this example, we simply use the line number as the event id.
+As we assume no loops and recursions, we establish the one-to-one mapping between each instruction $inst$ and the event $e_{inst}$.  We may use the two terms interchangeably.
+
+{\bf SSA Form\ }  We apply SSA transformation to the code of the branch structure and then update the instruction stored in the corresponding event.
+Our SSA transformation is similar to the standard SSA transformation but needs minor adaptations.
+\begin{itemize}
+\item {\bf Local primitive variables\ }  For the variable $z$ defined in an instruction $inst$ inside the branch, our SSA transformation replaces it as $z^{id}$, where $id$ comes from the event $e_{inst}$.
+Some local variables are defined out of the branch structure, e.g., the variable $x$ is defined at line 5 in Figure~\ref{fig:path}.
+We look up the SSA encoding history up to the branch event to find the symbol that replaces the variable most recently and use the symbol instead.
+For example, the variable $x$ is replaced to $x^5$ at the event $e_5$ according to the history, we simple replace the uses of $x$ as $x^5$.
+In addition, the SSA transformation may introduce Phi node in the form of $y=Phi(y_1, C_1, y_2, C_2, \dots y_i, C_i)$, meaning that $y$ is equal to $y_i$ in the condition $C_i$.
+\item {\bf Local Heap Accesses\ } Given the heap access of $x.f$, remember that the base variable $x$ is fully resolved to a concrete object $o$, we replace it as a local variable $l_{o.f}$ before the SSA transformation and treat it as a local primitive variable during the SSA transformation.
+\item {\bf Shared Heap Accesses\ } The SSA transformation does not encode the shared heap accesses.
+Similar to Section~\ref{sec:relax1}, we introduce the special symbols to denote the shared read/write.
+\end{itemize}
+
+The SSA form for the branch at line 6 is as follows, where the base variable $w$ at line 13 is replaced to an object $o2$ and the local heap access $w.f$ is encoded as a local primitive variable.
+
+\begin{figure}
+\centering
+\begin{tabular}{l}
+ {\tt 6: if($x^5$<10)}  \\ %{\tt if($x^5$<10)}\\
+{\tt 7: $W^7_{S}$=0;}   \\%{\tt $W^7_{S}$=0;}\\
+ {{\tt 8:  else}}   \\ %{\color{Gray}{\tt if($x^5\geq$10)}}\\
+ {{\tt 9: \ \ $z^9$=$R^{9}_{o1.f}$;}}  \\ %{\color{Gray}{\tt $z^{9}$=$R^{9}_{o1.f}$;}}\\
+ {{\tt 10: \ if($x^5$<$z^9$)}}  \\%{\color{Gray}{\tt if($x^5<z^{9}$)}}\\
+ {{\tt 11: \ \ $W^{11}_{S}$=0;}} \\%{\color{Gray}{\tt $W^{11}_{S}$=0;}}\\ 
+ {{\tt 12: \ else }}  \\%{\color{Gray}{\tt if($x^5\geq z^{10}$)}}\\
+ {{\tt 13: \ \ $W^{13}_{S}$=$l^{id}_{o2.f}$; }}   \\%{\color{Gray}{\tt $W^{13}_{S}$=$l^{id}_{o2.f}$;}}\\
+\end{tabular}
+\caption{Running Example (shared variables are in bold font). }
+\label{fig:pathssa}
+\end{figure}
+
+{\bf Constraints\ } During the traversal, we also record the shared access events.
+For the set of accesses $S_1$ of the location $\ell$, we also find the accesses $S_2$ of it from other threads.
+We assert the feasibility of each race pair $\in S_1 \times S_2$ by specifying constraints and checking them.
+
+{\bf Race Condition Constraint\ }  The first constraint is the race condition constraint, e.g., $O'_{4}=O'(e_11)$ specifies that the event at line 4 may form a race with the event at line 11, which is not executed in the original trace.
+
+{\bf Intra-thread Control Flow Constraint \ } Different from the strategy in Section~\ref{sec:relax1}, we do not assert the branch the predicted run will take.
+Instead, we assert that all branches inside the branch structure may be taken, which is trivially $true$.
+The solver automatically reasons about which branch should be taken, as indicated in the solution.
+Note that only one of the two branches can be taken because the conditions for them contradict.
+
+{\bf Intra-thread Value Constraint \ } Intra-thread value constraint is the core constraint that realizes the exploration different branches.
+It asserts the value constraints imposed by the {\sf assign/heapr/heapw} events take effect only when the guarding condition is satisfied.
+More formally, we have
+
+\[ cons(n) = \left\{ 
+  \begin{array}{l l}
+  (\bigwedge_{n_1\in n_{if}}{  (expr(header_n) \rightarrow  cons(n_1))}) \\
+  \bigwedge (\bigwedge_{n_2\in n_{else}}  { (\neg expr(header_n) \rightarrow  cons(n_2))}) \\
+   \quad  \quad  \quad \quad  \quad  \quad  \quad  \quad  \quad  \text{if n is a branch} \\
+  equation(n)      \\                                  
+   \quad  \quad  \quad \quad  \quad  \quad \quad  \quad  \quad  \text{otherwise} \\ 
+  \end{array} \right.\]
+Here, $expr(header)$ is the helper function that returns the condition associated with the header.
+$equation(n)$ returns the equation inside each event, which imposes the value constraint.
+In the first case when the instruction $n$ is a branch, we assert the value constraint imposed by each instruction in the if-branch can take effect only when the if-branch is taken, i.e., when $expr(header_n)$ is evaluated as true,  and the value constraint imposed by instruction in the else-branch can take effect only when the else-branch is taken.
+In the second case, when the instruction $n$ is a normal instruction that contains the equation, we assert the equation as the value constraint.
+Note that the constraint is defined recursively because a branch  may contain another branch structure.
+
+Consider the example in Figure~\ref{fig:pathssa},  the constraint is,
+$$
+\begin{array}{l}
+&(\theta'(x^5)<10 \rightarrow \theta'(W^7_{S})=0) \wedge (\theta'(x^5)\geq 10 \rightarrow \theta'(z^9)=\theta'(R^{9}_{o1.f}) 
+&\wedge  (\theta'(x^5)\geq 10 \rightarrow  
+& \ \ \ \ \  \ \ (\theta'(x^5)<\theta'(z^9) \rightarrow \theta'(W^{11}_{S})=0 
+& \ \ \ \ \  \ \ \wedge \theta'(x^5)\geq \theta'(z^9) \rightarrow \theta'(W^{13}_{S})=\theta'(l^{id}_{o2.f}))
+&)\\ 
+\end{array}
+$$
+
+{\bf Intra-thread Order Constraints\ } The intra-thread order constraint captures the program order imposed by the branch structure and the instruction sequence.
+More formally, for each branch $n$, we have,
+
+\[ cons(n) = \left\{ 
+  \begin{array}{l l}
+  (\bigwedge_{n_1\in n_{if}}{O'(header_n)<O'(n_1)<O'(end_n)}) \\
+   (\bigwedge_{n_1 and n_3 are adjacent in n_{if}}{O'(n_1)<O'(n_3)}) \\
+   (\bigwedge_{n_2\in n_{else}}{O'(header_n)<O'(n_2)<O'(end_n)}) \\
+    (\bigwedge_{n_2 and n_4 are adjacent in n_{else}}{O'(n_2)<O'(n_4)}) \\
+  \end{array} \right.\]
+Here, $end_n$ is a node that denotes the first event after the branch.
+The order constraints are imposed by the program, and therefore, are unconditionally valid.
+That is why we do not specify the order constraints under specific branch conditions.
+
+{\bf Inter-thread value constraints\ } The inter-thread value constraints are identical to those in Section~\ref{sec:relax1}.
+Interesting, we do not need to specify the value constraints under the specific branch condition.
+The underlying reason is that, the intra-thread value constraint already encodes the branch condition information, if the branch condition is not satisfied at runtime, the value constraint does not take effect and the read of a shared variable cannot propagate its value downstream, even if it is assigned to a value.
 
@@
 {\tt 2: {\bf Q}=101; } &  & {\tt $W^{2}_{Q}$=101;}\\
-{\tt 3: {\bf Q}=11; } & & {\tt $W^{3}_{Q}$=11;} \\
-{\tt 4: {\bf Q}=1; } & & {\tt $W^{4}_{Q}$=1;} \\
-{\tt 5: {\bf S}=1; } & & {\tt $W^{5}_{S}$=1;} \\
-& {\tt 6: x={\bf Q};} & {\tt $x^6$=$R^6_{Q}$;}\\
-& {\tt 7: if(x<10)} & {\tt if($x^6$<10)}\\
-& {\tt 8: \ {\bf S}=0;}  & {\tt $W^8_{S}$=0;}\\
-& {\color{Gray}{\tt 9:  else}} &  {\color{Gray}{\tt if($x^6\geq$10)}}\\
-& {\color{Gray}{\tt 10: \ z={\bf P}.f;}} & {\color{Gray}{\tt $z^{10}$=$R^{10}_{o1.f}$;}}\\
-& {\color{Gray}{\tt 11: \ if(x<z)}} & {\color{Gray}{\tt if($x^6<z^{10}$)}}\\
-& {\color{Gray}{\tt 12: \ \ {\bf S}=0;}}& {\color{Gray}{\tt $W^{12}_{S}$=0;}}\\ 
-& {\color{Gray}{\tt 13: \ else }}  & {\color{Gray}{\tt if($x^6\geq z^{10}$)}}\\
-& {\color{Gray}{\tt 14: \ \ w=new(); }}  & {\color{Gray}{\tt $w^{14}$=$o2$;}}\\
-& {\color{Gray}{\tt   \ \ \ \ \ \  //o2 }}  & {}\\
-& {\color{Gray}{\tt 15: \ \ w.f=0; }}  & {\color{Gray}{\tt $l^{15}_{o2.f}$=0;}}\\
-& {\color{Gray}{\tt 16: \ \ {\bf S}=w.f; }}  & {\color{Gray}{\tt $W^{16}_{S}$=$l^{15}_{o2.f}$;}}\\
+{\tt 3: {\bf Q}=1; } & & {\tt $W^{3}_{Q}$=1;} \\
+{\tt 4: {\bf S}=1; } & & {\tt $W^{4}_{S}$=1;} \\
+& {\tt 5: x={\bf Q};} & {\tt $x^5$=$R^5_{Q}$;}\\
+& {\tt 6: if(x<10)} & {\tt if($x^5$<10)}\\
+& {\tt 7: \ {\bf S}=0;}  & {\tt $W^7_{S}$=0;}\\
+& {\color{Gray}{\tt 8:  else}} &  \\ %{\color{Gray}{\tt if($x^5\geq$10)}}\\
+& {\color{Gray}{\tt 9: \ \ z={\bf P}.f;}} & \\ %{\color{Gray}{\tt $z^{9}$=$R^{9}_{o1.f}$;}}\\
+& {\color{Gray}{\tt 10: \ if(x<z)}} & \\%{\color{Gray}{\tt if($x^5<z^{9}$)}}\\
+& {\color{Gray}{\tt 11: \ \ {\bf S}=0;}}& \\%{\color{Gray}{\tt $W^{11}_{S}$=0;}}\\ 
+& {\color{Gray}{\tt 12: \ else }}  & \\%{\color{Gray}{\tt if($x^5\geq z^{10}$)}}\\
+& {\color{Gray}{\tt 13: \ \ {\bf S}=w.f; }}  & \\%{\color{Gray}{\tt $W^{13}_{S}$=$l^{id}_{o2.f}$;}}\\
 \end{tabular}
@@
 \end{figure}
-
-Consider the example in Figure~\ref{fig:path}, only the if-branch of $T_2$ is executed in the original trace $\tau$.
-Following the rules in Section~\ref{sec:relax1},   we derive a set of constraints to find the new trace $\tau'$ to witness the race pair ($e_5$, $e_8$). We list those specific to if-branch below.
-
-\begin{figure}
-	\begin{center}
-$$
-	\begin{array}{rcl}
-		&  O'_{e_5}=O'_{e_8} & \textbf{(race condition)} \\
-			& \theta'(x^6)<10 & \textbf{(Intra-thread control flow constraint)} \\
-	&  O'_{e_7}<O'_{e_8} & \textbf{(Intra-thread order constraint)} \\
-		& \theta'(W^8_S)=0 & \textbf{(Intra-thread value constraint)} \\
-	\end{array} 
-$$
-\end{center}
-caption{\label{fig:pathencode}
-\end{figure}
-
-We  replace these constraints to another set of constraints that account for both if-branch and else-branch from the event $e_7$.
+%
+%Consider the example in Figure~\ref{fig:path}, only the if-branch of $T_2$ is executed in the original trace $\tau$.
+%Following the rules in Section~\ref{sec:relax1},   we derive a set of constraints to find the new trace $\tau'$ to witness the race pair ($e_5$, $e_8$). We list those specific to if-branch below.
+%
+%\begin{figure}
+%	\begin{center}
+%$$
+%	\begin{array}{rcl}
+%		&  O'_{e_5}=O'_{e_8} & \textbf{(race condition)} \\
+%			& \theta'(x^6)<10 & \textbf{(Intra-thread control flow constraint)} \\
+%	&  O'_{e_7}<O'_{e_8} & \textbf{(Intra-thread order constraint)} \\
+%		& \theta'(W^8_S)=0 & \textbf{(Intra-thread value constraint)} \\
+%	\end{array} 
+%$$
+%\end{center}
+%caption{\label{fig:pathencode}
+%\end{figure}
+%
+%
+%We  replace these constraints to another set of constraints that account for both if-branch and else-branch from the event $e_7$.
 We refer to the branch event $e_7$ as the root branch event, denoted as $e_{root}$.
-
-First, we conduct the DFS traversal of the branch structure and produce an event, $<t, id, inst>$, to represent the potential execution of each instruction $inst$.
+%
+%
+%
+%First, we conduct the DFS traversal of the branch structure and produce an event, $<t, id, inst>$, to represent the potential execution of each instruction $inst$.
 Here, $t$ is the same as that of the root branch event as the exploration is conducted thread-locally, $id$ is a unique value different from existing ids.
 Specifically,
- we maintain an increasing counter to assign the id following the traversal order. For this example, we simply use the line number as the event id.
-
-similar to Section~\ref{sec:relax1}, we need the SSA form of
+% we maintain an increasing counter to assign the id following the traversal order. For this example, we simply use the line number as the event id.
+%
+%
+%
+%similar to Section~\ref{sec:relax1}, we need the SSA form of
 

### 2014-09-05 introduction.tex
intro+relaxI

@@
 & {\tt 4: if (y > 2)} \\
-& {\tt 5:~~print(1/x);} \\	
+& {\tt 5:~~z=1/x;} \\	
 & {\color{red} {\tt 6: else}} \\
-& {\color{red} {\tt 7:~~print(2/x);}}
+& {\color{red} {\tt 7:~~w=2/x;}}
 \end{tabular}
@@
 and some of the branch outcomes such that part of the thread local paths stay intact. The criterion is that
-we forbid \sysname to explore an unexecuted branch if the effect of the branch cannot be precisely modeled
+we forbid \sysname\ to explore an unexecuted branch if the effect of the branch cannot be precisely modeled
 (e.g. an array element is read while it was not defined in the original trace). The essence of our technique 
@@
 \begin{itemize}
-\item We propose to relax existing preditive analysis and identify a bound of relaxation (i.e. what can
+\item We propose to relax existing predictive analysis and identify a bound of relaxation (i.e. what can
 be relaxed).
-\item We develop a constraint encoding scheme that uniformally encodes the strict information (i.e. the
-part must be preserved) and the relaxed information. 
+\item We develop a constraint encoding scheme that encodes the necessary information (i.e. the
+part that must be preserved) and the relaxed information (i.e., the part that may be mutated) uniformly as part of a single logical representation.
 \item We develop a propotype \sysname. Our results show XXX.

### 2014-09-05 impl.tex
impl

@@
 \section{Implementation}\label{sec:impl}
-For array access event $arr[i]=x$, we treat them as the field access event $arr.i=x$, where the index $i$ is treated as a field.
 
-We distinguish the accesses of the shared fields and local fields.
-To collect the shared fields, we preprocess the trace and map each field to a set of accessing threads.
-If a field is written by two or more threads, it is shared.
-If a field is written by one thread and read by other threads, it is shared.
-Otherwise, it is a local field.
+\subsection{Time Window and Initial Value constraints}~\label{sec:timewindow}
+The trace is typically long and the constraint solver cannot scale to the whole trace. This is a known limitation to solver-based predictive analysis. Instead, we adopt the notion of time window. 
+We divide the trace into $N$ traces of the same length, where $N$ is configurable parameter (1000 in our experiment).
+To support the time window, we need to store the values of the shared variables and local variables at the end of a window and encode such values as the inputs of the next window.
+
+We maintain two maps,  $sstore$ and $lstore$, for the shared variables and local variables respectively.
+Suppose the trace for the last window is $\tau$.
+The following code illustrates how we maintain the $sstore$, i.e., storing the value written by the last write with each address.
+Beside, the variable $e.ins.left$ represents the left hand variable inside the instruction $e.ins$.
+
+\begin{algorithmic}[3]
+\For {$a \in \mathcal{A}$}
+ \State $\tau_a=W(a)$
+ \State $sstore.put(a,\tau_a[|\tau_a|-1].v)$ 
+\EndFor
+\end{algorithmic}
+
+%TODO define left:
+\begin{algorithmic}[3]
+\For {$e \in \tau$}
+  \If {$e.isLocalAcc()$}
+    \State $ lstore.put(e.\_def, e.v)$
+  \Else
+     \If{$e.isRead()$}      
+        \State $ lstore.put(e.\_def, e.v)$
+     \EndIf
+  \EndIf 
+\EndFor
+\end{algorithmic}
+
+When constructing the constraints for the next window, we specify the following constraints to encode the input values.
+
+The following formula enhances the thread interfernce constraint to include the case that the read reads from the initial values.
+
+$$\Phi_{TI}=\Phi_{TI} \wedge  R=sstore.get(e_R.a) \wedge \bigvee_{e\in S} O(e_R)<O(e)$$.
+
+The following formula encodes the input values of local variables.
+
+$$\Phi_{IN}=\bigvee_{l\in lstore.keys()} l=lstore.get(l)$$.
+
+{\bf Optimization\ }.
+Before the symbolic analysis, we conduct the backward slicing by treating the racy pair as the seeds.
+Only those selected in the slices are preserved while the rest events are removed.
+
+\subsection{Scalability Issues}
+Like existing predictive analyses, our approach suffers from scalability issues for two reasons.
+First, as we record local access events in the monitoring run,  the monitoring run takes long time, e.g., more than 2 hours, and our trace is typically large, e.g., at the GB level.
+To avoid the out of memory issue during the monitoring run, we store the event to the database on the fly, rather than storing them in a buffer.
+Besides, we terminate the monitoring run with a shutdown hook if it is beyond 30 minutes.
+Second,  the prediction phase needs to load the trace to memory, which easily causes out of memory problem.
+To solve the problem, we separate the analysis into several runs, each run restarting the JVM and resuming from the window of last run.
+Different runs communicate the window id using the external file.
+These runs are organized together automatically using the ant.
+The outputs of the files are merged into one summary.

### 2014-09-05 tech.tex
impl

@@
 Besides, as mentioned above, we need to know the address which guides the matching of reads and writes. As we assume the unexplored branch has no object creation, all addresses are fully resolved.
-
-\section{Implementation}\label{sec:impl}
-
-\subsection{Time Window and Initial Value constraints}~\label{sec:timewindow}
-The trace is typically long and the constraint solver cannot scale to the whole trace. Instead, we adopt the notion of time window. 
-We divide the trace into $N$ traces of the same length, where $N$ is configurable parameter (1000 in our experiment).
-To support the time window, we need to store the values of the shared variables and local variables at the end of a window and encode such values as the inputs of the next window.
-
-We maintain two maps,  $sstore$ and $lstore$, for the shared variables and local variables respectively.
-Suppose the trace for the last window is $\tau$.
-The following code illustrates how we maintain the $sstore$, i.e., storing the value written by the last write with each address.
-Beside, the variable $e.ins.left$ represents the left hand variable inside the instruction $e.ins$.
-
-\begin{algorithmic}[3]
-\For {$a \in \mathcal{A}$}
- \State $\tau_a=W(a)$
- \State $sstore.put(a,\tau_a[|\tau_a|-1].v)$ 
-\EndFor
-\end{algorithmic}
-
-%TODO define left:
-\begin{algorithmic}[3]
-\For {$e \in \tau$}
-  \If {$e.isLocalAcc()$}
-    \State $ lstore.put(e.\_def, e.v)$
-  \Else
-     \If{$e.isRead()$}      
-        \State $ lstore.put(e.\_def, e.v)$
-     \EndIf
-  \EndIf 
-\EndFor
-\end{algorithmic}
-
-When constructing the constraints for the next window, we specify the following constraints to encode the input values.
-
-The following formula enhances the thread interfernce constraint to include the case that the read reads from the initial values.
-
-$$\Phi_{TI}=\Phi_{TI} \wedge  R=sstore.get(e_R.a) \wedge \bigvee_{e\in S} O(e_R)<O(e)$$.
-
-The following formula encodes the input values of local variables.
-
-$$\Phi_{IN}=\bigvee_{l\in lstore.keys()} l=lstore.get(l)$$.
-
-{\bf Object Creation\ } Given the event in the form of $x=new Object()$, we encode the created object as a unique integer and treats the instruction as a normal integer assignment.
-The unique integer is the address of the object, calculated by $System.identityHashcode(x)$.
-
-{\bf Array access\ } For array access event $arr[i]=x$, we treat them as the field access event $arr.i=x$, where the index $i$ is treated as a field. 
-We ensure the predicted run sees the same index at the same event by specifying the additional constraint $i==v$, where $v$ is the concrete index value observed.
-
-{\bf Shared Analysis\ } We distinguish the accesses of the shared fields and local fields.
-To collect the shared fields, we preprocess the trace and map each field to a set of accessing threads.
-If a field is written by two or more threads, it is shared.
-If a field is written by one thread and read by other threads, it is shared.
-Otherwise, it is a local field.
-
-{\bf Optimization\ }.
-Before the symbolic analysis, we conduct the backward slicing by treating the racy pair as the seeds.
-Only those selected in the slices are preserved while the rest events are removed.
-
-\subsection{Scalability Issues}
-Our approach suffers from scalability issues for two reasons.
-First, as we record local access events in the monitoring run,  the monitoring run takes long time, e.g., more than 2 hours, and our trace is typically large, e.g., at the GB level.
-To avoid the out of memory issue during the monitoring run, we store the event to the database on the fly, rather than storing them in a buffer.
-Besides, we terminate the monitoring run with a shutdown hook if it is beyond 30 minutes.
-Second,  the prediction phase needs to load the trace to memory, which easily causes out of memory problem.
-To solve the problem, we separate the analysis into several runs, each run restarting the JVM and resuming from the window of last run.
-Different runs communicate the window id using the external file.
-These runs are organized together automatically using the ant.
-The outputs of the files are merged into one summary finally.
-Besides,

### 2014-09-05 eval.tex
eval - impl

@@
 We chose RV for two main reasons: (1) {\sf RV} is the only open source predictive analysis, (2) {\sf RV} represents the state-of-the-art approach, which is theoretically proven to have higher detection capability than other approaches.
+
+{\bf Prototype Implementation}  In our prototype implementation of \tool, we used the Z3 SMT solver for constraint resolution \cite{DeMoura:2008}.
+For static analysis, to discover unexecuted branches, we utilize the SOOT compiler framework \cite{Vallee-Rai:1999}.
+
+For scalability, as well as to constrain memory footprint, we have applied several engineering optimizations.
+In specific, (i) events are stored in a database rather than in an in-memory buffer; and (ii) the analysis is split into several segments, between which we restart the VM and resume the last snapshot, such that the different logs are fused together at the end.
+
+To address the scalability limitations of the constraint solver (Z3 in our case), we again apply a splitting optimization.
+The full trace is split into $N$ equal-length subtraces or ``time windows'' (in our experiments, $N=1,000$).
+The values of local and shared variables are then passed at the end of a given window as inputs to the next window.
+For lack of space, we omit the concrete details of this optimization, and instead refer the reader to Huang et al.
+\cite{HuangMR14}, who apply a similar optimization.
+
+The trace is typically long and the constraint solver cannot scale to the whole trace. This is a known limitation to solver-based predictive analysis. Instead, we adopt the notion of time window. 
+We divide the trace into $N$ traces of the same length, where $N$ is configurable parameter (1000 in our experiment).
+To support the time window, we need to store the values of the shared variables and local variables at the end of a window and encode such values as the inputs of the next window.
+
+We maintain two maps,  $sstore$ and $lstore$, for the shared variables and local variables respectively.
+Suppose the trace for the last window is $\tau$.
+The following code illustrates how we maintain the $sstore$, i.e., storing the value written by the last write with each address.
+Beside, the variable $e.ins.left$ represents the left hand variable inside the instruction $e.ins$.
+
+\begin{algorithmic}[3]
+	\For {$a \in \mathcal{A}$}
+	\State $\tau_a=W(a)$
+	\State $sstore.put(a,\tau_a[|\tau_a|-1].v)$ 
+	\EndFor
+\end{algorithmic}
+
+%TODO define left:
+\begin{algorithmic}[3]
+	\For {$e \in \tau$}
+	\If {$e.isLocalAcc()$}
+	\State $ lstore.put(e.\_def, e.v)$
+	\Else
+	\If{$e.isRead()$}      
+	\State $ lstore.put(e.\_def, e.v)$
+	\EndIf
+	\EndIf 
+	\EndFor
+\end{algorithmic}
+
+When constructing the constraints for the next window, we specify the following constraints to encode the input values.
+
+The following formula enhances the thread interfernce constraint to include the case that the read reads from the initial values.
+
+$$\Phi_{TI}=\Phi_{TI} \wedge  R=sstore.get(e_R.a) \wedge \bigvee_{e\in S} O(e_R)<O(e)$$.
+
+The following formula encodes the input values of local variables.
+
+$$\Phi_{IN}=\bigvee_{l\in lstore.keys()} l=lstore.get(l)$$.
 

### 2014-09-05 relax2.tex
intro

@@
 At the high level, we replace the constraints that model the executed branch to the constraints that conservatively model the possible executions in both branches.
-We leverage the static analysis to  realize this relaxation. We explain the details below.
+We leverage the static analysis to  realize this relaxation.
 
-As the static analysis has some known challenges, we make several simplifying assumptions: (1) the branch structure does not contain loops or recursions; (2) every base reference variables can be fully resolved to a concrete object, as required by our analysis (Section~\ref{sec:relax1}).
-% If the base reference variable is defined out of the branch and is not updated inside the unexplored branch, it is resolved to the object referenced by the variable at the beginning of the branch, which is computed from the trace.
-If the base reference variable is updated inside the unexplored branch to some shared access, the base reference variables cannot be fully resolved statically.
-However, if it is updated to an object created inside the unexplored branch, we can still model the object and treats the variable as fully resolved.
- (3) we assume there are no array accesses in the branch structure, because the array accesses often use the index expressions that cannot be fully resolved statically.
-Similarly, we assume there are no use of collections.
+We start with a branch event in the trace, without loss of generality, suppose its true branch is taken, we need to explore the false branch and add constraints to model it.
+As the static analysis has some known challenges, we make several simplifying assumptions: (1) the unexplored branch should not contain loops or recursions because the loop bound may depend on some shared variable which is statically unknown; (2) every base reference variable should be fully resolved to a concrete object, as required by our analysis (Section~\ref{sec:relax1}).
+We apply dataflow analysis (reaching definitions) to judge: If the base reference depends on some shared read in the explored branch, of which the value is statically unknown, it cannot be statically resolved to an object; If the base reference depends on multiple definitions from different paths, it cannot be statically resolved to an object.
+Otherwise, it can be resolved.
+(3) We assume the unexplored branch does not contain array access because the array accesses often use the index expressions that are statically unknown.
+Similarly, as the collections often use array for implementation, we assume they are also absent.
+(4) We inline the method calls (for two levels) in the unexplored branch, similar to the base object, we assume the caller object can be fully resolved statically.
 If any of the above assumptions is violated, we  avoid exploring the unexplored branches and adopt the purely dynamic solution in Section~\ref[sec:relax1}.
- %TODO inline
+We use the example in Figure~\ref{fig:running2}
 

### 2014-09-05 relax2.tex
relaxII done

@@
 
-We start with a branch event in the trace, without loss of generality, suppose its true branch is taken, we need to explore the false branch and add constraints to model it.
+We start with a branch event $e_b$ in the trace, without loss of generality, suppose its true branch is taken, we need to explore the false branch (thread-locally with the same thread as $e_b$) and add constraints to model it.
 As the static analysis has some known challenges, we make several simplifying assumptions: (1) the unexplored branch should not contain loops or recursions because the loop bound may depend on some shared variable which is statically unknown; (2) every base reference variable should be fully resolved to a concrete object, as required by our analysis (Section~\ref{sec:relax1}).
@@
 If any of the above assumptions is violated, we  avoid exploring the unexplored branches and adopt the purely dynamic solution in Section~\ref[sec:relax1}.
-We use the example in Figure~\ref{fig:running2}
-
- For presentation, we refer to the branch structure as, $B=<header, B_{if}, B_{else}>$, which includes the header instruction and two branches.
-Without loss of generality, we assume the original trace takes the branch $B_{if}$.
-Our running example is shown in Figure~\ref{fig:path}.
+We use the example in Figure~\ref{fig:running} for demonstration.
 
 {\bf Events \ }.
-We traverse the control flow graph of the branch $B$ and produce an event, $<t, id, inst>$, to represent the possible execution of each instruction $inst$.
-Here, $t$ is the same as that of the root branch event as the exploration is conducted thread-locally, $id$ is a unique value different from existing ids.
-Specifically,  we maintain an increasing counter to assign the id following the traversal order.
-For this example, we simply use the line number as the event id.
-As we assume no loops and recursions, we establish the one-to-one mapping between each instruction $inst$ and the event $e_{inst}$.  We may use the two terms interchangeably.
+We traverse the control flow graph of the unexplored branch and generate an event,  $<t, id, inst>$, to represent the possible execution of each instruction $inst$.
+The id can be any unique value, where the uniqueness is the requirement.
+We simply use the line number as the event id in our example.
+As the unexplored branch does not contain loop/recursion, we establish the one-to-one between each instruction $inst$ and the event $e_{inst}$.
 
-{\bf SSA Form\ }  We apply SSA transformation to the code of the branch structure and then update the instruction stored in the corresponding event.
-Our SSA transformation is similar to the standard SSA transformation but needs minor adaptations.
+{\bf SSA Form\ }  Similar to in Section~\ref{sec:relax1}, we need the SSA form of the instruction in each event.
+To achieve this, we apply SSA transformation to the code of  unexplored branch and combine its SSA form with the existing SSA form of the explored branch to produce the final SSA form.
+The SSA transformation is standard, but special issues need to be considered.
 \begin{itemize}
-\item {\bf Local primitive variables\ }  For the variable $z$ defined in an instruction $inst$ inside the branch, our SSA transformation replaces it as $z^{id}$, where $id$ comes from the event $e_{inst}$.
-Some local variables are defined out of the branch structure, e.g., the variable $x$ is defined at line 5 in Figure~\ref{fig:path}.
-We look up the SSA encoding history up to the branch event to find the symbol that replaces the variable most recently and use the symbol instead.
-For example, the variable $x$ is replaced to $x^5$ at the event $e_5$ according to the history, we simple replace the uses of $x$ as $x^5$.
-In addition, the SSA transformation may introduce Phi node in the form of $y=Phi(y_1, C_1, y_2, C_2, \dots y_i, C_i)$, meaning that $y$ is equal to $y_i$ in the condition $C_i$.
+\item {\bf Local primitive variables\ }  Some local variable is defined before the branch event $e_b$, of which the SSA form is already in the trace $\tau$, we reuse the SSA form.
+In addition, the SSA transformation may introduce Phi node, e.g., $y=Phi(y_1, C_1, y_2, C_2, \dots y_i, C_i)$, meaning that $y$ is equal to $y_i$ in the condition $C_i$.
+The combination of SSA forms of both branches may require further Phi node when both branches define a variable.
+We insert the Phi node to the trace and replace the following use accordingly.
 \item {\bf Local Heap Accesses\ } Given the heap access of $x.f$, remember that the base variable $x$ is fully resolved to a concrete object $o$, we replace it as a local variable $l_{o.f}$ before the SSA transformation and treat it as a local primitive variable during the SSA transformation.
-\item {\bf Shared Heap Accesses\ } The SSA transformation does not encode the shared heap accesses.
+\item {\bf Shared Heap Accesses\ } The standard SSA transformation does not encode the shared heap accesses.
 Similar to Section~\ref{sec:relax1}, we introduce the special symbols to denote the shared read/write.
@@
 
-The SSA form for the branch at line 6 is as follows, where the base variable $w$ at line 13 is replaced to an object $o2$ and the local heap access $w.f$ is encoded as a local primitive variable.
+The SSA form for the branch at line 4 is as follows.
 
@@
 \begin{tabular}{l}
- {\tt 6: if($x^5$<10)}  \\ %{\tt if($x^5$<10)}\\
+ {\tt 4: if($x^5$<10)}  \\ %{\tt if($x^5$<10)}\\
 {\tt 7: $W^7_{S}$=0;}   \\%{\tt $W^7_{S}$=0;}\\
@@
 
+{\bf Phi node}
+
 {\bf Intra-thread Order Constraints\ } The intra-thread order constraint captures the program order imposed by the branch structure and the instruction sequence.
@@
 
-\begin{figure}
-\centering
-\begin{tabular}{ll|l}
-\multicolumn{3}{c}{{\tt {\bf P} = new();//$o1$}} \\
-\multicolumn{1}{c}{$T_1$} & \multicolumn{1}{c}{$T_2$}  &  \multicolumn{1}{c}{$Trace$}\\
-{\tt 1: {\bf P}.f=100; } &  & {\tt $W^{1}_{o1.f}$=100;}\\
-{\tt 2: {\bf Q}=101; } &  & {\tt $W^{2}_{Q}$=101;}\\
-{\tt 3: {\bf Q}=1; } & & {\tt $W^{3}_{Q}$=1;} \\
-{\tt 4: {\bf S}=1; } & & {\tt $W^{4}_{S}$=1;} \\
-& {\tt 5: x={\bf Q};} & {\tt $x^5$=$R^5_{Q}$;}\\
-& {\tt 6: if(x<10)} & {\tt if($x^5$<10)}\\
-& {\tt 7: \ {\bf S}=0;}  & {\tt $W^7_{S}$=0;}\\
-& {\color{Gray}{\tt 8:  else}} &  \\ %{\color{Gray}{\tt if($x^5\geq$10)}}\\
-& {\color{Gray}{\tt 9: \ \ z={\bf P}.f;}} & \\ %{\color{Gray}{\tt $z^{9}$=$R^{9}_{o1.f}$;}}\\
-& {\color{Gray}{\tt 10: \ if(x<z)}} & \\%{\color{Gray}{\tt if($x^5<z^{9}$)}}\\
-& {\color{Gray}{\tt 11: \ \ {\bf S}=0;}}& \\%{\color{Gray}{\tt $W^{11}_{S}$=0;}}\\ 
-& {\color{Gray}{\tt 12: \ else }}  & \\%{\color{Gray}{\tt if($x^5\geq z^{10}$)}}\\
-& {\color{Gray}{\tt 13: \ \ {\bf S}=w.f; }}  & \\%{\color{Gray}{\tt $W^{13}_{S}$=$l^{id}_{o2.f}$;}}\\
-\end{tabular}
-\caption{Running Example (shared variables are in bold font). }
-\label{fig:path}
-\end{figure}
+%\begin{figure}
+%\centering
+%\begin{tabular}{ll|l}
+%\multicolumn{3}{c}{{\tt {\bf P} = new();//$o1$}} \\
+%\multicolumn{1}{c}{$T_1$} & \multicolumn{1}{c}{$T_2$}  &  \multicolumn{1}{c}{$Trace$}\\
+%{\tt 1: {\bf P}.f=100; } &  & {\tt $W^{1}_{o1.f}$=100;}\\
+%{\tt 2: {\bf Q}=101; } &  & {\tt $W^{2}_{Q}$=101;}\\
+%{\tt 3: {\bf Q}=1; } & & {\tt $W^{3}_{Q}$=1;} \\
+%{\tt 4: {\bf S}=1; } & & {\tt $W^{4}_{S}$=1;} \\
+%& {\tt 5: x={\bf Q};} & {\tt $x^5$=$R^5_{Q}$;}\\
+%& {\tt 6: if(x<10)} & {\tt if($x^5$<10)}\\
+%& {\tt 7: \ {\bf S}=0;}  & {\tt $W^7_{S}$=0;}\\
+%& {\color{Gray}{\tt 8:  else}} &  \\ %{\color{Gray}{\tt if($x^5\geq$10)}}\\
+%& {\color{Gray}{\tt 9: \ \ z={\bf P}.f;}} & \\ %{\color{Gray}{\tt $z^{9}$=$R^{9}_{o1.f}$;}}\\
+%& {\color{Gray}{\tt 10: \ if(x<z)}} & \\%{\color{Gray}{\tt if($x^5<z^{9}$)}}\\
+%& {\color{Gray}{\tt 11: \ \ {\bf S}=0;}}& \\%{\color{Gray}{\tt $W^{11}_{S}$=0;}}\\ 
+%& {\color{Gray}{\tt 12: \ else }}  & \\%{\color{Gray}{\tt if($x^5\geq z^{10}$)}}\\
+%& {\color{Gray}{\tt 13: \ \ {\bf S}=w.f; }}  & \\%{\color{Gray}{\tt $W^{13}_{S}$=$l^{id}_{o2.f}$;}}\\
+%\end{tabular}
+%\caption{Running Example (shared variables are in bold font). }
+%\label{fig:path}
+%\end{figure}
 %

### 2014-09-05 relax2.tex
relaxII

@@
 (4) We inline the method calls (for two levels) in the unexplored branch, similar to the base object, we assume the caller object can be fully resolved statically.
-If any of the above assumptions is violated, we  avoid exploring the unexplored branches and adopt the purely dynamic solution in Section~\ref[sec:relax1}.
+If any of the above assumptions is violated, we  avoid exploring the unexplored branches and adopt the purely dynamic solution in Section~\ref{sec:relax1}.
 We use the example in Figure~\ref{fig:running} for demonstration.
@@
 \begin{tabular}{l}
- {\tt 4: if($x^5$<10)}  \\ %{\tt if($x^5$<10)}\\
-{\tt 7: $W^7_{S}$=0;}   \\%{\tt $W^7_{S}$=0;}\\
- {{\tt 8:  else}}   \\ %{\color{Gray}{\tt if($x^5\geq$10)}}\\
- {{\tt 9: \ \ $z^9$=$R^{9}_{o1.f}$;}}  \\ %{\color{Gray}{\tt $z^{9}$=$R^{9}_{o1.f}$;}}\\
- {{\tt 10: \ if($x^5$<$z^9$)}}  \\%{\color{Gray}{\tt if($x^5<z^{9}$)}}\\
- {{\tt 11: \ \ $W^{11}_{S}$=0;}} \\%{\color{Gray}{\tt $W^{11}_{S}$=0;}}\\ 
- {{\tt 12: \ else }}  \\%{\color{Gray}{\tt if($x^5\geq z^{10}$)}}\\
- {{\tt 13: \ \ $W^{13}_{S}$=$l^{id}_{o2.f}$; }}   \\%{\color{Gray}{\tt $W^{13}_{S}$=$l^{id}_{o2.f}$;}}\\
+ {\tt 4: if($R^4_y$>2)}  \\ %{\tt if($x^5$<10)}\\
+{\tt 5: \ \ $z^5$=1+$R^5_x$;}   \\%{\tt $W^7_{S}$=0;}\\
+ {{\tt 6:  else}}   \\ %{\color{Gray}{\tt if($x^5\geq$10)}}\\
+ {{\tt 7: \ \ $w^7$=$2+R^{7}_{x}$;}}  \\ %{\color{Gray}{\tt $z^{9}$=$R^{9}_{o1.f}$;}}\\
 \end{tabular}
-\caption{Running Example (shared variables are in bold font). }
+\caption{SSA form of both true and false branches. }
 \label{fig:pathssa}
@@
 
-{\bf Race Condition Constraint\ }  The first constraint is the race condition constraint, e.g., $O'_{4}=O'(e_11)$ specifies that the event at line 4 may form a race with the event at line 11, which is not executed in the original trace.
+{\bf Race Condition Constraint\ }  The first type of constraint is the race condition constraint, e.g., $O'_{e_2}=O'(e_7)$ specifies that the event at line 2 may form a race with the event at line 7, which is actually not executed in the original trace.
 
 {\bf Intra-thread Control Flow Constraint \ } Different from the strategy in Section~\ref{sec:relax1}, we do not assert the branch the predicted run will take.
-Instead, we assert that all branches inside the branch structure may be taken, which is trivially $true$.
-The solver automatically reasons about which branch should be taken, as indicated in the solution.
-Note that only one of the two branches can be taken because the conditions for them contradict.
+Instead, we assert that both branches are possible, which is always $true$.
+The solver automatically reasons about which branch should be taken and encodes it in the solution.
+Note that only one of the two branches can be taken because the conditions for the branches contradict.
 
@@
 It asserts the value constraints imposed by the {\sf assign/heapr/heapw} events take effect only when the guarding condition is satisfied.
-More formally, we have
+Suppose $IF(e_b)$ denotes the event sequence that corresponds to the true branch, $ELSE(e_b)$ denotes the event sequence that corresponds to the false branch, the branch event $e_b$ is  $if(x_b<y_b)$, the event $e_m$ in $IF(e_b)$ is in the form of $x_m=y_m+z_m$ and the event $e_n$ in $ELSE(e_b)$ is in the form of $x_n=y_n+z_n$.
 
-\[ cons(n) = \left\{ 
-  \begin{array}{l l}
-  (\bigwedge_{n_1\in n_{if}}{  (expr(header_n) \rightarrow  cons(n_1))}) \\
-  \bigwedge (\bigwedge_{n_2\in n_{else}}  { (\neg expr(header_n) \rightarrow  cons(n_2))}) \\
-   \quad  \quad  \quad \quad  \quad  \quad  \quad  \quad  \quad  \text{if n is a branch} \\
-  equation(n)      \\                                  
-   \quad  \quad  \quad \quad  \quad  \quad \quad  \quad  \quad  \text{otherwise} \\ 
-  \end{array} \right.\]
-Here, $expr(header)$ is the helper function that returns the condition associated with the header.
-$equation(n)$ returns the equation inside each event, which imposes the value constraint.
-In the first case when the instruction $n$ is a branch, we assert the value constraint imposed by each instruction in the if-branch can take effect only when the if-branch is taken, i.e., when $expr(header_n)$ is evaluated as true,  and the value constraint imposed by instruction in the else-branch can take effect only when the else-branch is taken.
-In the second case, when the instruction $n$ is a normal instruction that contains the equation, we assert the equation as the value constraint.
-Note that the constraint is defined recursively because a branch  may contain another branch structure.
-
-Consider the example in Figure~\ref{fig:pathssa},  the constraint is,
+Consider our example, the corresponding value constraint is, 
 $$
 \begin{array}{l}
-&(\theta'(x^5)<10 \rightarrow \theta'(W^7_{S})=0) \wedge (\theta'(x^5)\geq 10 \rightarrow \theta'(z^9)=\theta'(R^{9}_{o1.f}) 
-&\wedge  (\theta'(x^5)\geq 10 \rightarrow  
-& \ \ \ \ \  \ \ (\theta'(x^5)<\theta'(z^9) \rightarrow \theta'(W^{11}_{S})=0 
-& \ \ \ \ \  \ \ \wedge \theta'(x^5)\geq \theta'(z^9) \rightarrow \theta'(W^{13}_{S})=\theta'(l^{id}_{o2.f}))
-&)\\ 
+\theta'(R^4_y)>2 \rightarrow \theta'(z^5)=1+\theta'(R^5_x) \\ 
+\wedge  \\
+\theta'(R^4_y)\leq 2 \rightarrow  \theta'(w^7)=2+\theta'(R^{7}_{x})\\
 \end{array}
@@
 
-{\bf Phi node}
+{\bf Assignment with {\sf Phi}\ }  For the assignment with the Phi, e.g., $y=Phi(y_1, x>0, y_2, x\leq 0)$, the encoding is similar.
+$$
+\begin{array}{l}
+\theta'(x)>0 \rightarrow \theta'(y)=\theta'(y_1) \wedge \theta'(x)\leq 0 \rightarrow \theta'(y)=\theta'(y_2)
+\end{array}
+$$
+
+In case that the unexplored branch contains another branch structure, the above encoding process is applied recursively and the resultant constraints are nested.
 
 {\bf Intra-thread Order Constraints\ } The intra-thread order constraint captures the program order imposed by the branch structure and the instruction sequence.
-More formally, for each branch $n$, we have,
+Suppose $IF(e_b)$=$e_{i_1}e_{i_2}\dots e_{i_m}$, $ELSE(e_b)=e_{j_1}e_{j_2}\dots e_{j_n}$, the first event after the branch structure is $e_{end}$.
+Then we have,
 
-\[ cons(n) = \left\{ 
-  \begin{array}{l l}
-  (\bigwedge_{n_1\in n_{if}}{O'(header_n)<O'(n_1)<O'(end_n)}) \\
-   (\bigwedge_{n_1 and n_3 are adjacent in n_{if}}{O'(n_1)<O'(n_3)}) \\
-   (\bigwedge_{n_2\in n_{else}}{O'(header_n)<O'(n_2)<O'(end_n)}) \\
-    (\bigwedge_{n_2 and n_4 are adjacent in n_{else}}{O'(n_2)<O'(n_4)}) \\
-  \end{array} \right.\]
-Here, $end_n$ is a node that denotes the first event after the branch.
-The order constraints are imposed by the program, and therefore, are unconditionally valid.
-That is why we do not specify the order constraints under specific branch conditions.
+$$
+\begin{array}{l}
+O'(e_b)<O'(e_{i_1})<O'(e_{i_2})<\dots <O'(e_{i_m})<O'(e_{end})\\
+\wedge\\
+O'(e_b)<O'(e_{j_1})<O'(e_{j_2})<\dots <O'(e_{j_m})<O'(e_{end})
+\end{array}
+$$
+
+The order constraints capture the program order, which are unconditionally valid. That is also why we do not specify the order constraints under specific branch conditions.
 
@@
 Interesting, we do not need to specify the value constraints under the specific branch condition.
-The underlying reason is that, the intra-thread value constraint already encodes the branch condition information, if the branch condition is not satisfied at runtime, the value constraint does not take effect and the read of a shared variable cannot propagate its value downstream, even if it is assigned to a value.
+The underlying reason is that, the shared read  $R$ in the instruction $x=R$ is isolated (meaning the value cannot be propagated downstream) if the instruction is disabled in the intra-thread value constraints due to the infeasible branch condition.
 

### 2014-09-06 overview.tex
minor

@@
 In our running example, this is essential to discover the race between lines {\tt 2} and {\tt 7}.
-We conduct the symbolic execution to exercise the unexplored branches, which returns a set of traces representing possible executions of the unexplored branch.
-Each trace is combined with the original trace to find races involving the accesses in the unexplored path.
-For the racy events, $e_2$ and $e_7$, we first specify the race condition constraint  $O'(e_2)=O'(e_7)$.
-In addition, we negate the path condition, thereby obtaining $\theta'(R_{\tt y}^4) > 2 \equiv {\bf false}$ in place of $\theta'(R_{\tt y}^4) > 2 \equiv {\bf true}$.
-We also model the other necessary constraints such as the intra-thread order constraint $O'(e_4) < O'(e_7)$, similar to the above analysis.
+We explore the unexecuted branch statically and collect constraints during the exploration.
+We then form a disjunction  of the original constraints and the collected constraints to include more possible  paths.
+The solver may pick up the unexplored branch to detect a race whenever it is necessary.
+Back to our example, the event $e_7$ is in the unexplored branch, our solver is able to identify the race between $e_7$ and $e_2$ after encoding the unexplored branch.
+
+For the racy events, $e_2$ and $e_7$, we first specify the race condition constraint  $\theta'(R^4_y)\leq 2\rightarrow O'_{e_2}=O'(e_7)$.
+Note that the race condition constraint associates the race with the branch condition that enables each event.
+Also we encode other necessary constraints such as the intra-thread order constraints, $O'(e_4)<O'(e_5)\wedge O'(e_4)<O'(e_7)$.
+The inter-thread
 

### 2014-09-06 eval.tex
changes

--- 2014-09-05 eval.tex
@@
 We chose RV for two main reasons: (1) {\sf RV} is the only open source predictive analysis, (2) {\sf RV} represents the state-of-the-art approach, which is theoretically proven to have higher detection capability than other approaches.
+
+{\bf Prototype Implementation}  In our prototype implementation of \tool, we used the Z3 SMT solver for constraint resolution \cite{DeMoura:2008}.
+For static analysis, to discover unexecuted branches, we utilize the SOOT compiler framework \cite{Vallee-Rai:1999}.
+
+For scalability, as well as to constrain memory footprint, we have applied several engineering optimizations.
+In specific, (i) events are stored in a database rather than in an in-memory buffer; and (ii) the analysis is split into several segments, between which we restart the VM and resume the last snapshot, such that the different logs are fused together at the end.
+
+To address the scalability limitations of the constraint solver (Z3 in our case), we again apply a splitting optimization.
+The full trace is split into $N$ equal-length subtraces or ``time windows'' (in our experiments, $N=1,000$).
+The values of local and shared variables are then passed at the end of a given window as inputs to the next window.
+For lack of space, we omit the concrete details of this optimization, and instead refer the reader to Huang et al.
+\cite{HuangMR14}, who apply a similar optimization.
 
@@
 16GB of RAM and 6M caches. The workstation runs version 12.04 of the Ubuntu Linux distribution, and has the Sun 
-64-Bit 1.6.0\_26 Java virtual machine (JVM) installed.\footnote{We note that Sun JDK 1.7 does not support the transformation of dacapo applications with {\sf tamiflex}.}
+64-Bit 1.6.0\_26 Java virtual machine (JVM) installed.
+%\footnote{We note that Sun JDK 1.7 does not support the transformation of dacapo applications with {\sf tamiflex}.}
 
@@
 Intuitively,  {\sf Recipe-s} predicts based on a single trace, while {\sf Recipe} predicts based on multiple traces containing different execution paths.
-We also compare between {\sf Recipe} and {\sf RV} version, as shown in the $Difference$ section, where $Diff$ shows the races found by {\sf Recipe} but missed by {\sf RV} while $Diff'$ shows the races found by {\sf RV} but missed by {\sf Recipe}.
+We also compare between {\sf Recipe} and {\sf RV} version, as shown in the $Difference$ section, where $Diff$ shows the races found by {\sf Recipe} but missed by {\sf RV} while $Diff'$ shows the races found by {\sf RV} but not {\sf Recipe}.
 
@@
 
-{\bf Summary of advantage and weakness\ } In general, our approach is flexible, which allows the relaxation of schedules and paths based on  fine-grained reasoning at the level of data values.
-As a result, it detects many more races compared to existing approaches.
-On the other hand, our approach is more heavyweight.
+To summarize, while our approach is able to relax schedules and paths based on  fine-grained reasoning at the level of data values, it also imposes higher overhead.
 An interesting topic for future research is how  to combine our approach with a lightweight approach with soundness guarantees like {\sf RV}.
-One option is to create a staged analysis, wherein {\sf RV} is first run.
+One option is staged analysis, wherein {\sf RV} is run.
 \tool\ is then left to run only on racy pairs not confirmed by {\sf RV}.
@@
 \centering
-\includegraphics[scale=0.7]{cases/Visio-bbuffer.pdf}
+\includegraphics[scale=0.45]{cases/Visio-bbuffer.pdf}
 \caption{Relaxation of Inter-thread dependence}\label{fig:relax1}
@@
 \centering
-\includegraphics[scale=0.7]{cases/Visio-msort.pdf}
+\includegraphics[scale=0.45]{cases/Visio-msort.pdf}
 \caption{Relaxation of Paths}\label{fig:relax2}

### 2014-09-06 introduction.tex
intro
changes

@@
 value from line {\tt 1} as it has the same effect on thread local execution (i.e. the same branch is taken). 
-However, the relaxation would allow us to execute lines {\tt 4} and {\tt 5} before line {\tt 2}, exposing 
+This relaxation would allow us to execute lines {\tt 4} and {\tt 5} before line {\tt 2}, thereby exposing 
 the first missing race. We can also allow the trace to take the else branch such that the second race is also exposed. 
-Note that even though the else branch was not executed originally, its effect can be precisely modeled as the value 
+Note that even though that branch was not executed originally, its effect can be precisely modeled as the value 
 of $x$ is known from the original trace. As such, we can still perform sound analysis.

### 2014-09-06 overview.tex
changes

--- 2014-09-05 overview.tex
@@
 \subsection{Preliminaries}
-To facilitate our presentation, we first introduce some terminologies used throughout this paper. Intuitively, a trace is a sequence of events recorded during the observation run.
+We first introduce some notation used throughout this paper.
+Intuitively, a trace is a sequence of events recorded during the observed run.
+An event, $e=<t, id, inst>$, is a concrete representation of the runtime execution of a static instruction $inst$, such that (i) $t$ refers to the thread that issues the event $e$ (denoted $t^e$); (ii) $id$ is a unique identifier of the event (unless spcified otherwise, we use the index of an event in the trace as its id); and (iii) $inst$ is a static three-address instruction (involving at most three operands, which modern compilers commonly support).
 
-{\bf Event\ } An event, $e=<t, id, inst>$, is a concrete representation of the runtime execution of a static instruction $inst$.
-
-\begin{itemize}
-\item $t$ refers to the thread that issues the event $e$, denoted as $t^e$.
-\item  $id$ refers to the id associated with each event.
-The key property of  $id$ is {\em uniqueness}, i.e., any two events in the trace own different ids.
-Unless otherwise specified, we use  the index of an event in the trace as its id, which satisfies the above property.
-Throughout this paper,  we denote an event as $e_{id}$ with the id as the subscript.
-%We may use the terms $e_{id}$ and $id$ interchangeably given their one-to-one correspondence, e.g., $t^{e_3}$ and $t^3$.
-\item $inst$ is the static instruction.
-The instructions are three-address instructions involving at most three operands, which modern compilers commonly support.
-Specifically, we are interested in the types of instructions listed in Table~\ref{Ta:syntax}.
+For this paper, we utilize the instructions listed in Table~\ref{Ta:syntax}.
 When the variable does not appear on the left hand of an equation, such as $y$ in $x.f=y$, it may refer to a variable, a constant or event object creation expression $new (...)$.
@@
 This standard form of simplification preserves all the data-race-related information contained in the original trace.
-\end{itemize}
 
@@
 
-{\bf Trace \ } A trace $\tau$ is a sequence of events. It can also be represented more comprehensively as,  $\tau=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \}, O, \theta>$.
+{\bf Trace Representation \ } In the following, we assume that trace $\tau$ is represented as $\tau=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \}, O, \theta>$, where
 \begin{itemize}
-\item  $\Gamma$ refers to the set of threads involved in the trace, which include $t_1$, $t_2$, \dots, $t_n$.
-\item   $\{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \}$ denotes the event sequences produced by each thread.  
-\item  $O$ is a function that assigns  to every event an integer value that denotes its scheduling order. For example, $e_i$ is scheduled earlier than $e_j$ if $O(e_i)<O(e_j)$.
-\item  $\theta$ is a function that maps every variable to a value.
-For example, given an event $e$ executing the instruction $x=y+z$, the mapping recorded in the trace may be $\theta(x)=3$,  $\theta(y)=2$,  $\theta(z)=1$.
-If the variable $x$ is defined in two events, the mapping for $x$ becomes ambiguous.
-To avoid the ambiguity, we conduct a pre-processing  (Section~\ref{sec:relax1}) that produces the SSA form of the trace, in which each variable is defined exactly once.
+\item  $\Gamma$ refers to the set of threads involved in the trace, which include $t_1$, $t_2$, \dots, $t_n$;
+\item   $\{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \}$ denotes the event sequences produced by each thread;
+\item  $O$ is a function that assigns  to every event an integer value that denotes its scheduling order: $e_i$ is scheduled earlier than $e_j$ iff $O(e_i)<O(e_j)$; and
+\item  $\theta$ is a function that maps every live variable to a value. 
+%For example, given an event $e$ executing the instruction $x=y+z$, the mapping recorded in the trace may be $\theta(x)=3$,  $\theta(y)=2$,  $\theta(z)=1$.  
 \end{itemize}
+To ensure the integrity of $\theta$ as well simplify downstream encoding steps performed by the analysis already at the syntactic level, we rewrite the raw input trace into SSA form, such that each variable is defined exactly once.
 
@@
 {\bf Race Condition\ } The potential race pair involves two accesses of the same shared location from different threads, where at least one access is a write.
-Given any potential race pair, e.g., ($e_2$, $e_5$), it is a real race if and only if the two accesses can occur at the same time (race condition), which is captured by the race condition constraint
-$$
-	O'(e_2) = O'(e_5)
-$$.
+Given any potential race pair, e.g., ($e_2$, $e_5$), it is a real race if and only if the two accesses can occur at the same time (race condition), which is captured by the race condition constraint:
+$O'(e_2) = O'(e_5)$.
 The race condition constraint--- together with the other constraints --- guarantees the feasibility of the predicted race if a solution is found for the overall constraint system.
@@
 \item {\bf Control Flow Constraints\ } Each thread takes the same control flows (or branch decisions) to reproduce the events.
- For the running example (ignoring the statements in red), this yields:
-$$
-	\theta'(R_{\tt y}^4) > 2 \equiv {\bf true}
-$$
+ For the running example (ignoring the statements in red), this yields: $\theta'(R_{\tt y}^4) > 2 \equiv {\bf true}$.
 That is, in the predicted trace $\tau'$, the value of variable {\tt y} read at $e_4$ should be greater than $2$.
@@
 This constraint is imposed by the fact that the two runs share the same instruction sequence and take the same control flows.
-We yield the following formula:
-$$
-	O'(e_1) < O'(e_2) < O'(e_3) \wedge O'(e_4) < O'(e_5)
-$$
+We obtain the  formula: $O'(e_1) < O'(e_2) < O'(e_3) \wedge O'(e_4) < O'(e_5)$.
 That is, in the predicted trace $\tau'$, the three events from $T_1$ are ordered in the same way as in $\tau$, and similarly, the branch event $e_4$ in $T_2$ still executes before the event $e_5$ inside the branch body.
@@
 
-{\bf Constraint Solving\ } The formuals from the different encoding steps are conjoined and sent to the off-the-shelf solver such as {\sf z3}.
-The race under analysis is real if the solver returns a solution, which includes the scheduling order and the values for the variables.
+{\bf Constraint Solving\ } The formuals from the different encoding steps are conjoined and sent to an off-the-shelf solver such as {\sf z3}.
+The race under analysis is guaranteed to be real if the solver returns a solution, which includes the scheduling order and the values for the variables.
 
-{\bf Highlights \ } In this example, our analysis finds two more pairs of races than existing analyses~\cite{yannis, jeff}, which include $(e_2, e_5)$ and $(e_2, e_7)$.
-The key insight for finding the first race is, we do not require $e_4$ to read from $e_3$ to retain the original value, as what existing approaches do, rather, we allow $e_4$ to read from $e_1$ while still preserving the feasibility of the branch.
+{\bf Highlights \ } In this example, our analysis finds two more pairs of races beyond existing analyses~\cite{yannis, jeff}: $(e_2, e_5)$ and $(e_2, e_7)$.
+For the first race, we do not require $e_4$ to read from $e_3$ to retain the original value, as in approaches.
+Rather, we allow $e_4$ to read from $e_1$ while still preserving the feasibility of the branch.
 This relaxation allows the events after $e_4$ to run concurrently with the events before $e_3$, leading to the detection of the first race.
-The key insight for finding the second race is, we explore the unexplored branches and ask the solver to compute a feasible schedule that exercises them.
+For the second race, we cover unexecuted branches and ask the solver to compute a feasible schedule that exercises them.
 The neighboring branches are likely to contain racy events if the executed branches contain racy events.

### 2014-09-06 related.tex
changes

@@
 
-Initial attempts to address the challenge of race detection focused on built-in synchronization primitives \cite{eraserPaper,Agarwal,vonPraun:2001,Choi:2002}.
+Initial attempts to address the challenge of race detection focused on built-in synchronization primitives \cite{eraser,SasturkarAWS05,vonPraun:2001,Choi:2002}.
 These include locks as well as the wait/notify and start/join scheduling controls.
-Notable among these efforts is \emph{lockset} analysis, which considers only locks \cite{eraserPaper}.
+Notable among these efforts is \emph{lockset} analysis, which considers only locks \cite{eraser}.
 Because the derived constraints are partial, permitting certain infeasible event reorderings, lockset analysis cannot guarantee soundness \cite{Naik:2006}.
 
-A different tradeoff is struck by the \emph{happens-before (HB)} approach \cite{Christiaens:2001,Dinning:1990,Mellor-Crummey:1991}, which is inspired by and based on Lamport's HB relation \cite{Lamport}.
+A different tradeoff is struck by the \emph{happens-before (HB)} approach \cite{Christiaens,Dinning:1990,Mellor-Crummey:1991}, which is founded on Lamport's HB relation \cite{Lamport}.
 In this style of analysis, all synchronization primitives are accounted for, though reordering constraints are conservative .
@@
 Recently there have been successful attempts to relax HB constraints.
-Among these are hybrid analysis \cite{XXX}, which permits both orderings of lock-synchronized blocks, as well as the \emph{Universal Causal Graph (UCG)} representation~\cite{ucg}, which also enables both orderings but only if these are consistent with wait/notify- and start/join-induced constraints.
+Among these are hybrid analysis \cite{hybrid}, which permits both orderings of lock-synchronized blocks, as well as the \emph{Universal Causal Graph (UCG)} representation~\cite{ucg}, which also enables both orderings but only if these are consistent with wait/notify- and start/join-induced constraints.
 
@@
 The first to propose PTA are Smagardakis et al.
-\cite{Smaragdakis:2012}.
+\cite{yannis}.
 In their original work, both synchronization constraints and inter-thread dependencies are preserved, where inter-thread dependencies are respected by

### 2014-09-06 relax1.tex
intro+relaxI

@@
 \section{Relaxation of Flow Dependencies}~\label{sec:relax1}
-Given the trace, $\tau=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \}, O, \theta>$, the main goal of our analysis is to derive a new trace,  $\tau'=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \}, O', \theta'>$, which reschedules the events from different threads (the order of events inside each thread should remain unchanged) to witness the race.
-The key insight underlying our first relaxation is that, the read events in $\tau'$ are allowed by our analysis to read different values from different writes (than in trace $\tau$), but they are required by existing approaches to read the same values as in $\tau$.
-Therefore, our relaxation allows more schedules, which are likely to expose more races.
-Note that we also need to compute the new value mapping $\theta'$ because the rescheduling may alter the value every variable is assigned to.
+Given the trace, $\tau=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \}, O, \theta>$, the main goal of our analysis is to derive a new trace,  $\tau'=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \}, O', \theta'>$, which reschedules the events from different threads (while preserving intra-thread event order) to prove the race.
+The key insight underlying the first relaxation is that a thread is allowed to read a different value, potentially from a different write, as long as branching decisions remain unchanged.
+Note that we also need to compute the new value mapping $\theta'$, because the rescheduling may alter the value every variable is assigned to.
 The computation of the new schedule and new value mapping is explained in the following.
-
-% % 
-% TODO: Explain the format of the raw trace in impl.tex.
-% %
 
@@
 This requirement is a prerequisite for the computation of the mapping, in which each variable is mapped to exactly one value.
-Another side effect is that the def/use chains become explicit in the SSA form, which simplifies the following analysis steps.
+Another side effect is that the def/use chains become explicit in  SSA form, which simplifies the following analysis steps.
 
@@
 \begin{itemize}
-\item {\bf Local Assignment\ } The SSA form for local assignment resembles the SSA form of static instructions in compiler optimization, except that the loops and recursions are fully resolved in a concrete trace, obviating the need for the {\sf Phi} node.
+\item {\bf Local Assignment\ } The SSA form for local assignment resembles the SSA form of static instructions in compiler optimization, except that the looping and recursion are fully resolved in a concrete trace, obviating the need for {\sf Phi} nodes.
 More concretely, we replace the variable $v$ defined in an event, as well as the following uses of the definition, to a new variable $v^{id}$, where $id$ is the unique id of the event.
 The uniqueness of the id guarantees that no two events define the same variable, i.e., each variable is defined exactly once.
-Consider the example in Figure~\ref{fig:running2}, where the line number is also used as event id,  the local variable $x$ defined at line 4 is renamed as $x^4$ in the trace, while $x$ defined at line 9 is renamed as $x^9$, which is used in the following two events.
+In the example in Figure~\ref{fig:running2}, for example,
+the local variable $x$ defined at line 4 is renamed as $x^4$.
+% where the line number is also used as event id,  the local variable $x$ defined at line 4 is renamed as $x^4$ in the trace, while $x$ defined at line 9 is renamed as $x^9$, which is used in the following two events.
 At line 5, where the object $o3$ is created,  we define a reference variable $w^5$.
 Internally,  the object $o3$ is encoded as a unique integer that indicates its heap address.
-\item {\bf Heap accesses\ } The def/use relation between local heap accesses can be locally determined.
+\item {\bf Heap accesses\ } The def/use relation between local heap accesses can be determined locally.
 Therefore, we treat them in the same way as local assignment.
@@
 Here  the subscript $o.f$ is interpreted as a heap location, i.e., the field $f$ of the object $o$.
-Consider the example in Figure~\ref{fig:running2}, at lines 6-7, we replace the local heap accesses as the local variable $l^6_{o3.f}$.
+Consider the example in Figure~\ref{fig:running2}.
+At lines 6-7, we replace the local heap accesses as the local variable $l^6_{o3.f}$.
 %An assumption here is the uses and the definition still refer to the same base object $o$ in the predicted run, which is enforced through the constraint specification %(Section~\ref{sec:relax1}).
-\item  {\bf Shared Heap Accesses\ }    Given the write $x.f=y$ or the read $y=x.f$ of a shared location at event $e_{id}$, we introduce two symbols, $W^{id}_{\tt o.f }$ and $R^{id}_{\tt  o.f }$ respectively, where $o$ is the runtime object referenced by the $x$.
+\item  {\bf Shared Heap Accesses\ }    Given the write $x.f=y$ or the read $y=x.f$ of a shared location at event $e_{id}$, we introduce two symbols, $W^{id}_{\tt o.f }$ and $R^{id}_{\tt  o.f }$ respectively, where $o$ is the runtime object referenced by $x$.
 Similarly, the subscript $o.f$ is interpreted as a heap location, i.e., the field $f$ of the object $o$.
@@
 Based on the SSA form of the original trace $\tau$, we build the constraints to compute a new trace $\tau'$ with the new schedule $O'$ and new value mapping $\theta'$.
-Consider the example in Figure~\ref{fig:running2}, in the original trace $\tau$, the event $e_8$ reads the value of 2 from $e_3$, therefore, $e_{11}$ must happen after $e_2$.
-Existing approaches enforce the same value dependence between $e_3$ and $e_8$, leading to the miss of the race between $e_2$ and $e_{11}$.
-Our analysis can derive a new trace $\tau'$ to find the race by relaxing the dependence.
-In $\tau'$, $e_8$ reads from a different event $e_1$, producing the new value mapping, $\theta'(y^8)=3$.
-Accordingly, at the event $e_9$, the value mapping for $x^9$ is also updated as 13 ($\theta'(y^8)$=3, $\theta'(z^7)=10$), enabling the true branch at line 10.
+Consider the example in Figure~\ref{fig:running2}.
+In the original trace $\tau$, the event $e_8$ reads the value of 2 from $e_3$.
+Therefore, $e_{11}$ must happen after $e_2$.
+Existing approaches enforce the same value dependence between $e_3$ and $e_8$, thereby missing the race between $e_2$ and $e_{11}$.
+Our analysis allows event $e_8$ to read from $e_1$, producing the new value mapping $\theta'(y^8)=3$.
+Accordingly, at $e_9$ the value mapping for $x^9$ is also updated as 13 ($\theta'(y^8)$=3, $\theta'(z^7)=10$), enabling the true branch at line 10.
 Finally, the events $e_2$ and $e_{11}$ are scheduled to run concurrently.
@@
 
-{\bf Race Condition\ } Following the standard definition, pair ($e_i$, $e_j$) of events forms a race iff (1) $e_i$ and $e_j$ are accesses of the same location $\ell$ by different threads, (2) at least one of them writes to $\ell$, and (3) $e_i$ and $e_j$ run concurrently.
+{\bf Race Condition\ } Following the standard definition, pair ($e_i$, $e_j$) of events forms a race iff (1) $e_i$ and $e_j$ access a mutual  location $\ell$, (2) at least one of them writes $\ell$, and (3) $e_i$ and $e_j$ run concurrently (being associated with different threads).
 We refer to the candidate pair throughout this section as (candidate) \emph{racy events}.
@@
 We then check each pair  separately.
-The checking is achieved by encoding all necessary constraints and invoking a constraint solver.
-The first constraint asserts the feasibility of the concurrent execution of the racy events:
-$$
-O'(e_i) = O'(e_{j})
-$$
-In Figure~\ref{fig:running2}, the race condition for the race pair, ($e_2$, $e_{11}$), is $O'(e_2)=O'(e_{11})$
+The checking is achieved by encoding all necessary constraints.
+The first constraint asserts the feasibility of the concurrent execution of the racy events: $O'(e_i) = O'(e_{j})$.
+%In Figure~\ref{fig:running2}, the race condition for the race pair, ($e_2$, $e_{11}$), is $O'(e_2)=O'(e_{11})$
 
-{\bf Intra-thread Constraints\ } Given the candidate race pair, $e_i$ and $e_j$, suppose $e_i$ is after $e_j$ in the original trace $\tau$.
-The predictive analysis reschedules the events in the prefix of $e_i$, i.e., prior to $e_i$, so that $e_i$ and $e_j$ can run concurrently.
+{\bf Intra-thread Constraints\ } Given racy events $e_i$ and $e_j$, suppose $e_i$ is after $e_j$ in the original trace $\tau$.
+The predictive analysis reschedules the events in the prefix up to $e_i$, so that $e_i$ and $e_j$ can run concurrently.
 We refer to the prefix as $prefix(\tau, e_i)$.
@@
 
-The predictive analysis by design requires that each thread should follow the same event sequence prior to $e$ as in the original run.
-The preservation of the event sequence for each thread requires the following intra-thread constraints:
+Preservation of the event sequence (up to $e$) for each thread requires the following intra-thread constraints:
 
@@
 Consider the example in Figure~\ref{fig:running2}, given the branch if($x^9$>11) at $e_10$, the original mapping is $\theta(x^9)=12$.
-Existing analyses require the same value mapping for $x^9$  in the predicted run, while we allow  a different value mapping, e.g., $\theta'(x^9)=13$, which retains the same truth value for the branch expression.
+We allow  a different value mapping, e.g., $\theta'(x^9)=13$, which retains the same truth value for the branch expression.
 \item {\bf Intra-thread Order Constraints\ } The events in the sequence should follow the same order as in the original trace. More formally, we require that

### 2014-09-06 relax2.tex
minor
-xz

@@
 \section{Exploration of Unexecuted Branches}~\label{sec:relax2}
-We now switch to the second feature of \tool, which is its ability to reason about neighboring branches that are not executed in the original trace $\tau$.
-At the high level, we replace the constraints that model the executed branch to the constraints that conservatively model the possible executions in both branches.
-We leverage the static analysis to  realize this relaxation.
+We now switch to the second feature of \tool, which is to reason about 
+neighboring executions involving unexecuted branches.
+%in the original trace $\tau$.
+At the high level, we enhance the constraints to model the unexecuted branches.
+However to ensure soundness, we cannot encode the unexecuted branches as
+normal static analyses do because they usually perform approximations such as bounded
+loop unrolling and heap abstraction. The basic idea is to {\em only encode the
+unexecuted branches when they can be precisely modeled based on our trace 
+$\tau$.} 
+%replace the constraints that model the executed branch to the constraints that conservatively model the possible executions in both branches.
+%We leverage the static analysis to  realize this relaxation.
 
-We start with a branch event $e_b$ in the trace, without loss of generality, suppose its true branch is taken, we need to explore the false branch (thread-locally with the same thread as $e_b$) and add constraints to model it.
-As the static analysis has some known challenges, we make several simplifying assumptions: (1) the unexplored branch should not contain loops or recursions because the loop bound may depend on some shared variable which is statically unknown; (2) every base reference variable should be fully resolved to a concrete object, as required by our analysis (Section~\ref{sec:relax1}).
-We apply dataflow analysis (reaching definitions) to judge: If the base reference depends on some shared read in the explored branch, of which the value is statically unknown, it cannot be statically resolved to an object; If the base reference depends on multiple definitions from different paths, it cannot be statically resolved to an object.
+In particular, we have the following criteria in determining if an 
+unexecuted branch should be included in the constraint system. These are 
+essentially the bound of our path relaxation.
+(1) If there is a loop inside the unexecuted branch, the loop bound needs to
+be resolved to a constant from the trace (otherwise we do not know how many
+times we need to unroll). (2) The branch
+is ignored if it contains recursive function(s). (3) Every base reference 
+variable in the branch needs to be fully resolved to a concrete object,
+as required by our earlier analysis (Section~\ref{sec:relax1}). This can be 
+done by backward traversal of the trace:
+if the base reference depends on some shared read in the explored 
+branch, of which the value is uncertain, it cannot be 
+resolved to an object; if the base reference depends on multiple definitions 
+from different paths, it cannot be resolved to an object.
 Otherwise, it can be resolved.
-(3) We assume the unexplored branch does not contain array access because the array accesses often use the index expressions that are statically unknown.
-Similarly, as the collections often use array for implementation, we assume they are also absent.
-(4) We inline the method calls (for two levels) in the unexplored branch, similar to the base object, we assume the caller object can be fully resolved statically.
-If any of the above assumptions is violated, we  avoid exploring the unexplored branches and adopt the purely dynamic solution in Section~\ref{sec:relax1}.
-We use the example in Figure~\ref{fig:running} for demonstration.
+(4) The indices of array read accesses in the branch 
+need to be resolved to constant values from the trace (with the preceding unexecuted branch extensions). This is to
+prevent reading an array element that was not defined. 
+(5) The branch is ignored if it uses collections. 
+(6) For method calls, we require the  caller object can be fully resolved from
+the trace.
 
-{\bf Events \ }.
-We traverse the control flow graph of the unexplored branch and generate an event,  $<t, id, inst>$, to represent the possible execution of each instruction $inst$.
-The id can be any unique value, where the uniqueness is the requirement.
-We simply use the line number as the event id in our example.
-As the unexplored branch does not contain loop/recursion, we establish the one-to-one between each instruction $inst$ and the event $e_{inst}$.
+%We start with a branch event $e_b$ in the trace, without loss of generality, 
+%suppose its true branch is taken, we need to explore the false branch 
+%(thread-locally with the same thread as $e_b$) and add constraints to model it.
+% As the static analysis has some known challenges, we make several 
+%simplifying assumptions: 
+%(1) the unexplored branch should not contain 
+%loops or recursions because the loop bound may depend on some shared variable 
+%which is statically unknown; (2) every base reference variable should be 
+%fully resolved to a concrete object, as required by our analysis 
+%(Section~\ref{sec:relax1}). We apply dataflow analysis (reaching definitions) 
+%to judge: If the base reference depends on some shared read in the explored 
+%branch, of which the value is statically unknown, it cannot be statically 
+%resolved to an object; If the base reference depends on multiple definitions 
+%from different paths, it cannot be statically resolved to an object. 
+%Otherwise, it can be resolved. (3) We assume the unexplored branch does not 
+%contain array access because the array accesses often use the index 
+%expressions that are statically unknown. Similarly, as the collections 
+%often use array for implementation, we assume they are also absent. (4) We 
+%inline the method calls (for two levels) in the unexplored branch, similar 
+%to the base object, we assume the caller object can be fully resolved 
+%statically.  If any of the above assumptions is violated, we  avoid exploring 
+%the unexplored branches and adopt the purely dynamic solution in Section~\ref{sec:relax1}. 
+Next, we will explain how to encode an unexecuted branch in details, 
+using the example in Figure~\ref{fig:running} for demonstration.
+We assume one of the branches of event $e_b$ needs to be extended in
+the following discussion.
 
-{\bf SSA Form\ }  Similar to in Section~\ref{sec:relax1}, we need the SSA form of the instruction in each event.
-To achieve this, we apply SSA transformation to the code of  unexplored branch and combine its SSA form with the existing SSA form of the explored branch to produce the final SSA form.
-The SSA transformation is standard, but special issues need to be considered.
+{\bf Events \ } We traverse the control flow graph of the unexplored branch 
+and generate an event,  $<t, id, inst>$, to represent 
+%the execution 
+%of 
+each instruction $inst$. The id can be any unique value.
+%, where the uniqueness is the requirement. 
+We simply use the line number as the event id in our example. As loops are
+unrolled,
+% the unexplored branch does not contain loop/recursion, 
+we establish the one-to-one mapping between each instruction $inst$ and 
+the event $e_{inst}$.
+
+{\bf SSA Form\ }  Similar to in Section~\ref{sec:relax1}, we need the SSA 
+form of the instruction in each event. To achieve this, we apply SSA 
+transformation to the code of unexecuted branch and combine its SSA form 
+with the existing SSA form of the corresponding executed branch to produce the final 
+SSA form. The SSA transformation is standard, but special considerations
+are needed.
+%issues need to be considered.
 \begin{itemize}
-\item {\bf Local primitive variables\ }  Some local variable is defined before the branch event $e_b$, of which the SSA form is already in the trace $\tau$, we reuse the SSA form.
-In addition, the SSA transformation may introduce Phi node, e.g., $y=Phi(y_1, C_1, y_2, C_2, \dots y_i, C_i)$, meaning that $y$ is equal to $y_i$ in the condition $C_i$.
-The combination of SSA forms of both branches may require further Phi node when both branches define a variable.
-We insert the Phi node to the trace and replace the following use accordingly.
-\item {\bf Local Heap Accesses\ } Given the heap access of $x.f$, remember that the base variable $x$ is fully resolved to a concrete object $o$, we replace it as a local variable $l_{o.f}$ before the SSA transformation and treat it as a local primitive variable during the SSA transformation.
-\item {\bf Shared Heap Accesses\ } The standard SSA transformation does not encode the shared heap accesses.
-Similar to Section~\ref{sec:relax1}, we introduce the special symbols to denote the shared read/write.
+\item {\bf Local primitive variables\ }  
+For local reads inside the extended branch whose definitions are
+%Some local variable is defined 
+before the branch event $e_b$, we reuse their SSA forms which 
+%of which the SSA form 
+are already in the trace $\tau$.
+%, we reuse the SSA form.  
+In addition, 
+%the SSA transformation 
+we encode conditional statements inside the extended branch using
+Phi node, e.g., $y=Phi(y_1, C_1, y_2, C_2, \dots y_i, C_i)$, 
+meaning that $y$ is equal to $y_i$ in the condition $C_i$.
+Note that we do not require paths inside the extended branch be fully
+resolved. 
+Phi nodes are also used to integrate the definitions in the extended
+branch with the executed branch.
+
+%both the original executed branch and 
+%the extended branch 
+%The combination of SSA forms of both branches may require further 
+%Phi node when both branches define a variable. We insert the Phi node 
+%to the trace and replace the following use accordingly.
+\item {\bf Local Heap Accesses\ } Given a heap access  $x.f$, 
+recall that the base variable $x$ is fully resolved to a concrete object
+ $o$, we replace it as a local variable $l_{o.f}$ before the SSA 
+transformation and treat it as a local primitive variable during the 
+SSA transformation.
+
+\item {\bf Shared Heap Accesses\ } The standard SSA transformation 
+does not encode the shared heap accesses. Similar to 
+Section~\ref{sec:relax1}, we introduce special symbols to 
+denote shared read/write.
 \end{itemize}
 
-The SSA form for the branch at line 4 is as follows.
+The SSA form for extending line 4 in Figure~\ref{fig:running}  is shown 
+in Figure \ref{fig:pathssa}. We omit the Phi nodes for $z$ and $w$ at 
+the end as they are never used later.
 
@@
 
-{\bf Constraints\ } During the traversal, we also record the shared access events.
-For the set of accesses $S_1$ of the location $\ell$, we also find the accesses $S_2$ of it from other threads.
-We assert the feasibility of each race pair $\in S_1 \times S_2$ by specifying constraints and checking them.
+{\bf Constraints\ } During traversal, we also record the shared access 
+events.  For the set of accesses $S_1$ of the location $\ell$, we also 
+find its accesses $S_2$ from other threads. We assert the feasibility 
+of each race pair $\in S_1 \times S_2$ by specifying constraints and 
+checking them.
 
-{\bf Race Condition Constraint\ }  The first type of constraint is the race condition constraint, e.g., $O'_{e_2}=O'(e_7)$ specifies that the event at line 2 may form a race with the event at line 7, which is actually not executed in the original trace.
+{\bf Race Condition Constraint\ }  The first type of constraint is the 
+race condition constraint, e.g., $O'(e_2)=O'(e_7)$ specifies that the event 
+at line 2 may form a race with the event at line 7, which is actually not 
+executed in the original trace. Importantly, we need to specify the race 
+condition under the specific branch condition, e.g., 
+$\theta'(R^4_y)\leq 2\rightarrow O'(e_2)=O'(e_7)$. The underlying reason is, 
+the race is valid, or $e_7$ can be executed,  only under the branch condition $\theta'(R^4_y)\leq 2$.
 
 {\bf Intra-thread Control Flow Constraint \ } Different from the strategy in Section~\ref{sec:relax1}, we do not assert the branch the predicted run will take.
-Instead, we assert that both branches are possible, which is always $true$.
+Instead, we assert that either branch is possible.
+%, which is always $true$. 
 The solver automatically reasons about which branch should be taken and encodes it in the solution. 
-Note that only one of the two branches can be taken because the conditions for the branches contradict.
+%Note that only one of the two branches can be taken because the conditions for the branches contradict.
 
-{\bf Intra-thread Value Constraint \ } Intra-thread value constraint is the core constraint that realizes the exploration different branches.
-It asserts the value constraints imposed by the {\sf assign/heapr/heapw} events take effect only when the guarding condition is satisfied.
-Suppose $IF(e_b)$ denotes the event sequence that corresponds to the true branch, $ELSE(e_b)$ denotes the event sequence that corresponds to the false branch, the branch event $e_b$ is  $if(x_b<y_b)$, the event $e_m$ in $IF(e_b)$ is in the form of $x_m=y_m+z_m$ and the event $e_n$ in $ELSE(e_b)$ is in the form of $x_n=y_n+z_n$.
+{\bf Intra-thread Value Constraint \ } Intra-thread value constraint is the 
+core constraint that realizes the exploration of different branches. It 
+asserts the value constraints imposed by the {\sf assign/heapr/heapw} events 
+take effect only when the guarding condition is satisfied. 
+Suppose $IF(e_b)$ denotes the event sequence that corresponds to the 
+true branch, $ELSE(e_b)$ denotes the event sequence that corresponds to 
+the false branch, the branch event $e_b$ is  $if(x_b<y_b)$, the event $e_m$ 
+in $IF(e_b)$ is in the form of $x_m=y_m+z_m$ and the 
+event $e_n$ in $ELSE(e_b)$ is in the form of $x_n=y_n+z_n$.
+
+{\bf [XZ: missing piece]}
 
@@
 
-{\bf Assignment with {\sf Phi}\ }  For the assignment with the Phi, e.g., $y=Phi(y_1, x>0, y_2, x\leq 0)$, the encoding is similar.
-$$
-\begin{array}{l}
-\theta'(x)>0 \rightarrow \theta'(y)=\theta'(y_1) \wedge \theta'(x)\leq 0 \rightarrow \theta'(y)=\theta'(y_2)
-\end{array}
-$$
+This is also how we encode Phi operations.
+%{\bf Assignment with {\sf Phi}\ }  For the assignment with the Phi, e.g., $y=Phi(y_1, x>0, y_2, x\leq 0)$, the encoding is similar.
+%$$
+%\begin{array}{l}
+%\theta'(x)>0 \rightarrow \theta'(y)=\theta'(y_1) \wedge \theta'(x)\leq 0 \rightarrow \theta'(y)=\theta'(y_2)
+%\end{array}
+%$$
 
-In case that the unexplored branch contains another branch structure, the above encoding process is applied recursively and the resultant constraints are nested.
+%In case that the unexplored branch contains another branch structure, the above encoding process is applied recursively and the resultant constraints are nested.
+
+{\bf Inter-thread value constraints\ } The inter-thread value constraints 
+are similarly extended. Details are elided.
 
@@
 
-{\bf Inter-thread value constraints\ } The inter-thread value constraints are identical to those in Section~\ref{sec:relax1}.
+%identical to those in Section~\ref{sec:relax1}.
 Interesting, we do not need to specify the value constraints under the specific branch condition.

### 2014-09-06 eval.tex

@@
 
-{\bf Prototype Implementation}  In our prototype implementation of \tool, we used the Z3 SMT solver for constraint resolution \cite{DeMoura:2008}.
-For static analysis, to discover unexecuted branches, we utilize the SOOT compiler framework \cite{Vallee-Rai:1999}.
+{\bf Prototype Implementation}  In our prototype implementation of \tool, we used the Z3 SMT solver for constraint resolution \cite{MouraB08}.
+For static analysis, to discover unexecuted branches, we utilize the SOOT compiler framework \cite{Vallee-RaiCGHLS99}.
 

### 2014-09-06 guarantee.tex
changes

@@
 \begin{mytheorem}[Soundness]
-The trace $\tau$ returned by our solver is a feasible trace for the concurrent program.
+The trace computed by the solver,  $\tau'=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \}, O', \theta'>$, is a feasible trace for the concurrent program, where $O'$ defines its schedule and $\theta'$ defines the value mapping.
 \end{mytheorem}
@@
 \begin{itemize}
-\item  $\tau$ is feasible.
-Researchers~\cite{chao,maximal} point out a trace is feasible iff it satisfies the sequential consistency, i.e., $\forall object\ o, \tau\downarrow_{o}$ satisfies the serial specification of the object $o$.
+\item  $\tau'$ is feasible.
+Researchers~\cite{chao,maximal} point out a trace is feasible iff it satisfies the sequential consistency, i.e., $\forall object\ o, \tau'\downarrow_{o}$ satisfies the serial specification of the object $o$.
 If  all the events considered are of the types in Table~\ref{Ta:syntax}, then the sequential consistency precisely means the following: (1) read-write consistency, i.e., each read event of a variable should contain the value written by the most recent write event, and (2) the synchronization consistency, e.g., the lock acquire and release of a lock should not be interleaved by the lock operations of the same lock, the begin event of a thread should follow the fork event of the parent thread.
@@
 The synchronization consistency is also captured as constraints in our solver.
-\item $\tau$ can be generated by the program, i.e., for each thread $t$, $\tau\downarrow_{t}$ can be generated by the corresponding thread code.
-This claim requires that,  for any two adjacent events $e$ and $e'$ in  $\tau\downarrow_{t}$, (1) if $e$ is not a branch, $inst^e$ and $inst^{e'}$ should be adjacent in the code; (2) if $e$ is a branch, $inst^e$ and $inst^{e'}$ should be adjacent and the $inst^e$ should be evaluated as the boolean value indicated by $inst^{e'}$.
+\item $\tau$ can be generated by the program, i.e., for each thread $t$, $\tau'\downarrow_{t}$ can be generated by the corresponding thread code.
+This claim requires that,  for any two adjacent events $e$ and $e'$ in  $\tau'\downarrow_{t}$, (1) if $e$ is not a branch, $inst^e$ and $inst^{e'}$ should be adjacent in the code; (2) if $e$ is a branch, $inst^e$ and $inst^{e'}$ should be adjacent and the $inst^e$ should be evaluated as the boolean value indicated by $inst^{e'}$.
 We derive the trace  through either concrete execution or symbolic execution, which satisfies the requirements and guarantees the correctness of the claim.
@@
 \begin{mydiscussion}[Detection Capability]
-Our technique has stronger detection capability than existing techniques~\cite{yannis,jeff,said}.
+Our technique has stronger detection capability than existing techniques~\cite{yannis,jeff, said}.
 \end{mydiscussion}
@@
 \centering
-\includegraphics[scale=0.7]{figs/Visio-discuss.pdf}
+\includegraphics[width=4cm,height=4cm]{figs/Visio-discuss.pdf}
 \caption{Illustration of Detection Capability.}\label{fig:discuss}

### 2014-09-06 overview.tex

@@
 
+The formuals from the different encoding steps are conjoined and sent to the off-the-shelf solver such as {\sf z3}.
+The race under analysis is real if the solver returns a solution, which includes the scheduling order and the values for the variables.
+
 {\bf Unexplored Branches\ } Beyond the encoding steps so far, which focus on the given trace, we can often encode constraints along unexplored branches.
@@
 Note that the race condition constraint associates the race with the branch condition that enables each event.
-Also we encode other necessary constraints such as the intra-thread order constraints, $O'(e_4)<O'(e_5)\wedge O'(e_4)<O'(e_7)$.
-The inter-thread
-
-{\bf Constraint Solving\ } The formuals from the different encoding steps are conjoined and sent to an off-the-shelf solver such as {\sf z3}.
-The race under analysis is guaranteed to be real if the solver returns a solution, which includes the scheduling order and the values for the variables.
+Also we need to encode (1) the intra-thread constraints, such as the  order constraints $O'(e_4)<O'(e_5)\wedge O'(e_4)<O'(e_7)$, and (2) the inter-thread constraints, which remain the same as before.
+With these constraints, the solver confirms the race pair, ($e_2$, $e_7$).
+More details are explained in Section~\ref{sec:relax2}.
 

### 2014-09-06 abstract.tex
abs

\begin{abstract}
Predictive analysis has recently emerged as a promising paradigm for sound detection of multithreading bugs (i.e., without false alarms). Intuitively, the idea is to start from an instrumented execution trace, and transform it based on the concrete information it provides (e.g., concrete values and memory accesses) while preserving feasibility (e.g., by reodering trace events iff flow dependencies are preserved).

Predictive analysis has been applied successfully to the problem of data-race detection, though state-of-the-art techniques are restrictive in the transformations they allow so as to guarantee feasibility and consequently also soundness. In particular, existing techniques (i) do not permit flow dependencies to be violated (except in very specific cases), and (ii) are unable to analyze execution paths beyond that of the input trace.

In this paper, we present \sysname, a predictive race detector that is able to relax both of these restrictions while guaranteeing soundness, as we prove formally. Thanks to its ability to explore event orderings that violate flow dependencies as well as unexecuted branches, \sysname\ is able to detect x$2.5$ more data races on a representative set of large-scale benchmarks compared to the state-of-the-art in predictive race detection.
\end{abstract}

### 2014-09-06 abstract.tex
abs

@@
 Predictive analysis has recently emerged as a promising paradigm for sound detection of multithreading bugs (i.e., without false alarms).
-Intuitively, the idea is to start from an instrumented execution trace, and transform it based on the concrete information it provides (e.g., concrete values and memory accesses) while preserving feasibility (e.g., by reodering trace events iff flow dependencies are preserved).
+The idea is to start from an instrumented execution trace, and transform it based on the concrete information it provides (e.g., concrete values and memory accesses) while preserving feasibility (e.g., by reodering trace events iff flow dependencies are preserved), such that bugs are exposed.
 
@@
 In this paper, we present \sysname, a predictive race detector that is able to relax both of these restrictions while guaranteeing soundness, as we prove formally.
-Thanks to its ability to explore event orderings that violate flow dependencies as well as unexecuted branches, \sysname\ is able to detect x$2.5$ more data races on a representative set of large-scale benchmarks compared to the state-of-the-art in predictive race detection.
+Thanks to its ability to explore event orderings that violate flow dependencies as well as unexecuted branches, \sysname\ is able to detect x$2.5$ more data races on a representative set of large-scale benchmarks compared to the state of the art in predictive race detection.
 \end{abstract}

### 2014-09-06 abstract.tex
abs

@@
 In this paper, we present \sysname, a predictive race detector that is able to relax both of these restrictions while guaranteeing soundness, as we prove formally.
-Thanks to its ability to explore event orderings that violate flow dependencies as well as unexecuted branches, \sysname\ is able to detect x$2.5$ more data races on a representative set of large-scale benchmarks compared to the state of the art in predictive race detection.
+Thanks to its ability to explore unexecuted branches as well as event orderings that violate flow dependencies, \sysname\ is able to detect x$2.5$ more data races on a representative set of large-scale benchmarks compared to the state of the art in predictive race detection.
 \end{abstract}

### 2014-09-06 overview.tex
-xz

@@
 \subsection{Preliminaries}
-We first introduce some notation used throughout this paper.
-Intuitively, a trace is a sequence of events recorded during the observed run.
-An event, $e=<t, id, inst>$, is a concrete representation of the runtime execution of a static instruction $inst$, such that (i) $t$ refers to the thread that issues the event $e$ (denoted $t^e$); (ii) $id$ is a unique identifier of the event (unless spcified otherwise, we use the index of an event in the trace as its id); and (iii) $inst$ is a static three-address instruction (involving at most three operands, which modern compilers commonly support).
+We first introduce some notation used throughout this paper. Intuitively, a 
+trace is a sequence of events recorded during the observed run. An 
+event, $e=<t, id, inst>$, is a concrete representation of the runtime 
+execution of a static instruction $inst$, such that (i) $t$ refers to the 
+thread; (ii) $id$ is a unique identifier of the event (unless specified 
+otherwise, we use the index of an event in the trace as its id); 
+and (iii) $inst$ is a static three-address instruction (involving at most three operands).
 
-For this paper, we utilize the instructions listed in Table~\ref{Ta:syntax}.
-When the variable does not appear on the left hand of an equation, such as $y$ in $x.f=y$, it may refer to a variable, a constant or event object creation expression $new (...)$.
-The $bop$ stands for the binary operator, which may refer to $+, -, *, /, \%, \wedge, \vee$ in the assignment, or refer to $<, >, =, \wedge, \vee$ in the branch.
-The target of the branch event is not important in our scope, therefore, we may abbreviate the branch instruction as the boolean expression afterwards.
-The listed instructions suffice to represent all trace events of interest.
-This is because a concrete finite execution trace can be reduced to a straight-line loop-free call-free path program  (argument passing of the method call is modeled as assignments and the virtual call resolution is modeled as branches).
-This standard form of simplification preserves all the data-race-related information contained in the original trace.
+For this paper, we utilize the instructions listed in 
+Table~\ref{Ta:syntax}. When the variable does not appear on the left hand 
+side of an equation, such as $y$ in $x.f=y$, it may refer to a variable, a 
+constant or an object creation expression $new (...)$.  Symbol $bop$ stands 
+for a binary operator, which may refer to $+, -, *, /, \%, \wedge, \vee$ in 
+an assignment, or  $<, >, =, \wedge, \vee$ in a branch instruction. 
+The target of the branch instruction is not important in our scope, 
+therefore, we may abbreviate a branch instruction as its boolean expression 
+afterwards. The listed instructions suffice to represent all trace events 
+of interest. 
+Note that 
+%This is because 
+a concrete finite execution trace can be reduced to a straight-line 
+loop-free call-free path program  (argument passing of a method call 
+is modeled as assignments and virtual call resolutions are modeled as 
+branches). 
+%This standard form of simplification preserves all the data-race-related information contained in the original trace.
 
@@
 	\caption{\label{Ta:syntax}Instruction Types}
+\vspace{-0.15in}
 \end{table}
 
-{\bf Trace Representation \ } In the following, we assume that trace $\tau$ is represented as $\tau=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \}, O, \theta>$, where
+{\bf Trace Representation \ } A trace $\tau$ is represented as $\tau=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \}, O, \theta>$, where
 \begin{itemize}
@@
 \end{itemize}
-To ensure the integrity of $\theta$ as well simplify downstream encoding steps performed by the analysis already at the syntactic level, we rewrite the raw input trace into SSA form, such that each variable is defined exactly once.
+To ensure the integrity of $\theta$ as well simplify downstream encoding 
+steps,
+% performed by the analysis already at the syntactic level, 
+we rewrite the raw input trace into SSA form, such that each variable is defined exactly once.
 
@@
 \subsection{Constraint System}
-The predictive analysis takes the program and a trace as the input.
-Suppose it starts with the following trace of the program in  Figure \ref{fig:running}:  $e_1 e_2 e_3 e_4 e_5$, where the line number is used as the event id (subscript).
-A potential race is between the pair, $(e_2, e_5)$, as they access the shared variable $x$ from different threads and one is a write.
+In this section, we are going to informally explain the two relaxations
+we make which allow us to catch the two races missed by existing 
+predictive analyses.
+Our analysis takes the program and a trace as the input.  
+Suppose it starts with the following trace of the program in  
+Figure \ref{fig:running}:  $e_1 e_2 e_3 e_4 e_5$, where the line number 
+is used as the event id (subscript).
 
-Given the trace, $\tau=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \}, O, \theta>$, the predictive analysis computes a new trace,  $\tau'=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \}, O', \theta'>$.
-$\tau'$ preserves the same event sequence for each thread, i.e., $T_1$ still executes $e_1 e_2 e_3$ and $T_2$ still executes $e_4 e_5$, but it reschedules the events from different threads so that the potential race pair runs concurrently, i.e., $e_2$ and $e_5$ run concurrently.
-When the schedule changes, the variables may get different values.
-Therefore, we need to compute both the new schedule $O'$ and the new value mapping $\theta'$, in order to witness a race.
-To guarantee that $\tau'$ is a feasible trace of the program $P$, the computation needs to respect a set of constraints, which include the value constraints, control flow constraints and synchronization constraints~\footnote{We omit the synchronization constraints in this paper, which are already well explained in existing techniques~\cite{yannis, pecan}.}.
-We explain them briefly in the following.
-The full details are explained in Section~\ref{sec:relax1} and the feasibility guarantee is explained in Section~\ref{sec:guarantee}.
+\smallskip
+\noindent
+{\em I. Value Relaxation\ }
+This allows us to catch the racy pair $(e_2, e_5)$.
+%A potential race is between 
+%the pair, $(e_2, e_5)$, as they access the shared variable $x$ from 
+%different threads and one is a write. 
+Given the trace, $\tau=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \},
+ O, \theta>$, our analysis computes a new trace,  
+$\tau'=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \}, O', \theta'>$.
+ $\tau'$ preserves the same event sequence for each thread, i.e., $T_1$ still 
+executes $e_1 e_2 e_3$ and $T_2$ still executes $e_4 e_5$, but it reschedules 
+the events from different threads so that the potential race pair runs 
+concurrently, i.e., $e_2$ and $e_5$ run concurrently. When the schedule 
+changes, the variables may get different values. Therefore, we need to 
+compute both the new schedule $O'$ and the new value mapping $\theta'$, in 
+order to witness a race. To guarantee that $\tau'$ is a feasible trace of 
+the program $P$, the computation needs to respect a set of constraints, 
+which include the value constraints, control flow constraints and 
+synchronization constraints~\footnote{We omit the synchronization constraints 
+in this paper, which are already well explained in existing 
+techniques~\cite{yannis, pecan}.}. We explain the constraints 
+briefly in the following. Details are in Section~\ref{sec:relax1} 
+and the soundness guarantee is explained in Section~\ref{sec:guarantee}.
 
@@
 
-We first preprocess the trace as follows.
-We introduce symbols to replace the shared accesses in the instructions, e.g., We use $W^{id}_{x}$/$R^{id}_{x}$ to replace the write/read of the shared variable $x$ by the event $e_{id}$.
-%The rationale is explained in Section~\ref{sec:relax1}.
+We first preprocess the trace as follows. We introduce symbols to replace 
+the shared accesses in the instructions, e.g., We use 
+$W^{id}_{x}$/$R^{id}_{x}$ to replace the write/read of the shared 
+variable $x$ by the 
+event $e_{id}$. %The rationale is explained in Section~\ref{sec:relax1}.
 
-{\bf Race Condition\ } The potential race pair involves two accesses of the same shared location from different threads, where at least one access is a write.
-Given any potential race pair, e.g., ($e_2$, $e_5$), it is a real race if and only if the two accesses can occur at the same time (race condition), which is captured by the race condition constraint:
-$O'(e_2) = O'(e_5)$.
-The race condition constraint--- together with the other constraints --- guarantees the feasibility of the predicted race if a solution is found for the overall constraint system.
+{\bf Race Condition\ } A potential race pair involves two accesses of 
+the same shared location from different threads, where at least one is write. 
+Given any potential race pair, e.g., ($e_2$, $e_5$), it is a real race 
+if and only if the two accesses can occur at the same time (race condition), 
+which is captured by the race condition constraint: $O'(e_2) = O'(e_5)$.
+The race condition constraint--- together with the other constraints --- 
+guarantees the feasibility of the predicted race if a solution is found 
+for the overall constraint system.
 
-{\bf Intra-thread Constraints\ } The predictive analysis needs to preserve the same event sequence for each thread, which further requires
+{\bf Intra-thread Constraints\ } Our analysis needs to preserve the 
+same event sequence for each thread, which requires
 
 \begin{itemize}
-\item {\bf Control Flow Constraints\ } Each thread takes the same control flows (or branch decisions) to reproduce the events.
- For the running example (ignoring the statements in red), this yields: $\theta'(R_{\tt y}^4) > 2 \equiv {\bf true}$.
-That is, in the predicted trace $\tau'$, the value of variable {\tt y} read at $e_4$ should be greater than $2$.
+\item {\bf Control Flow Constraints\ } Each thread takes the same control flow
+(or branch decisions) to reproduce the events.
+ For the running example (ignoring the statements in red), this 
+yields: $\theta'(R_{\tt y}^4) > 2 \equiv {\bf true}$.
+That is, in the predicted trace $\tau'$, the value of variable {\tt y} 
+read at $e_4$ should be greater than $2$.
 
-\item {\bf Intra-thread Order Constraints\ } The events should follow the same thread-local order as in the original trace.
-This constraint is imposed by the fact that the two runs share the same instruction sequence and take the same control flows.
+\item {\bf Intra-thread Order Constraints\ } The events should follow 
+the same thread-local order as in the original trace. This constraint 
+is imposed by the fact that the two runs share the same instruction 
+sequence and take the same control flow. 
 We obtain the  formula: $O'(e_1) < O'(e_2) < O'(e_3) \wedge O'(e_4) < O'(e_5)$.
-That is, in the predicted trace $\tau'$, the three events from $T_1$ are ordered in the same way as in $\tau$, and similarly, the branch event $e_4$ in $T_2$ still executes before the event $e_5$ inside the branch body.
+%That is, in the predicted trace $\tau'$, the three events from $T_1$ are 
+%ordered in the same way as in $\tau$, and similarly, the branch 
+%event $e_4$ in $T_2$ still executes before event $e_5$ inside the 
+%branch body.
 
-\item {\bf Intra-thread Value Constraints\ } The events should respect the value constraints imposed by each instruction.
+\item {\bf Intra-thread Value Constraints\ } The events should respect the 
+constraints imposed by each instruction.
  For our running example, we obtain:
@@
 
-{\bf Inter-thread Value Constraints\ } The inter-thread value constraints capture what writes the read events read from and under what scheduling condition the value flow occurs.
+{\bf Inter-thread Value Constraints\ } These constraints capture what writes the read events read from and under what scheduling condition the value flow occurs.
 The resulting formula for our example is
@@
 $$    
-Notice, importantly, that the formula associates the value constraints with the scheduling order constraints. 
+Notice, importantly, that the formula associates the value constraints 
+with the scheduling order constraints. 
 As an example, 
-$\theta'(R_{\tt y}^4)=\theta'(W_{\tt y}^1) \wedge O'(e_1) < O'(e_4) < O'(e_3)$ means that, in the predicted run, the read $e_4$ reads from the write $e_1$, under the condition that $e_4$ happens after $e_1$ and no other  writes (such as $e_3$) interleave them.
+$\theta'(R_{\tt y}^4)=\theta'(W_{\tt y}^1) \wedge O'(e_1) < O'(e_4) < O'(e_3)$ 
+means that, in the predicted run, the read $e_4$ reads from the write $e_1$, 
+under the condition that $e_4$ happens after $e_1$ and no other  
+writes (such as $e_3$) interleave them.
 
-The formuals from the different encoding steps are conjoined and sent to the off-the-shelf solver such as {\sf z3}.
-The race under analysis is real if the solver returns a solution, which includes the scheduling order and the values for the variables.
+The formulas from the different encoding steps are conjoined and sent to an 
+off-the-shelf solver such as {\sf z3}.  The race under analysis is real 
+if the solver returns a solution, which includes the scheduling order 
+and the values for the variables.
 
-{\bf Unexplored Branches\ } Beyond the encoding steps so far, which focus on the given trace, we can often encode constraints along unexplored branches.
-In our running example, this is essential to discover the race between lines {\tt 2} and {\tt 7}.
-We explore the unexecuted branch statically and collect constraints during the exploration.
-We then form a disjunction  of the original constraints and the collected constraints to include more possible  paths.
-The solver may pick up the unexplored branch to detect a race whenever it is necessary.
-Back to our example, the event $e_7$ is in the unexplored branch, our solver is able to identify the race between $e_7$ and $e_2$ after encoding the unexplored branch.
+\smallskip
+\noindent
+{\em II. Path Relaxation\ } 
+%Next, we show how we catch the second missing race
+%($e_2$, $e_7$) with another relaxation. 
+Besides the first relaxation, which focuses on the given trace, our
+technique also encodes constraints along unexplored branches. In our 
+running example, this is essential to discovering the race between 
+lines {\tt 2} and {\tt 7}. We explore the unexecuted branch statically 
+and collect constraints during the exploration. We then form a disjunction  
+of the original constraints and the collected constraints to include more 
+possible  paths. The solver may pick up the unexplored branch to detect 
+a race. Back to our example, event $e_7$ is in the unexplored branch, \tool
+is able to identify the race between $e_7$ and $e_2$ after encoding 
+the unexplored branch.
 
-For the racy events, $e_2$ and $e_7$, we first specify the race condition constraint  $\theta'(R^4_y)\leq 2\rightarrow O'_{e_2}=O'(e_7)$.
-Note that the race condition constraint associates the race with the branch condition that enables each event.
-Also we need to encode (1) the intra-thread constraints, such as the  order constraints $O'(e_4)<O'(e_5)\wedge O'(e_4)<O'(e_7)$, and (2) the inter-thread constraints, which remain the same as before.
+For the racy events, $e_2$ and $e_7$, we first specify the race condition 
+constraint  $\theta'(R^4_y)\leq 2\rightarrow O'_{e_2}=O'(e_7)$. Note that 
+the constraint associates the race with the branch condition that 
+enables each event.  Also we need to encode (1) the intra-thread constraints, 
+such as the  order constraints $O'(e_4)<O'(e_5)\wedge O'(e_4)<O'(e_7)$, 
+and (2) the inter-thread constraints, which remain the same as before. 
 With these constraints, the solver confirms the race pair, ($e_2$, $e_7$). 
-More details are explained in Section~\ref{sec:relax2}.
+More details are in Section~\ref{sec:relax2}.
 
-{\bf Highlights \ } In this example, our analysis finds two more pairs of races beyond existing analyses~\cite{yannis, jeff}: $(e_2, e_5)$ and $(e_2, e_7)$.
-For the first race, we do not require $e_4$ to read from $e_3$ to retain the original value, as in approaches.
-Rather, we allow $e_4$ to read from $e_1$ while still preserving the feasibility of the branch.
-This relaxation allows the events after $e_4$ to run concurrently with the events before $e_3$, leading to the detection of the first race.
-For the second race, we cover unexecuted branches and ask the solver to compute a feasible schedule that exercises them.
-The neighboring branches are likely to contain racy events if the executed branches contain racy events.
+\smallskip
+\noindent
+{\em Highlights \ } In this example, \tool finds two more pairs of races 
+beyond existing analyses~\cite{yannis, jeff}: $(e_2, e_5)$ and $(e_2, e_7)$.
+For the first race, we do not require $e_4$ to read from $e_3$ to 
+retain the original value (i.e. the true branch outcome), as in the
+existing approaches. Rather, we allow $e_4$ to read from $e_1$ while 
+still preserving the feasibility of the path. This relaxation allows 
+the events after $e_4$ to run concurrently with the events before 
+$e_3$, leading to the detection of the first race.  For the second race, 
+we cover unexecuted branches and query the solver to compute a feasible 
+schedule that exercises them. The neighboring branches are likely to 
+contain racy events if the executed branches contain racy events.
 

### 2014-09-06 eval.tex

@@
 This is because our approach needs to reason about the computation among variables inside the local access events, while {\sf RV} needs to only reason about the order relations among the events.
-
-To summarize, while our approach is able to relax schedules and paths based on  fine-grained reasoning at the level of data values, it also imposes higher overhead.
-An interesting topic for future research is how  to combine our approach with a lightweight approach with soundness guarantees like {\sf RV}.
-One option is staged analysis, wherein {\sf RV} is run.
-\tool\ is then left to run only on racy pairs not confirmed by {\sf RV}.
-In this way, we can also tackle misses due to practical limitations of the constraint solver.
 

### 2014-09-06 related.tex
related

@@
 
+Serbanuta et al.
+\cite{maximal} have recently shown that it is possible to build a model that is not only sound but also maximal: Any extension of the model with a new trace renders the model unsound.
+The theoretical framework created by Serbanuta et al.
+provides a foundation for different forms of PTA, including data races, but also atomicity, serializability and other properties.
+Serbanuta et al.
+do not report on any concrete instantiation of their framework.
+
 The first to propose PTA are Smagardakis et al.
@@
 Said et al.
-\cite{Said:2011} describe a PTA that is also sound albeit not maximal.
-Similarly to our analysis, Said et al.
+\cite{Said:2011} and Huang et al.
+\cite{HuangMR14} describe PTAs that are also sound albeit not maximal.
+Said et al.
 perform symbolic analysis of the input trace, and then utilize an SMT solver to search for interleaved schedules that establish the presence of data races.
 Soundsness is guaranteed by their ability to precisely encode the semantics of sequential consistency.
+Inspired by Said et al., Huang et al.
+present an improvement, whereby control-flow information is encoded into the constraint system so as to relax flow dependencies as long as control dependencies are preserved.
 ExceptionNULL \cite{Farzan:2012} is another example of a sound PTA without maximality guarantees, where the goal is to detect null dereferences rather than data races.
 
-Serbanuta et al.
-\cite{SerbanutaCR12} have recently shown that it is possible to build a model that is not only sound but also maximal: Any extension of the model with a new trace renders the model unsound.
-The theoretical framework created by Serbanuta et al.
-provides a foundation for different forms of PTA, including data races, but also atomicity, serializability and other properties.
-
-Inspired by this result and closer to \tool\ is the PTA technique of Huang et al.
-\cite{HuangMR14}, which --- like our technique --- is sound and also maximal, but only under the assumption that local assignments are not modeled.
-Huang et al.s technique, too, makes use of an SMT solver based on symbolic encoding of the input trace.
-As part of the encoding, control-flow information is taken into account to enable reorderings that do not violate control dependencies.
-Still, the two relaxations that \tool\ features, which enable (i) value- rather than dependence-based reasoning about data flow and (ii) consideration of unexplored branches, are strictly outside the scope of Huang et al.'s technique.
+\tool\ is similar to Said et al.
+and Huang et al.
+in that it also makes use of an SMT solver based on symbolic encoding of the input trace.
+Still, the two relaxations that \tool\ features, which enable (i) value- rather than dependence-based reasoning about data flow and (ii) consideration of unexplored branches, are strictly outside the scope of both of these techniques.
 In Section \ref{sec:eval}, we demonstrate via direct comparison with Huang et al.

### 2014-09-06 overview.tex

@@
 \noindent
-{\em I. Value Relaxation\ }
+{\em I. Value Relaxation\ } Our first relaxation is to allow variables
+to have different values as long as the same thread local path is preserved.
+It does not require data dependences or variable values as in~\cite{...}.
 This allows us to catch the racy pair $(e_2, e_5)$.
@@
 %different threads and one is a write. 
-Given the trace, $\tau=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \},
+Given a trace, $\tau=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \},
  O, \theta>$, our analysis computes a new trace,  
@@
 %($e_2$, $e_7$) with another relaxation. 
-Besides the first relaxation, which focuses on the given trace, our
-technique also encodes constraints along unexplored branches. In our 
+Besides the first relaxation, which focuses on the given trace, the second
+relaxation is to even reason about unexecuted branches as long as they
+can be precisely modeled. In our 
 running example, this is essential to discovering the race between 
@@
 \noindent
-{\em Highlights \ } In this example, \tool finds two more pairs of races 
+{\em Highlights \ } \tool finds two more pairs of races 
 beyond existing analyses~\cite{yannis, jeff}: $(e_2, e_5)$ and $(e_2, e_7)$.
 For the first race, we do not require $e_4$ to read from $e_3$ to 
-retain the original value (i.e. the true branch outcome), as in the
+retain the original value as in the
 existing approaches. Rather, we allow $e_4$ to read from $e_1$ while 
-still preserving the feasibility of the path. This relaxation allows 
+still preserving the thread local path. This relaxation allows 
 the events after $e_4$ to run concurrently with the events before 
@@
 we cover unexecuted branches and query the solver to compute a feasible 
-schedule that exercises them. The neighboring branches are likely to 
-contain racy events if the executed branches contain racy events.
+schedule that exercises them. 
+%The neighboring branches are likely to 
+%contain racy events if the executed branches contain racy events.
 

### 2014-09-06 related.tex
related

@@
 
-\subsection{Predictive Trace Analysis}
+\subsection{Predictive Trace Analysis (PTA)}
 
-Given concurrent execution trace $t$, a \emph{maximal and sound causal model} based on $t$ defines the set of all traces that a program that is able to generate $t$ can generate (maximality), and only those traces (soundness).
+Given concurrent execution trace $t$, PTA derives new traces to witness the data races.
 PTA is founded on the notion of sound causality, as it considers feasible reorderings of the input trace that prove a candidate data race as such.
-
-Serbanuta et al.
-\cite{maximal} have recently shown that it is possible to build a model that is not only sound but also maximal: Any extension of the model with a new trace renders the model unsound.
-The theoretical framework created by Serbanuta et al.
-provides a foundation for different forms of PTA, including data races, but also atomicity, serializability and other properties.
-Serbanuta et al.
-do not report on any concrete instantiation of their framework.
 
@@
 \cite{yannis}.
-In their original work, both synchronization constraints and inter-thread dependencies are preserved, where inter-thread dependencies are respected by
-only allowing reorderings that leave the dependence structure exhibited by the original trace in tact.
-This ensures that the values of shared memory locations remain the same, which secures the soundness argument, though maximality is not guaranteed.
+In their original work, both synchronization constraints and inter-thread dependencies are preserved, where inter-thread dependencies are respected by only allowing reorderings that leave the dependence structure exhibited by the original trace in tact.
+This ensures that the values of shared memory locations remain the same, which secures the soundness argument.
 
@@
 \cite{Said:2011} and Huang et al.
-\cite{HuangMR14} describe PTAs that are also sound albeit not maximal.
+\cite{HuangMR14} describe PTAs that are also sound.
 Said et al.
 perform symbolic analysis of the input trace, and then utilize an SMT solver to search for interleaved schedules that establish the presence of data races.
-Soundsness is guaranteed by their ability to precisely encode the semantics of sequential consistency.
+Soundness is guaranteed by their ability to precisely encode the semantics of sequential consistency.
 Inspired by Said et al., Huang et al.
 present an improvement, whereby control-flow information is encoded into the constraint system so as to relax flow dependencies as long as control dependencies are preserved.
-ExceptionNULL \cite{Farzan:2012} is another example of a sound PTA without maximality guarantees, where the goal is to detect null dereferences rather than data races.
+ExceptionNULL \cite{Farzan:2012} is another example of a sound PTA, of which the goal is to detect null dereferences rather than data races.
 
@@
 the dramatic improvement in coverage thanks to these relaxation methods, which we prove to handle in a sound manner.
+
+\subsection{Maximality Claim}
+Serbanuta et al. \cite{maximal} and Huang et al.~\cite{HuangMR14} both claim maximality of the detection capability, but under different assumptions. 
+Assuming the branch event is not modeled in the trace, Serbanuta et al.'s approach \cite{maximal} guarantees maximality.
+Given few information in the trace, it conservatively requires the shared read to obtain the same value in the predicted run to ensure soundness.
+However, if the branch event is modeled, the technique does not guarantee maximality.
+Assuming the branch events are modeled but local assignments are not recorded in the trace, Huang et al.'s approach~\cite{HuangMR14} guarantees the maximality.
+Compared to Serbanuta et al.'s technique \cite{maximal}, it allows some events to read different values if they do not affect the following branch decisions.
+However, it requires the events that affect the branch decisions to read exactly the same values.
+In practice, these assumptions are too restrictive  because the testers often have full access to the source/bytecode and can collect the local assignments freely.
+
+Assuming the local assignments are available in the trace, both techniques do not guarantee maximality.
+As illustrated in our example and discussed in Section~\ref{sec:guarantee}, our analysis allows higher detection capability, i.e., we allow the events to read different values even they flow into the branch events; we allow the exploration of un-executed paths, both of which are not allowed by the above techniques.

### 2014-09-06 relax1.tex
-xz

@@
 \section{Relaxation of Flow Dependencies}~\label{sec:relax1}
-Given the trace, $\tau=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \}, O, \theta>$, the main goal of our analysis is to derive a new trace,  $\tau'=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \}, O', \theta'>$, which reschedules the events from different threads (while preserving intra-thread event order) to prove the race.
-The key insight underlying the first relaxation is that a thread is allowed to read a different value, potentially from a different write, as long as branching decisions remain unchanged.
-Note that we also need to compute the new value mapping $\theta'$, because the rescheduling may alter the value every variable is assigned to.
-The computation of the new schedule and new value mapping is explained in the following.
+Our first relaxation is to allow variables to have different values 
+as long as thread local paths are preserved.
+Given a trace, $\tau=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \}, 
+O, \theta>$, the main goal of our analysis is to derive a new trace,
+  $\tau'=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \}, O', \theta'>$,
+ which reschedules the events from different threads (while preserving 
+intra-thread event order) to prove the race. The key insight underlying 
+this relaxation is that a thread is allowed to read a different value, 
+potentially from a different write, as long as branching decisions 
+remain unchanged. Note that we also need to compute the new value 
+mapping $\theta'$, because the rescheduling may alter the value every 
+variable is assigned to. The computation of the new schedule and new value 
+mapping is explained in the following.
 
@@
 \subsection{SSA Form of the Trace}~\label{sec:ssa}
-First of all, we rewrite the trace into SSA form, such that every variable in defined exactly once in the trace.
-This requirement is a prerequisite for the computation of the mapping, in which each variable is mapped to exactly one value.
-Another side effect is that the def/use chains become explicit in  SSA form, which simplifies the following analysis steps.
+First of all, we rewrite the trace into SSA form, such that every variable is
+defined exactly once in the trace. This requirement is a prerequisite for 
+the computation of the mapping, in which each variable is mapped to exactly 
+one value. Another side effect is that the def/use chains become explicit 
+in SSA form, which simplifies the following analysis steps.
 
-\tool\ handles the local assignment and heap accesses differently.
+\tool\ handles local assignments and heap accesses differently.
 \begin{itemize}
-\item {\bf Local Assignment\ } The SSA form for local assignment resembles the SSA form of static instructions in compiler optimization, except that the looping and recursion are fully resolved in a concrete trace, obviating the need for {\sf Phi} nodes.
-More concretely, we replace the variable $v$ defined in an event, as well as the following uses of the definition, to a new variable $v^{id}$, where $id$ is the unique id of the event.
-The uniqueness of the id guarantees that no two events define the same variable, i.e., each variable is defined exactly once.
-In the example in Figure~\ref{fig:running2}, for example,
+\item {\bf Local Assignment\ } The SSA form for local assignments 
+resembles the static SSA form in compiler optimizations, 
+except that loops and recursions are fully resolved in the trace, 
+obviating the need for {\sf Phi} nodes\footnote{While \tool in fact uses
+Phi nodes in the second relaxation, for simplicity during discussion, we 
+do not assume Phi nodes in this stage.} . More concretely, we replace the 
+variable $v$ defined in an event, as well as the following uses of the 
+definition, to a new variable $v^{id}$, where $id$ is the unique id 
+of the event. The uniqueness of the id guarantees that no two events 
+define the same variable, i.e., each variable is defined exactly once.  
+In Figure~\ref{fig:running2}, for example,
 the local variable $x$ defined at line 4 is renamed as $x^4$.
@@
 Internally,  the object $o3$ is encoded as a unique integer that indicates its heap address.
-\item {\bf Heap accesses\ } The def/use relation between local heap accesses can be determined locally.
-Therefore, we treat them in the same way as local assignment.
-The def/use relations between shared heap accesses are more complex.
-We may link a read with different writes under different schedules, e.g., the read access to {\tt y} at line 4 in Figure \ref{fig:running} can either obtain the definition before line 1 or at line 1 or at line 3.
-Therefore, we introduce two symbols to represent the shared read and shared write respectively and use them to establish different links flexibly.
+\item {\bf Heap accesses\ } The def/use relation between local heap accesses 
+can be determined locally. Therefore, we treat them in the same way as 
+local assignments. The def/use relations between shared heap accesses 
+are more complex. We may link a read with different writes under different 
+schedules, e.g., the read of {\tt y} at line 4 in Figure \ref{fig:running} can 
+be from before line 1, line 1, or line 3. Therefore, we introduce two symbols to represent the shared read and shared write respectively and use them 
+to reason about the possible options.
+%establish different links flexibly. 
 The details are as follows. 
 \begin{itemize}
-\item {\bf Local Heap Accesses\ } Given the write $x.f=y$ to a local heap location at event $e_{id}$,  we introduce a fresh local variable {\tt $l^{id}_{o.f}$} ($o$ is the runtime object referenced by $x$) to replace $x.f$.
-We also replace the uses of the location $o.f$ accordingly.
-Here  the subscript $o.f$ is interpreted as a heap location, i.e., the field $f$ of the object $o$.
-Consider the example in Figure~\ref{fig:running2}.
-At lines 6-7, we replace the local heap accesses as the local variable $l^6_{o3.f}$.
+\item {\bf Local Heap Accesses\ } Given the write $x.f=y$ to a local heap 
+location at event $e_{id}$,  we introduce a fresh local variable 
+{\tt $l^{id}_{o.f}$} ($o$ is the runtime object referenced by $x$) to 
+replace $x.f$. We also replace the uses of the location $o.f$ accordingly.
+Here  the subscript $o.f$ is interpreted as a heap location, i.e., the 
+field $f$ of the object $o$. Consider the example in 
+Figure~\ref{fig:running2}. At lines 6-7, we replace the local heap 
+accesses as the local variable $l^6_{o3.f}$.
 %An assumption here is the uses and the definition still refer to the same base object $o$ in the predicted run, which is enforced through the constraint specification %(Section~\ref{sec:relax1}).
@@
 Similarly, the subscript $o.f$ is interpreted as a heap location, i.e., the field $f$ of the object $o$.
+Note that in local heap encoding,
+both read and write are bounded to the same variable, whereas in shared
+heap encoding, we use different variables to denote read and write, allowing
+the solver to reason about the different dataflow.
 \end{itemize}
-We use the heap location such as $o.f$ in the the symbolic form of heap accesses.
-An underlying assumption is that the heap location remains unchanged in our predictive analysis, which we ensure through additional constraints (Section~\ref{sec:relax1}).
+We use the heap location such as $o.f$ in the symbolic form of heap accesses.
+ An underlying assumption is that the heap location remains unchanged, 
+which we ensure through additional constraints (Section~\ref{sec:relax1}).
 \end{itemize}
@@
 \begin{figure}
+{\small
 \centering
@@
 \label{fig:running2}
+}
 \end{figure}
@@
 \subsection{Constraint System}~\label{sec:constraints}
-Based on the SSA form of the original trace $\tau$, we build the constraints to compute a new trace $\tau'$ with the new schedule $O'$ and new value mapping $\theta'$.
-Consider the example in Figure~\ref{fig:running2}.
-In the original trace $\tau$, the event $e_8$ reads the value of 2 from $e_3$.
-Therefore, $e_{11}$ must happen after $e_2$.
-Existing approaches enforce the same value dependence between $e_3$ and $e_8$, thereby missing the race between $e_2$ and $e_{11}$.
-Our analysis allows event $e_8$ to read from $e_1$, producing the new value mapping $\theta'(y^8)=3$.
-Accordingly, at $e_9$ the value mapping for $x^9$ is also updated as 13 ($\theta'(y^8)$=3, $\theta'(z^7)=10$), enabling the true branch at line 10.
-Finally, the events $e_2$ and $e_{11}$ are scheduled to run concurrently.
-The relaxation needs to respect a set of constraints, which we explain in the following.
-We omit the synchronization constraints intentionally, as they are well explained in all existing predictive analysis techniques~\cite{yannis, pldi14}.
+Based on the SSA form of trace $\tau$, we build the constraints to 
+compute a new trace $\tau'$ with a new schedule $O'$ and a new 
+value mapping $\theta'$. Consider the example in Figure~\ref{fig:running2}. 
+In the original trace $\tau$, the event $e_8$ reads the value of 
+2 from $e_3$.  Existing 
+approaches enforce the same value dependence between $e_3$ and $e_8$.
+ Therefore, $e_{11}$ must happen after $e_2$ and the race between 
+them is missed.
+Our analysis allows event $e_8$ to read from $e_1$, producing the new 
+value mapping $\theta'(y^8)=3$. Accordingly, at $e_9$ the value mapping 
+for $x^9$ is also updated as 13 ($\theta'(y^8)$=3, $\theta'(z^7)=10$), 
+enabling the same true branch at line 10. Finally, the 
+events $e_2$ and $e_{11}$ are scheduled to run concurrently.  The 
+relaxation needs to respect a set of constraints, which we explain in 
+the following. We omit the synchronization constraints as they are well 
+explained in~\cite{yannis, pldi14}.
 
@@
 
-{\bf Race Condition\ } Following the standard definition, pair ($e_i$, $e_j$) of events forms a race iff (1) $e_i$ and $e_j$ access a mutual  location $\ell$, (2) at least one of them writes $\ell$, and (3) $e_i$ and $e_j$ run concurrently (being associated with different threads).
-We refer to the candidate pair throughout this section as (candidate) \emph{racy events}.
+{\bf Race Condition\ } An even pair 
+($e_i$, $e_j$) forms a race iff (1) $e_i$ and $e_j$ access a shared
+location $\ell$, (2) at least one of them writes $\ell$, and (3) $e_i$ and $e_j$ run concurrently (being associated with different threads). We refer to 
+the candidate pair throughout this section as (candidate) \emph{racy events}.
 
-We first identify all candidate pairs of events that access the same location from different threads (and at least one is a write).
-We then check each pair  separately.
-The checking is achieved by encoding all necessary constraints.
-The first constraint asserts the feasibility of the concurrent execution of the racy events: $O'(e_i) = O'(e_{j})$.
+We first identify all candidate pairs.
+%that access the same location from different threads (and at least one is a write). 
+We then check each pair separately.  The checking is achieved by encoding 
+all necessary constraints.  The first constraint asserts the
+concurrency:
+%feasibility 
+%f the concurrent execution of the racy events: 
+$O'(e_i) = O'(e_{j})$.
 %In Figure~\ref{fig:running2}, the race condition for the race pair, ($e_2$, $e_{11}$), is $O'(e_2)=O'(e_{11})$
 
-{\bf Intra-thread Constraints\ } Given racy events $e_i$ and $e_j$, suppose $e_i$ is after $e_j$ in the original trace $\tau$.
-The predictive analysis reschedules the events in the prefix up to $e_i$, so that $e_i$ and $e_j$ can run concurrently.
-We refer to the prefix as $prefix(\tau, e_i)$.
-Throughout this section, we only consider the events in the prefix.
+{\bf Intra-thread Constraints\ } Given racy events $e_i$ and $e_j$, suppose 
+$e_i$ is after $e_j$ in the original trace $\tau$. Our analysis reschedules 
+the events in the prefix up to $e_i$, so that $e_i$ and $e_j$ can run 
+concurrently. We refer to the prefix as $prefix(\tau, e_i)$. Throughout 
+this section, we only consider events in the prefix.
 
-Preservation of the event sequence (up to $e$) for each thread requires the following intra-thread constraints:
+Preservation of the event sequence (up to $e$) for each thread requires 
+the following intra-thread constraints:
 
 \begin{itemize}
-\item {\bf Control Flow Constraints\ } The branches in the predicted run should take the same decisions as in the original trace so that each thread reproduces the same set of events.
-Specifically, we only need to reason about the branches prior to $e_i$.
+\item {\bf Control Flow Constraints\ } The branches in the predicted run 
+should take the same decisions as in the original trace so that each thread 
+reproduces the same sequence of events\footnote{We allow path differences in
+the second relaxation. This is to avoid presenting the encoding
+of both relaxations together, which is overly complex.} 
+%the encodings of both relaxation together, .  
+%Specifically, we only need to reason about the branches prior to $e_i$.
 
@@
 $$ 
-The constraint specifies that the branch condition should be evaluated as the same boolean value in the predicted run $\tau'$ as in the original trace $\tau$.
- Importantly, Unlike existing analyses \cite{yannis,pldi14}, we do not pose the requirement that the values flowing into branching statements remain the same, but adopt the relaxed requirement that the evaluation of branching expressions remains the same.
-Consider the example in Figure~\ref{fig:running2}, given the branch if($x^9$>11) at $e_10$, the original mapping is $\theta(x^9)=12$.
-We allow  a different value mapping, e.g., $\theta'(x^9)=13$, which retains the same truth value for the branch expression.
-\item {\bf Intra-thread Order Constraints\ } The events in the sequence should follow the same order as in the original trace. More formally, we require that
+The constraint specifies that the branch condition should be evaluated as 
+the same boolean value in the predicted run $\tau'$ as in the original 
+trace $\tau$.
+ Importantly, Unlike existing analyses \cite{yannis,pldi14}, we do not 
+pose the requirement that the values flowing into branching statements 
+remain the same, but adopt the relaxed requirement that 
+the branch outcome remains the same.
+%the evaluation 
+%of branching expressions remains the same. 
+Consider the example in Figure~\ref{fig:running2}, given the branch 
+if($x^9$$>$11) at $e_{10}$, the original mapping is $\theta(x^9)=12$. We allow  
+a different value mapping, e.g., $\theta'(x^9)=13$, which retains the 
+same truth value for the branch.
+\item {\bf Intra-thread Order Constraints\ } The events in the sequence 
+should follow the same order as in the original trace. More formally, 
+we require that
 $$
@@
 $$ 
-This constraint specifies that two events from the same thread should follow the same order as reflected by the ids of the events.
-\item {\bf Intra-thread Value Constraints\ } The value mapping of the variables should not contradict the value constraints imposed by each instruction.
-Without loss of generality, suppose the instruction is a local assignment in the form of $x=y+z$, then we require that
+This constraint specifies that two events from the same thread should 
+follow the same order as reflected by the ids of the events.
+\item {\bf Intra-thread Value Constraints\ } The value mapping of the 
+variables should not contradict the constraints imposed by individual 
+instruction.   Without loss of generality, suppose the instruction is 
+a local assignment in the form of $x=y+z$, then we require that
 $$
@@
 $$
-Remember that every variable may obtain a different value in the new trace $\tau'$ because it may read from some shared variable affected by the rescheduling.
-The constraint specifies that, the variables should be re-assigned to some values that are consistent with the instruction.
-Consider the example in Figure~\ref{fig:running2}, given the event $x^9=y^8+z^7$ at line 9, if the mappings $\theta'(y^8)$ and $\theta'(z^7)$ change to 3 and 10 respectively, the mapping $\theta'(x^9)$ should change to 13 consistently.
-In the above, we use the local assignment event for demonstration, in general, the intra-thread value constraints should also be specified for the {\sf  heapr/heapw} events.
+Recall that every variable may obtain a different value in the new 
+trace $\tau'$ because it may read from some shared variable affected by 
+the rescheduling. The constraint specifies that, the variables should 
+be re-assigned to some values that are consistent with the instruction. 
+Consider the example in Figure~\ref{fig:running2}, given the 
+event $x^9=y^8+z^7$ at line 9, if the mappings $\theta'(y^8)$ and 
+$\theta'(z^7)$ change to 3 and 10 respectively, the 
+mapping $\theta'(x^9)$ should change to 13 to be consistent. 
+%In the above, we use the local assignment event for demonstration, in general, the intra-thread value constraints should also be specified for the {\sf  heapr/heapw} events.
 \end{itemize}
@@
 %Rescheduling through Relaxation of Flow Dependencies
-{\bf Inter-thread Value Constraints for Relaxation\ } We now move to the first novel feature of \tool, which is its ability to explore execution schedules that depart from the value flow exhibited in the original trace.
-More precisely, \tool\ is able to relax value flow dependencies in the original trace:  a read access may read a
-different value from other write events, as long as the read value enforces feasibility of the following execution.
- This is strictly beyond the coverage potential of existing predictive analyses, which restrict trace transformations to ones where any read access must read the same  value (often from the same write event) as in the original trace.
+{\bf Inter-thread Value Constraints for Relaxation\ } 
+%We now move to 
+%the
+%first novel feature of \tool, which is its 
+%explain our ability to explore execution schedules that depart from the 
+%value flow exhibited in the original trace. More precisely, 
+%\tool\ is able to relax value flow dependencies in the original trace:  a 
+%read access may read a different value from other write events, as 
+%long as the read value enforces feasibility of the following execution.
+% This is strictly beyond the coverage potential of existing predictive analyses, which restrict trace transformations to ones where any read access must read the same  value (often from the same write event) as in the original trace.
 
-To ensure feasibility under relaxation of flow dependencies, we need to secure the flow between the read/write with the execution schedule. 
-For example, in Figure \ref{fig:running2}, for the read at $e_8$ to read from the write t $e_1$, we need a schedule where 
+To ensure feasibility under value relaxation
+% of flow dependencies, 
+we need to ensure the flow between read and write with some execution schedule. 
+For example, in Figure \ref{fig:running2}, for the read at $e_8$ to read 
+from the write t $e_1$, we need a schedule where 
 $e_8$ happens after $e_1$ and other writes such as $e_3$ do not interleave them.
@@
 
-In general, the constraint formula, given read $R^{m}_{\ell}$ of location $\ell$ at the event $e_m$ with set ${\cal W}$ of matching write events (i.e., events including write access to $\ell$), takes the following form:
+In general, the constraint formula, given read $R^{m}_{\ell}$ of location 
+$\ell$ at event $e_m$ with set ${\cal W}$ the matching write events 
+(i.e., events including write access to $\ell$), takes the following form:
 $$
+{\small
 \begin{array}{rll}
@@
 \end{array}
+}
 $$
-This disjunctive formula iterates over all matching write events, and demands for each that (i) it occurs prior to the read event $O'(e_n) < O'(e_m)$ and (ii) all other write events either occur before $O'(e_p) < O'(e_n)$ it or after the read event
-$O'(e_m) < O'(e_p)$.
+This disjunctive formula iterates over all matching write events, and requires 
+for each write that (i) it occurs prior to the 
+read event $O'(e_n) < O'(e_m)$ and (ii) all other write events either 
+occur before $O'(e_p) < O'(e_n)$ it or after the read 
+event $O'(e_m) < O'(e_p)$.
 
-An important concern that arises due to relaxation of flow dependencies is that heap accesses may change their meaning, i.e., they involve different base objects and no longer match with each other.
-As an illustration, we refer to Figure \ref{fig:heapAccess}.
-While the read at the event $e_7$ appears to match the write at $e_5$, this is conditioned on the read at $e_6$ being linked to the assignment at $e_4$.
-Suppose  the event $e_6$ reads from $e_3$ due to the relaxation.
-Then $e_7$ and $e_5$ no longer share the same base object (or the same location).
-Even worse, we do not know which events $e_7$ matches, because the base object is no longer known.
+An important concern that arises due to value relaxation 
+%of flow dependencies 
+is that heap accesses may change their meaning, i.e., they involve 
+different base objects and no longer match with each other. As an 
+illustration, we refer to Figure \ref{fig:heapAccess}. While the read at 
+event $e_7$ appears to match the write at $e_5$, this is conditioned 
+on the read at $e_6$ being linked to the assignment at $e_4$ (
+object $o_2$ is the object denoted by $z$ according to the source code). 
+Suppose event $e_6$ reads from $e_3$ due to the relaxation. Then $e_7$ and 
+$e_5$ should not share the same base object. 
+%Even worse, we do not know which events $e_7$ matches, because the base object is no longer known.
 
 \begin{figure}
+{\small
 	\centering
@@
 	\end{tabular}
-	\caption{\label{fig:heapAccess} A trace.}
+	\caption{\label{fig:heapAccess} A trace. The source code of line 7 is
+``{\tt w=z.f}''.}
+}
+\vspace{-0.15in}
 \end{figure}
 
-{\bf Invariant Base Object Constraints\ } To address this challenge, we enhance the constraint system with the requirement that heap objects that are dereferenced in field access statements before the racy events retain their original address in the predicted trace.
-This achieves two guarantees: First, matching heap access events in the original trace are guaranteed to also match in the predicted trace.
-Second, candidate races in the original trace remain viable in the predicted trace as they still refer to same location.
-Third, sharing between threads of heap locations remains unchanged  because
-each location is accessed by the same number of threads given that the base object is the same.
+{\bf Invariant Base Object Constraints\ } To address this challenge, 
+we enhance the constraint system with the requirement that heap objects 
+that are dereferenced in field accesses before the racy events retain 
+their original address in the predicted trace. This provides two
+guarantees: First, matching heap access events in the original trace are 
+guaranteed to also match in the predicted trace. Second, candidate races 
+in the original trace remain viable in the predicted trace as they still 
+refer to same location. 
+%Third, sharing between threads of heap locations 
+%remains unchanged because each location is accessed by the same number 
+%of threads given that the base object is the same. 
+%XZ: I commented out the third as I think it is obvious and not that useful.
 

### 2014-09-06 introduction.tex
intro typos fixed

@@
 complexity of the paths. Path-insensitive analyses often do not have sufficient precision on the other hand. 
-Static analyses also have difficulty in modeling complex aliasing behavior. In contrast, dynamic analysese focus 
+Static analyses also have difficulty in modeling complex aliasing behavior. In contrast, dynamic analyses focus 
 on analyzing a few executions, usually just one. They are sound and easily scale to long and complex executions. But they 
@@
 
-Recently, a novel kind of analysis called predictive analysis~\cite{...} was proposed. It has the capabilities of 
+Recently, a novel kind of analysis called predictive analysis~\cite{yannis,pldi14,Said:2011} was proposed. It has the capabilities of 
 achieving soundness as dynamic analyses and reasoning about properties that did not occur during execution 
@@
 and aliasing information. Note that during execution, we know precisely which write access reaches any given read 
-access. While such information is encoded in a trace, predictive analysis reaons about mutation of the trace to 
+access. While such information is encoded in a trace, predictive analysis reasons about mutation of the trace to 
 expose problems or study properties that may not manifest during execution. It was used in sound data race detection. 
 In particular, it collects a trace of threaded execution that contains rich runtime information such as concrete values, 
-execution path, and memory accsses. It then leverages constraint solving to reason about if races can occur with
+execution path, and memory accesses. It then leverages constraint solving to reason about if races can occur with
 a different thread schedule, {\em while enforcing the critical runtime information (e.g. values, thread
-local paths, and memory accesses) remains the same during schedule pertubation}. The rationale is that 
+local paths, and memory accesses) remains the same during schedule perturbation}. The rationale is that 
 it becomes as difficult as static analysis if we allow these critical runtime information change. 
-Preditive analysis has been shown to be very effective. It can detect real data races from complex programs.
+Predictive analysis has been shown to be very effective. It can detect real data races from complex programs.
 And more importantly, it guarantees the results are sound (i.e. no false positives).
 
-Despite its effectiveness, we observe that existing predictive analyes are too restrictive. They 
+Despite its effectiveness, we observe that existing predictive analyses are too restrictive. They 
 require too much runtime information to remain unchanged during analysis such that coverage is unnecessarily
@@
 & {\tt 4: if (y > 2)} \\
-& {\tt 5:~~z=1/x;} \\	
+& {\tt 5:~~z=1+x;} \\	
 & {\color{red} {\tt 6: else}} \\
-& {\color{red} {\tt 7:~~w=2/x;}}
+& {\color{red} {\tt 7:~~w=2+x;}}
 \end{tabular}
@@
 
-Our contrituions are summarized as follows.
+Our contributions are summarized as follows.
 \begin{itemize}

### 2014-09-06 introduction.tex
intro, typos done

@@
 Recently, a novel kind of analysis called predictive analysis~\cite{yannis,pldi14,Said:2011} was proposed. It has the capabilities of 
-achieving soundness as dynamic analyses and reasoning about properties that did not occur during execution 
+achieving soundness as dynamic analyses and reasoning about properties that did not occur during the execution 
 just like in static analysis. The basic idea is to leverage dynamic analysis to provide the information that 
@@
 the first missing race. We can also allow the trace to take the else branch such that the second race is also exposed. 
-Note that even though that branch was not executed originally, its effect can be precisely modeled as the value 
+Note that even though that branch was not executed originally, its effect can be precisely modeled as the address 
 of $x$ is known from the original trace. As such, we can still perform sound analysis.

### 2014-09-06 eval.tex
-xz

@@
 \section{Evaluation}~\label{sec:eval}
-Our evaluation focuses on the effectiveness  and scalability of our approach.
-To measure the effectiveness, we compare with the predictive analysis, {\sf RV}~\cite{}.
-We chose RV for two main reasons: (1) {\sf RV} is the only open source predictive analysis, (2) {\sf RV} represents the state-of-the-art approach, which is theoretically proven to have higher detection capability than other approaches.
+Our evaluation focuses on the effectiveness  and scalability.
+To measure the effectiveness, we compare with the predictive analysis, 
+{\sf RV}~\cite{}. We chose RV for two main reasons: (1) {\sf RV} is the 
+only open source predictive analysis, (2) {\sf RV} represents the 
+state-of-the-art, which is theoretically proven to have higher detection 
+capability than other approaches.
 
-{\bf Prototype Implementation}  In our prototype implementation of \tool, we used the Z3 SMT solver for constraint resolution \cite{MouraB08}.
-For static analysis, to discover unexecuted branches, we utilize the SOOT compiler framework \cite{Vallee-RaiCGHLS99}.
+{\bf Prototype Implementation}  We use the Z3 SMT solver for constraint 
+resolution \cite{MouraB08}. For static analysis, to discover unexecuted 
+branches, we utilize the SOOT compiler framework \cite{Vallee-RaiCGHLS99}.
 
-For scalability, as well as to constrain memory footprint, we have applied several engineering optimizations.
-In specific, (i) events are stored in a database rather than in an in-memory buffer; and (ii) the analysis is split into several segments, between which we restart the VM and resume the last snapshot, such that the different logs are fused together at the end.
+For scalability, as well as to constrain memory footprint, we have 
+applied several engineering optimizations. In specific, (i) events are 
+stored in a database rather than in an in-memory buffer; and (ii) the 
+analysis is split into several segments, between which we restart 
+the VM and resume the last snapshot, such that the different logs 
+are fused together at the end.
 
-To address the scalability limitations of the constraint solver (Z3 in our case), we again apply a splitting optimization.
-The full trace is split into $N$ equal-length subtraces or ``time windows'' (in our experiments, $N=1,000$).
-The values of local and shared variables are then passed at the end of a given window as inputs to the next window.
-For lack of space, we omit the concrete details of this optimization, and instead refer the reader to Huang et al.
-\cite{HuangMR14}, who apply a similar optimization.
+To improve the scalability in constraint solving, we again apply 
+a splitting optimization. The full trace is split into $N$ equal-length 
+subtraces or ``time windows'' (in our experiments, $N=1,000$). The values 
+of local and shared variables are then passed at the end of a 
+given window as the inputs to the next window. For lack of space, we 
+omit the details, and instead refer the reader to 
+Huang et al. \cite{HuangMR14}, who apply a similar optimization.
 
-{\bf Evaluation Method} We conducted our experiments on a set of 16 applications, which were also used to evaluate {\sf RV}.
-Specifically, the set includes large applications such as {\sf Jigsaw}, {\sf Xalan} and {\sf Lusearch}.
-To handle large applications that make use of reflection, we applied the {\sf Tamiflex} tool~\cite{}, which replaces reflective calls with the concrete method invocations recorded in the observed run for the purpose of effective static analysis.
-We omit the benchmark {\sf eclipse} because the current version of {\sf Tamiflex}  leads to abnormal execution after the instrumentation, which throws an exception when the main service is started.
-By applying our tool to such abnormal execution, we identify only 3 races, similarly to the report of RV~\cite{}.
-In addition, the {\sf montecarlo} benchmark requires an external input that we downloaded from the internet and simplified.
-For the large applications, we utilized the most lightweight available configuration.
+{\bf Evaluation Method} We conducted our experiments on a set of 16 
+applications, which were also used to evaluate {\sf RV}. Specifically, 
+the set includes large applications such as {\sf Jigsaw}, 
+{\sf Xalan} and {\sf Lusearch}. To handle large applications that make 
+use of reflection, we applied the {\sf Tamiflex} tool~\cite{}, which 
+replaces reflective calls with the concrete method invocations recorded 
+in the observed run.
+% for the purpose of effective static analysis.  
+We omit the benchmark {\sf eclipse} because the current version 
+of {\sf Tamiflex}  leads to abnormal execution after the instrumentation, 
+which throws an exception when the main service is started. 
+By applying \tool to the abnormal execution, we identify only 3 races, 
+similarly to the report of RV~\cite{}.
+In addition, the {\sf montecarlo} benchmark requires an external input 
+that we downloaded from the internet and simplified. For the large 
+applications, we utilized the most lightweight available configuration.
 
- As the detection capability  depends on the observed run, for fair comparison we monitor the execution once by recording all necessary information required by both approaches, and then apply both techniques to the monitored run.
-Besides, our reported data for RV may be different from the original report for two reasons: (1) the different observed runs lead to different sets of races, (2) the original implementation of RV contains a bug in identification of the branches, which leads to  misses of many branches and  incorrect reduction of dependences during the analysis.
-We confirmed the bug with the author and fixed the bug in our experiments.\footnote{We have created an in-depth report on the bug, along with a witness test case.
-(See:  \url{https://sites.google.com/site/recipe3141/}.)}
+As the detection capability  depends on the observed run, for fair 
+comparison we monitor the execution once by recording all necessary 
+information required by both approaches, and then apply both techniques 
+to the monitored run. Besides, our reported data for RV may be different 
+from the original report for two reasons: (1) the different observed runs 
+lead to different sets of races, (2) the original implementation of RV 
+contains a bug in branch identification, which leads to  misses of 
+many branches and  incorrect reduction of dependences during the analysis. 
+We confirmed the bug with the author and fixed the bug in our 
+experiments.\footnote{We have created an in-depth report on the bug, 
+along with a witness test case. (See:  \url{https://sites.google.com/site/recipe3141/}.)}
 
-Our experiments and measurements 
-were all conducted on an x86 64 Thinkpad W530 worksta- 
-tion with eight 2.30GHz Intel Core i7-3610QM processors, 
-16GB of RAM and 6M caches. The workstation runs version 12.04 of the Ubuntu Linux distribution, and has the Sun 
-64-Bit 1.6.0\_26 Java virtual machine (JVM) installed.
+Our experiments and measurements were all conducted on an x86 64 
+Thinkpad W530 workstation with eight 2.30GHz Intel Core i7-3610QM, 
+16GB of RAM and 6M cache. The workstation runs Ubuntu 12.04,
+% of the Ubuntu Linux distribution, 
+and has the Sun 64-Bit 1.6.0\_26 JVM installed.
 %\footnote{We note that Sun JDK 1.7 does not support the transformation of dacapo applications with {\sf tamiflex}.}
@@
 
-The Trace section includes the number of threads ($Th$), the number of shared reads ($Reads$) and shared writes ($Writes$), the subset of shared reads that read the base object references  ($Base$), where the  base object references are references used as the base/target in the following field reference or the method invocation,  the number of branches ($Br$), the synchronization events ($Sync$) and the total number of events ($Total$), which includes local accesses in addition to the aforementioned events.
-Specifically, branch events are reported in the form $A/B$, where $A$ refers to the number of branches used by our analysis and $B$ refers to the number of branches used by RV.
-To ensure that the predicted run sees the same base object at each shared read/write, {\sf RV} inserts the artificial branch immediately in front of each shared access (and array accesses).
+The Trace section includes the number of threads ($Th$), the number of 
+shared reads ($Reads$) and shared writes ($Writes$), the subset of shared 
+reads that read the base object references  ($Base$), where the  base object 
+references are references used as the base/target in the following field 
+reference or method invocation,  the number of branches ($Br$), the 
+synchronization events ($Sync$) and the total number of events ($Total$), 
+which includes local accesses in addition to the aforementioned events. 
+Specifically, branch events are reported in the form $A/B$, where $A$ refers 
+to the number of branches used by our analysis and $B$ refers to the number 
+of branches used by RV. To ensure that the predicted run sees the same 
+base object at each shared read/write, {\sf RV} inserts an artificial 
+branch immediately in front of each shared access (and array accesses). 
 We do not use such artificial branches.
 
-We make interesting observations about the Trace section.
-The non-local events, i.e., all the events listed in Table~\ref{tab:main}, occupy around 30\% of the total trace in the first 7 benchmarks, but occupy less than 1\% of the trace in the rest of the benchmarks, which have relatively more complex logic.
-Reads of the shared base objects occupy a small portion ($\frac{1}{10}$-$\frac{1}{3}$) of the total shared reads.
-The rest of the shared reads read only primitive values or
-references not involved in field accesses (e.g., references involved in nullness checking).
+We make a few interesting observations about the Trace section. 
+The non-local events, i.e., all the events listed in Table~\ref{tab:main}, 
+occupy around 30\% of the total trace in the first 7 benchmarks, 
+but occupy less than 1\% of the trace in the rest of the benchmarks, 
+which have relatively more complex logic. Reads of the shared base objects 
+occupy a small portion ($\frac{1}{10}$-$\frac{1}{3}$) of the total 
+shared reads. The rest of the shared reads read only primitive values 
+or references not involved in field accesses (e.g., references involved 
+in nullness checking).
 Besides, our analysis involves significantly less branch events as compared to the {\sf RV} approach.
@@
 
-The $Races$ section shows the number of races detected by {\sf Recipe-s}, i.e., {\sf Recipe} without exploring un-executed paths; {\sf Recipe}, i.e., the full-fledged version; and {\sf RV}.
-Comparison between {\sf Recipe-s} and {\sf Recipe} reveals that {\sf Recipe} finds $>$100 more races, which demonstrates the strength of exploring  un-executed paths.
-Intuitively,  {\sf Recipe-s} predicts based on a single trace, while {\sf Recipe} predicts based on multiple traces containing different execution paths.
-We also compare between {\sf Recipe} and {\sf RV} version, as shown in the $Difference$ section, where $Diff$ shows the races found by {\sf Recipe} but missed by {\sf RV} while $Diff'$ shows the races found by {\sf RV} but not {\sf Recipe}.
+The $Races$ section shows the number of races detected by {\sf Recipe-s}, 
+i.e., {\sf Recipe} without exploring un-executed paths; {\sf Recipe}, i.e., the full-fledged version; and {\sf RV}. Comparison between {\sf Recipe-s} 
+and {\sf Recipe} reveals that {\sf Recipe} finds $>$100 more races, 
+which demonstrates the strength of exploring  un-executed paths. 
+Intuitively,  {\sf Recipe-s} predicts based on a single trace, while 
+{\sf Recipe} predicts based on multiple traces containing different 
+execution paths. We also compare between {\sf Recipe} and {\sf RV} version, 
+as shown in the $Difference$ section, where $Diff$ shows the races found 
+by {\sf Recipe} but missed by {\sf RV} while $Diff'$ shows the races 
+found by {\sf RV} but not {\sf Recipe}.
 
-First of all, {\sf Recipe} finds $>$150 more races than {\sf RV}.
-There are several reasons for that: (1) {\sf Recipe} can reason about the accesses in the un-executed paths, while {\sf RV} and {\sf Recipe-s} can only reason about the accesses in the executed paths.
-(2) {\sf Recipe} or {\sf Recipe-s} relax the scheduling even if it breaks the inter-thread read/write dependence structure in the observed run.
-{\sf Recipe} allows the majority of the shared reads, i.e., the reads of primitive values, to freely read a different value from a different write as long as the value leads to the same branch decisions.
-{\sf RV}, however, requires them to read the same values as in the observed run.
-(3) An critical optimization proposed by {\sf RV} is to preserve dependences only for the reads before the preceding branches  of the racy events, rather than all the reads.
-However, this optimization is underplayed by the fact that {\sf RV} introduces huge amount of artificial branches, i.e., one before each shared field access, which ensures the use of the same base objects.
-We get rid of such artificial branches and instead, rely on the small amount of base read events to ensure the use of the same base objects (Section~\ref{}).
-Our strategy reduces the number of branches greatly and amplifies the effectiveness of the optimization.
-We also carried out case studies (Section~\ref{}) to better illustrate the scenarios.
+First of all, {\sf Recipe} finds $>$150 more races than {\sf RV}. There 
+are several reasons for that: (1) {\sf Recipe} can reason about the 
+accesses in the un-executed paths, while {\sf RV} and {\sf Recipe-s} can 
+only reason about the accesses in the executed paths. (2) {\sf Recipe} 
+or {\sf Recipe-s} relax the scheduling even if it 
+%breaks the inter-thread read/write dependence structure 
+changes variable values in the observed run. {\sf Recipe} allows the 
+majority of the shared reads, i.e., the reads of primitive values, 
+to freely read a different value from a different write 
+as long as the value leads to the same branch decisions. {\sf RV}, however, 
+requires them to read the same values as in the observed run. (3) An 
+critical optimization proposed by {\sf RV} is to preserve dependences 
+only for the reads before the preceding branches of the racy events, 
+rather than all the reads. However, this optimization is underplayed 
+by the fact that {\sf RV} introduces huge amount of artificial 
+branches, i.e., one before each shared field access, which ensures 
+the use of the same base objects.
+We get rid of such artificial branches and instead, rely on the small 
+amount of base read events to ensure the use of the same base 
+objects (Section~\ref{}). Our strategy reduces the number of branches 
+greatly and amplifies the effectiveness of the optimization. We also 
+carried out case studies (Section~\ref{}) to better illustrate the scenarios.
 
-Another interesting observation is that although {\sf Recipe} should produce all races found by {\sf RV} in theory, {\sf Recipe} may miss races found by {\sf RV} in practice (Column $Diff'$), i.e., {\sf Recipe} is not strictly more effective than {\sf RV} in practice.
-The reason for this is limitations of the underlying constraint solver: (1) the solver cannot compute constraints with very complex arithmatic operations, and (2) the solver does not support some program constants such as the scientific notation $3E-10$.
-In our empirical evaluation, we encounted only one benchmark, account, where these issues manifested.
+Another interesting observation is that although {\sf Recipe} should 
+produce all races found by {\sf RV} in theory, {\sf Recipe} may miss 
+races found by {\sf RV} in practice (Column $Diff'$), i.e., {\sf Recipe} 
+is not strictly more effective than {\sf RV} in practice. The reason  is 
+the limitations of the underlying constraint solver: (1) the solver cannot 
+compute constraints with very complex arithmatic operations, and (2) the 
+solver does not support some program constants such as the scientific 
+notation $3E-10$. In our empirical evaluation, we encounted only one 
+benchmark, account, where these issues manifested.
 
-The last section, Running time, compares the analysis time for both approaches.
-We find our approach is significantly slower than {\sf RV}, e.g., {\sf RV} often finishes within 200 seconds, while our approach may take more than 1 hour.
-This is because our approach needs to reason about the computation among variables inside the local access events, while {\sf RV} needs to only reason about the order relations among the events.
+The last section, Running time, compares the analysis time for both 
+approaches. We find our approach is significantly slower than {\sf RV}, 
+e.g., {\sf RV} often finishes within 200 seconds, while our approach may 
+take more than 1 hour. This is because {\sf Recipe} has a larger search 
+space and needs to reason about the computation among variables inside 
+the local access events, while {\sf RV} needs to only reason about 
+the order relations among the events.
 
@@
 
-{\bf Relaxing Inter-thread dependencies\ } Figure~\ref{fig:relax1} demonstrates the relaxation of inter-thread dependence enabled by our approach. 
-The code is from the benchmark {\sf bbuffer}, where the line number is marked.
-Huang et al.~\cite{} detect the race between lines 291 and 400, but fail to detect the race between lines 294 and 400.
-The reason is that in the observed run, the execution follows the order 400, 291 and 294.
-For the event at line 294, its preceding branch at line 291 reads from line 400.
-Therefore, Huang et al.~\cite{}
- require the predicted run to preserve the dependence between line 291 and line 400, so that line 291 reads exactly the same value and the branching expression evaluates to the same truth value.
-The dependence enforces the order constraint, $400 \rightarrow 291$, which further enforces the order $400\rightarrow 291 \rightarrow 294$.
-Our approach does not require the existence of such dependence.
-Specifically, we allow line 291 to happen before line 400 in the predicted run as long as the value read by it leads to the same branch decision, which is true in this case.
-As a result, there is no order constraint between line 294 and line 400, and the two form a race.
-We were able to replay this race  easily using the eclipse IDE breakpoints.
+{\bf Relaxing Values\ } Figure~\ref{fig:relax1} demonstrates the relaxation of inter-thread dependence enabled by our approach. 
+The code is from the benchmark {\sf bbuffer}, where the line numbers are
+marked. Huang et al.~\cite{} detect the race between lines 291 and 400, 
+but fail to detect the race between lines 294 and 400. The reason is 
+that in the observed run, the execution follows the order 400, 291 and 294. 
+For the event at line 294, its preceding branch at line 291 reads 
+from line 400. Therefore, Huang et al.~\cite{}
+require the predicted run to preserve the dependence between line 291 and 
+line 400, so that line 291 reads exactly the same value and the branching 
+expression evaluates to the same truth value. The dependence enforces 
+the order constraint, $400 \rightarrow 291$, which further enforces 
+the order $400\rightarrow 291 \rightarrow 294$.    Our approach does not 
+require the existence of such dependence. Specifically, we allow line 
+291 to happen before line 400 in the predicted run as long as the value 
+read by it leads to the same branch decision, which is true in this case.  
+As a result, there is no order constraint between line 294 and 
+line 400, and the two form a race. We were able to replay this race  
+easily using the eclipse IDE breakpoints.
 
@@
 \centering
-\includegraphics[scale=0.45]{cases/Visio-bbuffer.pdf}
+\includegraphics[scale=0.45]{figs/Visio-bbuffer.pdf}
 \caption{Relaxation of Inter-thread dependence}\label{fig:relax1}
@@
 
-{\bf Relaxing the Paths\ } Figure~\ref{fig:relax2} demonstrates how our approach relaxes the scheduling to account for  unexecuted paths. 
-All threads invoke the method {\sf Sorting}, which recursively starts two child threads if there are two more available entries in the thread pool (lines 8-11), or  starts one child thread if there is only one available entry (lines 4-5).
-The availability is computed through the static method {\sf available} with the constant {\tt total}, which is equal to 5.
-The shared variable {\tt alive} denotes the used entries.
+{\bf Relaxing the Paths\ } Figure~\ref{fig:relax2} demonstrates how our 
+approach relaxes the scheduling to account for  unexecuted paths in 
+{\tt mergesort}. 
+All threads invoke the method {\sf Sorting}, which recursively starts 
+two child threads if there are two more available entries in the 
+thread pool (lines 8-11), or starts one child thread if there is only 
+one available entry (lines 4-5).  The availability is computed through 
+the static method {\sf available} with the constant {\tt total}, which is equal to 5. The shared variable {\tt alive} denotes the used entries.
 
-Initially there is one sorting thread. After it starts Thread 1 and Thread 2, there are three threads alive and only two more entries are available.
-Suppose that in the observed run Thread 2 consumes both thread entries and starts the child Thread 3 and Thread 4 (not shown), updating {\tt alive} to 5.
-Then Thread 1 cannot execute the branch at lines 4-5.
+Initially there is one sorting thread. After it starts Thread 1 and 
+Thread 2, there are three threads alive and only two more 
+entries are available. Suppose that in the observed run Thread 2 
+consumes both thread entries and starts the child Thread 3 
+and Thread 4 (not shown), updating {\tt alive} to 5. Then Thread 1 
+cannot execute the branch at lines 4-5.
 Huang et al.~\cite{} require the predicted run to preserve the dependence denoted by the dotted line since this dependence affects the branch condition at line 2.
@@
 \centering
-\includegraphics[scale=0.45]{cases/Visio-msort.pdf}
+\includegraphics[scale=0.45]{figs/Visio-msort.pdf}
 \caption{Relaxation of Paths}\label{fig:relax2}
@@
 \section{Threats to Validity}
-A Java method can consist of up to 65535 bytecode instructions.
-Therefore, we count the number of bytecode instructions inside each method.
-If the resulting number exceeds 65525, we avoid instrumenting the method.
-The consequence is that we will potentially races inside a method for which we skipped instrumentation..
-In this case, we specify the variables read from the method to be equal to their concrete values in the constraints.
-The constraint solver can proceed safely without being affected by such methods.
+A Java method can consist of up to 65535 bytecode instructions. Therefore,
+ we count the number of bytecode instructions inside each method. If the 
+resulting number exceeds 65525, we avoid instrumenting the method. 
+The consequence is that we will potentially miss races inside a method for 
+which we skipped instrumentation. In this case, we specify the variables 
+read from the method to be equal to their concrete values in the constraints. 
+The constraint solver can proceed safely without being affected by 
+such methods.
 
-A second point regards support for certain operators.
-In our current implementation, we do not support the boolean operators, including e.g.
-$\&$, as well as the bit operator $<<$, which accounts for most of our misses.
+\tool lacks support for certain operators. In our current 
+implementation, we do not support bit operations such as $\&$ and $<<$, which accounts for most of our misses.

### 2014-09-06 overview.tex

@@
 execution of a static instruction $inst$, such that (i) $t$ refers to the 
-thread; (ii) $id$ is a unique identifier of the event (unless specified 
+thread; (ii) $id$ is a {\em unique} identifier of the event (unless specified 
 otherwise, we use the index of an event in the trace as its id); 
@@
 
-%TODO this section is inconsistent with others. change it!
 \subsection{Constraint System}
@@
 to have different values as long as the same thread local path is preserved.
-It does not require data dependences or variable values as in~\cite{...}.
+It does not require data dependences or same variable values as in~\cite{yannis, pldi14, Said:2011}.
 This allows us to catch the racy pair $(e_2, e_5)$.
@@
 possible  paths. The solver may pick up the unexplored branch to detect 
-a race. Back to our example, event $e_7$ is in the unexplored branch, \tool
+a race. Back to our example, event $e_7$ is in the unexplored branch, \tool\ 
 is able to identify the race between $e_7$ and $e_2$ after encoding 
@@
 \noindent
-{\em Highlights \ } \tool finds two more pairs of races 
-beyond existing analyses~\cite{yannis, jeff}: $(e_2, e_5)$ and $(e_2, e_7)$.
+{\em Highlights \ } \tool\  finds two more pairs of races 
+beyond existing analyses~\cite{yannis, pldi14}: $(e_2, e_5)$ and $(e_2, e_7)$.
 For the first race, we do not require $e_4$ to read from $e_3$ to 

### 2014-09-06 abstract.tex
-xz

--- 2014-09-06 abstract.tex
@@
 \begin{abstract}
-Predictive analysis has recently emerged as a promising paradigm for sound detection of multithreading bugs (i.e., without false alarms).
-Intuitively, the idea is to start from an instrumented execution trace, and transform it based on the concrete information it provides (e.g., concrete values and memory accesses) while preserving feasibility (e.g., by reodering trace events iff flow dependencies are preserved).
-
-Predictive analysis has been applied successfully to the problem of data-race detection, though state-of-the-art techniques are restrictive in the transformations they allow so as to guarantee feasibility and consequently also soundness.
+Predictive analysis has recently emerged as a promising technique for 
+sound detection of multithreading bugs (i.e., without false alarms). 
+The idea is to start from an instrumented execution trace, and transform 
+it based on the concrete information it provides, such as concrete values 
+and memory accesses, to expose bugs, while preserving feasibility. For
+example, it may reoder trace events to simulate different thread schedules, 
+as long as flow dependencies are preserved.
+Predictive analysis has been successfully applied to data-race detection, 
+although the state-of-the-art techniques are overly restrictive in the 
+transformations they allow such that their analysis coverage is 
+unnecessarily limited. 
+%so as to guarantee feasibility 
+%and consequently also soundness.
 In particular, existing techniques (i) do not permit flow dependencies to be violated (except in very specific cases), and (ii) are unable to analyze execution paths beyond that of the input trace.
-
-In this paper, we present \sysname, a predictive race detector that is able to relax both of these restrictions while guaranteeing soundness, as we prove formally.
-Thanks to its ability to explore event orderings that violate flow dependencies as well as unexecuted branches, \sysname\ is able to detect x$2.5$ more data races on a representative set of large-scale benchmarks compared to the state-of-the-art in predictive race detection.
+In this paper, we present \sysname, a predictive race detector that has two
+important relaxations: value and path relaxations without affecting the
+soundness guarantee. The former allows 
+variables to have different values and the latter allows exploring 
+unexecuted paths that are neighbors to the traced path.
+% both of these restrictions while guaranteeing soundness, as we prove formally. Thanks to its ability to explore unexecuted branches as well as event orderings that violate flow dependencies, 
+Our results show that \sysname\ is able to detect x$2.5$ more data races on a 
+representative set of large-scale benchmarks compared to the state-of-the-art.
+%in predictive race detection.
 \end{abstract}

### 2014-09-06 guarantee.tex
-xz

@@
 \begin{itemize}
-\item  $\tau'$ is feasible.
-Researchers~\cite{chao,maximal} point out a trace is feasible iff it satisfies the sequential consistency, i.e., $\forall object\ o, \tau'\downarrow_{o}$ satisfies the serial specification of the object $o$.
-If  all the events considered are of the types in Table~\ref{Ta:syntax}, then the sequential consistency precisely means the following: (1) read-write consistency, i.e., each read event of a variable should contain the value written by the most recent write event, and (2) the synchronization consistency, e.g., the lock acquire and release of a lock should not be interleaved by the lock operations of the same lock, the begin event of a thread should follow the fork event of the parent thread.
-Our inter-thread data flow constraints guarantee the read-write consistency among shared locations.
-The SSA form encodes the read-write consistency among local variables.
-The synchronization consistency is also captured as constraints in our solver.
-\item $\tau$ can be generated by the program, i.e., for each thread $t$, $\tau'\downarrow_{t}$ can be generated by the corresponding thread code.
-This claim requires that,  for any two adjacent events $e$ and $e'$ in  $\tau'\downarrow_{t}$, (1) if $e$ is not a branch, $inst^e$ and $inst^{e'}$ should be adjacent in the code; (2) if $e$ is a branch, $inst^e$ and $inst^{e'}$ should be adjacent and the $inst^e$ should be evaluated as the boolean value indicated by $inst^{e'}$.
-We derive the trace  through either concrete execution or symbolic execution, which satisfies the requirements and guarantees the correctness of the claim.
+\item  $\tau'$ is feasible. Researchers~\cite{chao,maximal} point out a 
+trace is feasible iff it satisfies the sequential consistency, 
+i.e., $\forall object\ o, \tau'\downarrow_{o}$ satisfies the serial 
+specification of the object $o$. If  all the events considered are of 
+the types in Table~\ref{Ta:syntax}, then the sequential consistency 
+precisely means the following: (1) read-write consistency, i.e., each 
+read event of a variable should contain the value written by the most 
+recent write event, and (2) the synchronization consistency, e.g., the 
+acquisition and release of a lock should not interleave with other 
+lock operations of the same lock, the start event of a thread should 
+follow the fork event of the parent thread. Our inter-thread data 
+flow constraints guarantee the read-write consistency among shared 
+locations. The SSA form encodes the read-write consistency among local 
+variables. The synchronization consistency is also captured as constraints 
+in our solver. 
+\item $\tau$ can be generated by the program, i.e., for each 
+thread $t$, $\tau'\downarrow_{t}$ can be generated by the corresponding 
+thread code. This claim requires that,  for any two adjacent 
+events $e$ and $e'$ in  $\tau'\downarrow_{t}$, (1) if $e$ is not a 
+branch, $inst^e$ and $inst^{e'}$ should be adjacent in the code; (2) 
+if $e$ is a branch, $inst^e$ and $inst^{e'}$ should be adjacent in terms
+of control flow and the $inst^e$ should be evaluated as the boolean 
+value implied by $inst^{e'}$. We derive the trace  through either 
+concrete execution or precise modeling of unexplored branches, 
+which satisfies the requirements and guarantees the correctness of the claim.
 \end{itemize}
@@
 
-First, to the best of our knowledge, none of the existing techniques automatically explore the un-executed branches by relaxing the schedules, while ours enables this.
-Second, assume we disable the exploration of un-executed branches and all techniques start with the same trace. Our technique still explores more traces. 
-Figure~\ref{fig:discuss} illustrates the difference pictorially, where two vertical lines represent the progress of two threads and the circles with numbers denote the events with ids. 
-Existing techniques require the reads to read the same values, therefore, enforce the schedule order $e_5 \rightarrow e_6$ and $e_2\rightarrow e_3$.
-We allow the reads to read different values from different writes as long as the feasibility is preserved.
-Comparatively, we  enforce a weaker schedule order  $e_4 \rightarrow e_6$ and $e_2\rightarrow e_3$, which allows more scheduling, e.g., the concurrent execution of $e_7$ (or some event after $e_7$) and $e_5$.
+First, to the best of our knowledge, none of the existing techniques 
+automatically explore the un-executed branches by relaxing the schedules, whereas \tool enables this.
+Second, assume we disable the exploration of un-executed branches and all 
+techniques start with the same trace. Our technique still explores more 
+traces. Figure~\ref{fig:discuss} illustrates the difference, where two 
+vertical lines represent the progress of two threads and the circles 
+with numbers denote events with ids. 
+Existing techniques require the reads to read the same values, and hence
+enforce the schedule order $e_5 \rightarrow e_6$ and $e_2\rightarrow e_3$. 
+We allow the reads to read different values from different writes as long as 
+the feasibility is preserved. Comparatively, we enforce a weaker 
+schedule order  $e_4 \rightarrow e_6$ and $e_2\rightarrow e_3$, which 
+allows more scheduling, e.g., the concurrent execution of $e_7$ (or 
+some event after $e_7$) and $e_5$.
 
-However, our technique does not identify the maximal set of traces from a single trace.
-The underlying reason is that, we still enforce the schedule order for preserving the base object, e.g., $e_2\rightarrow e_3$.
-It is possible to further relax the schedule so that $e_3$ reads from $e_1$, but this requires challenging reasoning of the field accesses as their base objects become unknown.
-We plan to explore this idea in future work.
+However, \tool has not reached the maximum bound of sound exploration.
+% does not identify the maximal set of traces 
+%from a single trace. 
+The underlying reason is that, we still enforce the schedule for 
+preserving the base object, e.g., $e_2\rightarrow e_3$. It is possible 
+to further relax the schedule so that $e_3$ reads from $e_1$, but 
+this requires challenging reasoning of the field accesses as their base 
+objects become uncertain. We plan to explore this idea in future work.
 

### 2014-09-06 related.tex
rw

@@
 
-The first to propose PTA are Smagardakis et al.
+Wang et al.~\cite{chao} pioneere the symbolic predictive analysis, which our approach also belongs to.
+However, the approach lacks support of heap encoding, therefore, cannot apply to real-world applications.
+The first PTA that applies to real-world applications are proposed by Smagardakis et al.
 \cite{yannis}.
-In their original work, both synchronization constraints and inter-thread dependencies are preserved, where inter-thread dependencies are respected by only allowing reorderings that leave the dependence structure exhibited by the original trace in tact.
+In their original work, both synchronization constraints and inter-thread dependencies are preserved, where inter-thread dependencies are respected by only allowing reorderings that leave the dependence structure exhibited by the original trace intact.
 This ensures that the values of shared memory locations remain the same, which secures the soundness argument.

### 2014-09-06 related.tex
xz

@@
 
-In this section, we survey related research on race detection with special emphasis on predictive trace analysis (PTA).
+%In this section, we survey related research on race detection with special emphasis on predictive trace analysis (PTA).
 
@@
 
-Initial attempts to address the challenge of race detection focused on built-in synchronization primitives \cite{eraser,SasturkarAWS05,vonPraun:2001,Choi:2002}.
-These include locks as well as the wait/notify and start/join scheduling controls.
-Notable among these efforts is \emph{lockset} analysis, which considers only locks \cite{eraser}.
-Because the derived constraints are partial, permitting certain infeasible event reorderings, lockset analysis cannot guarantee soundness \cite{Naik:2006}.
+Initial attempts to address the challenge of race detection focused on 
+built-in synchronization primitives \cite{eraser,SasturkarAWS05,vonPraun:2001,
+Choi:2002}. These include locks as well as the wait/notify and start/join 
+primitives. Notable among these efforts is the \emph{lockset} analysis, 
+which considers only locks \cite{eraser}. Because the derived constraints 
+are partial, permitting certain infeasible event reorderings, lockset 
+analysis cannot guarantee soundness \cite{Naik:2006}.
 
-A different tradeoff is struck by the \emph{happens-before (HB)} approach \cite{Christiaens,Dinning:1990,Mellor-Crummey:1991}, which is founded on Lamport's HB relation \cite{Lamport}.
-In this style of analysis, all synchronization primitives are accounted for, though reordering constraints are conservative .
-As an example, HB inhibits reordering of two synchronized blocks governed by the same lock.
+A different tradeoff is struck by the \emph{happens-before (HB)} 
+approach \cite{Christiaens,Dinning:1990,Mellor-Crummey:1991}, which is 
+founded on Lamport's HB relation \cite{Lamport}. In HB analysis, all 
+synchronization primitives are accounted for, though reordering constraints 
+are overly conservative. 
+%As an example, HB inhibits reordering of two synchronized 
+%blocks governed by the same lock.
 
@@
 
-Recently there have been successful attempts to relax HB constraints.
-Among these are hybrid analysis \cite{hybrid}, which permits both orderings of lock-synchronized blocks, as well as the \emph{Universal Causal Graph (UCG)} representation~\cite{ucg}, which also enables both orderings but only if these are consistent with wait/notify- and start/join-induced constraints.
+Recently there have been successful attempts to relax HB constraints, including
+%Among these are 
+the hybrid analysis \cite{hybrid}, which permits both orderings of 
+lock-synchronized blocks, and the \emph{Universal Causal Graph (UCG)} 
+representation~\cite{ucg}, which also enables both orderings but only 
+if these are consistent with wait/notify- and start/join-induced constraints.
 
@@
 
-Given concurrent execution trace $t$, PTA derives new traces to witness the data races.
-PTA is founded on the notion of sound causality, as it considers feasible reorderings of the input trace that prove a candidate data race as such.
+Given a concurrent execution trace $t$, PTA derives new traces to witness 
+the data races. PTA is founded on the notion of sound causality, as it 
+considers feasible reorderings of the input trace that prove a candidate 
+data race as such.
 
-Wang et al.~\cite{chao} pioneere the symbolic predictive analysis, which our approach also belongs to.
-However, the approach lacks support of heap encoding, therefore, cannot apply to real-world applications.
-The first PTA that applies to real-world applications are proposed by Smagardakis et al.
-\cite{yannis}.
-In their original work, both synchronization constraints and inter-thread dependencies are preserved, where inter-thread dependencies are respected by only allowing reorderings that leave the dependence structure exhibited by the original trace intact.
-This ensures that the values of shared memory locations remain the same, which secures the soundness argument.
+Wang et al.~\cite{chao} pioneere the symbolic predictive analysis, 
+which our approach also belongs to. However, the approach lacks support 
+of heap encoding, therefore, cannot be applied to real-world applications. 
+The first PTA that applies to real-world applications is proposed by 
+Smagardakis et al. \cite{yannis}. In their original work, both synchronization 
+constraints and inter-thread dependencies are preserved, where inter-thread 
+dependencies are respected by only allowing reorderings that leave the 
+dependence structure exhibited by the original trace intact. This ensures 
+that the values of shared memory locations remain the same, which secures 
+the soundness argument.
 
-Said et al.
-\cite{Said:2011} and Huang et al.
-\cite{HuangMR14} describe PTAs that are also sound.
-Said et al.
-perform symbolic analysis of the input trace, and then utilize an SMT solver to search for interleaved schedules that establish the presence of data races.
-Soundness is guaranteed by their ability to precisely encode the semantics of sequential consistency.
-Inspired by Said et al., Huang et al.
-present an improvement, whereby control-flow information is encoded into the constraint system so as to relax flow dependencies as long as control dependencies are preserved.
-ExceptionNULL \cite{Farzan:2012} is another example of a sound PTA, of which the goal is to detect null dereferences rather than data races.
+Said et al. \cite{Said:2011} and Huang et al. \cite{HuangMR14} proposed
+PTAs that are also sound. Said et al. perform symbolic analysis of the 
+input trace, and then utilize an SMT solver to search for schedules that 
+establish the presence of data races. Soundness is guaranteed by their 
+ability to precisely encode the semantics of sequential consistency. 
+Inspired by Said et al., Huang et al. present an improvement, whereby 
+control-flow information is encoded into the constraint system so as 
+to relax flow dependencies as long as control dependencies are 
+preserved. ExceptionNULL \cite{Farzan:2012} is another example of 
+sound PTA. Its goal is to detect null dereferences rather than data races.
 
-\tool\ is similar to Said et al.
-and Huang et al.
-in that it also makes use of an SMT solver based on symbolic encoding of the input trace.
-Still, the two relaxations that \tool\ features, which enable (i) value- rather than dependence-based reasoning about data flow and (ii) consideration of unexplored branches, are strictly outside the scope of both of these techniques.
-In Section \ref{sec:eval}, we demonstrate via direct comparison with Huang et al.
-the dramatic improvement in coverage thanks to these relaxation methods, which we prove to handle in a sound manner.
+\tool\ is similar to Said et al. and Huang et al. in that it also makes 
+use of an SMT solver based on symbolic encoding of the input 
+trace. Still, the two sound relaxations that \tool\ features, which 
+enable (i) value- rather than dependence-based reasoning about data 
+flow and (ii) consideration of unexplored branches, are strictly beyond
+the scope of both techniques. In Section \ref{sec:eval}, we demonstrate 
+via direct comparison with Huang et al. the substantial improvement 
+in coverage thanks to these relaxation methods.
+%, which we prove to handle 
+%in a sound manner.
 
 \subsection{Maximality Claim}
-Serbanuta et al. \cite{maximal} and Huang et al.~\cite{HuangMR14} both claim maximality of the detection capability, but under different assumptions. 
-Assuming the branch event is not modeled in the trace, Serbanuta et al.'s approach \cite{maximal} guarantees maximality.
-Given few information in the trace, it conservatively requires the shared read to obtain the same value in the predicted run to ensure soundness.
-However, if the branch event is modeled, the technique does not guarantee maximality.
-Assuming the branch events are modeled but local assignments are not recorded in the trace, Huang et al.'s approach~\cite{HuangMR14} guarantees the maximality.
-Compared to Serbanuta et al.'s technique \cite{maximal}, it allows some events to read different values if they do not affect the following branch decisions.
-However, it requires the events that affect the branch decisions to read exactly the same values.
-In practice, these assumptions are too restrictive  because the testers often have full access to the source/bytecode and can collect the local assignments freely.
+Serbanuta et al. \cite{maximal} and Huang et al.~\cite{HuangMR14} both 
+claim maximality of the detection capability, but under different assumptions. 
+Assuming branch events are not modeled in the trace, \cite{maximal} guarantees maximality. Given few 
+information in the trace, it conservatively requires the shared read to 
+obtain the same value in the predicted run to ensure soundness. However, if 
+branch events are modeled, the technique does not guarantee maximality.  
+Assuming the branch events are modeled but local assignments are not recorded, 
+\cite{HuangMR14} guarantees maximality. 
+Compared to \cite{maximal}, it allows some events to read different 
+values if they do not affect the following branch decisions. However, 
+it requires the events that affect the branch decisions to read exactly 
+the same values. In practice, these assumptions are too restrictive  
+because local assignments can be easily collected in practice.
+%the testers often have full access to the source/bytecode and 
+%can collect local assignments freely.
 
-Assuming the local assignments are available in the trace, both techniques do not guarantee maximality.
-As illustrated in our example and discussed in Section~\ref{sec:guarantee}, our analysis allows higher detection capability, i.e., we allow the events to read different values even they flow into the branch events; we allow the exploration of un-executed paths, both of which are not allowed by the above techniques.
+Assuming local assignments are available in the trace, both techniques 
+cannot guarantee maximality. As illustrated in our example and discussed 
+in Section~\ref{sec:guarantee}, our analysis allows higher detection capability
+%, i.e., we allow the events to read different values even they flow into the branch events; we allow the exploration of un-executed paths, both of which are not allowed by the above techniques.

### 2014-09-06 relax1.tex
relax I typos fixed

@@
 the computation of the mapping, in which each variable is mapped to exactly 
-one value. Another side effect is that the def/use chains become explicit 
-in SSA form, which simplifies the following analysis steps.
+one value.
 
@@
 except that loops and recursions are fully resolved in the trace, 
-obviating the need for {\sf Phi} nodes\footnote{While \tool in fact uses
+obviating the need for {\sf Phi} nodes\footnote{While \tool\  in fact uses
 Phi nodes in the second relaxation, for simplicity during discussion, we 
@@
 the second relaxation. This is to avoid presenting the encoding
-of both relaxations together, which is overly complex.} 
+of both relaxations together, which is overly complex.}.
 %the encodings of both relaxation together, .  
@@
 Consider the example in Figure~\ref{fig:running2}, given the branch 
-if($x^9$$>$11) at $e_{10}$, the original mapping is $\theta(x^9)=12$. We allow  
+if($x^9$$>$11$) at $e_{10}$, the original mapping is $\theta(x^9)=12$. We allow  
 a different value mapping, e.g., $\theta'(x^9)=13$, which retains the 
@@
 $$ 
-This constraint specifies that two events from the same thread should 
-follow the same order as reflected by the ids of the events.
+This constraint specifies that two events from the same thread ($t^{e_m}$ denotes the thread of $e_m$) should 
+follow the same order as in the original trace.
 \item {\bf Intra-thread Value Constraints\ } The value mapping of the 
@@
 For example, in Figure \ref{fig:running2}, for the read at $e_8$ to read 
-from the write t $e_1$, we need a schedule where 
+from the write  $e_1$, we need a schedule where 
 $e_8$ happens after $e_1$ and other writes such as $e_3$ do not interleave them.
@@
 Suppose event $e_6$ reads from $e_3$ due to the relaxation. Then $e_7$ and 
-$e_5$ should not share the same base object. 
+$e_5$ should not match with each other as they do not share the same base object. 
 %Even worse, we do not know which events $e_7$ matches, because the base object is no longer known.

### 2014-09-06 relax2.tex
relax2 typo fixed

@@
 be resolved to a constant from the trace (otherwise we do not know how many
-times we need to unroll). (2) The branch
+times we need to unroll). (2) The unexecuted branch
 is ignored if it contains recursive function(s). (3) Every base reference 
-variable in the branch needs to be fully resolved to a concrete object,
+variable in the unexecuted branch needs to be fully resolved to a concrete object,
 as required by our earlier analysis (Section~\ref{sec:relax1}). This can be 
 done by backward traversal of the trace:
-if the base reference depends on some shared read in the explored 
+if the base reference depends on some shared read in the unexplored 
 branch, of which the value is uncertain, it cannot be 
@@
 Otherwise, it can be resolved.
-(4) The indices of array read accesses in the branch 
-need to be resolved to constant values from the trace (with the preceding unexecuted branch extensions). This is to
+(4) The indices of array read accesses in the unexecuted branch 
+need to be resolved to constant values from the trace. This is to
 prevent reading an array element that was not defined. 
-(5) The branch is ignored if it uses collections. 
-(6) For method calls, we require the  caller object can be fully resolved from
+(5) The unexecuted branch is ignored if it uses collections. 
+(6) For method calls in the unexecuted branch, we require the  caller object can be fully resolved from
 the trace.
@@
 using the example in Figure~\ref{fig:running} for demonstration.
-We assume one of the branches of event $e_b$ needs to be extended in
+Given a branch event $e_b$, we assume one of its branches is unexecuted and needs to be (statically) extended in
 the following discussion.
@@
 meaning that $y$ is equal to $y_i$ in the condition $C_i$.
-Note that we do not require paths inside the extended branch be fully
-resolved. 
 Phi nodes are also used to integrate the definitions in the extended
@@
 in $IF(e_b)$ is in the form of $x_m=y_m+z_m$ and the 
-event $e_n$ in $ELSE(e_b)$ is in the form of $x_n=y_n+z_n$.
-
-{\bf [XZ: missing piece]}
+event $e_n$ in $ELSE(e_b)$ is in the form of $x_n=y_n+z_n$. Then, we have,
+$$
+\begin{array}{l}
+\bigwedge\limits_{e_m \in IF(e_b)} (\theta'(x_b) < \theta'(y_b) \rightarrow  \theta'(x_m)=\theta'(y_m)+\theta'(z_m)) \\ 
+\quad \quad  \bigwedge  \\
+\bigwedge\limits_{e_n \in ELSE(e_b)} (\neg(\theta'(x_b)  < \theta'(y_b)) \rightarrow  \theta'(x_n)=\theta'(y_n)+\theta'(z_n)) \\ 
+\end{array}
+$$
 
@@
 \wedge\\
-O'(e_b)<O'(e_{j_1})<O'(e_{j_2})<\dots <O'(e_{j_m})<O'(e_{end})
+O'(e_b)<O'(e_{j_1})<O'(e_{j_2})<\dots <O'(e_{j_n})<O'(e_{end})
 \end{array}

### 2014-09-06 eval.tex
relax I typos fixed
citations fixed

@@
 To measure the effectiveness, we compare with the predictive analysis, 
-{\sf RV}~\cite{}. We chose RV for two main reasons: (1) {\sf RV} is the 
+{\sf RV}~\cite{pldi14}. We chose RV for two main reasons: (1) {\sf RV} is the 
 only open source predictive analysis, (2) {\sf RV} represents the 
@@
 omit the details, and instead refer the reader to 
-Huang et al. \cite{HuangMR14}, who apply a similar optimization.
+Huang et al. \cite{pldi14}, who apply a similar optimization.
 
@@
 {\sf Xalan} and {\sf Lusearch}. To handle large applications that make 
-use of reflection, we applied the {\sf Tamiflex} tool~\cite{}, which 
+use of reflection, we applied the {\sf Tamiflex} tool~\cite{tamiflex}, which 
 replaces reflective calls with the concrete method invocations recorded 
@@
 which throws an exception when the main service is started. 
-By applying \tool to the abnormal execution, we identify only 3 races, 
-similarly to the report of RV~\cite{}.
+By applying \tool\  to the abnormal execution, we identify only 3 races, 
+similarly to the report of RV~\cite{pldi14}.
 In addition, the {\sf montecarlo} benchmark requires an external input 
@@
 amount of base read events to ensure the use of the same base 
-objects (Section~\ref{}). Our strategy reduces the number of branches 
+objects (Section~\ref{sec:relax1} and Section~\ref{sec:relax2}). Our strategy reduces the number of branches 
 greatly and amplifies the effectiveness of the optimization. We also 
-carried out case studies (Section~\ref{}) to better illustrate the scenarios.
+carried out case studies (Section~\ref{sec:cases}) to better illustrate the scenarios.
 
@@
 
-\subsection{Case studies}
+\subsection{Case studies}~\label{sec:cases}
 We manually inspected the reported races over small applications to gain better understanding about our approach.
@@
 The code is from the benchmark {\sf bbuffer}, where the line numbers are
-marked. Huang et al.~\cite{} detect the race between lines 291 and 400, 
+marked. Huang et al.~\cite{pldi14} detect the race between lines 291 and 400, 
 but fail to detect the race between lines 294 and 400. The reason is 
@@
 For the event at line 294, its preceding branch at line 291 reads 
-from line 400. Therefore, Huang et al.~\cite{}
+from line 400. Therefore, Huang et al.~\cite{pldi14}
 require the predicted run to preserve the dependence between line 291 and 
@@
 cannot execute the branch at lines 4-5.
-Huang et al.~\cite{} require the predicted run to preserve the dependence denoted by the dotted line since this dependence affects the branch condition at line 2.
+Huang et al.~\cite{pldi14} require the predicted run to preserve the dependence denoted by the dotted line since this dependence affects the branch condition at line 2.
 As a result, the predicted run follows the same branch decision and cannot reason about the code at lines 4-5.

### 2014-09-06 overview.tex
citations fixed

@@
 in this paper, which are already well explained in existing 
-techniques~\cite{yannis, pecan}.}. We explain the constraints 
+techniques~\cite{yannis, pldi14}.}. We explain the constraints 
 briefly in the following. Details are in Section~\ref{sec:relax1} 

### 2014-09-06 related.tex
citations fixed

@@
 
-Said et al. \cite{Said:2011} and Huang et al. \cite{HuangMR14} proposed
+Said et al. \cite{Said:2011} and Huang et al. \cite{pldi14} proposed
 PTAs that are also sound. Said et al. perform symbolic analysis of the 
@@
 \subsection{Maximality Claim}
-Serbanuta et al. \cite{maximal} and Huang et al.~\cite{HuangMR14} both 
+Serbanuta et al. \cite{maximal} and Huang et al.~\cite{pldi14} both 
 claim maximality of the detection capability, but under different assumptions. 
@@
 Assuming the branch events are modeled but local assignments are not recorded, 
-\cite{HuangMR14} guarantees maximality. 
+\cite{pldi14} guarantees maximality. 
 Compared to \cite{maximal}, it allows some events to read different 

### 2014-09-06 relax2.tex
titles changed to value relaxation + path relaxation

@@
-\section{Exploration of Unexecuted Branches}~\label{sec:relax2}
+\section{Path Relaxation}~\label{sec:relax2}
 We now switch to the second feature of \tool, which is to reason about 

### 2014-09-06 guarantee.tex
guarantee typos fixed

@@
 \begin{mytheorem}[Soundness]
-The trace computed by the solver,  $\tau'=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \}, O', \theta'>$, is a feasible trace for the concurrent program, where $O'$ defines its schedule and $\theta'$ defines the value mapping.
+The trace $\tau'$ computed by the solver is a feasible trace for the concurrent program.
 \end{mytheorem}
 
-\begin{proof}
+\begin{proof}\let\qed\relax
 The proof consists of two parts: (1) the trace is feasible, and (2) the trace can be generated by the program. We sketch the proof as follows.
@@
 in our solver. 
-\item $\tau$ can be generated by the program, i.e., for each 
+\item $\tau'$ can be generated by the program, i.e., for each 
 thread $t$, $\tau'\downarrow_{t}$ can be generated by the corresponding 
@@
 \begin{mydiscussion}[Detection Capability]
-Our technique has stronger detection capability than existing techniques~\cite{yannis,jeff, said}.
+Our technique has stronger detection capability than existing techniques~\cite{yannis,pldi14,Said:2011}.
 \end{mydiscussion}

### 2014-09-06 abstract.tex
good shape now

@@
 example, it may reoder trace events to simulate different thread schedules, 
-as long as flow dependencies are preserved.
+as long as value flow dependencies are preserved.
+
 Predictive analysis has been successfully applied to data-race detection, 
@@
 unnecessarily limited.
-%so as to guarantee feasibility 
-%and consequently also soundness.
-In particular, existing techniques (i) do not permit flow dependencies to be violated (except in very specific cases), and (ii) are unable to analyze execution paths beyond that of the input trace.
+
 In this paper, we present \sysname, a predictive race detector that has two
@@
 unexecuted paths that are neighbors to the traced path.
-% both of these restrictions while guaranteeing soundness, as we prove formally. Thanks to its ability to explore unexecuted branches as well as event orderings that violate flow dependencies, 
-Our results show that \sysname\ is able to detect x$2.5$ more data races on a 
-representative set of large-scale benchmarks compared to the state-of-the-art.
+Our results show that \sysname\ is able to detect $2.5X$ as many data races on a 
+representative set of large-scale benchmarks as compared to the state-of-the-art.
 %in predictive race detection.

### 2014-09-06 eval.tex
good shape now

@@
 are fused together at the end.
+
+%Besides, we terminate the monitoring run with a shutdown hook if it is beyond 30 minutes.
 
@@
 applications, we utilized the most lightweight available configuration.
+To further control the cost of our experiments, we monitored each execution for at most 10 minutes.
 
@@
 
-First of all, {\sf Recipe} finds $>$150 more races than {\sf RV}. There 
+First of all, {\sf Recipe} finds $>$250 more races than {\sf RV}. There 
 are several reasons for that: (1) {\sf Recipe} can reason about the 
@@
 as long as the value leads to the same branch decisions. {\sf RV}, however, 
-requires them to read the same values as in the observed run. (3) An 
+requires them to read the same values as in the observed run. (3) A
 critical optimization proposed by {\sf RV} is to preserve dependences 
@@
 the limitations of the underlying constraint solver: (1) the solver cannot 
-compute constraints with very complex arithmatic operations, and (2) the 
+compute constraints with very complex arithmetic operations, and (2) the 
 solver does not support some program constants such as the scientific 
-notation $3E-10$. In our empirical evaluation, we encounted only one 
-benchmark, account, where these issues manifested.
+notation $3E-10$. In our empirical evaluation, we encountered only one 
+benchmark, {\sf account}, where these issues affect our detection capability.
 
@@
 
-%TODO Reorder the columns: RV ReceipeS Recipe
-
 \begin{table*}[htbp]
 \caption{Relaxed Analysis}
-\begin{flushleft}
 \begin{tabular}{|l|r|r|r|r|l|r|r|r|r|r|r|r|r|r|}
@@
  & \multicolumn{ 7}{c|}{Trace} & \multicolumn{ 3}{c|}{Races} & \multicolumn{ 2}{c|}{Difference} & \multicolumn{ 2}{c|}{Running time (sec)} \\ \hline
-Benchmarks & \multicolumn{1}{l|}{Th} & \multicolumn{1}{l|}{Reads} & \multicolumn{1}{l|}{Writes} & \multicolumn{1}{l|}{Base} & Br & \multicolumn{1}{l|}{sync} & \multicolumn{1}{l|}{Total} & \multicolumn{1}{l|}{Recipe-s} & \multicolumn{1}{l|}{Recipe} & \multicolumn{1}{l|}{RV} & \multicolumn{1}{l|}{Diff} & \multicolumn{1}{l|}{Diff'} & \multicolumn{1}{l|}{Recipe} & \multicolumn{1}{l|}{RV} \\ \hline
-critical & 3 & 13 & 7 & 5 & 2/13 & 6 & 78 & 8 & 8 & 8 & \multicolumn{1}{c|}{0} & 0 & 8 & \multicolumn{1}{c|}{2} \\ \hline
+Benchmarks & \multicolumn{1}{l|}{Th} & \multicolumn{1}{l|}{Reads} & \multicolumn{1}{l|}{Writes} & \multicolumn{1}{l|}{Base} & Br & \multicolumn{1}{l|}{sync} & \multicolumn{1}{l|}{Total} & \multicolumn{1}{l|}{RV} & \multicolumn{1}{l|}{Recipe-s} & \multicolumn{1}{l|}{Recipe} & \multicolumn{1}{l|}{Diff} & \multicolumn{1}{l|}{Diff'} & \multicolumn{1}{l|}{Recipe} & \multicolumn{1}{l|}{RV} \\ \hline
+critical & 3 & 13 & 7 & 5 & 2/13 & 6 & 78 & 8 & 8 & 8 & 0 & 0 & 8 & 2 \\ \hline
 airline & 11 & 45 & 15 & 4 & 32/82 & 30 & 317 & 9 & 9 & 9 & 0 & 0 & 490 & 4 \\ \hline
-account & 3 & 46 & 21 & 10 & 3/47 & 6 & 227 & 2 & 5 & 5 & 3 & 3 & 41 & 4 \\ \hline
+account & 3 & 46 & 21 & 10 & 3/47 & 6 & 227 & 5 & 2 & 5 & 3 & 3 & 41 & 4 \\ \hline
 pingpong & 4 & 7 & 7 & 3 & 0/15 & 6 & 111 & 1 & 1 & 1 & 0 & 0 & 19 & 1 \\ \hline
-bbuffer & 4 & 640 & 118 & 10 & 634/1.1K & 217 & 3.3K & 13 & 25 & 9 & 21 & 5 & 62 & 5 \\ \hline
+bbuffer & 4 & 640 & 118 & 10 & 634/1.1K & 217 & 3.3K & 9 & 13 & 25 & 21 & 5 & 62 & 5 \\ \hline
 bubblesort & 26 & 1.3k & 966 & 121 & 155/2.8K & 322 & 8.4K & 7 & 7 & 7 & 0 & 0 & 3295 & 3 \\ \hline
-bufwriter & 5 & 165 & 52 & 75 & 16/130 & 44 & 525 & 4 & 10 & 2 & 8 & 0 & 63 & 9 \\ \hline
-mergesort & 5 & 38 & 33 & 5 & 15/472 & 28 & 1.7K & 3 & 10 & 3 & 10 & 0 & 37 & 5 \\ \hline
-raytracer & 2 & 31 & 5 & 12 & 314/8,2K & 676 & 94.5K & 4 & 6 & 4 & 2 & 0 & 47 & 2 \\ \hline
-montecarlo & 2 & 5 & 86 & 2 & 1.9K/38.2K & 21.1K & 1.9M & 1 & 4 & 1 & 3 & 0 & 1 & 17 \\ \hline
-moldyn & 2 & 605 & 61 & 104 & 19.6K/52.6K & 62 & 203.4K & 6 & 14 & 2 & 12 & 0 & 2842 & 1 \\ \hline
-ftpserver & 28 & 684 & 299 & 71 & 4.4K/233.3K & 78.2K & 3.9M & 99 & 152 & 57 & 108 & 13 & 811 & 153 \\ \hline
-jigsaw & 12 & 525 & 702 & 211 & 63.2K/467.9K & 86.7K & 5.5M & 17 & 23 & 8 & 15 & 0 & 33 & 7 \\ \hline
-sunflow & 9 & \multicolumn{1}{l|}{2.1K} & \multicolumn{1}{l|}{1.3K} & 473 & 201.3K/827.0K & 50K & \multicolumn{1}{l|}{7.1M} & 38 & 78 & 20 & 69 & 11 & 4520 & 22 \\ \hline
-xalan & 9 & \multicolumn{1}{l|}{1.4K} & \multicolumn{1}{l|}{0.9K} & 209 & 15.7K/103.2K & \multicolumn{1}{l|}{190.1K} & \multicolumn{1}{l|}{6.6M} & 2 & 6 & 2 & 4 & 0 & 5317 & 10 \\ \hline
-lusearch & 10 & \multicolumn{1}{l|}{2.3K} & \multicolumn{1}{l|}{0.5K} & 715 & 22.2K/164K & \multicolumn{1}{l|}{93.2K} & \multicolumn{1}{l|}{9.1M} & 27 & 49 & 14 & 38 & 5 & 5430 & 8 \\ \hline
+bufwriter & 5 & 165 & 52 & 75 & 16/130 & 44 & 525 & 2 & 4 & 10 & 8 & 0 & 63 & 9 \\ \hline
+mergesort & 5 & 38 & 33 & 5 & 15/472 & 28 & 1.7K & 3 & 3 & 10 & 7 & 0 & 37 & 5 \\ \hline
+raytracer & 2 & 31 & 5 & 12 & 314/8,2K & 676 & 94.5K & 4 & 4 & 6 & 2 & 0 & 47 & 2 \\ \hline
+montecarlo & 2 & 5 & 86 & 2 & 1.9K/38.2K & 21.1K & 1.9M & 1 & 1 & 4 & 3 & 0 & 1 & 17 \\ \hline
+moldyn & 2 & 605 & 61 & 104 & 19.6K/52.6K & 62 & 203.4K & 2 & 6 & 14 & 12 & 0 & 2842 & 1 \\ \hline
+ftpserver & 28 & 684 & 299 & 71 & 4.4K/233.3K & 78.2K & 3.9M & 57 & 99 & 152 & 108 & 13 & 811 & 153 \\ \hline
+jigsaw & 12 & 525 & 702 & 211 & 63.2K/467.9K & 86.7K & 5.5M & 8 & 17 & 23 & 15 & 0 & 33 & 7 \\ \hline
+sunflow & 9 & 2.1K & 1.3K & 473 & 201.3K/827.0K & 50K & 7.1M & 20 & 38 & 78 & 69 & 11 & 4520 & 22 \\ \hline
+xalan & 9 & 1.4K & 0.9K & 209 & 15.7K/103.2K & 190.1K & 6.6M & 2 & 2 & 6 & 4 & 0 & 5317 & 10 \\ \hline
+lusearch & 10 & 2.3K & 0.5K & 715 & 22.2K/164K & 93.2K & 9.1M & 14 & 27 & 49 & 38 & 3 & 5430 & 8 \\ \hline
 \end{tabular}
-\end{flushleft}
 \label{tab:main}
 \end{table*}
+
+%\begin{table*}[htbp]
+%\caption{Relaxed Analysis}
+%\begin{flushleft}
+%\begin{tabular}{|l|r|r|r|r|l|r|r|r|r|r|r|r|r|r|}
+%\hline
+% & \multicolumn{ 7}{c|}{Trace} & \multicolumn{ 3}{c|}{Races} & \multicolumn{ 2}{c|}{Difference} & \multicolumn{ 2}{c|}{Running time (sec)} \\ \hline
+%Benchmarks & \multicolumn{1}{l|}{Th} & \multicolumn{1}{l|}{Reads} & \multicolumn{1}{l|}{Writes} & \multicolumn{1}{l|}{Base} & Br & \multicolumn{1}{l|}{sync} & \multicolumn{1}{l|}{Total} & \multicolumn{1}{l|}{Recipe-s} & \multicolumn{1}{l|}{Recipe} & \multicolumn{1}{l|}{RV} & \multicolumn{1}{l|}{Diff} & \multicolumn{1}{l|}{Diff'} & \multicolumn{1}{l|}{Recipe} & \multicolumn{1}{l|}{RV} \\ \hline
+%critical & 3 & 13 & 7 & 5 & 2/13 & 6 & 78 & 8 & 8 & 8 & \multicolumn{1}{c|}{0} & 0 & 8 & \multicolumn{1}{c|}{2} \\ \hline
+%airline & 11 & 45 & 15 & 4 & 32/82 & 30 & 317 & 9 & 9 & 9 & 0 & 0 & 490 & 4 \\ \hline
+%account & 3 & 46 & 21 & 10 & 3/47 & 6 & 227 & 2 & 5 & 5 & 3 & 3 & 41 & 4 \\ \hline
+%pingpong & 4 & 7 & 7 & 3 & 0/15 & 6 & 111 & 1 & 1 & 1 & 0 & 0 & 19 & 1 \\ \hline
+%bbuffer & 4 & 640 & 118 & 10 & 634/1.1K & 217 & 3.3K & 13 & 25 & 9 & 21 & 5 & 62 & 5 \\ \hline
+%bubblesort & 26 & 1.3K & 966 & 121 & 155/2.8K & 322 & 8.4K & 7 & 7 & 7 & 0 & 0 & 3295 & 3 \\ \hline
+%bufwriter & 5 & 165 & 52 & 75 & 16/130 & 44 & 525 & 4 & 10 & 2 & 8 & 0 & 63 & 9 \\ \hline
+%mergesort & 5 & 38 & 33 & 5 & 15/472 & 28 & 1.7K & 3 & 10 & 3 & 10 & 0 & 37 & 5 \\ \hline
+%raytracer & 2 & 31 & 5 & 12 & 314/8,2K & 676 & 94.5K & 4 & 6 & 4 & 2 & 0 & 47 & 2 \\ \hline
+%montecarlo & 2 & 5 & 86 & 2 & 1.9K/38.2K & 21.1K & 1.9M & 1 & 4 & 1 & 3 & 0 & 1 & 17 \\ \hline
+%moldyn & 2 & 605 & 61 & 104 & 19.6K/52.6K & 62 & 203.4K & 6 & 14 & 2 & 12 & 0 & 2842 & 1 \\ \hline
+%ftpserver & 28 & 684 & 299 & 71 & 4.4K/233.3K & 78.2K & 3.9M & 99 & 152 & 57 & 108 & 13 & 811 & 153 \\ \hline
+%jigsaw & 12 & 525 & 702 & 211 & 63.2K/467.9K & 86.7K & 5.5M & 17 & 23 & 8 & 15 & 0 & 33 & 7 \\ \hline
+%sunflow & 9 & \multicolumn{1}{l|}{2.1K} & \multicolumn{1}{l|}{1.3K} & 473 & 201.3K/827.0K & 50K & \multicolumn{1}{l|}{7.1M} & 38 & 78 & 20 & 69 & 11 & 4520 & 22 \\ \hline
+%xalan & 9 & \multicolumn{1}{l|}{1.4K} & \multicolumn{1}{l|}{0.9K} & 209 & 15.7K/103.2K & \multicolumn{1}{l|}{190.1K} & \multicolumn{1}{l|}{6.6M} & 2 & 6 & 2 & 4 & 0 & 5317 & 10 \\ \hline
+%lusearch & 10 & \multicolumn{1}{l|}{2.3K} & \multicolumn{1}{l|}{0.5K} & 715 & 22.2K/164K & \multicolumn{1}{l|}{93.2K} & \multicolumn{1}{l|}{9.1M} & 27 & 49 & 14 & 38 & 5 & 5430 & 8 \\ \hline
+%\end{tabular}
+%\end{flushleft}
+%\label{tab:main}
+%\end{table*}
 
@@
 All threads invoke the method {\sf Sorting}, which recursively starts 
-two child threads if there are two more available entries in the 
+two children threads if there are two more available entries in the 
 thread pool (lines 8-11), or starts one child thread if there is only 
@@
  we count the number of bytecode instructions inside each method. If the 
-resulting number exceeds 65525, we avoid instrumenting the method. 
+resulting number exceeds 65535, we avoid instrumenting the method. 
 The consequence is that we will potentially miss races inside a method for 
@@
 
-\tool lacks support for certain operators. In our current 
+\tool\  lacks support for certain operators. In our current 
 implementation, we do not support bit operations such as $\&$ and $<<$, which accounts for most of our misses.

### 2014-09-06 guarantee.tex
good shape now

@@
 First, to the best of our knowledge, none of the existing techniques 
-automatically explore the un-executed branches by relaxing the schedules, whereas \tool enables this.
+automatically explore the un-executed branches by relaxing the schedules, whereas \tool\  enables this.
 Second, assume we disable the exploration of un-executed branches and all 
@@
 
-However, \tool has not reached the maximum bound of sound exploration.
+However, \tool\  has not reached the maximum bound of sound exploration.
 % does not identify the maximal set of traces 

### 2014-09-06 introduction.tex
good shape now

@@
 enlarged search space, our results show that a race detector based on our relaxed predictive analysis 
-can detect XXX times more races than existing predictive analysis, while guaranteeing soundness.
+can detect $2.5X$ as many races as the existing predictive analysis, while guaranteeing soundness.
 
@@
 part that must be preserved) and the relaxed information (i.e., the part that may be mutated) uniformly as part of a single logical representation.
-\item We develop a propotype \sysname. Our results show XXX.
+\item We develop a propotype \sysname.
+Our results on 16 benchmarks show \sysname is significantly more effective than existing predictive analyses.
+We also conducted case studies to confirm the effectiveness.
 \end{itemize}

### 2014-09-06 overview.tex
submitted version

@@
 \begin{table}
+{ 
 	\begin{center}
@@
 	\end{center}
+	}
 	\caption{\label{Ta:syntax}Instruction Types}
-\vspace{-0.15in}
+\vspace{-3em}
 \end{table}
@@
 $$
+{\small 
 \begin{array}{l}
@@
 	\end{array}
+}
 $$
@@
 $$
+{\small 
 \begin{array}{lcl}
@@
 \end{array}
+}
 $$    

### 2014-09-06 related.tex
submitted version

@@
 
-Given a concurrent execution trace $t$, PTA derives new traces to witness 
+Given a concurrent execution trace $\tau$, PTA derives new traces to witness 
 the data races. PTA is founded on the notion of sound causality, as it 
@@
 
-Wang et al.~\cite{chao} pioneere the symbolic predictive analysis, 
+Wang et al.~\cite{chao} pioneer the symbolic predictive analysis, 
 which our approach also belongs to. However, the approach lacks support 
@@
 control-flow information is encoded into the constraint system so as 
-to relax flow dependencies as long as control dependencies are 
-preserved. ExceptionNULL \cite{Farzan:2012} is another example of 
+to relax value flow dependencies as long as the values do not flow into the control-flow event.
+ ExceptionNULL \cite{Farzan:2012} is another example of 
 sound PTA. Its goal is to detect null dereferences rather than data races.
@@
 Compared to \cite{maximal}, it allows some events to read different 
-values if they do not affect the following branch decisions. However, 
+values if the values do not flow into the following branch conditions, as they do not affect the following branch decisions. However, 
 it requires the events that affect the branch decisions to read exactly 
@@
 cannot guarantee maximality. As illustrated in our example and discussed 
-in Section~\ref{sec:guarantee}, our analysis allows higher detection capability
+in Section~\ref{sec:guarantee}, our analysis allows higher detection capability.
 %, i.e., we allow the events to read different values even they flow into the branch events; we allow the exploration of un-executed paths, both of which are not allowed by the above techniques.

### 2014-09-06 abstract.tex
minor

@@
 and memory accesses, to expose bugs, while preserving feasibility. For
-example, it may reoder trace events to simulate different thread schedules, 
+example, the analysis may reorder trace events to simulate different thread schedules, 
 as long as value flow dependencies are preserved.
@@
 although the state-of-the-art techniques are overly restrictive in the 
-transformations they allow such that their analysis coverage is 
-unnecessarily limited.
+transformations they allow, sacrificing coverage in return for a conservative soundness guarantee.
 

### 2014-09-06 abstract.tex
minor

@@
 Predictive analysis has recently emerged as a promising technique for 
-sound detection of multithreading bugs (i.e., without false alarms). 
+sound detection of concurrency bugs (i.e., without false alarms). 
 The idea is to start from an instrumented execution trace, and transform 

### 2014-09-06 eval.tex
submitted version
minor

@@
 applications, we utilized the most lightweight available configuration.
-To further control the cost of our experiments, we monitored each execution for at most 10 minutes.
 
 As the detection capability  depends on the observed run, for fair 
-comparison we monitor the execution once by recording all necessary 
+comparison we monitor the execution once (for a period of 10 minutes) by recording all necessary 
 information required by both approaches, and then apply both techniques 
@@
 \label{tab:main}
+\vspace{-2em}
 \end{table*}
@@
 \caption{Relaxation of Inter-thread dependence}\label{fig:relax1}
+\vspace{-1em}
 \end{figure}
@@
 \caption{Relaxation of Paths}\label{fig:relax2}
+\vspace{-2em}
 \end{figure}
